### 2015-03-25 tex/project.tex
Added proposal tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2014 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[icml2014,epsf,natbib]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{amssymb,amsthm,amsmath}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2014} with
% \usepackage[nohyperref]{icml2014} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage[accepted]{icml2014} 
% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
%\usepackage[accepted]{icml2014}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
%\icmltitlerunning{Project Proposal for CS6780}

\begin{document}

\twocolumn[
\icmltitle{Local Search Approximation Algorithms for Partitional and Hierarchical Clustering}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2014
% package.
\icmlauthor{Alice Paul (ajp336), Calvin Wylie (cjw278), David Lingenbrink (dal299)}{}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boring formatting information, machine learning, ICML}

\vskip 0.3in
]

\section{Introduction}
As the data we have access to grows in variety and size and becomes cheaper to store, we need methods to efficiently learn from this data. 
Partitioning data (clustering) into similar groups (clusters) is one of the key tools to learn about a set of unlabeled data such as a set of customers who have visited a website or a set of handwritten documents.
Clustering in these situations helps us learn underlying structure and classifications of our data, and it presents large data sets in a simplified and compressed view \cite{Jain}.

There are two main branches of clustering: partitional (or flat) clustering and hierarchical clustering. In partitional clustering we want to partition the data into disjoint sets whereas in hierarchical clustering we are interested in finding a nested tree of partitions. Often these problems are solved using heuristic algorithms which have no guarantee on the solution produced but tend to be fast and easy to implement. We will look at how local search approximation algorithms, which do have some guarantee on the constructed solution and are less computationally expensive than other approximation methods, compare in practice to the most used machine learning algorithms both in the quality of the solution and in computation time.

\section{Literature Review}

Clustering algorithms data back to over 50 years ago, and since then there have been many variations introduced \cite{Jain}. However, clustering algorithms are often formulated as NP-hard problems and are inherently hard to solve \cite{Jain}. When comparing algorithms, we need to consider the stability of the solutions, the computation time, and the proposed objective.

\subsection{Partitional Clustering}

For partitional clustering, we are often given a number $k$, representing the number of clusters, and distances between the data, representing a measure of dissimilarity between points, such as Euclidean distance. In addition, many problems require that this distance satisfy the triangle inequality. The objective is then formulated as identifying $k$ clusters so as to minimize some function of the distances.

One popular objective is that of $k$-means in which we want to choose $k$ cluster centers to minimize the sum of squared distances from each point to its cluster center \cite{Jain}. The most used algorithm for this problem is Lloyd's $k$-means algorithm which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}. This algorithm is only guaranteed to reach a local optimum and can be very sensitive to outliers and the initial points chosen \cite{Kanungo}. There also exists a local search algorithm for $k$-means which guarantees to be within a factor of $(9+\varepsilon)$ of the optimal solution that performs well in practice when combined with Lloyd's algorithm \cite{Kanungo}. A relaxation of this problem is fuzzy $c$-means which has the same objective but allows points to belong partially to multiple clusters \cite{ESL}.

Variations on $k$-means attempt to reduce the sensitivity and worst-case behavior. One other objective is to minimize the maximum distance from a point to its cluster center. This problem can be thought of as minimizing the worst case and is called $k$-center clustering. Here, a simple greedy algorithm is guaranteed to come within a factor of $2$ of the optimal solution as long as the given distances satisfy the triangle inequality \cite{Hopcroft}. Furthermore, this is best approximation ratio possible unless P=NP \cite{Shmoys}.  However, $k$-center does not perform well in practice compared to other methods \cite{Murphy}. This may be because it focuses too much on the worst case distance.

An objective function in between $k$-means and $k$-center is to minimize the sum of distances from each point to the median point of its cluster. This problem is called $k$-median or $k$-medoids clustering. Another way of defining this problem is choosing a set of $k$ data points to be centers so as to minimize the sum of distances from each point to their closest center.

There are two common heuristic algorithms for this problem. The first is a local search algorithm called partition around medoids (PAM) which starts with $k$ randomly chosen centers and repeatedly looks for a single point swap to improve the objective function until no such move is possible \cite{ESL}. 
Another common heuristics for this problem (often called just $k$-medoids) is similar to that of $k$-means by assigning each point to their closest center and then greedily updating the center within each cluster \cite{Park}. The algorithm repeats this until the solution becomes stable and is more computationally efficient than PAM \cite{Park}. Additionally, Velmurugan and Santhanam showed that on uniform and normally distributed data the $k$-medoids algorithm was empirically faster than that of $k$-means \cite{Velmurugan}.

If we relax the constraint to have exactly $k$ clusters, then a recent algorithm called affinity propogation has shown lower error and computation time than the algorithms above \cite{Frey}. Instead of specifying a number of clusters, the user inputs ``preferences'' on the data points to be cluster centers. This algorithm relies on passing messages between data points. For each pair of data points $i, k \in X$, ``availability'' $a(i,k)$ represents the evidence for having $i$ choose $k$ to be it's cluster center and ``responsibility'' $r(i,k)$ reflects how well-suited $k$ is to serve $i$. These are messages are then updated iteratively. Once the messages converge (this is not guaranteed) then $\max_{k \in X} a(i,k) + r(i,k)$ will identify the cluster center for $i$, and these decisions will be consistent \cite{Frey}.

The $k$-median problem is also a fundamental problem in approximation algorithms. The problem is formulated as above: choose $k$ cluster centers from a set $X$ so as to minimize the sum of distances from each point to their closest center. The best known approximation algorithm for this problem guarantees a solution within a factor of 2.592 of the optimal solution \cite{Wu}. However, this algorithm uses linear programming and is computationally intensive compared to the local search and greedy algorithms above. On the other hand, it is known that a local search algorithm that considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}. Note that for $p=1$, this is the same as PAM. As far as we know, nobody has implemented this algorithm on clustering data to compare the relative trade-off of solution quality and computation time as $p$ increases.

Additionally, center-based clustering in general can be seen as minimizing a supermodular decreasing function subject to a cardinality constraint, a problem related to maximizing a submodular function that is well-studied in approximation algorithms. Furthermore, the approximation algorithms in this area are constructed using local-search or greedy methods and are therefore computationally tractable on large data. In addition, using supermodular functions means our distance measure does not have to satisfy the triangle inequality, which can be useful when considering non-vector data.

While there exist many other methods for clustering such as spectral clustering and mixture models \cite{Murphy}, we will restrict the comparison of our proposed algorithm to the center-based algorithms above which are widely used and relate in objective function.

\subsection{Hierarchical Clustering}

In hierarchical clustering, on the other hand, we do not specify a number of clusters. Instead, we want to find a hierarchy of clusters so that the clusters at each level are formed by merging clusters at a lower level. This gives a bigger picture about the dissimilarity between different groups of points and can allow the user to choose a ``natural'' level of clustering \cite{ESL}.

There are two main basic methods for hierarchical clustering: agglomerative clustering and divisive clustering. In agglomerative clustering we start with each data point in its own cluster. We then repeatedly merge two clusters to form the next level until the data is in a single cluster.

There are three choices of criteria for choosing which clusters to merge, which reflect the relative benefits of $k$-means, $k$-median, and $k$-center. 
In single linkage clustering (also called nearest neighbor), we measure the distance between two clusters to be the smallest pairwise distance and then merge the two closest clusters. The main potential drawback to this method is that two points within a cluster can be very far apart. Complete linkage clustering on the other hand measures the distance between two clusters to be the largest pairwise distance and then merges the two closest clusters. While points within a single cluster will be quite similar using this metric, points may not belong to their closest cluster. 
Lastly, there is group average linkage which looks at the average pairwise distance between clusters and merges the two closest clusters. This strikes a balance between the closeness and compactness of the two techniques above and gives more reasonable results in practice \cite{ESL}.

The other basic method for clustering is divisive clustering which starts with all points in a single cluster and repeatedly divides a cluster into two parts either using a $k$-means or $k$-medoids algorithm with $k=2$ or by greedily separating the cluster so that all points are closer to their part of separated cluster. The cluster to split is chosen to be the cluster with highest diameter or average dissimilarity. This method has not been studied nearly as much as agglomerative clustering \cite{ESL} but has had relative success in classifying document datasets \cite{Zhao}.

Hierarchical clustering has also been studied in approximation algorithms. In this problem, we need to identify a hierarchy of clusters with their corresponding cluster centers so as to minimize the ratio of the cost of our solution on any level with $k$ clusters to the optimal $k$-median solution. Here, the cost of our solution is given by the sum of the distance from each point to its cluster center (i.e. the $k$-median objective we described above). The hierarchical $k$-median problem was originally introduced by Plaxton who gave a 238.88 approximation algorithm \cite{Plaxton}. This was improved by Lin et al.~\cite{Lin} who gave an incremental approximation algorithm framework. In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to construct a hierarchical clustering by increasing the cost by at most a factor of $20.71$ (or a factor of $10.03$ when randomized). Using the $(3+\frac{2}{p})$-approximation algorithm as the black-box, this translates into a $20.71 (3+\frac{2}{p})$-approximation algorithm (or $10.03 (3+\frac{2}{p})$ when randomized). These ratios can be improved by using a Lagrangian multiplier preserving $2$-approximation algorithm but the local search algorithm performed comparably in experimental tests \cite{Nagarajan}.

Again, several algorithms have been built upon agglomerative clustering to incorporate non-spherical cluster shapes and categorical data such as the CURE, ROCK, and Chameleon algorithms \cite{Guha_CURE, Guha_ROCK, Karypis} but do not relate as much to our proposed algorithm in objective.

\section{Partitional Clustering Approximation Algorithms}

\subsection{Local Search with Multiswaps}
In the $k$-median problem, we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$. The goal is to find a set $S \subseteq X$ such that $|S| \leq k$ so as to minimize 
\[ f(S) = \sum_{ i\in X} \min_{j \in S} d(i,j). \]
Consider a local search algorithm with starts with a random subset $S$ of $k$ points. In any iteration, the algorithm looks for a set $A \subseteq S$ and a set $B \subseteq X -S$ such that $|A| = |B| \leq p$ and
\[ f(S- B \cup A) \geq f(S). \]
If the algorithm finds such a set, it swaps out $B$ from $S$ and adds $A$. If no such set exists, the algorithm outputs $S$. Note that for $p$ independent of $k$ and $n$ and if we ensure that the algorithm makes at least $\delta = \varepsilon/n$ progress in each iteration, this is a polynomial-time algorithm. Furthermore, the local search algorithm outputs $S \subseteq X$ such that $|S| = k$ and
\[ f(S) \leq \left (3 + \frac{2}{p}+\varepsilon \right ) f(O) \]
for all $O \subseteq X$. Therefore, there is a trade-off between the computation time of this algorithm and the guaranteed quality of the solution.

\subsection{Supermodular Function Minimization}
Given a set $X$, a function $f:2^{X} \mapsto \mathbb{R}$ is supermodular, if for all $T \subseteq S \subseteq X$ and $j \notin S$, 
\[ f(T \cup \{j\}) - f(T) \leq f(S \cup \{j\}) - f(S). \]
Furthermore, $f$ is non-increasing if for all $T \subseteq S \subseteq X$, $f(S) \leq f(T)$.

Suppose we have a set of data $X$ of which we want to choose a subset $S$ such that $|S| \leq k$ so as to minimize 
\[ f(S) = \sum_{i \in X} \min_{j \in S} d(i,j) ,\]
where we can think of $d$ as some measure of dissimilarity (for example squared distance or absolute distance). For completeness, we define $f(\emptyset) = \min_{j \in S} f(\{j\})$. 
Note that this generalizes the $k$-median problem and can capture a variant of $k$-means in which we are restricted to choosing data points as centers.

It is clear that $f$ is non-increasing since for all $T \subseteq S \subseteq X$ and $i \in X$, $\min_{j \in S} d(i,j) \leq \min_{j \in T} d(i,j)$. Furthermore, for $j \notin S$, adding $j$ to $T$ will decrease the sum of the distances more than adding $j$ to $S$ so $f$ is supermodular. In general, there is no known approximation algorithm for minimizing a non-increasing supermodular function subject to a cardinality constraint. However, if we can place some properties on the ``steepnesss'' of $f$ then there is a known optimal approximation algorithm \cite{Sviridenko}.

As in \cite{Sviridenko}, we define the \emph{total curvature} $c$ of a non-increasing supermodular function $f$ as 
\[ c = 1 - \min_{j \in X}  \frac{ f(X) - f(X- \{j\})}{ f(\{j\}) - f(\emptyset)}.\]
Then, for every $\varepsilon >0$ and $c \in [0,1)$, there exists a local search algorithm that produces a set $S \subseteq X$ such that $|S| \leq k$ in polynomial time such that 
\[ f(S) \leq \left( 1 + \frac{c}{1-c} \cdot e^{-1} + \frac{1}{1-c} \cdot O(\varepsilon) \right) f(O) \]
for all $O \subseteq X$ such that $|O| \leq k$, with high probability \cite{Sviridenko}. Note that $c$ will vary depending on our dataset so while in some cases we may have a very good approximation ratio it can worsen as our distances become non-smooth (for example distances that do not form a metric).

This algorithm first defines linear and submodular functions on $X$:
\begin{align*}
l(A) &= \sum_{j \in A} f(\{j\}) - f(\emptyset), \\
g(A) & = - l(A) - f(X \setminus A).
\end{align*}
Furthermore, we let $\hat{v}_g = \max_{j \in X} g(\{j\})$, $\hat{v}_l = \max_{j \in X} |l(\{j\})|$, and $\hat{v} = \max(\hat{v}_g , \hat{v}_l)$.

Given a set $A \subseteq X$, we consider a two-step random sampling. We first draw a value $p \in [0,1]$ using the density function $\frac{{e}^p}{e-1}$. Then, we choose a subset $B \in A$ by selecting each $x \in A$ to be in $B$ independently with probability $p$. The expected value of $g(B)$ is given by
\[  h(A) = \sum_{B \subseteq A} g(B) \cdot \int_{0}^1 \frac{e^p}{e-1} \cdot p^{|B|-1} (1-p)^{|A|-|B|} dp. \]
Then, we define a potential function
\[ \psi(A) = (1-e^{-1}) h(A) + l(A). \]

We are now ready to describe how the algorithm works. The algorithm sets $\delta = \frac{\varepsilon}{n}$ to be the minimum step size and starts with a set $S$ of size $k$. Then, while there exists $a \in S$ and $b \in X \setminus S$ such that 
\[ \psi(S -a +b) \geq \psi(S) + \delta, \]
the algorithm removes $a$ from $S$ and adds $b$.

Note that $\psi$ requires summing an exponential number of terms. However, we can estimate $h(A)$ (and therefore $\psi(A)$) by randomly sampling $g(B)$ in the above fashion. If $\hat{\psi}(A)$ is an estimate of $\psi(A)$ sampled from $\Omega (\varepsilon^{-2} n^4 \ln^4 n \ln M)$ samples, then 
\[ Pr[ | \hat{\psi}(A) - \psi(A) | \geq \delta ] = O(M^{-1}). \]
By setting $M$ large enough (polynomially large), the algorithm will converge in polynomial time and will return a solution satisying the above guarantee with high probability \cite{Sviridenko}.

\subsection{Implementation}

\subsection{Results}

\section{Hierarchical Clustering Approximation Algorithms}
Again, suppose we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$. In the hierarchical setting, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$. Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$. We define
\[ \mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i)) . \]

Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if 
\begin{enumerate}
\item $S_1 \subset S_2$, 
\item  $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, and 
\item $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
\end{enumerate}
 A solution to the hierarchical clustering problem is therefore a chain of nested solutions. However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the algorithm.  
Consider any $k$-median algorithm $\mathcal{A}$, such as the algorithms above. After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce  augmentation step that will nest these clusterings so that they fit into the hierarchical framework and only increase the costs by a bounded factor.

First, we set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$). Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$. The cost of these solutions is determined by assigning each point to the nearest cluster center. We then order these solutions according to their cost into buckets of the form $[0,\beta_0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.

From each non-empty bucket, pick the solution with the smallest number of cluster centers and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r = V_n$. We then inductively build up our nested solutions in the following manner. Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_{i+1}$ by iteratively adding the closest point $j \in S_{i-1}$ to $i$ for all $i \in \overline{V}_i$. Note that $|S_{i}| \leq |\overline{V}_i|$.

We then define an assignment $a_i$. For each $j \in X$ such that $a_{i-1}(j) \in S_{i}$, we set $a_i(j) = a_{i-1}(j)$. Otherwise, we set $a_i(j)$ to be the facility $i \in S_i$ such that assigning all points in $a_{i-1}^{-1}(j)$ (i.e. all points in $j$'s cluster in $S_{i-1}$) would have minimum cost. Using this assignment, $(a_i, S_i)$ is nested in $(a_{i-1}, S_{i-1})$.

We continue this nesting until $|S_i| = 1$ and we have reached the top level of the hierarchy. After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantee that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq n-k+1 \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).

We can test this hierarchical framework for both the multiswap and supermodular local search algorithms above.

\subsection{Implementation}

\subsection{Results}

\
\label{submission}

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite

\bibliography{projectBib}
\bibliographystyle{icml2014}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was
% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
% slightly modified from the 2009 version by Kiri Wagstaff and 
% Sam Roweis's 2008 version, which is slightly modified from 
% Prasad Tadepalli's 2007 version which is a lightly 
% changed version of the previous year's version by Andrew Moore, 
% which was in turn edited from those of Kristian Kersting and 
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2015-03-26 src/kmeans-demo.py
Added kmeans-demo.py and requirments.txt

@@ -1,0 +1,99 @@
+from time import time
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn import metrics
+from sklearn.cluster import KMeans
+from sklearn.datasets import load_digits
+from sklearn.decomposition import PCA
+from sklearn.preprocessing import scale
+
+np.random.seed(42)
+
+digits = load_digits()
+data = scale(digits.data)
+
+n_samples, n_features = data.shape
+n_digits = len(np.unique(digits.target))
+labels = digits.target
+
+sample_size = 300
+
+print("n_digits: %d, \t n_samples %d, \t n_features %d"
+      % (n_digits, n_samples, n_features))
+
+
+print(79 * '_')
+print('% 9s' % 'init'
+      '    time  inertia    homo   compl  v-meas     ARI AMI  silhouette')
+
+
+def bench_k_means(estimator, name, data):
+    t0 = time()
+    estimator.fit(data)
+    print('% 9s   %.2fs    %i   %.3f   %.3f   %.3f   %.3f   %.3f    %.3f'
+          % (name, (time() - t0), estimator.inertia_,
+             metrics.homogeneity_score(labels, estimator.labels_),
+             metrics.completeness_score(labels, estimator.labels_),
+             metrics.v_measure_score(labels, estimator.labels_),
+             metrics.adjusted_rand_score(labels, estimator.labels_),
+             metrics.adjusted_mutual_info_score(labels,  estimator.labels_),
+             metrics.silhouette_score(data, estimator.labels_,
+                                      metric='euclidean',
+                                      sample_size=sample_size)))
+
+bench_k_means(KMeans(init='k-means++', n_clusters=n_digits, n_init=10),
+              name="k-means++", data=data)
+
+bench_k_means(KMeans(init='random', n_clusters=n_digits, n_init=10),
+              name="random", data=data)
+
+# in this case the seeding of the centers is deterministic, hence we run the
+# kmeans algorithm only once with n_init=1
+pca = PCA(n_components=n_digits).fit(data)
+bench_k_means(KMeans(init=pca.components_, n_clusters=n_digits, n_init=1),
+              name="PCA-based",
+              data=data)
+print(79 * '_')
+
+###############################################################################
[…]

### 2015-03-27 src/affinity-propogation-demo.py
Added affinity propogation demo

@@ -1,0 +1,62 @@
+"""
+=================================================
+Demo of affinity propagation clustering algorithm
+=================================================
+Reference:
+Brendan J. Frey and Delbert Dueck, "Clustering by Passing Messages
+Between Data Points", Science Feb. 2007
+"""
+
+from sklearn.cluster import AffinityPropagation
+from sklearn import metrics
+from sklearn.datasets.samples_generator import make_blobs
+from sklearn.datasets import load_digits
+from sklearn.preprocessing import scale
+
+##############################################################################
+# Generate sample data
+centers = [[1, 1], [-1, -1], [1, -1]]
+X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
+                            random_state=999)
+
+##############################################################################
+# Compute Affinity Propagation
+af = AffinityPropagation(preference=-50).fit(X)
+cluster_centers_indices = af.cluster_centers_indices_
+labels = af.labels_
+
+n_clusters_ = len(cluster_centers_indices)
+
+print('Estimated number of clusters: %d' % n_clusters_)
+print("Homogeneity: %0.3f" % metrics.homogeneity_score(labels_true, labels))
+print("Completeness: %0.3f" % metrics.completeness_score(labels_true, labels))
+print("V-measure: %0.3f" % metrics.v_measure_score(labels_true, labels))
+print("Adjusted Rand Index: %0.3f"
+      % metrics.adjusted_rand_score(labels_true, labels))
+print("Adjusted Mutual Information: %0.3f"
+      % metrics.adjusted_mutual_info_score(labels_true, labels))
+print("Silhouette Coefficient: %0.3f"
+      % metrics.silhouette_score(X, labels, metric='sqeuclidean'))
+
+##############################################################################
+# Plot result
+import matplotlib.pyplot as plt
+from itertools import cycle
+
+plt.close('all')
+plt.figure(1)
+plt.clf()
+
+colors = cycle('bgrcmykbgrcmykbgrcmykbgrcmyk')
+for k, col in zip(range(n_clusters_), colors):
+    class_members = labels == k
+    cluster_center = X[cluster_centers_indices[k]]
+    plt.plot(X[class_members, 0], X[class_members, 1], col + '.')
+    plt.plot(cluster_center[0], cluster_center[1], 'o', markerfacecolor=col,
+             markeredgecolor='k', markersize=14)
+    for x in X[class_members]:
[…]

### 2015-03-27 src/kmeans-demo.py
Modified kmeans-demo.py to use make_blobs random data

@@ -7,20 +7,26 @@ from sklearn.cluster import KMeans
 from sklearn.datasets import load_digits
 from sklearn.decomposition import PCA
 from sklearn.preprocessing import scale
+from sklearn.datasets.samples_generator import make_blobs
 
 np.random.seed(42)
 
-digits = load_digits()
-data = scale(digits.data)
+# digits = load_digits()
+# data = scale(digits.data)
 
-n_samples, n_features = data.shape
-n_digits = len(np.unique(digits.target))
-labels = digits.target
+# n_samples, n_features = data.shape
+# n_digits = len(np.unique(digits.target))
+# labels = digits.target
+# Generate sample data
+centers = [[1, 1], [-1, -1], [1, -1]]
+data, labels = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
+                            random_state=999)
+n_digits = 3
 
 sample_size = 300
 
-print("n_digits: %d, \t n_samples %d, \t n_features %d"
-      % (n_digits, n_samples, n_features))
+# print("n_digits: %d, \t n_samples %d, \t n_features %d"
+#       % (n_digits, n_samples, n_features))
 
 
 print(79 * '_')
@@ -50,16 +56,17 @@ bench_k_means(KMeans(init='random', n_clusters=n_digits, n_init=10),
 
 # in this case the seeding of the centers is deterministic, hence we run the
 # kmeans algorithm only once with n_init=1
-pca = PCA(n_components=n_digits).fit(data)
-bench_k_means(KMeans(init=pca.components_, n_clusters=n_digits, n_init=1),
-              name="PCA-based",
-              data=data)
+# pca = PCA(n_components=n_digits).fit(data)
+# bench_k_means(KMeans(init=pca.components_, n_clusters=n_digits, n_init=1),
+#               name="PCA-based",
+#               data=data)
 print(79 * '_')
 
 ###############################################################################
 # Visualize the results on PCA-reduced data
 
-reduced_data = PCA(n_components=2).fit_transform(data)
+# reduced_data = PCA(n_components=2).fit_transform(data)
+reduced_data = data
 kmeans = KMeans(init='k-means++', n_clusters=n_digits, n_init=10)
 kmeans.fit(reduced_data)
 
@@ -67,8 +74,8 @@ kmeans.fit(reduced_data)
[…]

### 2015-04-06 tex/project.tex
More background and updated bib

@@
 
-If we relax the constraint to have exactly $k$ clusters, then a recent algorithm called affinity propogation has shown lower error and computation time than the algorithms above \cite{Frey}.
+More recently, spectral based methods have become popular for clustering.
+These algorithms construct a graph based on the similarity matrix and compute the eigenvalues of the normalized graph Laplacian matrix \cite{Luxburg}.
+These eigenvectors are either used to split the current cluster in half using the second eigenvector (continuing recursively until there are $k$ clusters) or the eigenvectors are used to project the data points into a lower dimensional space on which the $k$-means algorithm above is run \cite{Verma}.
+We will base our comparisons against the algorithm introduced by \cite{Meila} which uses the latter method and performed best in practice on several datasets \cite{Verma}.
+
+If we relax the constraint to have exactly $k$ clusters, then a recent algorithm called affinity propagation has shown lower error and computation time than the algorithms above \cite{Frey}.
 Instead of specifying a number of clusters, the user inputs ``preferences'' on the data points to be cluster centers.
@@
 
-While there exist many other methods for clustering such as spectral clustering and mixture models \cite{Murphy}, we will restrict the comparison of our proposed algorithm to the center-based algorithms above which are widely used and relate in objective function.
+There exist many other methods for clustering.
+Methods such as MeanShift and DBSCAN focus on finding clusters of high density \cite{Comaniciu, Ester}.
+We choose not to compare against density based methods since their objective differs widely from the center based methods above.
+In addition, we will focus on test data in which center-based methods seem appropriate.
+We also choose not to compare our proposed methods against Bayesian clustering methods such as Gaussian mixture models since these methods require many more parameters (representing the prior information).
+In addition, the $k$-means algorithm above can be thought of as a special case of the Gaussian mixture model when all the covariances are equal \cite{ESL}.
+Therefore, this algorithm can serve as a basis of comparison when both algorithms have the same available information.
 
@@
 
-There are three choices of criteria for choosing which clusters to merge, which reflect the relative benefits of $k$-means, $k$-median, and $k$-center. 
+There are four choices of criteria for choosing which clusters to merge. 
 In single linkage clustering (also called nearest neighbor), we measure the distance between two clusters to be the smallest pairwise distance and then merge the two closest clusters.
@@
 While points within a single cluster will be quite similar using this metric, points may not belong to their closest cluster.
-Lastly, there is group average linkage which looks at the average pairwise distance between clusters and merges the two closest clusters.
+A method between these two extremes is group average linkage which looks at the average pairwise distance between clusters and merges the two closest clusters.
 This strikes a balance between the closeness and compactness of the two techniques above and gives more reasonable results in practice \cite{ESL}.
+Lastly, there is Ward's method which merges the two clusters so as to minimize the sum of each cluster's variance \cite{Ward}.
+All four of these methods can be implemented in the Lance-Williams recursive framework which updates the cluster distances efficiently \cite{Lance}.
 
@@
 Again, several algorithms have been built upon agglomerative clustering to incorporate non-spherical cluster shapes and categorical data such as the CURE, ROCK, and Chameleon algorithms \cite{Guha_CURE, Guha_ROCK, Karypis} but do not relate as much to our proposed algorithm in objective.
+We also choose not compare our proposed method with Bayesian hierarchical methods for the same reasons as above.
 
@@
 
+We will compare the above algorithms against $k$-means, $k$-medoids, PAM, affinity propagation, and Shi and Meila's spectral algorithm.
+All these methods except $k$-medoids and PAM are implemented in python's scikit-learn package.
+For comparisons against affinity propagation, we will first observe how many clusters the algorithm outputs and use that as input for our proposed algorithms.
+We implement the above local search algorithm for $p=2, 3,4$ so the computation time is reasonable.
+
+Given a true clustering, we will use the adjusted mutual information (AMI) measure to score the accuracy of each method.
+AMI measures the agreement of two assignments and is normalized to reflect chance so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation \cite{Vinh}.
+For each data set, we will also report the $k$-median objective function value for that clustering, the guaranteed approximation ratio, and the computation time.
+
+The data sets we chose to use for comparison were
+
+NEED TO DECIDE:
+OR library $p$-median.
+
 \subsection{Results}

### 2015-04-14 src/kmedoids.py
multiswaps.py, kmedoids.py

@@ -1,0 +1,60 @@
+#!/usr/bin/env python
+
+import numpy as np
+import random
+
+from sklearn.datasets.samples_generator import make_blobs
+from sklearn.metrics.pairwise import pairwise_distances
+
+def cluster(distances, k=3):
+
+    m = distances.shape[0] # number of points
+
+    # Pick k random medoids.
+    curr_medoids = np.array([-1]*k)
+    while not len(np.unique(curr_medoids)) == k:
+        curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
+    old_medoids = np.array([-1]*k) # Doesn't matter what we initialize these to.
+    new_medoids = np.array([-1]*k)
+   
+    # Until the medoids stop updating, do the following:
+    while not ((old_medoids == curr_medoids).all()):
+        # Assign each point to cluster with closest medoid.
+        clusters = assign_points_to_clusters(curr_medoids, distances)
+
+        # Update cluster medoids to be lowest cost point. 
+        for curr_medoid in curr_medoids:
+            cluster = np.where(clusters == curr_medoid)[0]
+            new_medoids[curr_medoids == curr_medoid] = compute_new_medoid(cluster, distances)
+
+        old_medoids[:] = curr_medoids[:]
+        curr_medoids[:] = new_medoids[:]
+
+    return clusters, curr_medoids
+
+def assign_points_to_clusters(medoids, distances):
+    distances_to_medoids = distances[:,medoids]
+    clusters = medoids[np.argmin(distances_to_medoids, axis=1)]
+    clusters[medoids] = medoids
+    return clusters
+
+def compute_new_medoid(cluster, distances):
+    mask = np.ones(distances.shape)
+    mask[np.ix_(cluster,cluster)] = 0.
+    cluster_distances = np.ma.masked_array(data=distances, mask=mask, fill_value=10e9)
+    costs = cluster_distances.sum(axis=1)
+    return costs.argmin(axis=0, fill_value=10e9)
+
+if __name__ == '__main__':
+    centers = [[1, 1], [-1, -1], [1, -1]]
+    X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
+                                random_state=999)
+    distances = pairwise_distances(X)
+
+    clusters, curr_medoids = cluster(distances, k=3)
+    print(curr_medoids)
+
+
+
+
+

### 2015-04-14 src/multiswaps.py

@@ -1,0 +1,64 @@
+import random
+import itertools
+
+import numpy as np
+
+from sklearn.datasets.samples_generator import make_blobs
+from sklearn.metrics.pairwise import pairwise_distances
+
+import matplotlib.pyplot as plt
+from itertools import cycle
+
+num_clusters = 3
+p = 2
+epsilon = 5
+
+def find_closest_median(i, medoids, distances):
+    return medoids[np.argmin(distances[i, medoids])]
+
+def objective(medoids, distances):
+    n = distances.shape[0]
+    s = 0
+    for i in range(0, n):
+        s += distances[i, find_closest_median(i, medoids, distances)]
+    return s
+
+def swap(medoids, A, B):
+    tmp = medoids
+    for i, j in zip(A, B):
+        tmp = swap_single(tmp, i, j)
+    return tmp
+
+def swap_single(medoids, a, b):
+    return [b if i == a else i for i in medoids]
+
+def search_for_swap(medoids, distances):
+    prev_objective = objective(medoids, distances)
+    improvement = False
+    for A in itertools.combinations(medoids, p):
+        for B in itertools.combinations(range(n), p):
+            new_medoids = swap(medoids, A, B)
+            if objective(new_medoids, distances) < prev_objective - epsilon:
+                improvement = True
+                print("Improvement: %f" % (prev_objective - objective(new_medoids, distances)))
+                return new_medoids, improvement
+    return medoids, improvement
+
+if __name__ == '__main__':
+    centers = [[1, 1], [-1, -1], [1, -1]]
+    X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
+                                random_state=999)
+    distances = pairwise_distances(X)
+    n = distances.shape[0]
+    
+    #Choose random medoids
+    # medoids = random.sample(range(n), num_clusters)
+
+    medoids = [0, 1, 2]
+    print(medoids, objective(medoids, distances))
+
+    improvement = True
+    while improvement:
+        medoids, improvement = search_for_swap(medoids, distances)
+
+    print(medoids, objective(medoids, distances))

### 2015-04-17 tex/project.tex
hierarchical cluster approximation alg

@@
 There also exists a local search algorithm for $k$-means which guarantees to be within a factor of $(9+\varepsilon)$ of the optimal solution that performs well in practice when combined with Lloyd's algorithm \cite{Kanungo}.
+This algorithm is a variant of the multiswap method proposed below but with respect to a $k$-means objective function and only implemented for $p=1$.
 A relaxation of this problem is fuzzy $c$-means which has the same objective but allows points to belong partially to multiple clusters \cite{ESL}.
@@
 This problem can be thought of as minimizing the worst case and is called $k$-center clustering.
-Here, a simple greedy algorithm is guaranteed to come within a factor of $2$ of the optimal solution as long as the given distances satisfy the triangle inequality \cite{Hopcroft}.
+Here, a simple greedy algorithm is guaranteed to come within a factor of $2$ of the optimal solution as long as the given distances satisfy the triangle inequality \cite{Gonzalez}.
 Furthermore, this is best approximation ratio possible unless P=NP \cite{Shmoys}.
@@
 The first is a local search algorithm called partition around medoids (PAM) which starts with $k$ randomly chosen centers and repeatedly looks for a single point swap to improve the objective function until no such move is possible \cite{ESL}.
-Another common heuristics for this problem (often called just $k$-medoids) is similar to that of $k$-means by assigning each point to their closest center and then greedily updating the center within each cluster \cite{Park}.
+Another common heuristic for this problem (often called just $k$-medoids) is similar to that of $k$-means by assigning each point to their closest center and then greedily updating the center within each cluster \cite{Park}.
 The algorithm repeats this until the solution becomes stable and is more computationally efficient than PAM \cite{Park}.
@@
 This algorithm relies on passing messages between data points.
-For each pair of data points $i, k \in X$, ``availability'' $a(i,k)$ represents the evidence for having $i$ choose $k$ to be it's cluster center and ``responsibility'' $r(i,k)$ reflects how well-suited $k$ is to serve $i$.
+For each pair of data points $i, j \in X$, ``availability'' $a(i,j)$ represents the evidence for having $i$ choose $j$ to be its cluster center and ``responsibility'' $r(i,j)$ reflects how well-suited $j$ is to serve $i$.
 These are messages are then updated iteratively.
-Once the messages converge (this is not guaranteed) then $\max_{k \in X} a(i,k) + r(i,k)$ will identify the cluster center for $i$, and these decisions will be consistent \cite{Frey}.
+Once the messages converge (this is not guaranteed) then $\max_{j \in X} (a(i,j) + r(i,j))$ will identify the cluster center for $i$, and these decisions will be consistent \cite{Frey}.
 
@@
 Methods such as MeanShift and DBSCAN focus on finding clusters of high density \cite{Comaniciu, Ester}.
-We choose not to compare against density based methods since their objective differs widely from the center based methods above.
+We choose not to compare against densit- based methods since their objective differs widely from the center-based methods above.
 In addition, we will focus on test data in which center-based methods seem appropriate.
@@
 There are two main basic methods for hierarchical clustering: agglomerative clustering and divisive clustering.
-In agglomerative clustering we start with each data point in its own cluster.
-We then repeatedly merge two clusters to form the next level until the data is in a single cluster.
+Agglomerative clustering starts with each data point in its own cluster.
+It then repeatedly merge two clusters to form the next level until the data is in a single cluster.
+On the other hand, divisive clustering starts with all points in a single cluster and repeatedly divides a cluster into two parts either using a $k$-means or $k$-medoids algorithm with $k=2$ or by greedily separating the cluster so that all points are closer to their part of separated cluster.
+The cluster to split is chosen to be the cluster with highest diameter or average dissimilarity.
+This method has not been studied nearly as much as agglomerative clustering \cite{ESL} but has had relative success in classifying document datasets \cite{Zhao}.
 
@@
 In single linkage clustering (also called nearest neighbor), we measure the distance between two clusters to be the smallest pairwise distance and then merge the two closest clusters.
-The main potential drawback to this method is that two points within a cluster can be very far apart.
+The main drawback to this method is that two points within a cluster can be very far apart.
 Complete linkage clustering on the other hand measures the distance between two clusters to be the largest pairwise distance and then merges the two closest clusters.
@@
 All four of these methods can be implemented in the Lance-Williams recursive framework which updates the cluster distances efficiently \cite{Lance}.
-
-The other basic method for clustering is divisive clustering which starts with all points in a single cluster and repeatedly divides a cluster into two parts either using a $k$-means or $k$-medoids algorithm with $k=2$ or by greedily separating the cluster so that all points are closer to their part of separated cluster.
-The cluster to split is chosen to be the cluster with highest diameter or average dissimilarity.
-This method has not been studied nearly as much as agglomerative clustering \cite{ESL} but has had relative success in classifying document datasets \cite{Zhao}.
 
@@
 We are now ready to describe how the algorithm works.
-The algorithm sets $\delta = \frac{\varepsilon}{n}$ to be the minimum step size and starts with a set $S$ of size $k$.
+The algorithm sets $\delta = \frac{\varepsilon}{n}$ to be the minimum step size and starts with a set $S$ of size $n-k$.
 Then, while there exists $a \in S$ and $b \in X \setminus S$ such that
 \[ \psi(S -a +b) \geq \psi(S) + \delta, \]
-the algorithm removes $a$ from $S$ and adds $b$.
+the algorithm removes $a$ from $S$ and adds $b$. If there is no swap that increases $\psi$, then the algorithm outputs $X \setminus S$.
 

### 2015-04-17 src/hierarchical-approx.py

@@ -1,0 +1,121 @@
+import numpy as np
+from sklearn import metrics
+import kmedoids
+import math
+
+from sklearn.datasets.samples_generator import make_blobs
+from sklearn.metrics.pairwise import pairwise_distances
+
+# Calls k-median algorithm
+def k_cluster_alg(k, distances, alg = "k-medoids"):
+    if alg == "k-medoids":
+        clusters, centers = kmedoids.cluster(distances, k)
+        distances_to_centers = distances[:,centers]
+        cost = sum( np.min(distances_to_centers, axis=1))
+              
+    return clusters, centers, cost
+
+# Finds the solution with cost in bucket [lb, ub] with the fewest
+# cluster centers
+def find_next_solution(k, lb, ub, distances, alg):
+    clusters = []
+    centers = []
+    
+    while True:
+        k -= 1
+        n_clusters, n_centers, n_cost = k_cluster_alg(k, distances, alg)
+        if n_cost > ub:
+            break
+        else:
+            clusters = n_clusters
+            centers = n_centers
+            
+    return centers, clusters
+
+def replacement_center(centers, points, distances):
+    i = np.argmin([ sum(distances[points, p]) for p in centers])
+    return centers[i]
+    
+
+# Nests the new cluster centers with the old ones
+def nest(curr_centers, curr_clusters, next_centers, distances, Z):
+    n = len(curr_clusters)
+    
+    closest_centers_i = np.argmin(distances[next_centers,:][:,curr_centers], 1)
+    closest_centers = [curr_centers[i] for i in closest_centers_i]
+    new_centers = list(set(closest_centers))
+
+    rem_centers = [x for x in curr_centers if x not in new_centers]
+    new_clusters = curr_clusters
+    k = max(curr_clusters)
+    
+    for y in rem_centers:
+        k+=1
+        # Find points to merge
+        c1 = curr_clusters[y]
+        points1 = [i for i in range(n) if curr_clusters[i] == c1]
+        
+        x = replacement_center(new_centers, points1, distances)
+        c2 = curr_clusters[x]
+        points2 = [i for i in range(n) if curr_clusters[i] == c2]
+
+        # Relabel with k
+        for p in points1:
+            curr_clusters[p] = k
+        for p in points2:
+            curr_clusters[p] = k
+        Z[k-n] = [c1, c2]
+        
+    return new_centers, new_clusters, Z
+        
+    
+    
+def cluster(distances, alg = "k-medoids", random = False):
[…]

### 2015-04-17 src/hierarchical-clustering-demo.py

@@ -1,0 +1,54 @@
+import numpy as np
+import scipy.cluster.hierarchy as hac
+
+
+a = np.array([[0.1,   2.5],
+              [1.5,   .4 ],
+              [0.3,   1  ],
+              [1  ,   .8 ],
+              [0.5,   0  ],
+              [0  ,   0.5],
+              [0.5,   0.5],
+              [2.7,   2  ],
+              [2.2,   3.1],
+              [3  ,   2  ],
+              [3.2,   1.3]])
+
+fig, axes23 = plt.subplots(2, 3)
+
+for method, axes in zip(['ward', 'average'], axes23):
+    z = hac.linkage(a, method=method)
+    den = hac.dendrogram(z)
+
+    # Plotting
+    axes[0].plot(range(1, len(z)+1), z[::-1, 2])
+    knee = np.diff(z[::-1, 2], 2)
+    axes[0].plot(range(2, len(z)), knee)
+
+    num_clust1 = knee.argmax() + 2
+    knee[knee.argmax()] = 0
+    num_clust2 = knee.argmax() + 2
+
+    axes[0].text(num_clust1, z[::-1, 2][num_clust1-1], 'possible\n<- knee point')
+
+    part1 = hac.fcluster(z, num_clust1, 'maxclust')
+    part2 = hac.fcluster(z, num_clust2, 'maxclust')
+
+    clr = ['#2200CC' ,'#D9007E' ,'#FF6600' ,'#FFCC00' ,'#ACE600' ,'#0099CC' ,
+    '#8900CC' ,'#FF0000' ,'#FF9900' ,'#FFFF00' ,'#00CC01' ,'#0055CC']
+
+    for part, ax in zip([part1, part2], axes[1:]):
+        for cluster in set(part):
+            ax.scatter(a[part == cluster, 0], a[part == cluster, 1], 
+                       color=clr[cluster])
+
+    m = '\n(method: {})'.format(method)
+    plt.setp(axes[0], title='Screeplot{}'.format(m), xlabel='partition',
+             ylabel='{}\ncluster distance'.format(m))
+    plt.setp(axes[1], title='{} Clusters'.format(num_clust1))
+    plt.setp(axes[2], title='{} Clusters'.format(num_clust2))
+
+plt.tight_layout()
+plt.show()
+
+

### 2015-04-17 src/opt_solns.py
Gurobi code

@@ -1,0 +1,29 @@
+import numpy as np
+from gurobipy import *
+
+    model = Model("k-median")
+    n = np.shape(distances)[0]
+    y,x = {}, {}
+    
+    for j in range(n):
+        y[j] = model.addVar(obj=0, vtype="B", name="y[%s]"%j)
+        for i in range(n):
+    model.update()
+    
+    for i in range(n):
+        model.addConstr(LinExpr(coef,var), "=", 1, name="Assign[%s]"%i)
+        
+    for j in range(n):
+        for i in range(n):
+            model.addConstr(x[i,j], "<", y[j], name="Strong[%s,%s]"%(i,j))
+            
+    model.addConstr(LinExpr(coef,var), "=", rhs=k, name="k_median")
+    
+    model.update()
+    model.__data = x,y
+    model.optimize()
+
+    if model.status == GRB.status.OPTIMAL:
+    else:
+        return 0
+

### 2015-04-17 src/opt_solns.py
Gurobi code

@@ -1,6 +1,7 @@
 import numpy as np
 from gurobipy import *
 
+def kmedian_opt(distances, k):
     model = Model("k-median")
     n = np.shape(distances)[0]
     y,x = {}, {}
@@ -8,15 +9,20 @@ from gurobipy import *
     for j in range(n):
         y[j] = model.addVar(obj=0, vtype="B", name="y[%s]"%j)
         for i in range(n):
+            x[i,j] = model.addVar(obj=distances[i,j], vtype="B", name="x[%s,%s]"%(i,j))
     model.update()
     
     for i in range(n):
+        coef = [1 for j in range(n)]
+        var = [x[i,j] for j in range(n)]
         model.addConstr(LinExpr(coef,var), "=", 1, name="Assign[%s]"%i)
         
     for j in range(n):
         for i in range(n):
             model.addConstr(x[i,j], "<", y[j], name="Strong[%s,%s]"%(i,j))
             
+    coef = [1 for j in range(n)]
+    var = [y[j] for j in range(n)]
     model.addConstr(LinExpr(coef,var), "=", rhs=k, name="k_median")
     
     model.update()
@@ -24,6 +30,6 @@ from gurobipy import *
     model.optimize()
 
     if model.status == GRB.status.OPTIMAL:
+        return model.objVal
     else:
         return 0
-

### 2015-04-17 src/submodular-demo.py
finishing submodular

@@ -18,7 +18,7 @@ import random
 
 np.random.seed(17)
 
-H_PERC = 10000
+H_PERC = 100
 
 EPSILON =80
 N = 20
@@ -66,7 +66,6 @@ def without(S, x):
 def curvatureValue(X, j, fX, fEmpty):
   num = (fX - f(without(X,j),X))
   denom = (f(frozenset([j]),X)-fEmpty)
-  print num, denom
   if denom == 0: 
     return float("inf")
   else:
@@ -77,7 +76,6 @@ def totalCurvature(X):
   #expects X to be a frozenset
   fX = f(X,X)
   fEmpty = f(frozenset(),X)
-  print fX, fEmpty
   vals = [curvatureValue(X,j, fX, fEmpty) for j in X]
   return 1- min(vals)
 
@@ -147,7 +145,7 @@ def supermodular(X):
   while True:
     newS = updateS(S,Sc,X)
     if newS == S:
-      print "Same value!"
+      print "Finished with f(S) = ", f(Sc, X)
       return Sc
     else:
       S = newS
@@ -155,7 +153,7 @@ def supermodular(X):
 
 print supermodular(toFrozenSet(X))
 
-
+print "curvature is ", totalCurvature(toFrozenSet(X))
 
 
 

### 2015-04-19 src/LoadpMedian.py
Updated Code
Fixed bugs to some algs

@@ -1,0 +1,33 @@
+import numpy as np
+import networkx as nx
+
+def LoadpMedian(filepath):
+    f = open(filepath)
+    dims = f.readline().split()
+    n = int(dims[0])
+    m = int(dims[1])
+    k = int(dims[2])
+
+    G = nx.Graph()
+    G.add_nodes_from([i+1 for i in range(n)])
+
+    for j in range(m):
+        line = f.readline().split()
+        u = int(line[0])
+        v = int(line[1])
+        w = int(line[2])
+        G.add_edge(u, v, weight = w)
+
+    DistDict = nx.floyd_warshall(G)
+    Distances = np.zeros((n,n))
+
+    for i in range(n):
+        for j in range(n):
+            d = DistDict[i+1][j+1]
+            Distances[i][j] = d
+            Distances[j][i] = d
+
+    return Distances, n, k
+    
+
+        

### 2015-04-19 src/hierarchical-approx.py

@@ -1,36 +1,53 @@
 import numpy as np
 from sklearn import metrics
 import kmedoids
+import multiswaps
 import math
 
 from sklearn.datasets.samples_generator import make_blobs
 from sklearn.metrics.pairwise import pairwise_distances
 
 # Calls k-median algorithm
-def k_cluster_alg(k, distances, alg = "k-medoids"):
-    if alg == "k-medoids":
+def k_cluster_alg(k, distances, alg = ["k-medoids",None]):
+    if alg[0] == "multiswaps":
+        centers, cost = multiswaps.cluster(distances, k, alg[1])
+    else:
         clusters, centers = kmedoids.cluster(distances, k)
         distances_to_centers = distances[:,centers]
         cost = sum( np.min(distances_to_centers, axis=1))
               
-    return clusters, centers, cost
+    return centers, cost
 
-# Finds the solution with cost in bucket [lb, ub] with the fewest
+# Finds the solutions with costs in buckets [lb, ub] with the fewest
 # cluster centers
-def find_next_solution(k, lb, ub, distances, alg):
-    clusters = []
-    centers = []
+def bucketed_solns(distances, beta, beta_0, alg):
+    n = distances.shape[0]
+    all_centers = [[] for i in range(n)]
+    all_costs = [0 for i in range(n)]
     
-    while True:
-        k -= 1
-        n_clusters, n_centers, n_cost = k_cluster_alg(k, distances, alg)
-        if n_cost > ub:
-            break
+    for k in range(1, n):
+        k_centers, k_cost = k_cluster_alg(k, distances, alg)
+        if k != 1 and k_cost > all_costs[k-1]:
+            all_centers[k] = all_centers[k-1]
+            all_costs[k] = all_costs[k-1]
         else:
-            clusters = n_clusters
-            centers = n_centers
-            
-    return centers, clusters
+            all_centers[k] = k_centers
+            all_costs[k] = k_cost
+    
+    bucketed_centers = []
+    lb = 0
+    ub = beta_0
+    k = n
+    while k>1:
+        next_bucket = [i for i in range(1,n) if all_costs[i]<= ub and \
+                       all_costs[i] > lb]
+        if len(next_bucket) > 0 and len(next_bucket) != n:
+            k = min(next_bucket)
+            bucketed_centers.append(all_centers[k])
+        lb = ub
+        ub = ub*beta
+
+    return bucketed_centers
 
 def replacement_center(centers, points, distances):
[…]

### 2015-04-19 src/kmedoids.py

@@ -7,15 +7,16 @@ from sklearn.datasets.samples_generator import make_blobs
 from sklearn.metrics.pairwise import pairwise_distances
 
 def cluster(distances, k=3):
-
     m = distances.shape[0] # number of points
 
     # Pick k random medoids.
-    curr_medoids = np.array([-1]*k)
+    curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
     while not len(np.unique(curr_medoids)) == k:
         curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
+
     old_medoids = np.array([-1]*k) # Doesn't matter what we initialize these to.
     new_medoids = np.array([-1]*k)
+    clusters = assign_points_to_clusters(curr_medoids, distances)
    
     # Until the medoids stop updating, do the following:
     while not ((old_medoids == curr_medoids).all()):
@@ -51,7 +52,7 @@ if __name__ == '__main__':
                                 random_state=999)
     distances = pairwise_distances(X)
 
-    clusters, curr_medoids = cluster(distances, k=3)
+    clusters, curr_medoids = cluster(distances, k=1)
     print(curr_medoids)
 
 

### 2015-04-19 src/multiswaps.py

@@ -9,9 +9,6 @@ from sklearn.metrics.pairwise import pairwise_distances
 import matplotlib.pyplot as plt
 from itertools import cycle
 
-num_clusters = 3
-p = 2
-epsilon = 5
 
 def find_closest_median(i, medoids, distances):
     return medoids[np.argmin(distances[i, medoids])]
@@ -32,7 +29,8 @@ def swap(medoids, A, B):
 def swap_single(medoids, a, b):
     return [b if i == a else i for i in medoids]
 
-def search_for_swap(medoids, distances):
+def search_for_swap(medoids, distances, p=2, epsilon=5):
+    n = distances.shape[0]
     prev_objective = objective(medoids, distances)
     improvement = False
     for A in itertools.combinations(medoids, p):
@@ -44,6 +42,15 @@ def search_for_swap(medoids, distances):
                 return new_medoids, improvement
     return medoids, improvement
 
+def cluster(distances, k, p):
+    n = distances.shape[0]
+    medoids = random.sample(range(n), k)
+    improvement = True
+    while improvement:
+        medoids, improvement = search_for_swap(medoids, distances, p)
+
+    return medoids, objective(medoids, distances)
+
 if __name__ == '__main__':
     centers = [[1, 1], [-1, -1], [1, -1]]
     X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
@@ -53,12 +60,13 @@ if __name__ == '__main__':
     
     #Choose random medoids
     # medoids = random.sample(range(n), num_clusters)
-
+    num_clusters = 3
+    p = 2
     medoids = [0, 1, 2]
     print(medoids, objective(medoids, distances))
 
     improvement = True
     while improvement:
-        medoids, improvement = search_for_swap(medoids, distances)
+        medoids, improvement = search_for_swap(medoids, distances, p)
 
     print(medoids, objective(medoids, distances))

### 2015-04-19 src/opt_solns.py

@@ -1,15 +1,26 @@
 import numpy as np
+import os
+from LoadpMedian import *
+from LoadData import *
 from gurobipy import *
+from sklearn.metrics.pairwise import pairwise_distances
 
-def kmedian_opt(distances, k):
+def kmedian_opt(distances, IP, k1, k2):
     model = Model("k-median")
     n = np.shape(distances)[0]
     y,x = {}, {}
-    
+
+    if IP:
+        var_type = GRB.BINARY
+    else:
+        var_type = GRB.CONTINUOUS
+        
     for j in range(n):
-        y[j] = model.addVar(obj=0, vtype="B", name="y[%s]"%j)
+        y[j] = model.addVar(obj=0, vtype=var_type, name="y[%s]"%j)
         for i in range(n):
-            x[i,j] = model.addVar(obj=distances[i,j], vtype="B", name="x[%s,%s]"%(i,j))
+            x[i,j] = model.addVar(obj=distances[i,j], vtype=var_type, \
+                                  name="x[%s,%s]"%(i,j))
+                
     model.update()
     
     for i in range(n):
@@ -23,13 +34,50 @@ def kmedian_opt(distances, k):
             
     coef = [1 for j in range(n)]
     var = [y[j] for j in range(n)]
-    model.addConstr(LinExpr(coef,var), "=", rhs=k, name="k_median")
-    
-    model.update()
-    model.__data = x,y
-    model.optimize()
 
-    if model.status == GRB.status.OPTIMAL:
-        return model.objVal
-    else:
-        return 0
+    model.setParam( 'OutputFlag', False )
+    model.__data = x,y
+    outputs = [0 for i in range(n+1)]
+    model.update()
+    
+    for k in range(k1,k2+1):
+        k_constr = model.addConstr(LinExpr(coef,var), ">=", rhs=k, name="k_median")
+        model.update()
+        model.optimize()
+
+        if model.status == GRB.status.OPTIMAL:
+            outputs[k] =  model.objVal
+        else:
+            outpus[k] = 0
+            
+    return outputs
+
+def write_opt_bounds(distances, filepath, IP=1):
+    f = open(filepath, 'w+')
+    n = np.shape(distances)[0]
+
+    bounds = kmedian_opt(distances, IP)
+    for k in range(1,n+1):
+        f.write(str(k)+" "+str(bound[k])+"\n")
+
+def write_opt_pmedians(path_files, file_bounds):
[…]

### 2015-04-19 src/opt_solns.py
Optimal Flat Clusterings

@@ -5,7 +5,7 @@ from LoadData import *
 from gurobipy import *
 from sklearn.metrics.pairwise import pairwise_distances
 
-def kmedian_opt(distances, IP, k1, k2):
+def kmedian_opt(distances, IP, k):
     model = Model("k-median")
     n = np.shape(distances)[0]
     y,x = {}, {}
@@ -37,18 +37,19 @@ def kmedian_opt(distances, IP, k1, k2):
 
     model.setParam( 'OutputFlag', False )
     model.__data = x,y
-    outputs = [0 for i in range(n+1)]
+    outputs = []
     model.update()
     
-    for k in range(k1,k2+1):
-        k_constr = model.addConstr(LinExpr(coef,var), ">=", rhs=k, name="k_median")
-        model.update()
-        model.optimize()
+    k_constr = model.addConstr(LinExpr(coef,var), "=", rhs=k, name="k_median")
+    model.update()
+    model.optimize()
 
-        if model.status == GRB.status.OPTIMAL:
-            outputs[k] =  model.objVal
-        else:
-            outpus[k] = 0
+    if model.status == GRB.status.OPTIMAL:
+        print model.objVal
+        outputs.append(model.objVal)
+    else:
+        print "not optimal"
+        output.append(0)
             
     return outputs
 
@@ -66,17 +67,17 @@ def write_opt_pmedians(path_files, file_bounds):
     for f in os.listdir(path_files):
         print f
         distances, n, k = LoadpMedian(path_files+"\\"+f)
-        bound = kmedian_opt(distances, 1, k, k)
+        bound = kmedian_opt(distances, 1, k)
         g.write(f+" "+str(bound)+"\n")
 
-def write_opt_data(path_files, file_bounds, name):
+def write_opt_data(path_files, file_bounds):
     g = open(file_bounds, 'w+')
     
     for f in os.listdir(path_files):
         print f
-        X, n, k = LoadData(path_files+"\\"+f, name)
+        X, y, n, k = LoadData(path_files+"\\"+f)
         distances = pairwise_distances(X)
-        bound = kmedian_opt(distances, 1, k, k)
+        bound = kmedian_opt(distances, 1, k)
         g.write(f+" "+str(bound)+"\n")
         
     

### 2015-04-20 src/opt_solns.py
Added all presentation files and associated images

@@ -5,7 +5,7 @@ from LoadData import *
 from gurobipy import *
 from sklearn.metrics.pairwise import pairwise_distances
 
-def kmedian_opt(distances, IP, k):
+def kmedian_opt(distances, IP, k1, k2):
     model = Model("k-median")
     n = np.shape(distances)[0]
     y,x = {}, {}
@@ -31,25 +31,25 @@ def kmedian_opt(distances, IP, k):
     for j in range(n):
         for i in range(n):
             model.addConstr(x[i,j], "<", y[j], name="Strong[%s,%s]"%(i,j))
-            
-    coef = [1 for j in range(n)]
-    var = [y[j] for j in range(n)]
 
     model.setParam( 'OutputFlag', False )
     model.__data = x,y
     outputs = []
     model.update()
-    
-    k_constr = model.addConstr(LinExpr(coef,var), "=", rhs=k, name="k_median")
-    model.update()
-    model.optimize()
 
-    if model.status == GRB.status.OPTIMAL:
-        print model.objVal
-        outputs.append(model.objVal)
-    else:
-        print "not optimal"
-        output.append(0)
+    for k in range(k1, k2):
+        coef = [1 for j in range(n)]
+        var = [y[j] for j in range(n)]
+        if k > k1:
+            model.remove(k_constr)
+        k_constr = model.addConstr(LinExpr(coef,var), "<", rhs=k)
+        model.update()
+        model.optimize()
+
+        if model.status == GRB.status.OPTIMAL:
+            outputs.append(model.objVal)
+        else:
+            outputs.append(0)
             
     return outputs
 
@@ -57,9 +57,9 @@ def write_opt_bounds(distances, filepath, IP=1):
     f = open(filepath, 'w+')
     n = np.shape(distances)[0]
 
-    bounds = kmedian_opt(distances, IP)
+    bounds = kmedian_opt(distances, IP, 1, n+1)
     for k in range(1,n+1):
-        f.write(str(k)+" "+str(bound[k])+"\n")
+        f.write(str(k)+" "+str(bounds[k-1])+"\n")
 
 def write_opt_pmedians(path_files, file_bounds):
     g = open(file_bounds, 'w+')
@@ -67,7 +67,7 @@ def write_opt_pmedians(path_files, file_bounds):
     for f in os.listdir(path_files):
         print f
         distances, n, k = LoadpMedian(path_files+"\\"+f)
[…]

### 2015-04-21 src/kmedoids.py
Merge

@@ -5,6 +5,9 @@ import random
 
 from sklearn.datasets.samples_generator import make_blobs
 from sklearn.metrics.pairwise import pairwise_distances
+from sklearn.metrics import adjusted_mutual_info_score
+
+from LoadData import LoadData
 
 def cluster(distances, k=3):
     m = distances.shape[0] # number of points
@@ -47,13 +50,16 @@ def compute_new_medoid(cluster, distances):
     return costs.argmin(axis=0, fill_value=10e9)
 
 if __name__ == '__main__':
-    centers = [[1, 1], [-1, -1], [1, -1]]
-    X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
-                                random_state=999)
+    # centers = [[1, 1], [-1, -1], [1, -1]]
+    # X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
+    #                             random_state=999)
+    # X, y, n, k = LoadData("data/Gaussian/Gauss_10_5_0.txt")
+    X, y, n, k = LoadData("data/Real Data Sets/iris.data.txt", cluster_loc=0, split=0)
     distances = pairwise_distances(X)
 
-    clusters, curr_medoids = cluster(distances, k=1)
-    print(curr_medoids)
+    clusters, curr_medoids = cluster(distances, k=k)
+    print(clusters, curr_medoids)
+    print(adjusted_mutual_info_score(y, clusters))
 
 
 

### 2015-04-21 src/multiswaps.py

@@ -5,9 +5,12 @@ import numpy as np
 
 from sklearn.datasets.samples_generator import make_blobs
 from sklearn.metrics.pairwise import pairwise_distances
+from sklearn.metrics import adjusted_mutual_info_score
 
 import matplotlib.pyplot as plt
 from itertools import cycle
+
+from LoadData import LoadData
 
 
 def find_closest_median(i, medoids, distances):
@@ -29,7 +32,10 @@ def swap(medoids, A, B):
 def swap_single(medoids, a, b):
     return [b if i == a else i for i in medoids]
 
-def search_for_swap(medoids, distances, p=2, epsilon=5):
+def score(labels_true, labels_pred):
+    return adjusted_mutual_info_score(labels_true, labels_pred)
+
+def search_for_swap(medoids, distances, p=2, epsilon=0):
     n = distances.shape[0]
     prev_objective = objective(medoids, distances)
     improvement = False
@@ -51,18 +57,28 @@ def cluster(distances, k, p):
 
     return medoids, objective(medoids, distances)
 
-if __name__ == '__main__':
-    centers = [[1, 1], [-1, -1], [1, -1]]
-    X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
-                                random_state=999)
-    distances = pairwise_distances(X)
+def predict(medoids, distances):
     n = distances.shape[0]
+    return [ np.argmin(distances[i, medoids]) for i in range(n) ]
+
+if __name__ == '__main__':
+    # centers = [[1, 1], [-1, -1], [1, -1]]
+    # n = 100
+    # X, y = make_blobs(n_samples=n, centers=centers, cluster_std=0.5,
+    #                             random_state=999)
+    # k = 3
+
+
+    # X, y, n, k = LoadData("data/Gaussian/Gauss_10_5_0.txt")
+    X, y, n, k = LoadData("data/Real Data Sets/iris.data.txt", cluster_loc=0, split=0)
+    # print(X, y, n, k)
+
+    distances = pairwise_distances(X)
+    # n = distances.shape[0]
     
     #Choose random medoids
-    # medoids = random.sample(range(n), num_clusters)
-    num_clusters = 3
+    medoids = random.sample(range(n), k)
     p = 2
-    medoids = [0, 1, 2]
     print(medoids, objective(medoids, distances))
 
     improvement = True
@@ -70,3 +86,9 @@ if __name__ == '__main__':
[…]

### 2015-04-21 tex/presentation/presentation.tex

\documentclass[usenames,dvipsnames]{beamer}
\usetheme{Boadilla}
\usepackage{subfig} 
\usepackage{multicol}
\usepackage{color}
\usepackage[english]{babel}

\title[Clustering Approx Algs]{Local Search Approximation Algorithms for Clustering}
\author{Dave Lingenbrink, Alice Paul, Calvin Wylie}
\date{April 23, 2015}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command: 
%\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Motivation}
Look for structure in unlabeled data.
\begin{itemize}
\item Social networks
\item Handwritten letters
\item Grouping customer classes
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{$k$-Means Algorithm and Local Search}
Objective: partition the data into sets $S_1, S_2, \ldots, S_k$ with means $\mu_1, \ldots, \mu_k$ to minimize $\sum_{i=1}^k \sum_{x \in S_i} ||x-\mu_i||^2$.

\begin{center}
\includegraphics<1>[width=3.75in]{k-means-1}
\includegraphics<2>[width=3.75in]{k-means-2}
\hspace{0.6in}\includegraphics<3>[width=3in]{k-means-3}
\includegraphics<4>[width=3in]{k-means-4}
\end{center}
\end{frame}

\begin{frame}
\frametitle{$k$-Median Objective Function}
Objective: find a subset of centers $S \subseteq X$ such that $|S| \leq k$ to minimize $\sum_{x \in X} \min_{s \in S} d(x - s)$.

\begin{center}
\includegraphics<1>[width=3in]{k-median-1}
\includegraphics<2>[width=3in]{k-median-2}
\includegraphics<3>[width=3in]{k-median-3}
\includegraphics<4>[width=3in]{k-median-4}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Proposed Algorithms}
\begin{enumerate}
\item Multiswap Algorithm: look at local swaps of size $p$. 
\[ \left(3+ \frac{2}{p} \right) \text{-approximation ratio}\]
\item Minimizing a monotone decreasing supermodular function: look at local swaps using a potential function $\psi$.
\[ \left(1 + \frac{c}{1-c} \cdot e^{-1} + \frac{1}{1-c} \cdot O(\varepsilon) \right) \text{-approximation ratio (whp)}\]
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Hiearchical Clustering}
Look for layered structure in unlabeled data and finding a natural level of clustering.
\begin{center}
\includegraphics[width=2.7in]{hierarchical-clustering-example}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Agglomerative Clustering}
Repeatedly merge the two closest clusters according to some distance function $d(C_1, C_2)$. 
\begin{center}
\includegraphics[width=2in]{agglomerative-clustering}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Hiearchical $k$-Median Objective Function}
Find subsets $S_1, S_2, \ldots, S_n$ ($|S_i| = i$) and assignments $a_1, \ldots, a_n$ such that the clusters are nested:
\begin{enumerate}
\item $S_i \subseteq S_{i+1}$ for $i = 1, \ldots, n-1$,
\item $\forall j \in X$, if $a_{i+1}(j) \in S_i$, then $a_i(j) = a_{i+1}(j)$, and
\item $\forall j, k \in X$, if $a_{i+1}(j) = a_{i+1}(k)$, then $a_{i}(j) = a_{i}(k)$. 
\end{enumerate}

\vspace{0.5in}
There exists an approximation algorithm that given $k$-median solutions for $k=1, \ldots, n$, will generate a hierarchical clustering such that the $k$-median objective function on any level increases by at most a factor of 20.71 (or a factor of 10.03 in expectation when randomized).
\end{frame}

\begin{frame}
\frametitle{Proposed Algorithm}
\begin{center}
\includegraphics<1>[width=3in]{hierarchical-1}
\includegraphics<2>[width=3in]{hierarchical-2}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Data Sets and Preliminary Results}
\begin{enumerate}
\item Iris and soybean data sets from UCI Machine Learning Repository
\item Generated Gaussian data
\item OR Library's $p$-median instances
\end{enumerate}

Comparisons in adjusted mutual information (weighted for hierarchical results), computation time, and comparison to optimal $k$-median solutions.
\end{frame}

\begin{frame}
Questions?

\vspace{0.5in}
Learn more about the proposed algorithms:
\begin{itemize}
\item Arya et al. Local search heuristics for k-median and facility location problems. \emph{SIAM Journal on Computing}, 33(3):544-562, 2004.
\item Lin et al. A general approach for incremental approximation and hierarchical clustering. In \emph{Proceedings of the Seventeenth Annual ACM-SIAM Symposium on Discrete Algorithms}, SODA `06, pp.1147-1156, Philadelphia, PA, USA, 2006. 
\item Sviridenko, Maxim and Ward, Justin. Tight bounds for submodular and supermodular optimization with bounded curvature. arXiv:1311.4728v3 [cs.DS] 2014.
\end{itemize}
\end{frame}

\end{document}

### 2015-04-21 src/submodular-demo.py
added capability to return which center is closest

@@ -1,13 +1,4 @@
-"""
-=================================================
-Demo of affinity propagation clustering algorithm
-=================================================
-Reference:
-Brendan J. Frey and Delbert Dueck, "Clustering by Passing Messages
-Between Data Points", Science Feb. 2007
-"""
 
-from sklearn.cluster import AffinityPropagation
 from sklearn import metrics
 from sklearn.datasets.samples_generator import make_blobs
 from sklearn.datasets import load_digits
@@ -21,7 +12,7 @@ np.random.seed(17)
 H_PERC = 100
 
 EPSILON =80
-N = 20
+N = 10
 delta = EPSILON/N
 
 ##############################################################################
@@ -133,8 +124,6 @@ def updateS(S,Sc, X):
           print "old ", psiS, " new ", psiNew, "difference ", psiNew - psiS
           print "old fvalue ", f(Sc, X), "f value ", f(newSc,X)
           return newS
-  print "no change"
-  print "final f value ", f(Sc,X)
   return S
           
 
@@ -151,7 +140,27 @@ def supermodular(X):
       S = newS
       Sc = X.difference(S)
 
-print supermodular(toFrozenSet(X))
+def supermodular_list(data):
+  L = list(data)
+  print data
+  print L
+  centers = supermodular(toFrozenSet(data))
+  print centers
+  closestCenter =[]
+
+  for x in L:
+    currentCenter = 0
+    currentDistance = float("inf")
+    for y in centers:
+      if(d(x,y)<currentDistance):
+        currentCenter=y
+        currentDistance=d(x,y)
+    closestCenter+=[currentCenter]
+  return closestCenter
+
+
+# print supermodular(toFrozenSet(X))
+print supermodular_list(X)
 
 print "curvature is ", totalCurvature(toFrozenSet(X))
 

### 2015-04-21 src/kmedoids.py
Cleaned up kmedoids and multiswaps

@@ -3,12 +3,6 @@
 import numpy as np
 import random
 
-from sklearn.datasets.samples_generator import make_blobs
-from sklearn.metrics.pairwise import pairwise_distances
-from sklearn.metrics import adjusted_mutual_info_score
-
-from LoadData import LoadData
-
 def cluster(distances, k=3):
     m = distances.shape[0] # number of points
 
@@ -48,18 +42,6 @@ def compute_new_medoid(cluster, distances):
     cluster_distances = np.ma.masked_array(data=distances, mask=mask, fill_value=10e9)
     costs = cluster_distances.sum(axis=1)
     return costs.argmin(axis=0, fill_value=10e9)
-
-if __name__ == '__main__':
-    # centers = [[1, 1], [-1, -1], [1, -1]]
-    # X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
-    #                             random_state=999)
-    # X, y, n, k = LoadData("data/Gaussian/Gauss_10_5_0.txt")
-    X, y, n, k = LoadData("data/Real Data Sets/iris.data.txt", cluster_loc=0, split=0)
-    distances = pairwise_distances(X)
-
-    clusters, curr_medoids = cluster(distances, k=k)
-    print(clusters, curr_medoids)
-    print(adjusted_mutual_info_score(y, clusters))
 
 
 

### 2015-04-21 src/multiswaps.py

@@ -1,17 +1,6 @@
 import random
 import itertools
-
 import numpy as np
-
-from sklearn.datasets.samples_generator import make_blobs
-from sklearn.metrics.pairwise import pairwise_distances
-from sklearn.metrics import adjusted_mutual_info_score
-
-import matplotlib.pyplot as plt
-from itertools import cycle
-
-from LoadData import LoadData
-
 
 def find_closest_median(i, medoids, distances):
     return medoids[np.argmin(distances[i, medoids])]
@@ -32,9 +21,6 @@ def swap(medoids, A, B):
 def swap_single(medoids, a, b):
     return [b if i == a else i for i in medoids]
 
-def score(labels_true, labels_pred):
-    return adjusted_mutual_info_score(labels_true, labels_pred)
-
 def search_for_swap(medoids, distances, p=2, epsilon=0):
     n = distances.shape[0]
     prev_objective = objective(medoids, distances)
@@ -48,47 +34,18 @@ def search_for_swap(medoids, distances, p=2, epsilon=0):
                 return new_medoids, improvement
     return medoids, improvement
 
-def cluster(distances, k, p):
-    n = distances.shape[0]
-    medoids = random.sample(range(n), k)
-    improvement = True
-    while improvement:
-        medoids, improvement = search_for_swap(medoids, distances, p)
-
-    return medoids, objective(medoids, distances)
-
 def predict(medoids, distances):
     n = distances.shape[0]
-    return [ np.argmin(distances[i, medoids]) for i in range(n) ]
+    return [ find_closest_median(i, medoids, distances) for i in range(n) ]
 
-if __name__ == '__main__':
-    # centers = [[1, 1], [-1, -1], [1, -1]]
-    # n = 100
-    # X, y = make_blobs(n_samples=n, centers=centers, cluster_std=0.5,
-    #                             random_state=999)
-    # k = 3
-
-
-    # X, y, n, k = LoadData("data/Gaussian/Gauss_10_5_0.txt")
-    X, y, n, k = LoadData("data/Real Data Sets/iris.data.txt", cluster_loc=0, split=0)
-    # print(X, y, n, k)
-
-    distances = pairwise_distances(X)
-    # n = distances.shape[0]
-    
-    #Choose random medoids
+def cluster(distances, k, p, epsilon=0):
+    n = distances.shape[0]
     medoids = random.sample(range(n), k)
-    p = 2
-    print(medoids, objective(medoids, distances))
-
     improvement = True
[…]

### 2015-04-21 src/testing.py
Initial testing framework

@@ -1,0 +1,67 @@
+import numpy as np
+
+import multiswaps
+import kmedoids
+import submodular_demo
+from LoadData import LoadData
+
+from sklearn.metrics.pairwise import pairwise_distances
+from sklearn.metrics import adjusted_mutual_info_score
+
+def objective(distances, assignment):
+    n = distances.shape[0]
+    return np.sum([ distances[i, assignment[i]] for i in range(n) ])
+
+def test_clustering(cluster_alg, distances, labels_true, N=10):
+    obj_results = []
+    ami_results = []
+    for _ in range(N):
+        assignment, _ = cluster_alg()
+        obj_results.append(objective(distances, assignment))
+        ami_results.append(adjusted_mutual_info_score(y, assignment))
+
+    avg_obj = np.mean(obj_results)
+    avg_ami = np.mean(ami_results)
+    min_obj = np.min(obj_results)
+    min_ami = np.min(ami_results)
+    max_obj = np.max(obj_results)
+    max_ami = np.max(ami_results)
+
+    return avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami
+
+if __name__ == '__main__':
+    X, y, n, k = LoadData("data/Real Data Sets/soybean-small.data.txt", cluster_loc=0, split=0)
+    distances = pairwise_distances(X)
+
+    p = 3
+    multi_swaps = lambda: multiswaps.cluster(distances, k, p, epsilon=1)
+    avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami = test_clustering(multi_swaps, distances, y, N=10)
+    print("%s-swap Average: Objective: %f AMI: %f" % (p, avg_obj, avg_ami))
+    print("%s-swap Best Case: Objective: %f AMI: %f" % (p, min_obj, max_ami))
+    print("%s-swap Worst Case: Objective: %f AMI: %f" % (p, max_obj, min_ami))
+
+    # k_medoids = lambda: kmedoids.cluster(distances, k=k)
+    # avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami = test_clustering(k_medoids, distances, y, N=100)
+    # print("K-Medoids Average: Objective: %f AMI: %f" % (avg_obj, avg_ami))
+    # print("K-Medoids Best Case: Objective: %f AMI: %f" % (min_obj, max_ami))
[…]

### 2015-04-21 src/submodular-demo.py
removed global vars from supermodular

@@ -7,21 +7,19 @@ import numpy as np
 import math
 import random
 
-np.random.seed(17)
+# np.random.seed(17)
 
-H_PERC = 100
+# H_PERC = 100
 
-EPSILON =80
-N = 10
-delta = EPSILON/N
+
 
 ##############################################################################
 # Generate sample data
-centers = [[1, 1], [-1, -1], [1, -1]]
-X, labels_true = make_blobs(n_samples=N, centers=centers, cluster_std=0.5,
-                            random_state=999)
+# centers = [[1, 1], [-1, -1], [1, -1]]
+# X, labels_true = make_blobs(n_samples=10, centers=centers, cluster_std=0.5,
+#                             random_state=999)
 
-k = len(centers)
+
 ##############################################################################
 # Define Functions
 def toFrozenSet(X):
@@ -91,19 +89,19 @@ def drawB(A):
   return frozenset(B)
 
 HDict ={}
-def estimateH(A,X):
+def estimateH(A,X, h_perc):
   if A in HDict:
     return HDict[A]
   Sum = 0
-  for x in range(H_PERC):
+  for x in range(h_perc):
     Sum+= g(drawB(A),X)
-  HDict[A] = Sum/H_PERC
-  return Sum/H_PERC
+  HDict[A] = Sum/h_perc
+  return Sum/h_perc
 
-def psi(A,X):
-  return (1-1/math.e)*estimateH(A,X) + l(A,X)
+def psi(A,X, h_perc):
+  return (1-1/math.e)*estimateH(A,X,h_perc) + l(A,X)
 
-def initialS(X):
+def initialS(X, k, N):
   i = N-k
   S = set()
   for x in X:
@@ -113,13 +111,14 @@ def initialS(X):
     i-=1
   return frozenset(S)
 
-def updateS(S,Sc, X):
-  psiS = psi(S, X)
+def updateS(S,Sc, X, epsilon,h_perc):
+  delta = epsilon/len(X)
+  psiS = psi(S, X,h_perc)
   for a in S:
       for b in Sc:
         newS = S.difference([a]).union([b])
         newSc = Sc.difference([b]).union([a])
-        psiNew = psi(newS,X)
+        psiNew = psi(newS,X,h_perc)
         if psiNew>= psiS+delta:
           print "old ", psiS, " new ", psiNew, "difference ", psiNew - psiS
           print "old fvalue ", f(Sc, X), "f value ", f(newSc,X)
[…]

### 2015-04-21 src/submodular-demo.py
made epison in submodular have a default value

@@ -126,7 +126,7 @@ def updateS(S,Sc, X, epsilon,h_perc):
   return S
           
 
-def supermodular(X, k, epsilon,h_perc=100):
+def supermodular(X, k, epsilon=10,h_perc=100):
   S = initialS(X, k, len(X))
   Sc = X.difference(S)
   print "initial f ", f(Sc,X)
@@ -139,7 +139,7 @@ def supermodular(X, k, epsilon,h_perc=100):
       S = newS
       Sc = X.difference(S)
 
-def supermodular_list(data, k, epsilon, h_perc=100):
+def supermodular_list(data, k, epsilon=10, h_perc=100):
   L = list(data)
   print data
   print L
@@ -159,9 +159,9 @@ def supermodular_list(data, k, epsilon, h_perc=100):
 
 
 # print supermodular(toFrozenSet(X))
-print supermodular_list(X,3,80)
+# print supermodular_list(X,3,80)
 
-print "curvature is ", totalCurvature(toFrozenSet(X))
+# print "curvature is ", totalCurvature(toFrozenSet(X))
 
 
 

### 2015-04-22 src/submodular_demo.py
fixed submodular

@@ -124,7 +124,15 @@ def updateS(S,Sc, X, epsilon,h_perc):
           print "old fvalue ", f(Sc, X), "f value ", f(newSc,X)
           return newS
   return S
-          
+
+def centersToIndex(X, centers):
+  centersDic = {}
+  for c in centers:
+    for i in range(len(X)):
+      if c == x:
+        centersDic[c] = i
+        break
+
 
 def supermodular(X, k, epsilon=10,h_perc=100):
   S = initialS(X, k, len(X))
@@ -144,9 +152,8 @@ def supermodular_list(data, k, epsilon=10, h_perc=100):
   print data
   print L
   centers = supermodular(toFrozenSet(data), k, epsilon, h_perc)
-  print centers
+  centersDic = centersToIndex(L, centers)
   closestCenter =[]
-
   for x in L:
     currentCenter = 0
     currentDistance = float("inf")
@@ -154,7 +161,7 @@ def supermodular_list(data, k, epsilon=10, h_perc=100):
       if(d(x,y)<currentDistance):
         currentCenter=y
         currentDistance=d(x,y)
-    closestCenter+=[currentCenter]
+    closestCenter+=[centersDic[currentCenter]]
   return closestCenter
 
 

### 2015-04-22 src/submodular_demo.py
further fixed submodular

@@ -14,10 +14,7 @@ import random
 
 
 ##############################################################################
-# Generate sample data
-# centers = [[1, 1], [-1, -1], [1, -1]]
-# X, labels_true = make_blobs(n_samples=10, centers=centers, cluster_std=0.5,
-#                             random_state=999)
+
 
 
 ##############################################################################
@@ -129,9 +126,10 @@ def centersToIndex(X, centers):
   centersDic = {}
   for c in centers:
     for i in range(len(X)):
-      if c == x:
+      if c == tuple(X[i]):
         centersDic[c] = i
         break
+  return centersDic
 
 
 def supermodular(X, k, epsilon=10,h_perc=100):
@@ -149,8 +147,6 @@ def supermodular(X, k, epsilon=10,h_perc=100):
 
 def supermodular_list(data, k, epsilon=10, h_perc=100):
   L = list(data)
-  print data
-  print L
   centers = supermodular(toFrozenSet(data), k, epsilon, h_perc)
   centersDic = centersToIndex(L, centers)
   closestCenter =[]
@@ -164,8 +160,11 @@ def supermodular_list(data, k, epsilon=10, h_perc=100):
     closestCenter+=[centersDic[currentCenter]]
   return closestCenter
 
-
-# print supermodular(toFrozenSet(X))
+# Generate sample data
+# centers = [[1, 1], [-1, -1], [1, -1]]
+# X, labels_true = make_blobs(n_samples=10, centers=centers, cluster_std=0.5,
+#                             random_state=999)
+# # print supermodular(toFrozenSet(X))
 # print supermodular_list(X,3,80)
 
 # print "curvature is ", totalCurvature(toFrozenSet(X))

### 2015-04-22 src/multiswaps.py
Update

@@ -30,7 +30,7 @@ def search_for_swap(medoids, distances, p=2, epsilon=0):
             new_medoids = swap(medoids, A, B)
             if objective(new_medoids, distances) < prev_objective - epsilon:
                 improvement = True
-                print("Improvement: %f" % (prev_objective - objective(new_medoids, distances)))
+                # print("Improvement: %f" % (prev_objective - objective(new_medoids, distances)))
                 return new_medoids, improvement
     return medoids, improvement
 

### 2015-04-22 src/submodular_demo.py

@@ -120,8 +120,8 @@ def updateS(S,Sc, X, epsilon,h_perc):
         newSc = Sc.difference([b]).union([a])
         psiNew = psi(newS,X,h_perc)
         if psiNew>= psiS+delta:
-          print "old ", psiS, " new ", psiNew, "difference ", psiNew - psiS
-          print "old fvalue ", f(Sc, X), "f value ", f(newSc,X)
+          print("old ", psiS, " new ", psiNew, "difference ", psiNew - psiS)
+          print("old fvalue ", f(Sc, X), "f value ", f(newSc,X))
           return newS
   return S
 
@@ -129,19 +129,20 @@ def centersToIndex(X, centers):
   centersDic = {}
   for c in centers:
     for i in range(len(X)):
-      if c == x:
+      if c == X:
         centersDic[c] = i
         break
+  return centersDic
 
 
 def supermodular(X, k, epsilon=10,h_perc=100):
   S = initialS(X, k, len(X))
   Sc = X.difference(S)
-  print "initial f ", f(Sc,X)
+  print("initial f ", f(Sc,X))
   while True:
     newS = updateS(S,Sc,X, epsilon,h_perc)
     if newS == S:
-      print "Finished with f(S) = ", f(Sc, X)
+      print("Finished with f(S) = ", f(Sc, X))
       return Sc
     else:
       S = newS
@@ -149,8 +150,8 @@ def supermodular(X, k, epsilon=10,h_perc=100):
 
 def supermodular_list(data, k, epsilon=10, h_perc=100):
   L = list(data)
-  print data
-  print L
+  # print data
+  # print L
   centers = supermodular(toFrozenSet(data), k, epsilon, h_perc)
   centersDic = centersToIndex(L, centers)
   closestCenter =[]
@@ -176,46 +177,3 @@ def supermodular_list(data, k, epsilon=10, h_perc=100):
 # print totalCurvature(toFrozenSet(X))
 # print D
 # print X
-
-
-##############################################################################
-# Compute Affinity Propagation
-# af = AffinityPropagation(preference=-50).fit(X)
-# cluster_centers_indices = af.cluster_centers_indices_
-# labels = af.labels_
-
-# n_clusters_ = len(cluster_centers_indices)
-
-# print('Estimated number of clusters: %d' % n_clusters_)
-# print("Homogeneity: %0.3f" % metrics.homogeneity_score(labels_true, labels))
[…]

### 2015-04-22 src/testing.py

@@ -4,6 +4,7 @@ import multiswaps
 import kmedoids
 import submodular_demo
 from LoadData import LoadData
+from LoadpMedian import LoadpMedian
 
 from sklearn.metrics.pairwise import pairwise_distances
 from sklearn.metrics import adjusted_mutual_info_score
@@ -31,26 +32,29 @@ def test_clustering(cluster_alg, distances, labels_true, N=10):
 
 if __name__ == '__main__':
     X, y, n, k = LoadData("data/Real Data Sets/soybean-small.data.txt", cluster_loc=0, split=0)
+    # X, y, n, k = LoadData("data/Real Data Sets/iris.data.txt", cluster_loc=0, split=0)
+    # X, y, n, k = LoadData("data/Gaussian2/Gauss_7_5_5.txt", cluster_loc=1, split=1)
     distances = pairwise_distances(X)
+    # distances, n, k = LoadpMedian("data/p-median Instances/pmed1.txt")
 
-    p = 3
-    multi_swaps = lambda: multiswaps.cluster(distances, k, p, epsilon=1)
-    avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami = test_clustering(multi_swaps, distances, y, N=10)
-    print("%s-swap Average: Objective: %f AMI: %f" % (p, avg_obj, avg_ami))
-    print("%s-swap Best Case: Objective: %f AMI: %f" % (p, min_obj, max_ami))
-    print("%s-swap Worst Case: Objective: %f AMI: %f" % (p, max_obj, min_ami))
+    # for p in [1,2]: #[1,2,3]:
+    #     multi_swaps = lambda: multiswaps.cluster(distances, k, p, epsilon=1)
+    #     avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami = test_clustering(multi_swaps, distances, y, N=10)
+    #     print("%s-swap Average: Objective: %f AMI: %f" % (p, avg_obj, avg_ami))
+    #     print("%s-swap Best Case: Objective: %f AMI: %f" % (p, min_obj, max_ami))
+    #     print("%s-swap Worst Case: Objective: %f AMI: %f" % (p, max_obj, min_ami))
 
     # k_medoids = lambda: kmedoids.cluster(distances, k=k)
-    # avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami = test_clustering(k_medoids, distances, y, N=100)
[…]

### 2015-04-24 tex/presentation/presentation.tex
Optimal Solutions
Calculated for all values of k except for pmed35 - pmed40 which are too
slow

@@
 \begin{itemize}
-\item Social networks
-\item Handwritten letters
-\item Grouping customer classes
+\item Group plants by species
+\item Find similar gene sequences to find gene families across species
+\item Infer population structures
+\item Crime analysis
+\item PET scans to differentiate different tissues
+\item Social networks and recognizing communities
+\item Image segmentation and object regonition
+\item Grouping customer classes and market segmentation
 \end{itemize}
@@
 \begin{frame}
-\frametitle{$k$-Means Algorithm and Local Search}
+\frametitle{$k$-Means Algorithm}
 Objective: partition the data into sets $S_1, S_2, \ldots, S_k$ with means $\mu_1, \ldots, \mu_k$ to minimize $\sum_{i=1}^k \sum_{x \in S_i} ||x-\mu_i||^2$.

### 2015-04-24 src/submodular_demo.py
Corrected use of L(X) and f(emptyset)

@@ -35,8 +35,8 @@ def f(S,X):
     return D[S]
   if S == frozenset():
     values = [f(frozenset([a]), X) for a in X]
-    D[S]= min(values)
-    return min(values)
+    D[S]= max(values)
+    return max(values)
   Sum = 0
   for x in X:
     if x not in S:
@@ -52,6 +52,7 @@ def without(S, x):
 def curvatureValue(X, j, fX, fEmpty):
   num = (fX - f(without(X,j),X))
   denom = (f(frozenset([j]),X)-fEmpty)
+  print num, denom
   if denom == 0: 
     return float("inf")
   else:
@@ -63,12 +64,13 @@ def totalCurvature(X):
   fX = f(X,X)
   fEmpty = f(frozenset(),X)
   vals = [curvatureValue(X,j, fX, fEmpty) for j in X]
+  print vals
   return 1- min(vals)
 
 def l(A, X):
   fEmpty = f(frozenset(),X)
   vals = [f(frozenset([j]), X) for j in A]
-  return sum(vals)-fEmpty
+  return sum(vals)-fEmpty*len(A)
 
 def g(A,X):
   return -l(A,X) - f(X.difference(A),X)
@@ -146,6 +148,9 @@ def supermodular(X, k, epsilon=10,h_perc=100):
       Sc = X.difference(S)
 
 def supermodular_list(data, k, epsilon=10, h_perc=100):
+  c = totalCurvature(toFrozenSet(data))
+  print "curvature is ", c
+  print "ratio is (", (1 + c/(1-c)*math.e), " + ", (1/(1-c)), "O(epsilon))"
   L = list(data)
   centers = supermodular(toFrozenSet(data), k, epsilon, h_perc)
   centersDic = centersToIndex(L, centers)
@@ -161,13 +166,13 @@ def supermodular_list(data, k, epsilon=10, h_perc=100):
   return closestCenter
 
 # Generate sample data
-# centers = [[1, 1], [-1, -1], [1, -1]]
-# X, labels_true = make_blobs(n_samples=10, centers=centers, cluster_std=0.5,
-#                             random_state=999)
-# # print supermodular(toFrozenSet(X))
-# print supermodular_list(X,3,80)
+centers = [[1, 1], [-1, -1], [1, -1]]
+X, labels_true = make_blobs(n_samples=10, centers=centers, cluster_std=0.5,
+                            random_state=999)
+# print supermodular(toFrozenSet(X))
[…]

### 2015-04-26 src/kmedoids.py
Testing
Added time to testing.py, updated k-medoids to have a max-iter, obtained
testing results for k-medoids, k-means, spectral, and affinity prop

@@ -3,20 +3,25 @@
 import numpy as np
 import random
 
-def cluster(distances, k=3):
+def cluster(distances, k=3, max_iter=1000):
     m = distances.shape[0] # number of points
 
     # Pick k random medoids.
     curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
-    while not len(np.unique(curr_medoids)) == k:
+    i = 0
+    while (len(np.unique(curr_medoids)) < k) and (i<max_iter):
+        i+=1
+        if i > max_iter:
+            curr_medoids = np.array([i for i in range(k)])
         curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
 
     old_medoids = np.array([-1]*k) # Doesn't matter what we initialize these to.
     new_medoids = np.array([-1]*k)
     clusters = assign_points_to_clusters(curr_medoids, distances)
+    i = 0
    
     # Until the medoids stop updating, do the following:
-    while not ((old_medoids == curr_medoids).all()):
+    while not ((old_medoids == curr_medoids).all()) and i < max_iter:
         # Assign each point to cluster with closest medoid.
         clusters = assign_points_to_clusters(curr_medoids, distances)
 
@@ -27,6 +32,7 @@ def cluster(distances, k=3):
 
         old_medoids[:] = curr_medoids[:]
         curr_medoids[:] = new_medoids[:]
+        i += 1
 
     return clusters, curr_medoids
 

### 2015-04-26 src/testing.py

@@ -1,9 +1,10 @@
 import numpy as np
-
+import timeit
 import multiswaps
 import kmedoids
 import submodular_demo
 from LoadData import LoadData
+from LoadpMedian import LoadpMedian
 
 from sklearn.metrics.pairwise import pairwise_distances
 from sklearn.metrics import adjusted_mutual_info_score
@@ -13,32 +14,41 @@ def objective(distances, assignment):
     return np.sum([ distances[i, assignment[i]] for i in range(n) ])
 
 def test_clustering(cluster_alg, distances, labels_true, N=10):
+    time_results = []
     obj_results = []
     ami_results = []
     for _ in range(N):
+        start_time = timeit.default_timer()
         assignment, _ = cluster_alg()
+        time_results.append(timeit.default_timer()-start_time)
         obj_results.append(objective(distances, assignment))
-        ami_results.append(adjusted_mutual_info_score(y, assignment))
+        ami_results.append(adjusted_mutual_info_score(labels_true, assignment))
 
+    avg_tim = np.mean(time_results)
     avg_obj = np.mean(obj_results)
     avg_ami = np.mean(ami_results)
+    min_tim = np.min(time_results)
     min_obj = np.min(obj_results)
     min_ami = np.min(ami_results)
+    max_tim = np.max(time_results)
     max_obj = np.max(obj_results)
     max_ami = np.max(ami_results)
 
-    return avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami
+    return avg_tim, avg_obj, avg_ami, min_tim, min_obj, min_ami, \
+           max_tim, max_obj, max_ami
 
-if __name__ == '__main__':
-    X, y, n, k = LoadData("data/Real Data Sets/soybean-small.data.txt", cluster_loc=0, split=0)
-    distances = pairwise_distances(X)
+# if __name__ == '__main__':
+    
+    # X, y, n, k = LoadData("C:\\Users\\ajp336\\Dropbox\\approx-clustering\\data\\Real Data Sets\\soybean-small.data.txt", cluster_loc=0, split=0)
+    # distances = pairwise_distances(X)
 
-    p = 3
-    multi_swaps = lambda: multiswaps.cluster(distances, k, p, epsilon=1)
[…]

### 2015-04-30 src/kmedoids.py
Submodular changes and kmedoids warm start

@@ -3,18 +3,22 @@
 import numpy as np
 import random
 
-def cluster(distances, k=3, max_iter=1000):
+def cluster(distances, k=3, max_iter=1000, warm_start = None):
     m = distances.shape[0] # number of points
 
     # Pick k random medoids.
-    curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
-    i = 0
-    while (len(np.unique(curr_medoids)) < k) and (i<max_iter):
-        i+=1
-        if i > max_iter:
-            curr_medoids = np.array([i for i in range(k)])
+    if warm_start is None:
         curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
-
+        i = 0
+        while (len(np.unique(curr_medoids)) < k):
+            i+=1
+            if i > max_iter:
+                curr_medoids = np.array([i for i in range(k)])
+                break
+            curr_medoids = np.array([random.randint(0, m - 1) for _ in range(k)])
+    else:
+        curr_medoids = warm_start
+        
     old_medoids = np.array([-1]*k) # Doesn't matter what we initialize these to.
     new_medoids = np.array([-1]*k)
     clusters = assign_points_to_clusters(curr_medoids, distances)

### 2015-04-30 src/submodular_demo.py

@@ -1,8 +1,7 @@
 
-from sklearn import metrics
 from sklearn.datasets.samples_generator import make_blobs
-from sklearn.datasets import load_digits
-from sklearn.preprocessing import scale
+from sklearn.metrics.pairwise import pairwise_distances
+
 import numpy as np
 import math
 import random
@@ -19,207 +18,204 @@ import random
 
 ##############################################################################
 # Define Functions
-def toFrozenSet(X):
-  XList = X.tolist()
-  XListOfTuples = [tuple(l) for l in XList]
-  XFrozenSet = frozenset(XListOfTuples)
-  return XFrozenSet
+  
+# Curvature Calculation
 
-def d(x,y):
-  return np.linalg.norm(np.array(x)-np.array(y))
-
-D = {}
-
-def f(S,X):
-  if S in D:
-    return D[S]
-  if S == frozenset():
-    values = [f(frozenset([a]), X) for a in X]
-    D[S]= max(values)
-    return max(values)
-  Sum = 0
-  for x in X:
-    if x not in S:
-      distances = [d(x, s) for s in S]
-      Sum+=min(distances)
-  D[S]=Sum
-  return Sum
-
-def without(S, x):
-  #returns a copy of S without x without changing S
-  return S.difference([x])
-
-def curvatureValue(X, j, fX, fEmpty):
-  num = (fX - f(without(X,j),X))
-  denom = (f(frozenset([j]),X)-fEmpty)
-  print num, denom
-  if denom == 0: 
-    return float("inf")
-  else:
-    return num/denom
-
-
-def totalCurvature(X):
-  #expects X to be a frozenset
-  fX = f(X,X)
-  fEmpty = f(frozenset(),X)
-  vals = [curvatureValue(X,j, fX, fEmpty) for j in X]
-  print vals
-  return 1- min(vals)
-
-def l(A, X):
-  fEmpty = f(frozenset(),X)
-  vals = [f(frozenset([j]), X) for j in A]
-  return sum(vals)-fEmpty*len(A)
-
-def g(A,X):
-  return -l(A,X) - f(X.difference(A),X)
-
-def drawP():
-  uniformRandom = np.random.random_sample()
-  return math.log(uniformRandom*(math.e-1)+1)
-
-def drawB(A):
-  p = drawP()
-  B = set()
-  for x in A:
-    if np.random.random_sample()<p:
-      B.add(x)
[…]

### 2015-04-30 tex/project.tex
Paper Error
Fixed typo in paper
added writeup on datasets.  Still needs more after I talk to alice

@@
 \[ f(S) = \sum_{i \in X} \min_{j \in S} d(i,j) ,\]
-where we can think of $d$ as some measure of dissimilarity (for example squared distance or absolute distance). For completeness, we define $f(\emptyset) = \min_{j \in S} f(\{j\})$. 
+where we can think of $d$ as some measure of dissimilarity (for example squared distance or absolute distance). For completeness, we define $f(\emptyset) = \max_{j \in S} f(\{j\})$. 
 Note that this generalizes the $k$-median problem and can capture a variant of $k$-means in which we are restricted to choosing data points as centers.
@@
 \begin{align*}
-l(A) &= \sum_{j \in A} f(\{j\}) - f(\emptyset), \\
+l(A) &= \sum_{j \in A} \left [ f(\{j\}) - f(\emptyset) \right], \\
 g(A) & = - l(A) - f(X \setminus A).
@@
 
+\section{Data Sets}
+In order to empirically test these clustering methods, we used four different data sets.  
+\begin{description}
+\item[Generated Gaussian Data] We wanted to test our clustering algorithms on data that had data nicely clustered around $k$ centers.
+So, we generated synthetic $k$-center Gaussian data algorithmically and used this for testing.
+\item[$P$-median]
+\item[UCI Iris Dataset] This is one of the most used data sets in pattern recognition (\cite{Iris}).
+It contains data on 150 instances of flowers, each of with belongs to one of three species of iris.
+One of the species is linearly seperable from the others, but the other two are not.
+In total, there are 150 instances and 4 attributes.
+\item[UCI Soybean Dataset] This is another small, real data set from UCI.  This, too, is widely used in machine learning.  It has 47 instances, with 35 attributes.
+\end{description}
+
 \

### 2015-05-02 src/submodular_demo.py
Results excel

@@ -19,31 +19,6 @@ import random
 ##############################################################################
 # Define Functions
   
-# Curvature Calculation
-
-def curvatureValue(X, j, fX, fEmpty, distances, FDict = {}):
-    n = distances.shape[0]
-    Xj = X.difference([j])
-    
-    num = (fX - f(Xj,distances,FDict))
-    denom = (f(frozenset([j]),distances,FDict)-fEmpty)
-    
-    if denom == 0: 
-        return float("inf")
-    else:
-        return num/float(denom)
-
-
-def totalCurvature(distances, F = {}):
-    n = distances.shape[0]
-    X = frozenset([i for i in range(n)])
-  
-    fX = f(X,distances,F)
-    fEmpty = f(frozenset(),distances,F)
-    vals = [curvatureValue(X,j, fX, fEmpty, distances,F) for j in X]
-    return 1- min(vals)
-
-
 
 # Find closest median in S to i
 def find_closest_median(i, S, distances):
@@ -59,8 +34,9 @@ def find_closest_median(i, S, distances):
 
 # Returns the objective function for S
 def f(S, distances, F):
-    if S in F:
-        return F[S]
+    if F is not None:
+        if S in F:
+            return F[S]
     
     n = distances.shape[0]
     if S == frozenset():
@@ -71,9 +47,33 @@ def f(S, distances, F):
             if x not in S:
                 distx = [distances[x][s] for s in S]
                 Sum+=min(distx)
-            
-    F[S] = Sum
+    if F is not None:      
+        F[S] = Sum
     return Sum
+
+# Curvature Calculation
+
+def curvatureValue(X, j, fX, fEmpty, distances, FDict = None):
+    n = distances.shape[0]
+    Xj = X.difference([j])
+    
+    num = (fX - f(Xj,distances,FDict))
+    denom = (f(frozenset([j]),distances,FDict)-fEmpty)
+    
+    if denom == 0: 
+        return float("inf")
+    else:
+        return num/float(denom)
+
+
+def totalCurvature(distances, F = None):
+    n = distances.shape[0]
+    X = frozenset([i for i in range(n)])
+  
+    fX = f(X,distances,F)
+    fEmpty = f(frozenset(),distances,F)
+    vals = [curvatureValue(X,j, fX, fEmpty, distances,F) for j in X]
+    return 1- min(vals)
 
 # The linear part of the submodular function
 def l(A, distances, F):
[…]

### 2015-05-04 src/multiswaps.py
multiswap sped up

@@ -1,16 +1,35 @@
 import random
 import itertools
 import numpy as np
+from scipy.special import binom
+
+from sklearn.datasets.samples_generator import make_blobs
+from sklearn.metrics.pairwise import pairwise_distances
 
 def find_closest_median(i, medoids, distances):
-    return medoids[np.argmin(distances[i, medoids])]
+    closest = None
+    closest_dist = float("inf")
+    
+    for j in medoids:
+        if distances[i][j] < closest_dist:
+            closest_dist = distances[i][j]
+            closest = j
+        
+    return closest
 
-def objective(medoids, distances):
+def objective(medoids, distances, FDict):
     n = distances.shape[0]
-    s = 0
-    for i in range(0, n):
-        s += distances[i, find_closest_median(i, medoids, distances)]
-    return s
+
+    if medoids in FDict:
+        return FDict[medoids]
+        
+    Sum = 0
+    for x in range(n):
+        if x not in medoids:
+            distx = [distances[x][s] for s in medoids]
+            Sum+=min(distx)
+    FDict[medoids] = Sum
+    return Sum
 
 def swap(medoids, A, B):
     tmp = medoids
@@ -21,31 +40,68 @@ def swap(medoids, A, B):
 def swap_single(medoids, a, b):
     return [b if i == a else i for i in medoids]
 
-def search_for_swap(medoids, distances, p=2, epsilon=0):
+def search_for_swap(medoids, obj, distances, FDict, p=2, delta=0):
     n = distances.shape[0]
-    prev_objective = objective(medoids, distances)
+    prev_objective = obj
+    medoids_list = list(medoids)
+    medoids_c_list = [i for i in range(n) if i not in medoids]
+    medoids_c = frozenset(medoids_c_list)
+    
     improvement = False
-    for A in itertools.combinations(medoids, p):
-        for B in itertools.combinations(range(n), p):
-            new_medoids = swap(medoids, A, B)
-            if objective(new_medoids, distances) < prev_objective - epsilon:
-                improvement = True
-                print("Improvement: %f" % (prev_objective - objective(new_medoids, distances)))
-                return new_medoids, improvement
-    return medoids, improvement
+    for q in range(1,p+1):
+        if q == 1:
+            for A in itertools.combinations(medoids, q):
+                for B in itertools.combinations(medoids_c, q):
+                    new_medoids = medoids.difference(A).union(B)
+                    new_objective = objective(new_medoids, distances, FDict)
+                    if new_objective < (1.0-delta)*prev_objective:
[…]

### 2015-05-06 src/multiswaps.py
All Flat Results

@@ -2,6 +2,8 @@ import random
 import itertools
 import numpy as np
 from scipy.special import binom
+import LoadpMedian
+import kmedoids
 
 from sklearn.datasets.samples_generator import make_blobs
 from sklearn.metrics.pairwise import pairwise_distances
@@ -49,27 +51,29 @@ def search_for_swap(medoids, obj, distances, FDict, p=2, delta=0):
     
     improvement = False
     for q in range(1,p+1):
-        if q == 1:
-            for A in itertools.combinations(medoids, q):
-                for B in itertools.combinations(medoids_c, q):
-                    new_medoids = medoids.difference(A).union(B)
-                    new_objective = objective(new_medoids, distances, FDict)
-                    if new_objective < (1.0-delta)*prev_objective:
-                        improvement = True
-                        #print("Improvement: %f" % (prev_objective - new_objective))
-                        return new_medoids, new_objective, improvement
-        else:
-            iters = 0
-            while iters < n*q:
-                iters += 1
-                A = np.random.choice(medoids_list,size=q,replace=False)
-                B = np.random.choice(medoids_c_list,size=q,replace=False)
-                new_medoids = medoids.difference(A).union(B)
-                new_objective = objective(new_medoids, distances, FDict)
-                if new_objective < (1.0-delta)*prev_objective:
-                    improvement = True
-                    #print("Improvement: %f" % (prev_objective - new_objective))
-                    return new_medoids, new_objective, improvement
+        if q > len(medoids):
+            break
+        #    for A in itertools.combinations(medoids, q):
+        #        for B in itertools.combinations(medoids_c, q):
+        #            new_medoids = medoids.difference(A).union(B)
+        #            new_objective = objective(new_medoids, distances, FDict)
+        #            if new_objective < (1.0-delta)*prev_objective:
+        #                improvement = True
+        #                #print("Improvement: %f" % (prev_objective - new_objective))
+        #                return new_medoids, new_objective, improvement
+        #else:
+        #print q
+        iters = 0
+        while iters < n*q:
+            iters += 1
+            A = np.random.choice(medoids_list,size=q,replace=False)
+            B = np.random.choice(medoids_c_list,size=q,replace=False)
[…]

### 2015-05-06 src/submodular_demo.py

@@ -149,25 +149,18 @@ def updateS(S, psiOld, distances, sorted_distances, H, F, delta, h_perc):
     return list(S), psiOld, False
 
 # find a random initial set represented as a frozenset
-def initialS(k,n,max_iter=1000):
-    X = frozenset([i for i in range(n)])
-    S = frozenset()
-    i = 0
-    while (len(S) < k):
-        j = np.random.choice(list(X))
-        X = X.difference([j])
-        S = S.union([j])
-    return S
+def initialS(k,n):
+    S = np.random.choice([i for i in range(n)], size=k, replace=False)
+    return frozenset(S)
 
 
 # supermodular looks for a locally optimal solution with a potential function
-def cluster(distances, k, warm_start= None, epsilon=5, h_perc=50, \
-                 max_iter=100):
+def cluster(distances, k, warm_start= None, epsilon=5, h_perc=50):
     # Find Initial S
     n = distances.shape[0]
     X = frozenset([i for i in range(n)])
     if warm_start is None:
-        S = initialS(n-k,n,max_iter)
+        S = initialS(n-k,n)
     else:
         S = X.difference(frozenset(warm_start))
 
@@ -213,7 +206,7 @@ if __name__ == '__main__':
     X, labels_true = make_blobs(n_samples=100, centers=centers, cluster_std=0.5,
                             random_state=999)
     distances = pairwise_distances(X)
-    print cluster(distances,3,None,10,100,100)
+    print cluster(distances,3,None,10,100)
 
     print "curvature is ", totalCurvature(distances)
 

### 2015-05-06 src/multiswaps.py
Fix print statement

@@ -56,7 +56,7 @@ def search_for_swap(medoids, obj, distances, FDict, p=2, delta=0):
                     new_objective = objective(new_medoids, distances, FDict)
                     if new_objective < (1.0-delta)*prev_objective:
                         improvement = True
-                        #print("Improvement: %f" % (prev_objective - new_objective))
+                        # print("Improvement: %f" % (prev_objective - new_objective))
                         return new_medoids, new_objective, improvement
         else:
             iters = 0
@@ -68,7 +68,7 @@ def search_for_swap(medoids, obj, distances, FDict, p=2, delta=0):
                 new_objective = objective(new_medoids, distances, FDict)
                 if new_objective < (1.0-delta)*prev_objective:
                     improvement = True
-                    #print("Improvement: %f" % (prev_objective - new_objective))
+                    # print("Improvement: %f" % (prev_objective - new_objective))
                     return new_medoids, new_objective, improvement
                 
     return medoids, prev_objective, improvement
@@ -101,7 +101,7 @@ if __name__ == '__main__':
     X, labels_true = make_blobs(n_samples=900, centers=centers, cluster_std=0.5,
                             random_state=999)
     distances = pairwise_distances(X)
-    print cluster(distances,3,2,[0,1,2],0.1)
+    print(cluster(distances,3,2,[0,1,2],0.1))
 
 
 

### 2015-05-06 src/submodular_demo.py

@@ -212,6 +212,6 @@ if __name__ == '__main__':
     X, labels_true = make_blobs(n_samples=100, centers=centers, cluster_std=0.5,
                             random_state=999)
     distances = pairwise_distances(X)
-    print cluster(distances,3,None,10,100,100)
+    print(cluster(distances,3,None,10,100,100))
 
-    print "curvature is ", totalCurvature(distances)
+    print("curvature is ", totalCurvature(distances))

### 2015-05-06 src/testing.py

@@ -39,16 +39,17 @@ def test_clustering(cluster_alg, distances, labels_true, N=10):
 
 if __name__ == '__main__':
     
-    X, y, n, k = LoadData("data/Real Data Sets/soybean-small.data.txt", cluster_loc=0, split=0)
+    X, y, n, k = LoadData("data/Gaussian2/Gauss_7_5_10.txt", cluster_loc=1, split=1)
+    # X, y, n, k = LoadData("data/Real Data Sets/soybean-small.data.txt", cluster_loc=0, split=0)
     distances = pairwise_distances(X)
 
-    # p = 3
-    # multi_swaps = lambda: multiswaps.cluster(distances, k, p, epsilon=1)
-    # avg_tim, avg_obj, avg_ami, min_tim, min_obj, min_ami, max_tim, \
-    #          max_obj, max_ami = test_clustering(multi_swaps, distances, y, N=10)
-    # print("%s-swap Average: Time: %f Objective: %f AMI: %f" % (p, avg_tim, avg_obj, avg_ami))
-    # print("%s-swap Best Case: Time: %f Objective: %f AMI: %f" % (p, min_tim, min_obj, max_ami))
-    # print("%s-swap Worst Case: Time: %f Objective: %f AMI: %f" % (p, max_tim, max_obj, min_ami))
+    p = 3
+    multi_swaps = lambda: multiswaps.cluster(distances, k, p, epsilon=1)
+    avg_tim, avg_obj, avg_ami, min_tim, min_obj, min_ami, max_tim, \
+             max_obj, max_ami = test_clustering(multi_swaps, distances, y, N=1)
+    print("%s-swap Average: Time: %f Objective: %f AMI: %f" % (p, avg_tim, avg_obj, avg_ami))
+    print("%s-swap Best Case: Time: %f Objective: %f AMI: %f" % (p, min_tim, min_obj, max_ami))
+    print("%s-swap Worst Case: Time: %f Objective: %f AMI: %f" % (p, max_tim, max_obj, min_ami))
 
     # k_medoids = lambda: kmedoids.cluster(distances, k=k)
     # avg_obj, avg_ami, min_obj, min_ami, max_obj, max_ami = test_clustering(k_medoids, distances, y, N=10)
@@ -59,11 +60,11 @@ if __name__ == '__main__':
     # for i in range(2):
     #     submodular_demo.supermodular_list(X, k, h_perc=10)
 
[…]

### 2015-05-08 tex/project.tex

@@
 
+\begin{figure}[h]
+\begin{tabular}{ | l | l | l | l | l | }
+\hline
+	Data Set Name &Optimal Objective & $k$ & $n$ & dim   \\ \hline
+	Soybean & 113.8 & 4 & 47 & 35 \\ 
+	Iris & 98.2 & 3 & 150 & 4 \\ 
+	Gauss\_5\_5\_5 & 195.4 & 5 & 388 & 5 \\ 
+	Gauss\_5\_10\_5 & 270 & 10 & 544 & 5 \\ 
+	Gauss\_5\_50\_5 & 178.1 & 50 & 408 & 5 \\ 
+	pmed1 & 5819 & 5 & 100 & \  \\ 
+	\hline
+\end{tabular}
+\caption{Some specifics about our datasets}
+\end{figure}
+
+\begin{figure*}[h]
+\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset Name & Time & Time & Time & Time & Time & Time & Time & Time & Time \\ \hline
+	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap  \\ \hline
+	Soybean & 0.015 & 0.061 & 0.002 & 1.179 & 0.026 & 0.016 & 0.048 & 0.049 & 0.081 \\ 
+	Iris & 0.013 & 0.023 & 0.003 & 9.788 & 0.125 & 0.102 & 0.212 & 0.229 & 0.414 \\ 
+	Gauss\_5\_5\_5 & 0.025 & 0.056 & 0.032 & 75.262 & 1.159 & 1.506 & 2.188 & 2.382 & 3.71 \\ 
+	Gauss\_5\_10\_5 & 0.054 & 0.095 & 0.109 & 134.137 & 6.834 & 5.407 & 9.243 & 9.428 & 15.96 \\ 
+	Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 \\ 
+	pmed1 & N/A & N/A & 0.003 & 5.042 & 0.084 & 0.09 & 0.155 & 0.161 & 0.255 \\ \hline
+\end{tabular}
+
+\caption{Running Times For Algorithms}
+\end{figure*}
+
+\begin{figure*}[h]
+\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset Name & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
+	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap   \\ \hline
+	Soybean & 0.69 & 0.18 & 0.55 & 0.61 & 0.71 & 0.73 & 0.72 & 0.71 & 0.71  \\ 
+	Iris & 0.75 & 0.76 & 0.69 & 0.7 & 0.77 & 0.76 & 0.76 & 0.76 & 0.76  \\ 
+	Gauss\_5\_5\_5 & 0.74 & 0.71 & 0.69 & 0.66 & 0.74 & 0.75 & 0.74 & 0.74 & 0.74  \\ 
+	Gauss\_5\_10\_5 & 0.78 & 0.74 & 0.68 & 0.58 & 0.73 & 0.72 & 0.74 & 0.74 & 0.73  \\ 
+	Gauss\_5\_50\_5 & 0.65 & 0.6 & 0.62 & 0.57 & 0.62 & 0.62 & 0.63 & 0.63 & 0.62  \\ \hline
+\end{tabular}
+
+\caption{AMI For Algorithms}
+\end{figure*}
+
+\begin{figure*}[h]
+\begin{tabular}{ | l | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset Name & Approx & Approx & Approx & Approx & Approx & Approx & Approx  \\ \hline
+	& k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap   \\ \hline
+	Soybean & 1.17 & 1.2 & 1.03 & 1.01 & 1.02 & 1.01 & 1.02  \\ 
+	Iris & 1.08 & 1.37 & 1.02 & 1 & 1.01 & 1 & 1.01  \\ 
+	Gauss\_5\_5\_5 & 1.07 & 1.44 & 1.03 & 1 & 1.03 & 1.01 & 1.03  \\ 
+	Gauss\_5\_10\_5 & 1.08 & 1.39 & 1.02 & 1.01 & 1.02 & 1.01 & 1.02  \\ 
+	Gauss\_5\_50\_5 & 1.09 & 1.27 & 1.03 & 1.02 & 1.03 & 1.01 & 1.03  \\ 
+	pmed1 & 1.11 & 1.41 & 1.03 & 1.01 & 1.03 & 1.01 & 1.02  \\ \hline
+\end{tabular}
+
+\caption{Approximation Coeffiecients For Algorithms}
+\end{figure*}
+
 \section{Data Sets}
@@
 It contains data on 150 instances of flowers, each of with belongs to one of three species of iris.
-One of the species is linearly seperable from the others, but the other two are not.
+One of the species is linearly separable from the others, but the other two are not.
 In total, there are 150 instances and 4 attributes.

### 2015-05-08 tex/project.tex
Some implemenation and results work

@@
 
-\subsection{Implementation}
+\subsection{Implementation and Results}
 
-We will compare the above algorithms against $k$-means, $k$-medoids, PAM, affinity propagation, and Shi and Meila's spectral algorithm.
-All these methods except $k$-medoids and PAM are implemented in python's scikit-learn package.
-For comparisons against affinity propagation, we will first observe how many clusters the algorithm outputs and use that as input for our proposed algorithms.
-We implement the above local search algorithm for $p=2, 3,4$ so the computation time is reasonable.
+We implement the above local search algorithm for $p=1, 2, 3$ so the computation time is reasonable, and compare the above algorithms against $k$-medoids and Shi and Meila's spectral algorithm in time, AMI, and approximation ratio.
+Note that for $p=1$, the local search algorithm is equivalent to PAM.
+In order to calculate approximation ratios, Gurobi was used to compute the global optimum for each data set.
 
-Given a true clustering, we will use the adjusted mutual information (AMI) measure to score the accuracy of each method.
-AMI measures the agreement of two assignments and is normalized to reflect chance so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation \cite{Vinh}.
-For each data set, we will also report the $k$-median objective function value for that clustering, the guaranteed approximation ratio, and the computation time.
+To achieve reasonable computation times for the local search and supermodular function minimization algorithms, at each iteration we first consider nearby swaps, as well as placing
 
-The data sets we chose to use for comparison were
+We also considered a hybrid heuristic algorithm where we ran $k$-medoids on the result of the local search algorithm.
+This is motivated by the fact that, similar to Lloyd's algorithm, $k$-medoids will quickly converge to local optima.
+So
 
-NEED TO DECIDE:
-OR library $p$-median.
+When information on true labels is available, we use adjusted mutual information (AMI) measure to score accuracy of predicted clusters.
+AMI measures the agreement of two assignments and is normalized to reflect chance so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
+See \cite{Vinh} for more information.
+For each data set and each clustering method, we also report objective function values, the approximation ratio, and the computation time.
 
-\subsection{Results}
-
-\section{Hierarchical Clustering Approximation Algorithms}
-Again, suppose we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$.
-In the hierarchical setting, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
-Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$.
-We define
-\[ \mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i)) . \]
-
-Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if 
-\begin{enumerate}
-\item $S_1 \subset S_2$, 
-\item  $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, and 
-\item $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
-\end{enumerate}
- A solution to the hierarchical clustering problem is therefore a chain of nested solutions.
-However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the algorithm.
-Consider any $k$-median algorithm $\mathcal{A}$, such as the algorithms above.
-After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce  augmentation step that will nest these clusterings so that they fit into the hierarchical framework and only increase the costs by a bounded factor.
-
-First, we set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
-Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
-The cost of these solutions is determined by assigning each point to the nearest cluster center.
-We then order these solutions according to their cost into buckets of the form $[0,\beta_0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
-
-From each non-empty bucket, pick the solution with the smallest number of cluster centers and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r = V_n$.
-We then inductively build up our nested solutions in the following manner.
-Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_{i+1}$ by iteratively adding the closest point $j \in S_{i-1}$ to $i$ for all $i \in \overline{V}_i$.
-Note that $|S_{i}| \leq |\overline{V}_i|$.
-
-We then define an assignment $a_i$.
-For each $j \in X$ such that $a_{i-1}(j) \in S_{i}$, we set $a_i(j) = a_{i-1}(j)$.
-Otherwise, we set $a_i(j)$ to be the facility $i \in S_i$ such that assigning all points in $a_{i-1}^{-1}(j)$ (i.e.
-all points in $j$'s cluster in $S_{i-1}$) would have minimum cost.
-Using this assignment, $(a_i, S_i)$ is nested in $(a_{i-1}, S_{i-1})$.
-
-We continue this nesting until $|S_i| = 1$ and we have reached the top level of the hierarchy.
-After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantee that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq n-k+1 \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).
-
-We can test this hierarchical framework for both the multiswap and supermodular local search algorithms above.
-
-\subsection{Implementation}
-
-\subsection{Results}
+The data sets tested against were:
+\begin{description}
+\item[Gaussian] Synthetic $k$-center Gaussian data, generated in Python.  Add info on std deviation, size, etc.
+\item[P-median] ADD CITATION OR Library p-median data sets
+\item[UCI Iris] \cite{Iris} Contains 4 attributes on 150 instances of flowers, each of with belongs to one of three species of iris.
+One of the species is linearly separable from the others, but the other two are not.
+One of the most used data sets in pattern recognition.
+\item[UCI Soybean (Small)] \cite{Soybean} 47 instances, with 35 attributes data set of soybean disease from UCI.
+\end{description}
 
@@
 \end{figure}
+
+Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on an 8 core ....
 
@@
 
-\section{Data Sets}
-In order to empirically test these clustering methods, we used four different data sets.  
-\begin{description}
-\item[Generated Gaussian Data] We wanted to test our clustering algorithms on data that had data nicely clustered around $k$ centers.
-So, we generated synthetic $k$-center Gaussian data algorithmically and used this for testing.
-\item[$P$-median]
-\item[UCI Iris Dataset] This is one of the most used data sets in pattern recognition (\cite{Iris}).
-It contains data on 150 instances of flowers, each of with belongs to one of three species of iris.
-One of the species is linearly separable from the others, but the other two are not.
-In total, there are 150 instances and 4 attributes.
-\item[UCI Soybean Dataset] This is another small, real data set from UCI.  This, too, is widely used in machine learning.  It has 47 instances, with 35 attributes.
-\end{description}
+Comment on results here.
 
-\
+\section{Hierarchical Clustering Approximation Algorithms}
+Suppose we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$.
+In the hierarchical setting, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
+Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$.
+We define
+\[ \mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i)) . \]
+
+Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if 
+\begin{enumerate}
+\item $S_1 \subset S_2$, 
+\item  $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, and 
+\item $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
+\end{enumerate}
+ A solution to the hierarchical clustering problem is therefore a chain of nested solutions.
+However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the algorithm.
+Consider any $k$-median algorithm $\mathcal{A}$, such as the algorithms above.
+After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce  augmentation step that will nest these clusterings so that they fit into the hierarchical framework and only increase the costs by a bounded factor.
+
+First, we set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
+Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
+The cost of these solutions is determined by assigning each point to the nearest cluster center.
+We then order these solutions according to their cost into buckets of the form $[0,\beta_0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
+
+From each non-empty bucket, pick the solution with the smallest number of cluster centers and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r = V_n$.
+We then inductively build up our nested solutions in the following manner.
+Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_{i+1}$ by iteratively adding the closest point $j \in S_{i-1}$ to $i$ for all $i \in \overline{V}_i$.
+Note that $|S_{i}| \leq |\overline{V}_i|$.
+
+We then define an assignment $a_i$.
+For each $j \in X$ such that $a_{i-1}(j) \in S_{i}$, we set $a_i(j) = a_{i-1}(j)$.
+Otherwise, we set $a_i(j)$ to be the facility $i \in S_i$ such that assigning all points in $a_{i-1}^{-1}(j)$ (i.e.
+all points in $j$'s cluster in $S_{i-1}$) would have minimum cost.
+Using this assignment, $(a_i, S_i)$ is nested in $(a_{i-1}, S_{i-1})$.
+
+We continue this nesting until $|S_i| = 1$ and we have reached the top level of the hierarchy.
+After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantee that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq n-k+1 \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).
+
+We test this hierarchical framework for both the multiswap and supermodular local search algorithms above.
+
+\subsection{Implementation}
+
+\subsection{Results}
+
 \label{submission}

### 2015-05-08 tex/project.tex
More results and implementation

@@
 
-We implement the above local search algorithm for $p=1, 2, 3$ so the computation time is reasonable, and compare the above algorithms against $k$-medoids and Shi and Meila's spectral algorithm in time, AMI, and approximation ratio.
-Note that for $p=1$, the local search algorithm is equivalent to PAM.
+We implemented the local search algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compare against $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
+Note that for $p=1$, the local search algorithm is equivalent to the well-known PAM algorithm.
 In order to calculate approximation ratios, Gurobi was used to compute the global optimum for each data set.
 
-To achieve reasonable computation times for the local search and supermodular function minimization algorithms, at each iteration we first consider nearby swaps, as well as placing
+CITATION FOR K-MEDOIDS CODE?
 
-We also considered a hybrid heuristic algorithm where we ran $k$-medoids on the result of the local search algorithm.
-This is motivated by the fact that, similar to Lloyd's algorithm, $k$-medoids will quickly converge to local optima.
-So
+To achieve reasonable computation times for the local search and supermodular function minimization algorithms, at each iteration we first consider nearby swaps and ensure that some minimum level progress is achieved.
+We also implemented a maximum of 1000 iterations for the $k$-median algorithm.
 
-When information on true labels is available, we use adjusted mutual information (AMI) measure to score accuracy of predicted clusters.
-AMI measures the agreement of two assignments and is normalized to reflect chance so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
-See \cite{Vinh} for more information.
-For each data set and each clustering method, we also report objective function values, the approximation ratio, and the computation time.
+We also considered a hybrid heuristic algorithm (labelled $p$-swap+ in the results) where we ran $k$-medoids on the result of the local search ($p$-swap) algorithm.
+This is motivated by the fact that, similar to Lloyd's algorithm, $k$-medoids will quickly converge to local minima.
+Since $p$-swaps avoids getting stuck in local minima, the hybrid algorithm should in theory achieve good results with minimal extra computation time over the regular local search algorithm.
+
+For each data set and each clustering method, we report objective function values, the approximation ratio, computation time, and adjusted mutual information (AMI).
+When information on true labels is available, AMI is a measure which scores accuracy of predicted clusters.
+AMI measures the agreement of two assignments and is normalized to reflect chance, so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
+See \cite{Vinh} for further information.
+The results reported are average values across 10 trials.
 
@@
 
+\section{Conclusion}
+
+\section{Acknowledgements}
+David Williamson and Karthik Sridharan?
+
 \label{submission}

### 2015-05-08 tex/project.tex
Added data appendix

--- 2015-05-08 tex/project.tex
@@
 
+\section{Appendix 1: Full Results}
+
+\subsection{Gaussian Data}
+\begin{figure*}[h]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l | }
+\hline
+	Dataset & NameOpt Obj & k & n & dim  \\ \hline
+	Gauss\_3\_5\_0 & 195 & 5 & 542 & 3 \\ 
+	Gauss\_3\_5\_5 & 225 & 5 & 620 & 3 \\ 
+	Gauss\_3\_5\_10 & 158 & 5 & 441 & 3 \\
+	Gauss\_3\_10\_0 & 147 & 10 & 472 & 3 \\
+	Gauss\_3\_10\_5 & 142 & 10 & 480 & 3 \\ 
+	Gauss\_3\_10\_10 & 201 & 10 & 557 & 3 \\ 
+	Gauss\_3\_50\_0 & 108 & 50 & 471 & 3 \\
+	Gauss\_3\_50\_5 & 94 & 50 & 422 & 3 \\ 
+	Gauss\_3\_50\_10 & 95 & 50 & 408 & 3 \\ 
+	Gauss\_5\_5\_0 & 352 & 5 & 720 & 5 \\ 
+	Gauss\_5\_5\_5 & 195 & 5 & 388 & 5 \\ 
+	Gauss\_5\_5\_10 & 214 & 5 & 394 & 5 \\ 
+	Gauss\_5\_10\_0 & 346 & 10 & 703 & 5 \\ 
+	Gauss\_5\_10\_5 & 270 & 10 & 544 & 5 \\ 
+	Gauss\_5\_10\_10 & 292 & 10 & 572 & 5 \\ 
+	Gauss\_5\_50\_0 & 188 & 50 & 445 & 5 \\ 
+	Gauss\_5\_50\_5 & 178 & 50 & 408 & 5 \\ 
+	Gauss\_5\_50\_10 & 229 & 50 & 480 & 5 \\ 
+	Gauss\_7\_5\_0 & 349 & 5 & 567 & 7 \\ 
+	Gauss\_7\_5\_5 & 368 & 5 & 579 & 7 \\ 
+	Gauss\_7\_5\_10 & 356 & 5 & 527 & 7 \\ 
+	Gauss\_7\_10\_0 & 351 & 10 & 573 & 7 \\ 
+	Gauss\_7\_10\_5 & 302 & 10 & 478 & 7 \\ 
+	Gauss\_7\_10\_10 & 354 & 10 & 534 & 7 \\ 
+	Gauss\_7\_50\_0 & 257 & 50 & 451 & 7 \\ 
+	Gauss\_7\_50\_5 & 256 & 50 & 430 & 7 \\ 
+	Gauss\_7\_50\_10 & 298 & 50 & 480 & 7 \\ \hline
+\end{tabular}
+\end{center}
+
+\caption{Details For Gaussian Datasets}
+\end{figure*}
+
+\begin{figure*}[h]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset &Time & Time & Time & Time & Time & Time & Time & Time & Time & Time   \\ \hline
+	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-swaps+ & 3-Swap & 4-swaps   \\ 
+	Gauss\_3\_5\_0 & 0.024 & 0.056 & 0.066 & 131.305 & 2.707 & 3.195 & 4.779 & 4.138 & 7.683 & 11.325 \\ 
+	Gauss\_3\_5\_5 & 0.032 & 0.076 & 0.099 & 195.013 & 3.682 & 3.764 & 6.639 & 5.32 & 9.627 & 14.575 \\ 
+	Gauss\_3\_5\_10 & 0.033 & 0.055 & 0.047 & 80.439 & 2.244 & 1.802 & 2.806 & 3.024 & 5.107 & 7.457 \\ 
+	Gauss\_3\_10\_0 & 0.041 & 0.083 & 0.095 & 121.551 & 5.133 & 3.576 & 7.452 & 6.3 & 10.157 & 11.872 \\ 
+	Gauss\_3\_10\_5 & 0.048 & 0.1 & 0.117 & 101.072 & 5.026 & 4.96 & 8.57 & 6.952 & 11.948 & 12.225 \\ 
+	Gauss\_3\_10\_10 & 0.05 & 0.106 & 0.153 & 149.352 & 7.101 & 6.456 & 11.273 & 10.646 & 18.21 & 17.414 \\ 
+	Gauss\_3\_50\_0 & 0.154 & 0.324 & 0.435 & 175.827 & 29.201 & 34.228 & 43.65 & 49.964 & 48.259 & 66.413 \\ 
+	Gauss\_3\_50\_5 & 0.151 & 0.253 & 0.376 & 127.64 & 26.318 & 21.541 & 38.004 & 33.213 & 45.862 & 57.262 \\ 
+	Gauss\_3\_50\_10 & 0.153 & 0.265 & 0.301 & 132.119 & 21.341 & 22.556 & 35.389 & 32.139 & 44.361 & 50.143 \\ 
+	Gauss\_5\_5\_0 & 0.026 & 0.079 & 0.08 & 230.345 & 4.997 & 4.21 & 9.428 & 8.063 & 12.811 & 19.337 \\ 
+	Gauss\_5\_5\_5 & 0.025 & 0.056 & 0.032 & 75.262 & 1.159 & 1.506 & 2.188 & 2.382 & 3.71 & 5.797 \\ 
+	Gauss\_5\_5\_10 & 0.025 & 0.042 & 0.027 & 69.349 & 1.383 & 1.458 & 2.403 & 2.575 & 3.811 & 5.748 \\ 
+	Gauss\_5\_10\_0 & 0.044 & 0.1 & 0.18 & 227.747 & 11.633 & 12.129 & 17.724 & 18.654 & 26.563 & 33.646 \\ 
+	Gauss\_5\_10\_5 & 0.054 & 0.095 & 0.109 & 134.137 & 6.834 & 5.407 & 9.243 & 9.428 & 15.96 & 21.795 \\ 
+	Gauss\_5\_10\_10 & 0.045 & 0.092 & 0.117 & 127.614 & 7.639 & 6.456 & 11.807 & 12.186 & 18.339 & 23.554 \\ 
+	Gauss\_5\_50\_0 & 0.142 & 0.211 & 0.389 & 139.818 & 30.155 & 27.431 & 37.251 & 43.421 & 50.192 & 68.329 \\ 
+	Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 & 53.546 \\ 
+	Gauss\_5\_50\_10 & 0.149 & 0.219 & 0.417 & 170.431 & 37.132 & 35.119 & 45.341 & 47.002 & 56.683 & 69.712 \\ 
+	Gauss\_7\_5\_0 & 0.019 & 0.054 & 0.052 & 185.865 & 2.975 & 3.565 & 4.954 & 5.005 & 8.531 & 11.867 \\ 
+	Gauss\_7\_5\_5 & 0.031 & 0.078 & 0.047 & 154.1 & 2.603 & 3.04 & 5.45 & 5.738 & 8.056 & 12.505 \\ 
+	Gauss\_7\_5\_10 & 0.018 & 0.052 & 0.043 & 130.185 & 2.949 & 2.478 & 4.668 & 4.849 & 7.324 & 11.036 \\ 
+	Gauss\_7\_10\_0 & 0.045 & 0.071 & 0.097 & 185.521 & 8.015 & 7.338 & 12.495 & 10.474 & 16.174 & 20.671 \\ 
+	Gauss\_7\_10\_5 & 0.04 & 0.063 & 0.073 & 136.189 & 5.315 & 5.365 & 9.287 & 6.87 & 11.768 & 16.072 \\ 
+	Gauss\_7\_10\_10 & 0.048 & 0.09 & 0.1 & 168.962 & 7.047 & 6.538 & 9.783 & 9.322 & 13.581 & 19.41 \\ 
+	Gauss\_7\_50\_0 & 0.143 & 0.199 & 0.401 & 152.958 & 32.592 & 37.747 & 39.026 & 42.999 & 60.483 & 69.667 \\ 
+	Gauss\_7\_50\_5 & 0.147 & 0.212 & 0.327 & 136.496 & 27.8 & 26.288 & 38.683 & 34.335 & 47.779 & 59.414 \\ 
+	Gauss\_7\_50\_10 & 0.149 & 0.231 & 0.426 & 163.651 & 37.968 & 41.976 & 55.514 & 50.175 & 56.302 & 79.819 \\ \hline
+\end{tabular}
+
+\end{center}
+
+\caption{Running Times For Gaussian Datasets}
+\end{figure*}
+
+\begin{figure*}[h]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset&AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
+	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap & 4-swaps   \\ \hline
+	Gauss\_3\_5\_0 & 0.884 & 0.887 & 0.729 & 0.678 & 0.88 & 0.876 & 0.884 & 0.876 & 0.881 & 0.882 \\ 
+	Gauss\_3\_5\_5 & 0.7 & 0.62 & 0.637 & 0.545 & 0.683 & 0.68 & 0.685 & 0.69 & 0.682 & 0.686 \\ 
+	Gauss\_3\_5\_10 & 0.68 & 0.618 & 0.62 & 0.504 & 0.663 & 0.665 & 0.659 & 0.659 & 0.664 & 0.668 \\ 
+	Gauss\_3\_10\_0 & 0.693 & 0.641 & 0.699 & 0.619 & 0.692 & 0.692 & 0.688 & 0.696 & 0.685 & 0.685 \\ 
+	Gauss\_3\_10\_5 & 0.534 & 0.466 & 0.516 & 0.446 & 0.534 & 0.533 & 0.535 & 0.532 & 0.531 & 0.536 \\ 
+	Gauss\_3\_10\_10 & 0.669 & 0.56 & 0.616 & 0.516 & 0.651 & 0.653 & 0.652 & 0.657 & 0.657 & 0.653 \\ 
+	Gauss\_3\_50\_0 & 0.504 & 0.498 & 0.498 & 0.475 & 0.501 & 0.497 & 0.491 & 0.497 & 0.495 & 0.501 \\ 
+	Gauss\_3\_50\_5 & 0.407 & 0.412 & 0.402 & 0.396 & 0.392 & 0.385 & 0.389 & 0.386 & 0.387 & 0.385 \\ 
+	Gauss\_3\_50\_10 & 0.421 & 0.424 & 0.406 & 0.39 & 0.413 & 0.414 & 0.408 & 0.404 & 0.405 & 0.41 \\ 
+	Gauss\_5\_5\_0 & 0.99 & 0.99 & 0.858 & 0.719 & 0.988 & 0.994 & 0.989 & 0.994 & 0.989 & 0.985 \\ 
+	Gauss\_5\_5\_5 & 0.741 & 0.713 & 0.687 & 0.657 & 0.742 & 0.749 & 0.741 & 0.744 & 0.737 & 0.747 \\ 
+	Gauss\_5\_5\_10 & 0.833 & 0.83 & 227.742 & 0.663 & 0.76 & 0.779 & 0.806 & 0.747 & 0.789 & 0.77 \\ 
+	Gauss\_5\_10\_0 & 0.974 & 0.972 & 0.858 & 0.732 & 0.96 & 0.964 & 0.964 & 0.965 & 0.96 & 0.964 \\ 
+	Gauss\_5\_10\_5 & 0.783 & 0.736 & 0.677 & 0.575 & 0.729 & 0.725 & 0.737 & 0.737 & 0.729 & 0.725 \\ 
+	Gauss\_5\_10\_10 & 0.878 & 0.829 & 0.795 & 0.704 & 0.873 & 0.875 & 0.873 & 0.872 & 0.875 & 0.868 \\ 
+	Gauss\_5\_50\_0 & 0.723 & 0.669 & 0.668 & 0.622 & 0.695 & 0.692 & 0.692 & 0.7 & 0.696 & 0.692 \\ 
+	Gauss\_5\_50\_5 & 0.646 & 0.596 & 0.621 & 0.57 & 0.622 & 0.619 & 0.627 & 0.631 & 0.625 & 0.626 \\
+	Gauss\_5\_50\_10 & 0.669 & 0.639 & 0.626 & 0.546 & 0.64 & 0.641 & 0.639 & 0.65 & 0.643 & 0.637 \\ 
+	Gauss\_7\_5\_0 & 1 & 1 & 0.852 & 0.773 & 0.999 & 1 & 0.999 & 1 & 0.999 & 0.998 \\ 
+	Gauss\_7\_5\_5 & 0.847 & 0.832 & 0.86 & 0.72 & 0.852 & 0.85 & 0.849 & 0.847 & 0.847 & 0.847 \\ 
+	Gauss\_7\_5\_10 & 0.979 & 0.963 & 0.839 & 0.744 & 0.981 & 0.979 & 0.98 & 0.979 & 0.978 & 0.978 \\ 
+	Gauss\_7\_10\_0 & 0.995 & 0.995 & 0.869 & 0.785 & 0.988 & 0.984 & 0.987 & 0.984 & 0.982 & 0.976 \\ 
+	Gauss\_7\_10\_5 & 0.988 & 0.994 & 0.853 & 0.732 & 0.988 & 0.994 & 0.99 & 0.994 & 0.993 & 0.994 \\ 
+	Gauss\_7\_10\_10 & 0.89 & 0.81 & 0.845 & 0.724 & 0.851 & 0.856 & 0.854 & 0.854 & 0.85 & 0.851 \\ 
+	Gauss\_7\_50\_0 & 0.867 & 0.834 & 0.784 & 0.693 & 0.832 & 0.835 & 0.84 & 0.833 & 0.835 & 0.846 \\ 
+	Gauss\_7\_50\_5 & 0.852 & 0.808 & 0.727 & 0.647 & 0.813 & 0.815 & 0.814 & 0.815 & 0.812 & 0.807 \\ 
+	Gauss\_7\_50\_10 & 0.829 & 0.761 & 0.742 & 0.645 & 0.807 & 0.799 & 0.807 & 0.804 & 0.804 & 0.799 \\ \hline
+\end{tabular}
+
+\end{center}
+
+\caption{AMI For Gaussian Datasets}
+\end{figure*}
+
+\begin{figure*}[h]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset & Approx & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
+	& k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swap+ & 3-Swap & 4-swaps   \\ \hline
+	Gauss\_3\_5\_0 & 1.159 & 1.481 & 1.013 & 1 & 1.011 & 1 & 1.013 & 1.009 \\ 
+	Gauss\_3\_5\_5 & 1.057 & 1.4 & 1.005 & 1.001 & 1.006 & 1.001 & 1.01 & 1.007 \\ 
+	Gauss\_3\_5\_10 & 1.047 & 1.377 & 1.007 & 1.004 & 1.014 & 1.007 & 1.009 & 1.008 \\ 
+	Gauss\_3\_10\_0 & 1.029 & 1.404 & 1.018 & 1.01 & 1.013 & 1.01 & 1.022 & 1.019 \\ 
+	Gauss\_3\_10\_5 & 1.063 & 1.329 & 1.025 & 1.013 & 1.021 & 1.012 & 1.02 & 1.026 \\ 
+	Gauss\_3\_10\_10 & 1.06 & 1.369 & 1.017 & 1.009 & 1.018 & 1.007 & 1.016 & 1.016 \\ 
+	Gauss\_3\_50\_0 & 1.112 & 1.354 & 1.05 & 1.026 & 1.037 & 1.025 & 1.049 & 1.044 \\ 
+	Gauss\_3\_50\_5 & 1.124 & 1.327 & 1.045 & 1.025 & 1.041 & 1.019 & 1.044 & 1.04 \\ 
+	Gauss\_3\_50\_10 & 1.121 & 1.316 & 1.044 & 1.023 & 1.037 & 1.021 & 1.043 & 1.043 \\ 
+	Gauss\_5\_5\_0 & 1.105 & 1.576 & 1.013 & 1 & 1.011 & 1 & 1.016 & 1.015 \\ 
+	Gauss\_5\_5\_5 & 1.07 & 1.436 & 1.03 & 1.002 & 1.028 & 1.005 & 1.026 & 1.024 \\ 
+	Gauss\_5\_5\_10 & 0.003 & 1.369 & 1.021 & 1.004 & 1.016 & 1.007 & 1.016 & 1.015 \\ 
+	Gauss\_5\_10\_0 & 1.148 & 1.619 & 1.029 & 1 & 1.026 & 1.001 & 1.028 & 1.029 \\ 
+	Gauss\_5\_10\_5 & 1.08 & 1.389 & 1.017 & 1.011 & 1.018 & 1.009 & 1.017 & 1.019 \\ 
+	Gauss\_5\_10\_10 & 1.152 & 1.458 & 1.022 & 1.006 & 1.023 & 1.007 & 1.021 & 1.026 \\
+	Gauss\_5\_50\_0 & 1.096 & 1.271 & 1.031 & 1.017 & 1.033 & 1.015 & 1.031 & 1.029 \\ 
+	Gauss\_5\_50\_5 & 1.089 & 1.272 & 1.03 & 1.02 & 1.027 & 1.014 & 1.033 & 1.032 \\ 
+	Gauss\_5\_50\_10 & 1.09 & 1.299 & 1.029 & 1.017 & 1.036 & 1.016 & 1.028 & 1.036 \\ 
+	Gauss\_7\_5\_0 & 1.214 & 1.59 & 1.016 & 1 & 1.018 & 1 & 1.014 & 1.016 \\ 
+	Gauss\_7\_5\_5 & 1.046 & 1.446 & 1.017 & 1.009 & 1.014 & 1.004 & 1.014 & 1.019 \\ 
+	Gauss\_7\_5\_10 & 1.153 & 1.53 & 1.01 & 1 & 1.007 & 1 & 1.01 & 1.009 \\ 
+	Gauss\_7\_10\_0 & 1.132 & 1.504 & 1.032 & 1 & 1.023 & 1 & 1.032 & 1.034 \\ 
+	Gauss\_7\_10\_5 & 1.159 & 1.569 & 1.031 & 1 & 1.028 & 1 & 1.027 & 1.024 \\ 
+	Gauss\_7\_10\_10 & 1.074 & 1.56 & 1.024 & 1.003 & 1.019 & 1.005 & 1.02 & 1.02 \\ 
+	Gauss\_7\_50\_0 & 1.104 & 1.321 & 1.029 & 1.01 & 1.031 & 1.011 & 1.027 & 1.026 \\ 
+	Gauss\_7\_50\_5 & 1.132 & 1.326 & 1.033 & 1.013 & 1.031 & 1.013 & 1.026 & 1.036 \\ 
+	Gauss\_7\_50\_10 & 1.107 & 1.345 & 1.027 & 1.014 & 1.025 & 1.012 & 1.031 & 1.026 \\ \hline
+\end{tabular}
+
+\end{center}
+
+\caption{Approximation Constants For Gaussian Datasets}
+\end{figure*}
+
+\subsection{P-Median}
+
+\begin{figure*}[h]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | }
+\hline
+	Dataset & Optimal Objective & $k$ & $n$ \\ \hline
+	pmed1 & 5819 & 5 & 100 \\ 
+	pmed2 & 4093 & 10 & 100 \\ 
+	pmed3 & 4250 & 10 & 100 \\ 
+	pmed4 & 3034 & 20 & 100 \\
+	pmed5 & 1355 & 33 & 100 \\ 
+	pmed6 & 7824 & 5 & 200 \\ 
+	pmed7 & 5631 & 10 & 200 \\ 
+	pmed8 & 4445 & 20 & 200 \\ 
+	pmed9 & 2734 & 40 & 200 \\ 
+	pmed10 & 1255 & 67 & 200 \\ 
+	pmed11 & 7696 & 5 & 300 \\ 
+	pmed12 & 6634 & 10 & 300 \\ 
+	pmed13 & 4374 & 30 & 300 \\ 
+	pmed14 & 2968 & 60 & 300 \\ 
+	pmed15 & 1729 & 100 & 300 \\
+	pmed16 & 8162 & 5 & 400 \\ 
+	pmed17 & 6999 & 10 & 400 \\
+	pmed18 & 4809 & 40 & 400 \\ 
+	pmed19 & 2845 & 80 & 400 \\ 
+	pmed20 & 1789 & 133 & 400 \\ 
+	pmed21 & 9138 & 5 & 500 \\ 
+	pmed22 & 8579 & 10 & 500 \\ 
+	pmed23 & 4619 & 50 & 500 \\ 
+	pmed24 & 2961 & 100 & 500 \\ 
+	pmed25 & 1828 & 167 & 500 \\ 
+	pmed26 & 9917 & 5 & 600 \\ 
+	pmed27 & 8307 & 10 & 600 \\ 
+	pmed28 & 4498 & 60 & 600 \\ 
+	pmed29 & 3033 & 120 & 600 \\
+	pmed30 & 1989 & 200 & 600 \\ 
+	pmed31 & 10086 & 5 & 700 \\ 
+	pmed32 & 9297 & 10 & 700 \\ 
+	pmed33 & 4700 & 60 & 700 \\ 
+	pmed34 & 3013 & 140 & 700 \\ 
+	pmed35 & 10400 & 5 & 800 \\ 
+	pmed36 & 9934 & 10 & 800 \\ 
+	pmed37 & 5057 & 80 & 800 \\ 
+	pmed38 & 11060 & 5 & 900 \\ 
+	pmed39 & 9423 & 10 & 900 \\ 
+	pmed40 & 5128 & 90 & 900 \\ \hline
+\end{tabular}
+
+\end{center}
+
+\caption{Details for P-Median Datasets}
+\end{figure*}
+
+\begin{figure*}[h]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset & Time & Time & Time & Time & Time & Time & Time \\ \hline
+	&k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap   \\ \hline
+	pmed1 & 0.0032 & 5 & 0.1 & 0.1 & 0.2 & 0.2 & 0.3 \\ 
+	pmed2 & 0.0069 & 6.4 & 0.2 & 0.2 & 0.3 & 0.3 & 0.4 \\ 
+	pmed3 & 0.0067 & 7.5 & 0.2 & 0.2 & 0.3 & 0.3 & 0.4 \\ 
+	pmed4 & 0.0125 & 8 & 0.3 & 0.4 & 0.5 & 0.6 & 0.6 \\ 
+	pmed5 & 0.0381 & 8.3 & 0.5 & 0.6 & 0.7 & 0.7 & 0.9 \\ 
+	pmed6 & 0.0065 & 19.3 & 0.3 & 0.4 & 0.5 & 0.6 & 0.8 \\ 
+	pmed7 & 0.0117 & 23.9 & 0.7 & 0.9 & 0.9 & 1.3 & 1.5 \\ 
+	pmed8 & 0.027 & 27.2 & 1.4 & 1.8 & 2.2 & 2.6 & 3.1 \\ 
+	pmed9 & 0.0642 & 29 & 3.4 & 3.5 & 3.3 & 4.8 & 4.7 \\ 
+	pmed10 & 0.7329 & 39.8 & 4.4 & 5.8 & 6.2 & 7.3 & 8.2 \\ 
+	pmed11 & 0.0124 & 42 & 0.6 & 0.8 & 1.4 & 1.2 & 2.2 \\ 
+	pmed12 & 0.0186 & 45.7 & 1.7 & 2 & 2.3 & 2.4 & 4.3 \\ 
+	pmed13 & 0.0685 & 69.1 & 6.6 & 5.3 & 7 & 8.4 & 11 \\ 
+	pmed14 & 0.2238 & 76.3 & 13.6 & 14.2 & 17.1 & 17.4 & 25 \\ 
+	pmed15 & 1.0969 & 100.2 & 18.4 & 19.7 & 23.1 & 28.7 & 35.5 \\ 
+	pmed16 & 0.0251 & 87.1 & 1.3 & 1.2 & 2.3 & 2.4 & 3.8 \\ 
+	pmed17 & 0.0581 & 97.1 & 3.5 & 2.5 & 3.9 & 4.8 & 6.8 \\ 
+	pmed18 & 0.2596 & 134.7 & 15.3 & 16.5 & 23.1 & 23.3 & 31.3 \\ 
+	pmed19 & 1.0193 & 153.5 & 34.9 & 34.1 & 40.3 & 47.9 & 57.8 \\ 
+	pmed20 & 1.7947 & 175 & 52.2 & 45 & 68.6 & 68.2 & 79.3 \\ 
+	pmed21 & 0.0327 & 109.4 & 1.8 & 2.1 & 4.1 & 4.2 & 6.1 \\ 
+	pmed22 & 0.0786 & 113.1 & 4.3 & 5.6 & 8.3 & 7.5 & 12.7 \\ 
+	pmed23 & 0.5123 & 199.8 & 35.9 & 34.8 & 46.9 & 52.3 & 62.2 \\ 
+	pmed24 & 1.7861 & 245.9 & 66.6 & 90.3 & 99.2 & 88.5 & 125.5 \\ 
+	pmed25 & 3.1433 & 451.9 & 103 & 118 & 126.8 & 146.4 & 138.1 \\ 
+	pmed26 & 0.0503 & 159.8 & 2.8 & 2.4 & 4.7 & 5.6 & 6.9 \\ 
+	pmed27 & 0.1265 & 183.4 & 7.1 & 6.1 & 11.7 & 10.7 & 13 \\ 
+	pmed28 & 0.8738 & 263.3 & 60.3 & 71.3 & 95.6 & 98.5 & 89.3 \\ 
+	pmed29 & 2.8418 & 424.3 & 152.6 & 146.8 & 196.5 & 172.8 & 167 \\ 
+	pmed30 & 5.2891 & 571 & 175 & 207.5 & 254.2 & 249.7 & 255.7 \\ 
+	pmed31 & 0.0728 & 225.6 & 4.9 & 4.4 & 8.2 & 6.6 & 9.2 \\ 
+	pmed32 & 0.1528 & 249.5 & 11 & 7.8 & 14.3 & 12.9 & 19 \\ 
+	pmed33 & 1.4723 & 440.1 & 112.8 & 91.9 & 143.9 & 158.5 & 154.1 \\ 
+	pmed34 & 3.6669 & 591.7 & 234.8 & 217.8 & 233.1 & 287.8 & 256.8 \\ 
+	pmed35 & 0.1133 & 298.3 & 4.8 & 5.9 & 7.5 & 9.3 & 11.9 \\ 
+	pmed36 & 0.2221 & 259.8 & 13 & 11.4 & 15.9 & 24 & 23.6 \\ 
+	pmed37 & 2.4409 & 571 & 190.5 & 149.7 & 172.8 & 282.3 & 224.7 \\ 
+	pmed38 & 0.1326 & 349.9 & 7.1 & 6.5 & 9.8 & 12.2 & 14.5 \\ 
+	pmed39 & 0.2676 & 419.8 & 16.5 & 13.6 & 22.6 & 25.5 & 29.4 \\ 
+	pmed40 & 3.5269 & 849.9 & 251.3 & 264.9 & 250.4 & 391.7 & 350.2 \\ \hline
+\end{tabular}
+
+\end{center}
+
+\caption{Running Time for P-Median Datasets}
+\end{figure*}
+
+\begin{figure*}[h]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l | l | l | l | }
+\hline
+	Dataset & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
+	  & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap \\ \hline
+	pmed1 & 1.113 & 1.409 & 1.03 & 1.013 & 1.029 & 1.005 & 1.018 \\ 
+	pmed2 & 1.158 & 1.486 & 1.044 & 1.011 & 1.038 & 1.011 & 1.044 \\ 
+	pmed3 & 1.161 & 1.442 & 1.053 & 1.01 & 1.037 & 1.01 & 1.046 \\ 
+	pmed4 & 1.16 & 1.5 & 1.07 & 1.028 & 1.061 & 1.019 & 1.056 \\ 
+	pmed5 & 1.373 & 1.681 & 1.06 & 1.026 & 1.071 & 1.033 & 1.051 \\ 
+	pmed6 & 1.09 & 1.286 & 1.015 & 1.008 & 1.016 & 1.008 & 1.012 \\ 
+	pmed7 & 1.126 & 1.367 & 1.027 & 1.011 & 1.023 & 1.01 & 1.025 \\ 
+	pmed8 & 1.179 & 1.489 & 1.064 & 1.017 & 1.051 & 1.021 & 1.048 \\ 
+	pmed9 & 1.249 & 1.559 & 1.065 & 1.031 & 1.063 & 1.024 & 1.063 \\ 
+	pmed10 & 1.462 & 1.839 & 1.076 & 1.037 & 1.076 & 1.039 & 1.076 \\ 
+	pmed11 & 1.071 & 1.369 & 1.014 & 1.004 & 1.008 & 1.005 & 1.011 \\ 
+	pmed12 & 1.144 & 1.368 & 1.025 & 1.007 & 1.028 & 1.007 & 1.025 \\ 
+	pmed13 & 1.209 & 1.446 & 1.047 & 1.019 & 1.05 & 1.021 & 1.05 \\ 
+	pmed14 & 1.243 & 1.504 & 1.056 & 1.026 & 1.053 & 1.025 & 1.049 \\ 
+	pmed15 & 1.61 & 1.658 & 1.067 & 1.032 & 1.068 & 1.03 & 1.06 \\ 
+	pmed16 & 1.062 & 1.333 & 1.012 & 1.004 & 1.012 & 1.005 & 1.008 \\ 
+	pmed17 & 1.149 & 1.359 & 1.023 & 1.014 & 1.028 & 1.013 & 1.025 \\ 
+	pmed18 & 1.169 & 1.414 & 1.044 & 1.016 & 1.033 & 1.016 & 1.04 \\ 
+	pmed19 & 1.261 & 1.537 & 1.056 & 1.022 & 1.059 & 1.02 & 1.055 \\ 
+	pmed20 & 1.501 & 1.73 & 1.079 & 1.032 & 1.076 & 1.027 & 1.073 \\ 
+	pmed21 & 1.096 & 1.384 & 1.027 & 1.005 & 1.015 & 1.003 & 1.021 \\ 
+	pmed22 & 1.128 & 1.371 & 1.026 & 1.015 & 1.018 & 1.007 & 1.021 \\ 
+	pmed23 & 1.151 & 1.42 & 1.041 & 1.019 & 1.045 & 1.016 & 1.041 \\ 
+	pmed24 & 1.271 & 1.519 & 1.058 & 1.019 & 1.055 & 1.019 & 1.052 \\ 
+	pmed25 & 1.449 & 1.711 & 1.07 & 1.033 & 1.077 & 1.035 & 1.073 \\ 
+	pmed26 & 1.073 & 1.317 & 1.018 & 1.008 & 1.012 & 1.003 & 1.014 \\ 
+	pmed27 & 1.113 & 1.36 & 1.017 & 1.008 & 1.018 & 1.009 & 1.017 \\ 
+	pmed28 & 1.168 & 1.386 & 1.046 & 1.011 & 1.037 & 1.015 & 1.04 \\ 
+	pmed29 & 1.263 & 1.497 & 1.055 & 1.026 & 1.057 & 1.024 & 1.057 \\ 
+	pmed30 & 1.456 & 1.643 & 1.067 & 1.033 & 1.064 & 1.032 & 1.061 \\ 
+	pmed31 & 1.082 & 1.336 & 1.01 & 1.003 & 1.006 & 1.001 & 1.013 \\ 
+	pmed32 & 1.12 & 1.344 & 1.02 & 1.011 & 1.018 & 1.011 & 1.017 \\ 
+	pmed33 & 1.182 & 1.432 & 1.043 & 1.018 & 1.044 & 1.017 & 1.04 \\ 
+	pmed34 & 1.323 & 1.539 & 1.061 & 1.024 & 1.059 & 1.021 & 1.056 \\ 
+	pmed35 & 1.083 & 1.292 & 1.011 & 1.005 & 1.015 & 1.007 & 1.01 \\ 
+	pmed36 & 1.138 & 1.334 & 1.016 & 1.009 & 1.016 & 1.011 & 1.016 \\ 
+	pmed37 & 1.157 & 1.439 & 1.042 & 1.02 & 1.05 & 1.017 & 1.047 \\ 
+	pmed38 & 1.063 & 1.288 & 1.007 & 1.009 & 1.008 & 1.006 & 1.01 \\ 
+	pmed39 & 1.127 & 1.42 & 1.02 & 1.007 & 1.014 & 1.007 & 1.015 \\ 
+	pmed40 & 1.187 & 1.423 & 1.046 & 1.017 & 1.043 & 1.019 & 1.039 \\ \hline
+\end{tabular}
+
+\end{center}
+
+\caption{Approximation Constants for P-Median Datasets}
+\end{figure*}
+
 \label{submission}

### 2015-05-09 tex/project.tex
Updated project write-up

--- 2015-05-08 tex/project.tex
@@
 
-If we relax the constraint to have exactly $k$ clusters, then a recent algorithm called affinity propagation has shown lower error and computation time than the algorithms above \cite{Frey}.
-Instead of specifying a number of clusters, the user inputs ``preferences'' on the data points to be cluster centers.
-This algorithm relies on passing messages between data points.
-For each pair of data points $i, j \in X$, ``availability'' $a(i,j)$ represents the evidence for having $i$ choose $j$ to be its cluster center and ``responsibility'' $r(i,j)$ reflects how well-suited $j$ is to serve $i$.
-These are messages are then updated iteratively.
-Once the messages converge (this is not guaranteed) then $\max_{j \in X} (a(i,j) + r(i,j))$ will identify the cluster center for $i$, and these decisions will be consistent \cite{Frey}.
-
 The $k$-median problem is also a fundamental problem in approximation algorithms.
@@
 In addition, using supermodular functions means our distance measure does not have to satisfy the triangle inequality, which can be useful when considering non-vector data.
+This algorithm has also not been previously tested in practice.
 
@@
 Methods such as MeanShift and DBSCAN focus on finding clusters of high density \cite{Comaniciu, Ester}.
-We choose not to compare against densit- based methods since their objective differs widely from the center-based methods above.
+We choose not to compare against density based methods since their objective differs widely from the center-based methods above.
 In addition, we will focus on test data in which center-based methods seem appropriate.
@@
 Therefore, this algorithm can serve as a basis of comparison when both algorithms have the same available information.
+Lastly, if we relax the constraint to have exactly $k$ clusters and instead allow the user to input ``preferences'' on the data points being clusters, then a recent algorithm called affinity propagation has shown to perform particularly well \cite{Frey}.
+Again, because of the difference in available information, we do not to test against this algorithm.
 
@@
 the $k$-median objective we described above).
+Note that it may not always be possible to find a solution with approximation ratio 1.
 The hierarchical $k$-median problem was originally introduced by Plaxton who gave a 238.88 approximation algorithm \cite{Plaxton}.
@@
 These ratios can be improved by using a Lagrangian multiplier preserving $2$-approximation algorithm but the local search algorithm performed comparably in experimental tests \cite{Nagarajan}.
+While the authors tested the variants of this algorithm against other $k$-median approximation algorithms, it has not been tested against traditional clustering algorithms or on any labeled data sets.
 
@@
 \[ f(S) = \sum_{ i\in X} \min_{j \in S} d(i,j). \]
-Consider a local search algorithm with starts with a random subset $S$ of $k$ points.
+Consider a local search algorithm which starts with a random subset $S$ of $k$ points.
 In any iteration, the algorithm looks for a set $A \subseteq S$ and a set $B \subseteq X -S$ such that $|A| = |B| \leq p$ and
@@
 If no such set exists, the algorithm outputs $S$.
-Note that for $p$ independent of $k$ and $n$ and if we ensure that the algorithm makes at least $\delta = \varepsilon/n$ progress in each iteration, this is a polynomial-time algorithm.
+For $p$ independent of $k$ and $n$, let $\mathcal{N}$ be the size of the search neighborhood on each iteration.
+If we ensure that the algorithm improves by at least a factor of $(1- \delta) = (1- \varepsilon/\mathcal{N})$ in each iteration, this is a polynomial-time algorithm.
 Furthermore, the local search algorithm outputs $S \subseteq X$ such that $|S| = k$ and
-\[ f(S) \leq \left (3 + \frac{2}{p}+\varepsilon \right ) f(O) \]
+\[ f(S) \leq \frac{1}{1-\varepsilon} \left (3 + \frac{2}{p}) \right ) f(O) \]
 for all $O \subseteq X$. Therefore, there is a trade-off between the computation time of this algorithm and the guaranteed quality of the solution.
@@
 
-We implemented the local search algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compare against $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
-Note that for $p=1$, the local search algorithm is equivalent to the well-known PAM algorithm.
+We implemented the local search swapping algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compare against $k$-means, $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
+We also implemented the $k$-medoids algorithm with a maximum iteration of 1000.
+Note that for $p=1$, the multiswaps algorithm is equivalent to the well-known PAM algorithm.
 In order to calculate approximation ratios, Gurobi was used to compute the global optimum for each data set.
 
-CITATION FOR K-MEDOIDS CODE?
+To achieve reasonable computation times for the multiswaps and supermodular function minimization algorithms, at each iteration we first consider nearby swaps before sampling for other potential swaps.
+For the multiswaps algorithm, we sample $O(p\cdot n)$ potential swaps on each iteration, and for the supermodular function, sample $min(100,n^2)$ swaps where each potential function is calculated with by sampling 50 subsets.
+We also ensure that some minimum level progress is achieved for both algorithms by setting $\varepsilon = 0.1$ for the multiswaps algorithm and $\varepsilon = 10$ for the submodular function.
+These values were chosen by looking at the relative performance and computation time on random Gaussian data not tested below.
 
-To achieve reasonable computation times for the local search and supermodular function minimization algorithms, at each iteration we first consider nearby swaps and ensure that some minimum level progress is achieved.
-We also implemented a maximum of 1000 iterations for the $k$-median algorithm.
-
-We also considered a hybrid heuristic algorithm (labelled $p$-swap+ in the results) where we ran $k$-medoids on the result of the local search ($p$-swap) algorithm.
+We also considered a hybrid heuristic algorithm (labelled $p$-swap+ in the results) where we ran $k$-medoids for 10 iterations on the result of the multiswap search algorithm.
 This is motivated by the fact that, similar to Lloyd's algorithm, $k$-medoids will quickly converge to local minima.
@@
 For each data set and each clustering method, we report objective function values, the approximation ratio, computation time, and adjusted mutual information (AMI).
-When information on true labels is available, AMI is a measure which scores accuracy of predicted clusters.
-AMI measures the agreement of two assignments and is normalized to reflect chance, so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
+AMI is a measure which scores accuracy of predicted clusters compared to a true labeling by measuring the agreement of two assignments.
+The measure is normalized to reflect chance, so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
 See \cite{Vinh} for further information.
@@
 \begin{description}
-\item[Gaussian] Synthetic $k$-center Gaussian data, generated in Python.  Add info on std deviation, size, etc.
-\item[P-median] ADD CITATION OR Library p-median data sets
+\item[Gaussian] Synthetic $k$-center Gaussian data, generated in Python.
+In particular, Gauss\textunderscore d \textunderscore k \textunderscore e was generated by first choosing k random centers in $[-1,1]^d$.
+Around each center, we then generated $1000/k \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
+\item[OR $p$-median Library] From a collection of test data sets created by J. E. Beasley with data sets of size $100, 200, \ldots, 900$ and varying values of $k$ \cite{Beasley}. 
 \item[UCI Iris] \cite{Iris} Contains 4 attributes on 150 instances of flowers, each of with belongs to one of three species of iris.
@@
 
-Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on an 8 core ....
+Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core Intel processor.
 
@@
 Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
-The cost of these solutions is determined by assigning each point to the nearest cluster center.
-We then order these solutions according to their cost into buckets of the form $[0,\beta_0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
+The cost of these solutions is determined by assigning each point to the nearest cluster center and is scaled so that the minimum cost for $k<n$ is at least one.
+We then order these solutions according to their cost into buckets of the form $[0,0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
 
@@
 
-\subsection{Implementation}
-
-\subsection{Results}
+\subsection{Implementation and Results}
 
@@
 \section{Acknowledgements}
-David Williamson and Karthik Sridharan?
+We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
 
@@
 \bibliographystyle{icml2014}
+
+%\appendix
+%
+%\section{Appendix 1: Full Results}
+%
+%\subsection{Gaussian Data}
+%\begin{figure*}[h]
+%\begin{center}
+%\begin{tabular}{ | l | l | l | l | l | }
+%\hline
+%	Dataset & NameOpt Obj & k & n & dim  \\ \hline
+%	Gauss\_3\_5\_0 & 195 & 5 & 542 & 3 \\ 
+%	Gauss\_3\_5\_5 & 225 & 5 & 620 & 3 \\ 
+%	Gauss\_3\_5\_10 & 158 & 5 & 441 & 3 \\
+%	Gauss\_3\_10\_0 & 147 & 10 & 472 & 3 \\
+%	Gauss\_3\_10\_5 & 142 & 10 & 480 & 3 \\ 
+%	Gauss\_3\_10\_10 & 201 & 10 & 557 & 3 \\ 
+%	Gauss\_3\_50\_0 & 108 & 50 & 471 & 3 \\
+%	Gauss\_3\_50\_5 & 94 & 50 & 422 & 3 \\ 
+%	Gauss\_3\_50\_10 & 95 & 50 & 408 & 3 \\ 
+%	Gauss\_5\_5\_0 & 352 & 5 & 720 & 5 \\ 
+%	Gauss\_5\_5\_5 & 195 & 5 & 388 & 5 \\ 
+%	Gauss\_5\_5\_10 & 214 & 5 & 394 & 5 \\ 
+%	Gauss\_5\_10\_0 & 346 & 10 & 703 & 5 \\ 
+%	Gauss\_5\_10\_5 & 270 & 10 & 544 & 5 \\ 
+%	Gauss\_5\_10\_10 & 292 & 10 & 572 & 5 \\ 
+%	Gauss\_5\_50\_0 & 188 & 50 & 445 & 5 \\ 
+%	Gauss\_5\_50\_5 & 178 & 50 & 408 & 5 \\ 
+%	Gauss\_5\_50\_10 & 229 & 50 & 480 & 5 \\ 
+%	Gauss\_7\_5\_0 & 349 & 5 & 567 & 7 \\ 
+%	Gauss\_7\_5\_5 & 368 & 5 & 579 & 7 \\ 
+%	Gauss\_7\_5\_10 & 356 & 5 & 527 & 7 \\ 
+%	Gauss\_7\_10\_0 & 351 & 10 & 573 & 7 \\ 
+%	Gauss\_7\_10\_5 & 302 & 10 & 478 & 7 \\ 
+%	Gauss\_7\_10\_10 & 354 & 10 & 534 & 7 \\ 
+%	Gauss\_7\_50\_0 & 257 & 50 & 451 & 7 \\ 
+%	Gauss\_7\_50\_5 & 256 & 50 & 430 & 7 \\ 
+%	Gauss\_7\_50\_10 & 298 & 50 & 480 & 7 \\ \hline
+%\end{tabular}
+%\end{center}
+%
+%\caption{Details For Gaussian Datasets}
+%\end{figure*}
+%
+%\begin{figure*}[h]
+%\begin{center}
+%\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
+%\hline
+%	Dataset &Time & Time & Time & Time & Time & Time & Time & Time & Time & Time   \\ \hline
+%	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-swaps+ & 3-Swap & 4-swaps   \\ 
+%	Gauss\_3\_5\_0 & 0.024 & 0.056 & 0.066 & 131.305 & 2.707 & 3.195 & 4.779 & 4.138 & 7.683 & 11.325 \\ 
+%	Gauss\_3\_5\_5 & 0.032 & 0.076 & 0.099 & 195.013 & 3.682 & 3.764 & 6.639 & 5.32 & 9.627 & 14.575 \\ 
+%	Gauss\_3\_5\_10 & 0.033 & 0.055 & 0.047 & 80.439 & 2.244 & 1.802 & 2.806 & 3.024 & 5.107 & 7.457 \\ 
+%	Gauss\_3\_10\_0 & 0.041 & 0.083 & 0.095 & 121.551 & 5.133 & 3.576 & 7.452 & 6.3 & 10.157 & 11.872 \\ 
+%	Gauss\_3\_10\_5 & 0.048 & 0.1 & 0.117 & 101.072 & 5.026 & 4.96 & 8.57 & 6.952 & 11.948 & 12.225 \\ 
+%	Gauss\_3\_10\_10 & 0.05 & 0.106 & 0.153 & 149.352 & 7.101 & 6.456 & 11.273 & 10.646 & 18.21 & 17.414 \\ 
+%	Gauss\_3\_50\_0 & 0.154 & 0.324 & 0.435 & 175.827 & 29.201 & 34.228 & 43.65 & 49.964 & 48.259 & 66.413 \\ 
+%	Gauss\_3\_50\_5 & 0.151 & 0.253 & 0.376 & 127.64 & 26.318 & 21.541 & 38.004 & 33.213 & 45.862 & 57.262 \\ 
+%	Gauss\_3\_50\_10 & 0.153 & 0.265 & 0.301 & 132.119 & 21.341 & 22.556 & 35.389 & 32.139 & 44.361 & 50.143 \\ 
+%	Gauss\_5\_5\_0 & 0.026 & 0.079 & 0.08 & 230.345 & 4.997 & 4.21 & 9.428 & 8.063 & 12.811 & 19.337 \\ 
+%	Gauss\_5\_5\_5 & 0.025 & 0.056 & 0.032 & 75.262 & 1.159 & 1.506 & 2.188 & 2.382 & 3.71 & 5.797 \\ 
+%	Gauss\_5\_5\_10 & 0.025 & 0.042 & 0.027 & 69.349 & 1.383 & 1.458 & 2.403 & 2.575 & 3.811 & 5.748 \\ 
+%	Gauss\_5\_10\_0 & 0.044 & 0.1 & 0.18 & 227.747 & 11.633 & 12.129 & 17.724 & 18.654 & 26.563 & 33.646 \\ 
+%	Gauss\_5\_10\_5 & 0.054 & 0.095 & 0.109 & 134.137 & 6.834 & 5.407 & 9.243 & 9.428 & 15.96 & 21.795 \\ 
+%	Gauss\_5\_10\_10 & 0.045 & 0.092 & 0.117 & 127.614 & 7.639 & 6.456 & 11.807 & 12.186 & 18.339 & 23.554 \\ 
+%	Gauss\_5\_50\_0 & 0.142 & 0.211 & 0.389 & 139.818 & 30.155 & 27.431 & 37.251 & 43.421 & 50.192 & 68.329 \\ 
+%	Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 & 53.546 \\ 
+%	Gauss\_5\_50\_10 & 0.149 & 0.219 & 0.417 & 170.431 & 37.132 & 35.119 & 45.341 & 47.002 & 56.683 & 69.712 \\ 
+%	Gauss\_7\_5\_0 & 0.019 & 0.054 & 0.052 & 185.865 & 2.975 & 3.565 & 4.954 & 5.005 & 8.531 & 11.867 \\ 
+%	Gauss\_7\_5\_5 & 0.031 & 0.078 & 0.047 & 154.1 & 2.603 & 3.04 & 5.45 & 5.738 & 8.056 & 12.505 \\ 
+%	Gauss\_7\_5\_10 & 0.018 & 0.052 & 0.043 & 130.185 & 2.949 & 2.478 & 4.668 & 4.849 & 7.324 & 11.036 \\ 
+%	Gauss\_7\_10\_0 & 0.045 & 0.071 & 0.097 & 185.521 & 8.015 & 7.338 & 12.495 & 10.474 & 16.174 & 20.671 \\ 
+%	Gauss\_7\_10\_5 & 0.04 & 0.063 & 0.073 & 136.189 & 5.315 & 5.365 & 9.287 & 6.87 & 11.768 & 16.072 \\ 
+%	Gauss\_7\_10\_10 & 0.048 & 0.09 & 0.1 & 168.962 & 7.047 & 6.538 & 9.783 & 9.322 & 13.581 & 19.41 \\ 
+%	Gauss\_7\_50\_0 & 0.143 & 0.199 & 0.401 & 152.958 & 32.592 & 37.747 & 39.026 & 42.999 & 60.483 & 69.667 \\ 
+%	Gauss\_7\_50\_5 & 0.147 & 0.212 & 0.327 & 136.496 & 27.8 & 26.288 & 38.683 & 34.335 & 47.779 & 59.414 \\ 
+%	Gauss\_7\_50\_10 & 0.149 & 0.231 & 0.426 & 163.651 & 37.968 & 41.976 & 55.514 & 50.175 & 56.302 & 79.819 \\ \hline
+%\end{tabular}
+%
+%\end{center}
+%
+%\caption{Running Times For Gaussian Datasets}
+%\end{figure*}
+%
+%\begin{figure*}[h]
+%\begin{center}
+%\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
+%\hline
+%	Dataset&AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
+%	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap & 4-swaps   \\ \hline
+%	Gauss\_3\_5\_0 & 0.884 & 0.887 & 0.729 & 0.678 & 0.88 & 0.876 & 0.884 & 0.876 & 0.881 & 0.882 \\ 
+%	Gauss\_3\_5\_5 & 0.7 & 0.62 & 0.637 & 0.545 & 0.683 & 0.68 & 0.685 & 0.69 & 0.682 & 0.686 \\ 
+%	Gauss\_3\_5\_10 & 0.68 & 0.618 & 0.62 & 0.504 & 0.663 & 0.665 & 0.659 & 0.659 & 0.664 & 0.668 \\ 
+%	Gauss\_3\_10\_0 & 0.693 & 0.641 & 0.699 & 0.619 & 0.692 & 0.692 & 0.688 & 0.696 & 0.685 & 0.685 \\ 
+%	Gauss\_3\_10\_5 & 0.534 & 0.466 & 0.516 & 0.446 & 0.534 & 0.533 & 0.535 & 0.532 & 0.531 & 0.536 \\ 
+%	Gauss\_3\_10\_10 & 0.669 & 0.56 & 0.616 & 0.516 & 0.651 & 0.653 & 0.652 & 0.657 & 0.657 & 0.653 \\ 
+%	Gauss\_3\_50\_0 & 0.504 & 0.498 & 0.498 & 0.475 & 0.501 & 0.497 & 0.491 & 0.497 & 0.495 & 0.501 \\ 
+%	Gauss\_3\_50\_5 & 0.407 & 0.412 & 0.402 & 0.396 & 0.392 & 0.385 & 0.389 & 0.386 & 0.387 & 0.385 \\ 
+%	Gauss\_3\_50\_10 & 0.421 & 0.424 & 0.406 & 0.39 & 0.413 & 0.414 & 0.408 & 0.404 & 0.405 & 0.41 \\ 
+%	Gauss\_5\_5\_0 & 0.99 & 0.99 & 0.858 & 0.719 & 0.988 & 0.994 & 0.989 & 0.994 & 0.989 & 0.985 \\ 
+%	Gauss\_5\_5\_5 & 0.741 & 0.713 & 0.687 & 0.657 & 0.742 & 0.749 & 0.741 & 0.744 & 0.737 & 0.747 \\ 
+%	Gauss\_5\_5\_10 & 0.833 & 0.83 & 227.742 & 0.663 & 0.76 & 0.779 & 0.806 & 0.747 & 0.789 & 0.77 \\ 
+%	Gauss\_5\_10\_0 & 0.974 & 0.972 & 0.858 & 0.732 & 0.96 & 0.964 & 0.964 & 0.965 & 0.96 & 0.964 \\ 
+%	Gauss\_5\_10\_5 & 0.783 & 0.736 & 0.677 & 0.575 & 0.729 & 0.725 & 0.737 & 0.737 & 0.729 & 0.725 \\ 
+%	Gauss\_5\_10\_10 & 0.878 & 0.829 & 0.795 & 0.704 & 0.873 & 0.875 & 0.873 & 0.872 & 0.875 & 0.868 \\ 
+%	Gauss\_5\_50\_0 & 0.723 & 0.669 & 0.668 & 0.622 & 0.695 & 0.692 & 0.692 & 0.7 & 0.696 & 0.692 \\ 
+%	Gauss\_5\_50\_5 & 0.646 & 0.596 & 0.621 & 0.57 & 0.622 & 0.619 & 0.627 & 0.631 & 0.625 & 0.626 \\
+%	Gauss\_5\_50\_10 & 0.669 & 0.639 & 0.626 & 0.546 & 0.64 & 0.641 & 0.639 & 0.65 & 0.643 & 0.637 \\ 
+%	Gauss\_7\_5\_0 & 1 & 1 & 0.852 & 0.773 & 0.999 & 1 & 0.999 & 1 & 0.999 & 0.998 \\ 
+%	Gauss\_7\_5\_5 & 0.847 & 0.832 & 0.86 & 0.72 & 0.852 & 0.85 & 0.849 & 0.847 & 0.847 & 0.847 \\ 
+%	Gauss\_7\_5\_10 & 0.979 & 0.963 & 0.839 & 0.744 & 0.981 & 0.979 & 0.98 & 0.979 & 0.978 & 0.978 \\ 
+%	Gauss\_7\_10\_0 & 0.995 & 0.995 & 0.869 & 0.785 & 0.988 & 0.984 & 0.987 & 0.984 & 0.982 & 0.976 \\ 
+%	Gauss\_7\_10\_5 & 0.988 & 0.994 & 0.853 & 0.732 & 0.988 & 0.994 & 0.99 & 0.994 & 0.993 & 0.994 \\ 
+%	Gauss\_7\_10\_10 & 0.89 & 0.81 & 0.845 & 0.724 & 0.851 & 0.856 & 0.854 & 0.854 & 0.85 & 0.851 \\ 
+%	Gauss\_7\_50\_0 & 0.867 & 0.834 & 0.784 & 0.693 & 0.832 & 0.835 & 0.84 & 0.833 & 0.835 & 0.846 \\ 
+%	Gauss\_7\_50\_5 & 0.852 & 0.808 & 0.727 & 0.647 & 0.813 & 0.815 & 0.814 & 0.815 & 0.812 & 0.807 \\ 
+%	Gauss\_7\_50\_10 & 0.829 & 0.761 & 0.742 & 0.645 & 0.807 & 0.799 & 0.807 & 0.804 & 0.804 & 0.799 \\ \hline
+%\end{tabular}
+%
+%\end{center}
+%
+%\caption{AMI For Gaussian Datasets}
+%\end{figure*}
+%
+%
+%\begin{figure*}[h]
+%\begin{center}
+%\begin{tabular}{ | l | l | l | l | l | l | l | l | l | }
+%\hline
+%	Dataset & Approx & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
+%	& k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swap+ & 3-Swap & 4-swaps   \\ \hline
+%	Gauss\_3\_5\_0 & 1.159 & 1.481 & 1.013 & 1 & 1.011 & 1 & 1.013 & 1.009 \\ 
+%	Gauss\_3\_5\_5 & 1.057 & 1.4 & 1.005 & 1.001 & 1.006 & 1.001 & 1.01 & 1.007 \\ 
+%	Gauss\_3\_5\_10 & 1.047 & 1.377 & 1.007 & 1.004 & 1.014 & 1.007 & 1.009 & 1.008 \\ 
+%	Gauss\_3\_10\_0 & 1.029 & 1.404 & 1.018 & 1.01 & 1.013 & 1.01 & 1.022 & 1.019 \\ 
+%	Gauss\_3\_10\_5 & 1.063 & 1.329 & 1.025 & 1.013 & 1.021 & 1.012 & 1.02 & 1.026 \\ 
+%	Gauss\_3\_10\_10 & 1.06 & 1.369 & 1.017 & 1.009 & 1.018 & 1.007 & 1.016 & 1.016 \\ 
+%	Gauss\_3\_50\_0 & 1.112 & 1.354 & 1.05 & 1.026 & 1.037 & 1.025 & 1.049 & 1.044 \\ 
+%	Gauss\_3\_50\_5 & 1.124 & 1.327 & 1.045 & 1.025 & 1.041 & 1.019 & 1.044 & 1.04 \\ 
+%	Gauss\_3\_50\_10 & 1.121 & 1.316 & 1.044 & 1.023 & 1.037 & 1.021 & 1.043 & 1.043 \\ 
+%	Gauss\_5\_5\_0 & 1.105 & 1.576 & 1.013 & 1 & 1.011 & 1 & 1.016 & 1.015 \\ 
+%	Gauss\_5\_5\_5 & 1.07 & 1.436 & 1.03 & 1.002 & 1.028 & 1.005 & 1.026 & 1.024 \\ 
+%	Gauss\_5\_5\_10 & 0.003 & 1.369 & 1.021 & 1.004 & 1.016 & 1.007 & 1.016 & 1.015 \\ 
+%	Gauss\_5\_10\_0 & 1.148 & 1.619 & 1.029 & 1 & 1.026 & 1.001 & 1.028 & 1.029 \\ 
+%	Gauss\_5\_10\_5 & 1.08 & 1.389 & 1.017 & 1.011 & 1.018 & 1.009 & 1.017 & 1.019 \\ 
+%	Gauss\_5\_10\_10 & 1.152 & 1.458 & 1.022 & 1.006 & 1.023 & 1.007 & 1.021 & 1.026 \\
+%	Gauss\_5\_50\_0 & 1.096 & 1.271 & 1.031 & 1.017 & 1.033 & 1.015 & 1.031 & 1.029 \\ 
+%	Gauss\_5\_50\_5 & 1.089 & 1.272 & 1.03 & 1.02 & 1.027 & 1.014 & 1.033 & 1.032 \\ 
+%	Gauss\_5\_50\_10 & 1.09 & 1.299 & 1.029 & 1.017 & 1.036 & 1.016 & 1.028 & 1.036 \\ 
+%	Gauss\_7\_5\_0 & 1.214 & 1.59 & 1.016 & 1 & 1.018 & 1 & 1.014 & 1.016 \\ 
+%	Gauss\_7\_5\_5 & 1.046 & 1.446 & 1.017 & 1.009 & 1.014 & 1.004 & 1.014 & 1.019 \\ 
+%	Gauss\_7\_5\_10 & 1.153 & 1.53 & 1.01 & 1 & 1.007 & 1 & 1.01 & 1.009 \\ 
+%	Gauss\_7\_10\_0 & 1.132 & 1.504 & 1.032 & 1 & 1.023 & 1 & 1.032 & 1.034 \\ 
+%	Gauss\_7\_10\_5 & 1.159 & 1.569 & 1.031 & 1 & 1.028 & 1 & 1.027 & 1.024 \\ 
+%	Gauss\_7\_10\_10 & 1.074 & 1.56 & 1.024 & 1.003 & 1.019 & 1.005 & 1.02 & 1.02 \\ 
+%	Gauss\_7\_50\_0 & 1.104 & 1.321 & 1.029 & 1.01 & 1.031 & 1.011 & 1.027 & 1.026 \\ 
+%	Gauss\_7\_50\_5 & 1.132 & 1.326 & 1.033 & 1.013 & 1.031 & 1.013 & 1.026 & 1.036 \\ 
+%	Gauss\_7\_50\_10 & 1.107 & 1.345 & 1.027 & 1.014 & 1.025 & 1.012 & 1.031 & 1.026 \\ \hline
+%\end{tabular}
+%
+%\end{center}
+%
+%\caption{Approximation Constants For Gaussian Datasets}
+%\end{figure*}
+%
+%\subsection{P-Median}
+%
+%\begin{figure*}[h]
+%\begin{center}
+%\begin{tabular}{ | l | l | l | l | }
+%\hline
+%	Dataset & Optimal Objective & $k$ & $n$ \\ \hline
+%	pmed1 & 5819 & 5 & 100 \\ 
+%	pmed2 & 4093 & 10 & 100 \\ 
+%	pmed3 & 4250 & 10 & 100 \\ 
+%	pmed4 & 3034 & 20 & 100 \\
+%	pmed5 & 1355 & 33 & 100 \\ 
+%	pmed6 & 7824 & 5 & 200 \\ 
+%	pmed7 & 5631 & 10 & 200 \\ 
+%	pmed8 & 4445 & 20 & 200 \\ 
+%	pmed9 & 2734 & 40 & 200 \\ 
+%	pmed10 & 1255 & 67 & 200 \\ 
+%	pmed11 & 7696 & 5 & 300 \\ 
+%	pmed12 & 6634 & 10 & 300 \\ 
+%	pmed13 & 4374 & 30 & 300 \\ 
+%	pmed14 & 2968 & 60 & 300 \\ 
+%	pmed15 & 1729 & 100 & 300 \\
+%	pmed16 & 8162 & 5 & 400 \\ 
+%	pmed17 & 6999 & 10 & 400 \\
+%	pmed18 & 4809 & 40 & 400 \\ 
+%	pmed19 & 2845 & 80 & 400 \\ 
+%	pmed20 & 1789 & 133 & 400 \\ 
+%	pmed21 & 9138 & 5 & 500 \\ 
+%	pmed22 & 8579 & 10 & 500 \\ 
+%	pmed23 & 4619 & 50 & 500 \\ 
+%	pmed24 & 2961 & 100 & 500 \\ 
+%	pmed25 & 1828 & 167 & 500 \\ 
+%	pmed26 & 9917 & 5 & 600 \\ 
+%	pmed27 & 8307 & 10 & 600 \\ 
+%	pmed28 & 4498 & 60 & 600 \\ 
+%	pmed29 & 3033 & 120 & 600 \\
+%	pmed30 & 1989 & 200 & 600 \\ 
+%	pmed31 & 10086 & 5 & 700 \\ 
+%	pmed32 & 9297 & 10 & 700 \\ 
+%	pmed33 & 4700 & 60 & 700 \\ 
+%	pmed34 & 3013 & 140 & 700 \\ 
+%	pmed35 & 10400 & 5 & 800 \\ 
+%	pmed36 & 9934 & 10 & 800 \\ 
+%	pmed37 & 5057 & 80 & 800 \\ 
+%	pmed38 & 11060 & 5 & 900 \\ 
+%	pmed39 & 9423 & 10 & 900 \\ 
+%	pmed40 & 5128 & 90 & 900 \\ \hline
+%\end{tabular}
+%
+%
+%\end{center}
+%
+%\caption{Details for P-Median Datasets}
+%\end{figure*}
+%
+%\begin{figure*}[h]
+%\begin{center}
+%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
+%\hline
+%	Dataset & Time & Time & Time & Time & Time & Time & Time \\ \hline
+%	&k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap   \\ \hline
+%	pmed1 & 0.0032 & 5 & 0.1 & 0.1 & 0.2 & 0.2 & 0.3 \\ 
+%	pmed2 & 0.0069 & 6.4 & 0.2 & 0.2 & 0.3 & 0.3 & 0.4 \\ 
+%	pmed3 & 0.0067 & 7.5 & 0.2 & 0.2 & 0.3 & 0.3 & 0.4 \\ 
+%	pmed4 & 0.0125 & 8 & 0.3 & 0.4 & 0.5 & 0.6 & 0.6 \\ 
+%	pmed5 & 0.0381 & 8.3 & 0.5 & 0.6 & 0.7 & 0.7 & 0.9 \\ 
+%	pmed6 & 0.0065 & 19.3 & 0.3 & 0.4 & 0.5 & 0.6 & 0.8 \\ 
+%	pmed7 & 0.0117 & 23.9 & 0.7 & 0.9 & 0.9 & 1.3 & 1.5 \\ 
+%	pmed8 & 0.027 & 27.2 & 1.4 & 1.8 & 2.2 & 2.6 & 3.1 \\ 
+%	pmed9 & 0.0642 & 29 & 3.4 & 3.5 & 3.3 & 4.8 & 4.7 \\ 
+%	pmed10 & 0.7329 & 39.8 & 4.4 & 5.8 & 6.2 & 7.3 & 8.2 \\ 
+%	pmed11 & 0.0124 & 42 & 0.6 & 0.8 & 1.4 & 1.2 & 2.2 \\ 
+%	pmed12 & 0.0186 & 45.7 & 1.7 & 2 & 2.3 & 2.4 & 4.3 \\ 
+%	pmed13 & 0.0685 & 69.1 & 6.6 & 5.3 & 7 & 8.4 & 11 \\ 
+%	pmed14 & 0.2238 & 76.3 & 13.6 & 14.2 & 17.1 & 17.4 & 25 \\ 
+%	pmed15 & 1.0969 & 100.2 & 18.4 & 19.7 & 23.1 & 28.7 & 35.5 \\ 
+%	pmed16 & 0.0251 & 87.1 & 1.3 & 1.2 & 2.3 & 2.4 & 3.8 \\ 
+%	pmed17 & 0.0581 & 97.1 & 3.5 & 2.5 & 3.9 & 4.8 & 6.8 \\ 
+%	pmed18 & 0.2596 & 134.7 & 15.3 & 16.5 & 23.1 & 23.3 & 31.3 \\ 
+%	pmed19 & 1.0193 & 153.5 & 34.9 & 34.1 & 40.3 & 47.9 & 57.8 \\ 
+%	pmed20 & 1.7947 & 175 & 52.2 & 45 & 68.6 & 68.2 & 79.3 \\ 
+%	pmed21 & 0.0327 & 109.4 & 1.8 & 2.1 & 4.1 & 4.2 & 6.1 \\ 
+%	pmed22 & 0.0786 & 113.1 & 4.3 & 5.6 & 8.3 & 7.5 & 12.7 \\ 
+%	pmed23 & 0.5123 & 199.8 & 35.9 & 34.8 & 46.9 & 52.3 & 62.2 \\ 
+%	pmed24 & 1.7861 & 245.9 & 66.6 & 90.3 & 99.2 & 88.5 & 125.5 \\ 
+%	pmed25 & 3.1433 & 451.9 & 103 & 118 & 126.8 & 146.4 & 138.1 \\ 
+%	pmed26 & 0.0503 & 159.8 & 2.8 & 2.4 & 4.7 & 5.6 & 6.9 \\ 
+%	pmed27 & 0.1265 & 183.4 & 7.1 & 6.1 & 11.7 & 10.7 & 13 \\ 
+%	pmed28 & 0.8738 & 263.3 & 60.3 & 71.3 & 95.6 & 98.5 & 89.3 \\ 
+%	pmed29 & 2.8418 & 424.3 & 152.6 & 146.8 & 196.5 & 172.8 & 167 \\ 
+%	pmed30 & 5.2891 & 571 & 175 & 207.5 & 254.2 & 249.7 & 255.7 \\ 
+%	pmed31 & 0.0728 & 225.6 & 4.9 & 4.4 & 8.2 & 6.6 & 9.2 \\ 
+%	pmed32 & 0.1528 & 249.5 & 11 & 7.8 & 14.3 & 12.9 & 19 \\ 
+%	pmed33 & 1.4723 & 440.1 & 112.8 & 91.9 & 143.9 & 158.5 & 154.1 \\ 
+%	pmed34 & 3.6669 & 591.7 & 234.8 & 217.8 & 233.1 & 287.8 & 256.8 \\ 
+%	pmed35 & 0.1133 & 298.3 & 4.8 & 5.9 & 7.5 & 9.3 & 11.9 \\ 
+%	pmed36 & 0.2221 & 259.8 & 13 & 11.4 & 15.9 & 24 & 23.6 \\ 
+%	pmed37 & 2.4409 & 571 & 190.5 & 149.7 & 172.8 & 282.3 & 224.7 \\ 
+%	pmed38 & 0.1326 & 349.9 & 7.1 & 6.5 & 9.8 & 12.2 & 14.5 \\ 
+%	pmed39 & 0.2676 & 419.8 & 16.5 & 13.6 & 22.6 & 25.5 & 29.4 \\ 
+%	pmed40 & 3.5269 & 849.9 & 251.3 & 264.9 & 250.4 & 391.7 & 350.2 \\ \hline
+%\end{tabular}
+%
+%
+%\end{center}
+%
+%\caption{Running Time for P-Median Datasets}
+%\end{figure*}
+%
+%\begin{figure*}[h]
+%\begin{center}
+%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
+%\hline
+%	Dataset & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
+%	  & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap \\ \hline
+%	pmed1 & 1.113 & 1.409 & 1.03 & 1.013 & 1.029 & 1.005 & 1.018 \\ 
+%	pmed2 & 1.158 & 1.486 & 1.044 & 1.011 & 1.038 & 1.011 & 1.044 \\ 
+%	pmed3 & 1.161 & 1.442 & 1.053 & 1.01 & 1.037 & 1.01 & 1.046 \\ 
+%	pmed4 & 1.16 & 1.5 & 1.07 & 1.028 & 1.061 & 1.019 & 1.056 \\ 
+%	pmed5 & 1.373 & 1.681 & 1.06 & 1.026 & 1.071 & 1.033 & 1.051 \\ 
+%	pmed6 & 1.09 & 1.286 & 1.015 & 1.008 & 1.016 & 1.008 & 1.012 \\ 
+%	pmed7 & 1.126 & 1.367 & 1.027 & 1.011 & 1.023 & 1.01 & 1.025 \\ 
+%	pmed8 & 1.179 & 1.489 & 1.064 & 1.017 & 1.051 & 1.021 & 1.048 \\ 
+%	pmed9 & 1.249 & 1.559 & 1.065 & 1.031 & 1.063 & 1.024 & 1.063 \\ 
+%	pmed10 & 1.462 & 1.839 & 1.076 & 1.037 & 1.076 & 1.039 & 1.076 \\ 
+%	pmed11 & 1.071 & 1.369 & 1.014 & 1.004 & 1.008 & 1.005 & 1.011 \\ 
+%	pmed12 & 1.144 & 1.368 & 1.025 & 1.007 & 1.028 & 1.007 & 1.025 \\ 
+%	pmed13 & 1.209 & 1.446 & 1.047 & 1.019 & 1.05 & 1.021 & 1.05 \\ 
+%	pmed14 & 1.243 & 1.504 & 1.056 & 1.026 & 1.053 & 1.025 & 1.049 \\ 
+%	pmed15 & 1.61 & 1.658 & 1.067 & 1.032 & 1.068 & 1.03 & 1.06 \\ 
+%	pmed16 & 1.062 & 1.333 & 1.012 & 1.004 & 1.012 & 1.005 & 1.008 \\ 
+%	pmed17 & 1.149 & 1.359 & 1.023 & 1.014 & 1.028 & 1.013 & 1.025 \\ 
+%	pmed18 & 1.169 & 1.414 & 1.044 & 1.016 & 1.033 & 1.016 & 1.04 \\ 
+%	pmed19 & 1.261 & 1.537 & 1.056 & 1.022 & 1.059 & 1.02 & 1.055 \\ 
+%	pmed20 & 1.501 & 1.73 & 1.079 & 1.032 & 1.076 & 1.027 & 1.073 \\ 
+%	pmed21 & 1.096 & 1.384 & 1.027 & 1.005 & 1.015 & 1.003 & 1.021 \\ 
+%	pmed22 & 1.128 & 1.371 & 1.026 & 1.015 & 1.018 & 1.007 & 1.021 \\ 
+%	pmed23 & 1.151 & 1.42 & 1.041 & 1.019 & 1.045 & 1.016 & 1.041 \\ 
+%	pmed24 & 1.271 & 1.519 & 1.058 & 1.019 & 1.055 & 1.019 & 1.052 \\ 
+%	pmed25 & 1.449 & 1.711 & 1.07 & 1.033 & 1.077 & 1.035 & 1.073 \\ 
+%	pmed26 & 1.073 & 1.317 & 1.018 & 1.008 & 1.012 & 1.003 & 1.014 \\ 
+%	pmed27 & 1.113 & 1.36 & 1.017 & 1.008 & 1.018 & 1.009 & 1.017 \\ 
+%	pmed28 & 1.168 & 1.386 & 1.046 & 1.011 & 1.037 & 1.015 & 1.04 \\ 
+%	pmed29 & 1.263 & 1.497 & 1.055 & 1.026 & 1.057 & 1.024 & 1.057 \\ 
+%	pmed30 & 1.456 & 1.643 & 1.067 & 1.033 & 1.064 & 1.032 & 1.061 \\ 
+%	pmed31 & 1.082 & 1.336 & 1.01 & 1.003 & 1.006 & 1.001 & 1.013 \\ 
+%	pmed32 & 1.12 & 1.344 & 1.02 & 1.011 & 1.018 & 1.011 & 1.017 \\ 
+%	pmed33 & 1.182 & 1.432 & 1.043 & 1.018 & 1.044 & 1.017 & 1.04 \\ 
+%	pmed34 & 1.323 & 1.539 & 1.061 & 1.024 & 1.059 & 1.021 & 1.056 \\ 
+%	pmed35 & 1.083 & 1.292 & 1.011 & 1.005 & 1.015 & 1.007 & 1.01 \\ 
+%	pmed36 & 1.138 & 1.334 & 1.016 & 1.009 & 1.016 & 1.011 & 1.016 \\ 
+%	pmed37 & 1.157 & 1.439 & 1.042 & 1.02 & 1.05 & 1.017 & 1.047 \\ 
+%	pmed38 & 1.063 & 1.288 & 1.007 & 1.009 & 1.008 & 1.006 & 1.01 \\ 
+%	pmed39 & 1.127 & 1.42 & 1.02 & 1.007 & 1.014 & 1.007 & 1.015 \\ 
+%	pmed40 & 1.187 & 1.423 & 1.046 & 1.017 & 1.043 & 1.019 & 1.039 \\ \hline
+%\end{tabular}
+%
+%
+%\end{center}
+%
+%\caption{Approximation Constants for P-Median Datasets}
+%\end{figure*}
 

### 2015-05-09 tex/project.tex
added data tables for heirarchical

@@
 \hline
-	Data Set Name &Optimal Objective & $k$ & $n$ & dim   \\ \hline
-	Soybean & 113.8 & 4 & 47 & 35 \\ 
-	Iris & 98.2 & 3 & 150 & 4 \\ 
-	Gauss\_5\_5\_5 & 195.4 & 5 & 388 & 5 \\ 
-	Gauss\_5\_10\_5 & 270 & 10 & 544 & 5 \\ 
-	Gauss\_5\_50\_5 & 178.1 & 50 & 408 & 5 \\ 
-	pmed1 & 5819 & 5 & 100 & \  \\ 
-	\hline
+	  & Opt Obj & $k$ & $n$ & dim \\ \hline
+	Dataset Name & \  & \  & \  & \  \\ \hline
+	Soybean & 114 & 4 & 47 & 35 \\ 
+	Iris & 98 & 3 & 150 & 4 \\ 
+	Gaussian avg & 241 & 22 & 507 & 5 \\ 
+	pmed5 & 1355 & 33 & 100 & \  \\ 
+	pmed20 & 1789 & 133 & 400 & \  \\ 
+	pmed40 & 5128 & 90 & 900 & \  \\ 
+	pmedian avg & 5535 & 46 & 460 & \  \\ \hline
 \end{tabular}
+
 \caption{Some specifics about our datasets}
@@
 \begin{figure*}[h]
+\small
 \begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | }
 \hline
-	Dataset Name & Time & Time & Time & Time & Time & Time & Time & Time & Time \\ \hline
-	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap  \\ \hline
+	  & Time & Time & Time & Time & Time & Time & Time & Time & Time \\ \hline
+	Dataset Name & K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap \\ \hline
 	Soybean & 0.015 & 0.061 & 0.002 & 1.179 & 0.026 & 0.016 & 0.048 & 0.049 & 0.081 \\ 
 	Iris & 0.013 & 0.023 & 0.003 & 9.788 & 0.125 & 0.102 & 0.212 & 0.229 & 0.414 \\ 
-	Gauss\_5\_5\_5 & 0.025 & 0.056 & 0.032 & 75.262 & 1.159 & 1.506 & 2.188 & 2.382 & 3.71 \\ 
-	Gauss\_5\_10\_5 & 0.054 & 0.095 & 0.109 & 134.137 & 6.834 & 5.407 & 9.243 & 9.428 & 15.96 \\ 
-	Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 \\ 
-	pmed1 & N/A & N/A & 0.003 & 5.042 & 0.084 & 0.09 & 0.155 & 0.161 & 0.255 \\ \hline
+	Gaussian avg & 0.073 & 0.128 & 0.182 & 144.455 & 13.289 & 13.021 & 18.749 & 18.696 & 24.592 \\ 
+	pmed5 & 0.038 & 8.298 & 0.481 & 0.622 & 0.653 & 0.748 & 0.939 & \  & \  \\ 
+	pmed20 & 1.795 & 174.951 & 52.208 & 44.979 & 68.554 & 68.231 & 79.288 & \  & \  \\ 
+	pmed40 & 3.527 & 849.852 & 251.259 & 264.867 & 250.412 & 391.738 & 350.235 & \  & \  \\ 
+	pmedian avg & 0.807 & 203.308 & 40.386 & 40.347 & 48.264 & 56.76 & 56.19 & \  & \  \\ \hline
 \end{tabular}
@@
 \begin{figure*}[h]
-\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
+\small
+\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | }
 \hline
-	Dataset Name & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
-	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap   \\ \hline
-	Soybean & 0.69 & 0.18 & 0.55 & 0.61 & 0.71 & 0.73 & 0.72 & 0.71 & 0.71  \\ 
-	Iris & 0.75 & 0.76 & 0.69 & 0.7 & 0.77 & 0.76 & 0.76 & 0.76 & 0.76  \\ 
-	Gauss\_5\_5\_5 & 0.74 & 0.71 & 0.69 & 0.66 & 0.74 & 0.75 & 0.74 & 0.74 & 0.74  \\ 
-	Gauss\_5\_10\_5 & 0.78 & 0.74 & 0.68 & 0.58 & 0.73 & 0.72 & 0.74 & 0.74 & 0.73  \\ 
-	Gauss\_5\_50\_5 & 0.65 & 0.6 & 0.62 & 0.57 & 0.62 & 0.62 & 0.63 & 0.63 & 0.62  \\ \hline
+	  & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI \\ \hline
+	Dataset Name & K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap \\ \hline
+	Soybean & 0.686 & 0.178 & 0.55 & 0.612 & 0.708 & 0.732 & 0.717 & 0.708 & 0.712 \\ 
+	Iris & 0.748 & 0.757 & 0.69 & 0.7 & 0.772 & 0.76 & 0.763 & 0.757 & 0.761 \\ 
+	Gaussian avg & 0.777 & 0.744 & 9.112 & 0.623 & 0.76 & 0.761 & 0.762 & 0.761 & 0.76 \\ \hline
 \end{tabular}
@@
 \begin{figure*}[h]
-\begin{tabular}{ | l | l | l | l | l | l | l | l | l | }
+\small
+\begin{tabular}{ | l | l | l | l | l | l | l | l | }
 \hline
-	Dataset Name & Approx & Approx & Approx & Approx & Approx & Approx & Approx  \\ \hline
-	& k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap   \\ \hline
-	Soybean & 1.17 & 1.2 & 1.03 & 1.01 & 1.02 & 1.01 & 1.02  \\ 
-	Iris & 1.08 & 1.37 & 1.02 & 1 & 1.01 & 1 & 1.01  \\ 
-	Gauss\_5\_5\_5 & 1.07 & 1.44 & 1.03 & 1 & 1.03 & 1.01 & 1.03  \\ 
-	Gauss\_5\_10\_5 & 1.08 & 1.39 & 1.02 & 1.01 & 1.02 & 1.01 & 1.02  \\ 
-	Gauss\_5\_50\_5 & 1.09 & 1.27 & 1.03 & 1.02 & 1.03 & 1.01 & 1.03  \\ 
-	pmed1 & 1.11 & 1.41 & 1.03 & 1.01 & 1.03 & 1.01 & 1.02  \\ \hline
+	  & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
+	Dataset Name & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap \\ \hline
+	Soybean & 1.167 & 1.199 & 1.027 & 1.012 & 1.022 & 1.012 & 1.024 \\ 
+	Iris & 1.081 & 1.375 & 1.016 & 1.004 & 1.013 & 1.004 & 1.009 \\ 
+	Gaussian avg & 1.064 & 1.416 & 1.025 & 1.009 & 1.023 & 1.008 & 1.024 \\ 
+	pmed5 & 1.373 & 1.681 & 1.06 & 1.026 & 1.071 & 1.033 & 1.051 \\ 
+	pmed20 & 1.501 & 1.73 & 1.079 & 1.032 & 1.076 & 1.027 & 1.073 \\ 
+	pmed40 & 1.187 & 1.423 & 1.046 & 1.017 & 1.043 & 1.019 & 1.039 \\ 
+	pmedian avg & 1.203 & 1.456 & 1.041 & 1.017 & 1.039 & 1.016 & 1.037 \\ \hline
 \end{tabular}
@@
 \caption{Approximation Coeffiecients For Algorithms}
+\end{figure*}
+
+\begin{figure*}[h]
+\small
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+	   & Time & Time & Time & Time & Time & Time \\ \hline
+	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 0.0002 & 0.0002 & 0.2 & 0 & 0.2 & 0.2 \\ 
+	Iris & 0.0023 & 0.003 & 3.2 & 0.2 & 3.8 & 5.1 \\ 
+	Gaussian Average & 0.0858 & 0.0853 & 354.5 & 7.5 & 369.9 & 423.3 \\ 
+	pmed5 & 1.2 & 0.1 & 1.3 & 1.7 & \  & \  \\ 
+	pmed25 & 279.3 & 5 & 306.7 & 312.5 & \  & \  \\ 
+	pmedian avg & 87.4 & 2 & 94.5 & 101.6 & \  & \  \\ \hline
+\end{tabular}
+
+\caption{Running Times For Heirarchical Algorithms}
+\end{figure*}
+
+\begin{figure*}[h]
+\small
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+	   & F-Score & F-Score & F-Score & F-Score & F-Score & F-Score \\ \hline
+	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 0.6 & 0.5 & 0.8 & 0.6 & 0.7 & 0.6 \\ 
+	Iris & 0.5 & 0.5 & 0.9 & 0.8 & 0.8 & 0.7 \\ 
+	Gaussian Average & 0.4 & 0.4 & 0.6 & 0.6 & 0.6 & 0.6 \\ \hline
+\end{tabular}
+
+\caption{F-Score For Heirarchical Algorithms}
+\end{figure*}
+
+\begin{figure*}[h]
+\small
+\begin{tabular}{ | l | l | l | l | l | }
+\hline
+	   & Approx & Approx & Approx & Approx \\ \hline
+	Dataset & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 1.6 & 2.3 & 1.7 & 2 \\ 
+	Iris & 2560631.5 & 2550279.6 & 2787312 & 1623132.1 \\ 
+	Gaussian Average & 5.2 & 7.3 & 5.5 & 4.7 \\ 
+	pmed5 & 11.4 & 51.1 & 16.3 & 12 \\ 
+	pmed25 & 6.6 & 13.4 & 5.4 & 13.2 \\ 
+	pmedian avg & 10.3 & 19.8 & 10.5 & 8.6 \\ \hline
+\end{tabular}
+
+\caption{Approximation Coeffiecients For Heirarchical Algorithms}
 \end{figure*}

### 2015-05-10 tex/project.tex
Flat results and Hier Implementation

--- 2015-05-09 tex/project.tex
@@
 
-We implemented the local search swapping algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compare against $k$-means, $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
+We implemented the local search swapping algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python and compare against $k$-means, $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
 We also implemented the $k$-medoids algorithm with a maximum iteration of 1000.
@@
 
-For each data set and each clustering method, we report objective function values, the approximation ratio, computation time, and adjusted mutual information (AMI).
+For each data set and each clustering method, we report the approximation ratio, computation time, and adjusted mutual information (AMI).
 AMI is a measure which scores accuracy of predicted clusters compared to a true labeling by measuring the agreement of two assignments.
@@
 
-Comment on results here.
+In terms of computation time, our proposed algorithms were much slower than $k$-means, the spectral algorithm, or $k$-medoids.
+In general, the multiswaps algorithm was around two orders of magnitude slower with a small multiplicative increase as $p$ increased, and the supermodular algorithm was around four orders of magnitude slower.
+This was probably due to the extra sampling the algorithm had to do.
+
+In addition, the supermodular function did not perform as well in terms of AMI or approximation ratio compared to the other methods.
+Calculating the curvature, we saw that every  test set's curvature was almost equal to 1 so the guarantee of the algorithm was quite poor.
+Given that, the algorithm did maintain an approximation ratio below 2 for all instances and an AMI measure above 0.5 for almost all instances even averaging around 0.85 for the 7 dimensional Gaussian data.
+
+When increasing $p$ for the multiswaps algorithm, we did see a slight increase in time but the times remained on the same order of magnitude.
+The increase in $p$ did not seem to have any significant effect on the AMI measure or approximation ratio with the values being roughly equivalent.
+This may due to the fact the algorithm for $p=1$ performed quite well already with an approximation ratio very close to 1 that beat out $k$-medoids.
+In addition, PAM had the highest AMI measures for the iris and soybean data sets.
+The AMI measures for the Gaussian data set were higher than those of $k$-medoids in general but equivalent or only slightly weaker than those of the spectral method or $k$-means.
+Adding in a few rounds of $k$-medoids at the end of our multiswaps algorithm did not seem to have much of an effect on the AMI measure but did boost the approximation ratio, sometimes finding an optimal solution.
+
+It is interesting to note that higher dimensional Gaussian data seemed easier for the algorithms to classify along with smaller values of $k$ and lower percentage of noisy data, which makes sense based on our intuition of clustering.
+For the OR library instances, we saw that the approximation ratio tended to increase as $k$ grew larger but not as $n$ grew.
+
+Overall, the results of our experiments do not point to the fact that these proposed approximation algorithms can add any advantage over current methods.
+However, the results do point to the fact that we can use sampling to run PAM with lower computational cost while still maintaining good performance accuracy both in terms of AMI measure and approximation ratios.
+Adding in a few rounds of $k$-medoids seems also to be a good method to refine a given solution towards a local optimum.
 
@@
 \subsection{Implementation and Results}
+We implemented the nesting hierarchical framework above in Python offering options for both the randomized and deterministic versions.
+As a black-box $k$-median solver, we used $k$-medoids and the 1-swap+ algorithms above.
+We compared the results to hierarchical agglomerative clustering with both average linkage and Ward's linkage (implemented in the scipy package in Python).
 
-\section{Conclusion}
+We also tested a more computationally friendly heuristic of the hierarchical framework which just buckets solutions for values of $k$ that are a power of two.
+This yields many fewer $k$-median instances to solve and reduced the run-time significantly.
+This was motivated by the fact that with the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
+
+For each data set and each clustering method, we report the approximation ratio, computation time, and FScore measure.
+FScore is a measure that looks at the clusters represented on each level and how they represent the true clustering in terms of precision and recall.
+See \cite{Larsen} for more information.
+
+\section{Conclusion and Future Work}
 

### 2015-05-10 src/hierarchical_approx.py
Project write-up

@@ -1,11 +1,12 @@
 import numpy as np
-from sklearn import metrics
+#from sklearn import metrics
 import kmedoids
 import multiswaps
 import math
+import LoadData
 
-from sklearn.datasets.samples_generator import make_blobs
-from sklearn.metrics.pairwise import pairwise_distances
+#from sklearn.datasets.samples_generator import make_blobs
+#from sklearn.metrics.pairwise import pairwise_distances
 
 def next_assignment(Z, k, assignment):
     n= len(assignment)
@@ -56,14 +57,21 @@ def bucketed_solns(distances, beta, beta_0, alg, logn = False):
             k = 2*k
         else:
             k += 1
+
+    min_cost = (1.0/np.min([all_costs[i] for i in range(len(all_costs)) if all_costs[i] != 0]))
+    all_costs = [min_cost*all_costs[i] for i in range(len(all_costs))]
+    print min_cost
     
     bucketed_centers = []
     lb = 0
     ub = beta_0
     k = n
     while k>1:
-        next_bucket = [i for i in range(len(all_costs)) if all_costs[i]<= ub and \
-                       all_costs[i] > lb]
+        if k == n:
+            next_bucket = [i for i in range(len(all_costs)) if all_costs[i] == 0]           
+        else:
+            next_bucket = [i for i in range(len(all_costs)) if all_costs[i]< ub and \
+                           all_costs[i] >= lb]
         if len(next_bucket) > 0 and len(next_bucket) != n:
             x = min(next_bucket)
             k = len(all_centers[x])
@@ -134,17 +142,22 @@ def cluster(distances, alg = ["k-medoids", None], random = False, logn = False):
     
 if __name__ == '__main__':
     centers = [[1, 1], [-1, -1], [1, -1]]
-    X, labels_true = make_blobs(n_samples=50, centers=centers, cluster_std=0.5,
-                                random_state=999)
-    distances = pairwise_distances(X)
+    X, y,n,k = LoadData.LoadData("/Users/Alice/Desktop/approx-clustering/data/Real Data Sets/iris.data.txt",0,0)
+    
+    distances = np.zeros(shape=(n,n))
+    for i in range(n):
+        for j in range(n):
+            distances[i][j] = np.linalg.norm(X[i]-X[j])
 
-    Z= cluster(distances,["multiswaps",1],0,0)
-    #print Z
-    n=6
+    Z= cluster(distances,["k-medoids",None],0,0)
+    print Z
+    
[…]

### 2015-05-10 src/multiswaps.py

@@ -2,11 +2,11 @@ import random
 import itertools
 import numpy as np
 from scipy.special import binom
-import LoadpMedian
+#import LoadpMedian
 import kmedoids
 
-from sklearn.datasets.samples_generator import make_blobs
-from sklearn.metrics.pairwise import pairwise_distances
+#from sklearn.datasets.samples_generator import make_blobs
+#from sklearn.metrics.pairwise import pairwise_distances
 
 def find_closest_median(i, medoids, distances):
     closest = None

### 2015-05-10 tex/project.tex
Project write-up

@@
 
-We implemented the local search swapping algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python and compare against $k$-means, $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
+We implemented the local search swapping algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compare against $k$-means, $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
 We also implemented the $k$-medoids algorithm with a maximum iteration of 1000.
@@
 
-For each data set and each clustering method, we report the approximation ratio, computation time, and adjusted mutual information (AMI).
+For each data set and each clustering method, we report objective function values, the approximation ratio, computation time, and adjusted mutual information (AMI).
 AMI is a measure which scores accuracy of predicted clusters compared to a true labeling by measuring the agreement of two assignments.

### 2015-05-10 tex/project.tex
Hier Result Analysis

@@
 
+In general, the hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure even for our logarithmic version.
+The FScore measures were very low for the agglomerative clustering methods, often around 0.25, but shot up to closer to the AMI measures we saw in flat clustering for our implemented algorithms.
+
+For $k$-medoids in the non-logarithmic version and PAM in the logarithmic version, the hierarchical framework increased the computation time by three orders of magnitude, whereas this was decreased to two orders difference for $k$-medoids logarithmic version.
+Interestingly, the approximation factors did not seem to change between the randomized and deterministic versions for $k$-medoids.
+The approximation ratios seemed best as $n$ increased or as the dimension increased.
+In particular, for the Gaussian data in 7 dimensions, the approximation ratio was often around 3.
+Recall that this approximation measures the constructed $k$-median solution on every level against an optimal one so we expect the ratio to not come close to 1 in many cases.
+
 \section{Conclusion and Future Work}
+For partitional clustering, it seems that the approximation algorithms in the literature are in general slower and cannot improve much on the current heuristics used despite their approximation guarantees.
+This was not true in the hierarchical clustering case where we saw significant improvement in the FScore measures for our implemented methods and overall good approximation ratios where none were had before.
+In particular, our logarithmic nesting algorithm seems to provide the best balance between performance and computation.
+
+In the future, we hope to explore how linear programming based methods perform in this context as well and to implement our own agglomerative clustering so we can test it on the OR library instances.
+Lastly, when nesting solutions in the logarithmic framework, we could use agglomerative clustering to make smarter local merges.
 

### 2015-05-10 tex/project.tex
Updated bib further

@@
 The algorithm repeats this until the solution becomes stable and is more computationally efficient than PAM \cite{Park}.
-Additionally, Velmurugan and Santhanam showed that on uniform and normally distributed data the $k$-medoids algorithm was empirically faster than that of $k$-means \cite{Velmurugan}.
 
@@
 \begin{description}
-\item[Gaussian] Synthetic $k$-center Gaussian data, generated in Python.
-In particular, Gauss\textunderscore d \textunderscore k \textunderscore e was generated by first choosing k random centers in $[-1,1]^d$.
-Around each center, we then generated $1000/k \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
-\item[OR $p$-median Library] From a collection of test data sets created by J. E. Beasley with data sets of size $100, 200, \ldots, 900$ and varying values of $k$ \cite{Beasley}. 
+\item[Gaussian] 27 synthetic $k$-center Gaussian data sets, generated in Python.
+In particular, Gauss\textunderscore d\textunderscore k\textunderscore e was generated by first choosing k random centers in $[-1,1]^d$.
+Around each center, we then generated $(1000/k) \cdot \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
+\item[OR $p$-median Library] From a collection of test data sets created by \cite{Beasley} with 40 test sets of size $100, 200, \ldots, 900$ and varying values of $k$. 
 \item[UCI Iris] \cite{Iris} Contains 4 attributes on 150 instances of flowers, each of with belongs to one of three species of iris.
@@
 
-\section{Acknowledgements}
-We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
+%\section{Acknowledgements}
+%We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
 

### 2015-05-10 src/hierarchical_approx.py
Fixed Iris error

@@ -60,18 +60,18 @@ def bucketed_solns(distances, beta, beta_0, alg, logn = False):
 
     min_cost = (1.0/np.min([all_costs[i] for i in range(len(all_costs)) if all_costs[i] != 0]))
     all_costs = [min_cost*all_costs[i] for i in range(len(all_costs))]
-    print min_cost
     
     bucketed_centers = []
     lb = 0
     ub = beta_0
     k = n
     while k>1:
-        if k == n:
+        if k == n and np.min(all_costs) == 0:
             next_bucket = [i for i in range(len(all_costs)) if all_costs[i] == 0]           
         else:
             next_bucket = [i for i in range(len(all_costs)) if all_costs[i]< ub and \
                            all_costs[i] >= lb]
+            
         if len(next_bucket) > 0 and len(next_bucket) != n:
             x = min(next_bucket)
             k = len(all_centers[x])
@@ -140,24 +140,24 @@ def cluster(distances, alg = ["k-medoids", None], random = False, logn = False):
         
     return Z
     
-if __name__ == '__main__':
-    centers = [[1, 1], [-1, -1], [1, -1]]
-    X, y,n,k = LoadData.LoadData("/Users/Alice/Desktop/approx-clustering/data/Real Data Sets/iris.data.txt",0,0)
-    
-    distances = np.zeros(shape=(n,n))
-    for i in range(n):
-        for j in range(n):
-            distances[i][j] = np.linalg.norm(X[i]-X[j])
-
-    Z= cluster(distances,["k-medoids",None],0,0)
-    print Z
-    
-    assignment = [i for i in range(n)]
-    for k in range(n-1,145,-1):
-        assignment = next_assignment(Z,k,assignment)
-        cost = sum([distances[i][assignment[i]] for i in range(n)])
-        print assignment
-        print cost
+##if __name__ == '__main__':
+##    centers = [[1, 1], [-1, -1], [1, -1]]
+##    X, y,n,k = LoadData.LoadData("/Users/Alice/Desktop/approx-clustering/data/Real Data Sets/iris.data.txt",0,0)
+##    
+##    distances = np.zeros(shape=(n,n))
+##    for i in range(n):
+##        for j in range(n):
+##            distances[i][j] = np.linalg.norm(X[i]-X[j])
+##
+##    Z= cluster(distances,["k-medoids",None],0,0)
+##    print Z
+##    
+##    assignment = [i for i in range(n)]
+##    for k in range(n-1,145,-1):
[…]

### 2015-05-10 tex/project.tex
Further updated bib and down to 8 pages!

@@
 \hline
-	Data Set Name &Optimal Objective & $k$ & $n$ & dim   \\ \hline
-	Soybean & 113.8 & 4 & 47 & 35 \\ 
-	Iris & 98.2 & 3 & 150 & 4 \\ 
+ Dataset Name & Opt Obj & $k$ & $n$ & dim \\ \hline
+Soybean & 114 & 4 & 47 & 35 \\ 
+Iris & 98 & 3 & 150 & 4 \\ 
 Gauss\_5\_5\_5 & 195.4 & 5 & 388 & 5 \\ 
 Gauss\_5\_10\_5 & 270 & 10 & 544 & 5 \\ 
-	Gauss\_5\_50\_5 & 178.1 & 50 & 408 & 5 \\ 
-	pmed1 & 5819 & 5 & 100 & \  \\ 
-	\hline
+Gaussian avg & 241 & 22 & 507 & 5 \\ 
+pmed5 & 1355 & 33 & 100 & \  \\ 
+pmed20 & 1789 & 133 & 400 & \  \\ 
+% pmed40 & 5128 & 90 & 900 & \  \\ 
+pmedian avg & 5535 & 46 & 460 & \  \\ \hline
  \end{tabular}
@@
 
-Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core Intel processor.
+Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core Intel processor. Results are given in Figure 2  below with some results averaged due to space constraints.
 
 \begin{figure*}[h]
+\label{fig:flat_results}
+\small
 \begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | }
@@
 	Gauss\_5\_10\_5 & 0.054 & 0.095 & 0.109 & 134.137 & 6.834 & 5.407 & 9.243 & 9.428 & 15.96 \\ 
-	Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 \\ 
-	pmed1 & N/A & N/A & 0.003 & 5.042 & 0.084 & 0.09 & 0.155 & 0.161 & 0.255 \\ \hline
+	%Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 \\ 
+	Gaussian avg & 0.073 & 0.128 & 0.182 & 144.455 & 13.289 & 13.021 & 18.749 & 18.696 & 24.592 \\ 
+	pmed5 & 0.038 & 8.298 & 0.481 & 0.622 & 0.653 & 0.748 & 0.939 & \  & \  \\ 
+	pmed20 & 1.795 & 174.951 & 52.208 & 44.979 & 68.554 & 68.231 & 79.288 & \  & \  \\ 
+	% pmed40 & 3.527 & 849.852 & 251.259 & 264.867 & 250.412 & 391.738 & 350.235 & \  & \  \\ 
+	pmedian avg & 0.807 & 203.308 & 40.386 & 40.347 & 48.264 & 56.76 & 56.19 & \  & \  \\ \hline
 \end{tabular}
+\vspace{0.2 cm}
 
-\caption{Running Times For Algorithms}
-\end{figure*}
-
-\begin{figure*}[h]
 \begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
@@
 	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap   \\ \hline
-	Soybean & 0.69 & 0.18 & 0.55 & 0.61 & 0.71 & 0.73 & 0.72 & 0.71 & 0.71  \\ 
-	Iris & 0.75 & 0.76 & 0.69 & 0.7 & 0.77 & 0.76 & 0.76 & 0.76 & 0.76  \\ 
-	Gauss\_5\_5\_5 & 0.74 & 0.71 & 0.69 & 0.66 & 0.74 & 0.75 & 0.74 & 0.74 & 0.74  \\ 
-	Gauss\_5\_10\_5 & 0.78 & 0.74 & 0.68 & 0.58 & 0.73 & 0.72 & 0.74 & 0.74 & 0.73  \\ 
-	Gauss\_5\_50\_5 & 0.65 & 0.6 & 0.62 & 0.57 & 0.62 & 0.62 & 0.63 & 0.63 & 0.62  \\ \hline
+	Soybean & 0.686 & 0.178 & 0.55 & 0.612 & 0.708 & 0.732 & 0.717 & 0.708 & 0.712 \\ 
+	Iris & 0.748 & 0.757 & 0.69 & 0.7 & 0.772 & 0.760 & 0.763 & 0.757 & 0.761 \\ 
+	Gauss\_5\_5\_5 & 0.741 & 0.713 & 0.687 & 0.657 & 0.742 & 0.749 & 0.741 & 0.744 & 0.737  \\ 
+	Gauss\_5\_10\_5 & 0.783 & 0.736 & 0.677 & 0.575 & 0.729 & 0.725 & 0.737 & 0.737& 0.729  \\ 
+	%Gauss\_5\_50\_5 & 0.646 & 0.596 & 0.621 & 0.570 & 0.622 & 0.619 & 0.627 & 0.631 & 0.625  \\ 
+	Gaussian avg & 0.777 & 0.744 & 9.112 & 0.623 & 0.76 & 0.761 & 0.762 & 0.761 & 0.76 \\ \hline
 \end{tabular}
+\vspace{0.2 cm}
 
-\caption{AMI For Algorithms}
-\end{figure*}
-
-\begin{figure*}[h]
 \begin{tabular}{ | l | l | l | l | l | l | l | l | l | }
@@
 	& k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap   \\ \hline
-	Soybean & 1.17 & 1.2 & 1.03 & 1.01 & 1.02 & 1.01 & 1.02  \\ 
-	Iris & 1.08 & 1.37 & 1.02 & 1 & 1.01 & 1 & 1.01  \\ 
-	Gauss\_5\_5\_5 & 1.07 & 1.44 & 1.03 & 1 & 1.03 & 1.01 & 1.03  \\ 
-	Gauss\_5\_10\_5 & 1.08 & 1.39 & 1.02 & 1.01 & 1.02 & 1.01 & 1.02  \\ 
-	Gauss\_5\_50\_5 & 1.09 & 1.27 & 1.03 & 1.02 & 1.03 & 1.01 & 1.03  \\ 
-	pmed1 & 1.11 & 1.41 & 1.03 & 1.01 & 1.03 & 1.01 & 1.02  \\ \hline
+	Soybean & 1.167 & 1.199 & 1.027 & 1.012 & 1.022 & 1.012 & 1.024 \\ 
+	Iris & 1.081 & 1.375 & 1.016 & 1.004 & 1.013 & 1.004 & 1.009 \\ 
+	Gauss\_5\_5\_5 & 1.070 & 1.436 & 1.030 & 1.002 & 1.027 & 1.005 & 1.026  \\ 
+	Gauss\_5\_10\_5 & 1.080 & 1.389 & 1.017 & 1.011 & 1.018 & 1.009 & 1.017  \\ 
+	%Gauss\_5\_50\_5 & 1.0887 & 1.272 & 1.030 & 1.020 & 1.027 & 1.014 & 1.033  \\ 
+	Gaussian avg & 1.064 & 1.416 & 1.025 & 1.009 & 1.023 & 1.008 & 1.024 \\ 
+	pmed5 & 1.373 & 1.681 & 1.06 & 1.026 & 1.071 & 1.033 & 1.051 \\ 
+	pmed20 & 1.501 & 1.73 & 1.079 & 1.032 & 1.076 & 1.027 & 1.073 \\ 
+	%pmed40 & 1.187 & 1.423 & 1.046 & 1.017 & 1.043 & 1.019 & 1.039 \\ 
+	pmedian avg & 1.203 & 1.456 & 1.041 & 1.017 & 1.039 & 1.016 & 1.037 \\ \hline
 \end{tabular}
 
-\caption{Approximation Coeffiecients For Algorithms}
+\caption{Average Computation Time (sec), AMI measure, and Approximation Coeffiecients for Partitional Clustering Algorithms}
 \end{figure*}
@@
 In general, the multiswaps algorithm was around two orders of magnitude slower with a small multiplicative increase as $p$ increased, and the supermodular algorithm was around four orders of magnitude slower.
-This was probably due to the extra sampling the algorithm had to do.
+This was probably due to the extra potential function sampling.
 
@@
 Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$.
-We define
-\[ \mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i)) . \]
+We define $\mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i))$.
 
-Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if 
-\begin{enumerate}
-\item $S_1 \subset S_2$, 
-\item  $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, and 
-\item $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
-\end{enumerate}
+Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if  (1) $S_1 \subset S_2$, (2)  $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, and (3) $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
  A solution to the hierarchical clustering problem is therefore a chain of nested solutions.
@@
 Consider any $k$-median algorithm $\mathcal{A}$, such as the algorithms above.
-After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce  augmentation step that will nest these clusterings so that they fit into the hierarchical framework and only increase the costs by a bounded factor.
+After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce  augmentation step that will nest these clusterings into the hierarchical framework and only increase the costs by a bounded factor.
 
@@
 
-We test this hierarchical framework for both the multiswap and supermodular local search algorithms above.
-
 \subsection{Implementation and Results}
@@
 See \cite{Larsen} for more information.
+Results are in Figure 3.
+
+\begin{figure*}[h]
+\label{fig:hier_results}
+\small
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+	   & Time & Time & Time & Time & Time & Time \\ \hline
+	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 0.0002 & 0.0002 & 0.2 & 0 & 0.2 & 0.2 \\ 
+	Iris & 0.0023 & 0.003 & 3.2 & 0.2 & 3.8 & 5.1 \\ 
+	Gauss\_5\_5\_5 &  & & & & & \\
+	Gauss\_5\_10\_5 & & & & & & \\
+	%Gauss\_5\_50\_5 &  & & & & & \\
+	Gaussian Average & 0.0858 & 0.0853 & 354.5 & 7.5 & 369.9 & 423.3 \\ 
+	pmed5 & 1.2 & 0.1 & 1.3 & 1.7 & \  & \  \\ 
+	pmed25 & 279.3 & 5 & 306.7 & 312.5 & \  & \  \\ 
+	pmedian avg & 87.4 & 2 & 94.5 & 101.6 & \  & \  \\ \hline
+\end{tabular}
+\vspace{0.2 cm}
+
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+	   & F-Score & F-Score & F-Score & F-Score & F-Score & F-Score \\ \hline
+	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 0.6 & 0.5 & 0.8 & 0.6 & 0.7 & 0.6 \\ 
+	Iris & 0.5 & 0.5 & 0.9 & 0.8 & 0.8 & 0.7 \\ 
+	Gauss\_5\_5\_5 &  & & & & & \\
+	Gauss\_5\_10\_5 &  & & & & & \\
+	%Gauss\_5\_50\_5 &  & & & & & \\
+	Gaussian Average & 0.4 & 0.4 & 0.6 & 0.6 & 0.6 & 0.6 \\ \hline
+\end{tabular}
+\vspace{0.2 cm}
+
+\begin{tabular}{ | l | l | l | l | l | }
+\hline
+	   & Approx & Approx & Approx & Approx \\ \hline
+	Dataset & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 1.6 & 2.3 & 1.7 & 2 \\ 
+	Iris & 2560631.5 & 2550279.6 & 2787312 & 1623132.1 \\ 
+	Gauss\_5\_5\_5 &  & & &  \\
+	Gauss\_5\_10\_5 &  & & &  \\
+	%Gauss\_5\_50\_5 &  & & &  \\
+	Gaussian Average & 5.2 & 7.3 & 5.5 & 4.7 \\ 
+	pmed5 & 11.4 & 51.1 & 16.3 & 12 \\ 
+	pmed25 & 6.6 & 13.4 & 5.4 & 13.2 \\ 
+	pmedian avg & 10.3 & 19.8 & 10.5 & 8.6 \\ \hline
+\end{tabular}
+
+\caption{Average Computation Time (sec), FScore measure, and Approximation Coeffiecients for Hierarchical Clustering Algorithms}
+\end{figure*}
 
@@
 In particular, for the Gaussian data in 7 dimensions, the approximation ratio was often around 3.
-Recall that this approximation measures the constructed $k$-median solution on every level against an optimal one so we expect the ratio to not come close to 1 in many cases.
+Recall that this approximation measures the constructed $k$-median solution on every level against an optimal one so we expect the ratio be higher than 1.
+The high approximation ratio for the iris data set was due to the fact that the optimal solution for $k = n-1, n-2$ was almost equal to zero and our heuristic $k$-median algorithms only sometimes picked up on these solutions.
+The approximation factor on the lower levels was more consistent with the other results.
 
@@
 
-In the future, we hope to explore how linear programming based methods perform in this context as well and to implement our own agglomerative clustering so we can test it on the OR library instances.
-Lastly, when nesting solutions in the logarithmic framework, we could use agglomerative clustering to make smarter local merges.
+In the future, we hope to explore how linear programming based methods perform in both the flat and hierarchical context as well and to implement our own agglomerative clustering to test on the OR library instances and help make smarter local merges in the nesting framework.
 

### 2015-05-10 tex/project.tex
WIP

--- 2015-05-10 tex/project.tex
@@
 Partitioning data (clustering) into similar groups (clusters) is one of the key tools to learn about a set of unlabeled data such as a set of customers who have visited a website or a set of handwritten documents.
-Clustering in these situations helps us learn underlying structure and classifications of our data, and it presents large data sets in a simplified and compressed view \cite{Jain}.
+Clustering in these situations helps us learn underlying structure and classifications of our data, and it allows large data sets to be presented in a simplified and compressed view \cite{Jain}.
 
-There are two main branches of clustering: partitional (or flat) clustering and hierarchical clustering.
+There are two main branches of clustering: partitional (or flat) and hierarchical.
 In partitional clustering we want to partition the data into disjoint sets whereas in hierarchical clustering we are interested in finding a nested tree of partitions.
-Often these problems are solved using heuristic algorithms which have no guarantee on the solution produced but tend to be fast and easy to implement.
-We will look at how local search approximation algorithms, which do have some guarantee on the constructed solution and are less computationally expensive than other approximation methods, compare in practice to the most used machine learning algorithms both in the quality of the solution and in computation time.
+Often these problems are solved using heuristic algorithms which have no guarantee on the solution produced but are typically fast and easy to implement.
+We will look at how local search approximation algorithms, which do have some approximation guarantee on the constructed solution, and are less computationally expensive than other approximation methods, compare in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
 
@@
 
-Clustering algorithms data back to over 50 years ago, and since then there have been many variations introduced \cite{Jain}.
-However, clustering algorithms are often formulated as NP-hard problems and are inherently hard to solve \cite{Jain}.
-When comparing algorithms, we need to consider the stability of the solutions, the computation time, and the proposed objective.
+Clustering algorithms date back more than 50 years, and there have been many variations introduced since\cite{Jain}.
+However, clustering algorithms are often formulated as NP-hard problems with non-convex objective functions, which are inherently hard to solve \cite{Jain}.
+Below, we will review the widely used heuristics and known approximation algorithms for both partitional and hierarchical clustering.
 
@@
 
-For partitional clustering, we are often given a number $k$, representing the number of clusters, and distances between the data, representing a measure of dissimilarity between points, such as Euclidean distance.
-In addition, many problems require that this distance satisfy the triangle inequality.
-The objective is then formulated as identifying $k$ clusters so as to minimize some function of the distances.
+For partitional clustering, we often specify a number $k$, representing the number of clusters, and pairwise distances between the data points, representing a measure of dissimilarity, such as Euclidean distance.
+The objective is then to identify $k$ clusters so as to minimize some function of the distances.
 
-One popular objective is that of $k$-means in which we want to choose $k$ cluster centers to minimize the sum of squared distances from each point to its cluster center \cite{Jain}.
-The most used algorithm for this problem is Lloyd's $k$-means algorithm which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}.
-This algorithm is only guaranteed to reach a local optimum and can be very sensitive to outliers and the initial points chosen \cite{Kanungo}.
-There also exists a local search algorithm for $k$-means which guarantees to be within a factor of $(9+\varepsilon)$ of the optimal solution that performs well in practice when combined with Lloyd's algorithm \cite{Kanungo}.
-This algorithm is a variant of the multiswap method proposed below but with respect to a $k$-means objective function and only implemented for $p=1$.
-A relaxation of this problem is fuzzy $c$-means which has the same objective but allows points to belong partially to multiple clusters \cite{ESL}.
+Arguably the most common partitional clustering method is $k$-means, in which we want to choose $k$ cluster centers to minimize the sum of squared distances from each point to its cluster center \cite{Jain}.
+$k$-means is usually solved using Lloyd's algorithm (also referred to simply as the $k$-means algorithm), which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}.
+Lloyd's algorithm is only guaranteed to reach a local optimum and can be very sensitive to outliers and the initial points chosen \cite{Kanungo}.
 
-Variations on $k$-means attempt to reduce the sensitivity and worst-case behavior.
-One other objective is to minimize the maximum distance from a point to its cluster center.
-This problem can be thought of as minimizing the worst case and is called $k$-center clustering.
-Here, a simple greedy algorithm is guaranteed to come within a factor of $2$ of the optimal solution as long as the given distances satisfy the triangle inequality \cite{Gonzalez}.
+Several variations on $k$-means attempt to reduce the sensitivity and worst-case behavior, such as to minimize the maximum distance from a point to its cluster center, called $k$-center clustering.
+Here, a simple greedy algorithm is guaranteed to produce a solution within a factor of $2$ of the optimum, as long as the given distances satisfy the triangle inequality \cite{Gonzalez}.
 Furthermore, this is best approximation ratio possible unless P=NP \cite{Shmoys}.
 However, $k$-center does not perform well in practice compared to other methods \cite{Murphy}.
-This may be because it focuses too much on the worst case distance.
-
-An objective function in between $k$-means and $k$-center is to minimize the sum of distances from each point to the median point of its cluster.
-This problem is called $k$-median or $k$-medoids clustering.
-Another way of defining this problem is choosing a set of $k$ data points to be centers so as to minimize the sum of distances from each point to their closest center.
-
-There are two common heuristic algorithms for this problem.
-The first is a local search algorithm called partition around medoids (PAM) which starts with $k$ randomly chosen centers and repeatedly looks for a single point swap to improve the objective function until no such move is possible \cite{ESL}.
-Another common heuristic for this problem (often called just $k$-medoids) is similar to that of $k$-means by assigning each point to their closest center and then greedily updating the center within each cluster \cite{Park}.
-The algorithm repeats this until the solution becomes stable and is more computationally efficient than PAM \cite{Park}.
-Additionally, Velmurugan and Santhanam showed that on uniform and normally distributed data the $k$-medoids algorithm was empirically faster than that of $k$-means \cite{Velmurugan}.
 
 More recently, spectral based methods have become popular for clustering.
-These algorithms construct a graph based on the similarity matrix and compute the eigenvalues of the normalized graph Laplacian matrix \cite{Luxburg}.
-These eigenvectors are either used to split the current cluster in half using the second eigenvector (continuing recursively until there are $k$ clusters) or the eigenvectors are used to project the data points into a lower dimensional space on which the $k$-means algorithm above is run \cite{Verma}.
+These algorithms construct a graph based on the similarity matrix and compute the eigenvectors of the normalized graph Laplacian matrix \cite{Luxburg}.
+These eigenvectors are used to split the current cluster in half using the second eigenvector (continuing recursively until there are $k$ clusters).
+Another method is to use the eigenvectors to project the data points into a lower dimensional space on which the $k$-means algorithm is run \cite{Verma}.
 We will base our comparisons against the algorithm introduced by \cite{Meila} which uses the latter method and performed best in practice on several datasets \cite{Verma}.
 
-The $k$-median problem is also a fundamental problem in approximation algorithms.
-The problem is formulated as above: choose $k$ cluster centers from a set $X$ so as to minimize the sum of distances from each point to their closest center.
-The best known approximation algorithm for this problem guarantees a solution within a factor of 2.592 of the optimal solution \cite{Wu}.
-However, this algorithm uses linear programming and is computationally intensive compared to the local search and greedy algorithms above.
-On the other hand, it is known that a local search algorithm that considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
-Note that for $p=1$, this is the same as PAM.
-As far as we know, nobody has implemented this algorithm on clustering data to compare the relative trade-off of solution quality and computation time as $p$ increases.
+An objective function that is somewhat in between $k$-means and $k$-center is to minimize the sum of distances from each point to the median point of its cluster, called $k$-median or $k$-medoids clustering.
+There are two common heuristic algorithms for this problem.
+The first is a local search algorithm called partition around medoids (PAM), which starts with $k$ randomly chosen data points as centers, and looks to swap one of the centers with another data point to improve the objective value.
+This repeats until no more improving swaps exist \cite{ESL}.
+Another common heuristic (which we shall call $k$-medoids throughout this paper), which is more computationally efficient than PAM \cite{Park}, is a varitation on Lloyd's $k$-means algorithm.
+Starting with randomly chosen centers, we assign each data point to its closest center, and then update the center within each cluster \cite{Park}.
+The algorithm repeats this until the solution becomes stable.
+Additionally, Velmurugan and Santhanam showed that on uniform and normally distributed data, $k$-medoids was empirically faster than $k$-means \cite{Velmurugan}.
 
-Additionally, center-based clustering in general can be seen as minimizing a supermodular decreasing function subject to a cardinality constraint, a problem related to maximizing a submodular function that is well-studied in approximation algorithms.
-Furthermore, the approximation algorithms in this area are constructed using local-search or greedy methods and are therefore computationally tractable on large data.
-In addition, using supermodular functions means our distance measure does not have to satisfy the triangle inequality, which can be useful when considering non-vector data.
-This algorithm has also not been previously tested in practice.
+Various approximation algorithms for he $k$-median problem have been proposed, with a best known approximation ratio of 2.592 \cite{Wu}.
+However, this algorithm uses linear programming and is computationally intensive compared to local search and greedy algorithms.
+It is known that a local search algorithm that considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
+Note that for $p=1$, this equivalent to PAM.
+As far as we know, this algorithm has not been implemented or tested on clustering data, so the relative trade-off of solution quality and computation time as $p$ increases is unknown.
 
-There exist many other methods for clustering.
-Methods such as MeanShift and DBSCAN focus on finding clusters of high density \cite{Comaniciu, Ester}.
+Other partitional clustering methods include MeanShift and DBSCAN, which focus on finding clusters of high density \cite{Comaniciu, Ester}.
 We choose not to compare against density based methods since their objective differs widely from the center-based methods above.
-In addition, we will focus on test data in which center-based methods seem appropriate.
 We also choose not to compare our proposed methods against Bayesian clustering methods such as Gaussian mixture models since these methods require many more parameters (representing the prior information).
-In addition, the $k$-means algorithm above can be thought of as a special case of the Gaussian mixture model when all the covariances are equal \cite{ESL}.
-Therefore, this algorithm can serve as a basis of comparison when both algorithms have the same available information.
-Lastly, if we relax the constraint to have exactly $k$ clusters and instead allow the user to input ``preferences'' on the data points being clusters, then a recent algorithm called affinity propagation has shown to perform particularly well \cite{Frey}.
-Again, because of the difference in available information, we do not to test against this algorithm.
+Lastly, if we relax the constraint to identify exactly $k$ clusters, and instead allow the user to input ``preferences'' on the data points being clusters, a recent method called affinity propagation has been shown to perform particularly well \cite{Frey}.
+However, again, because of the difference in available information, we do not to test against this algorithm.
 
@@
 
-In hierarchical clustering, on the other hand, we do not specify a number of clusters.
-Instead, we want to find a hierarchy of clusters so that the clusters at each level are formed by merging clusters at a lower level.
-This gives a bigger picture about the dissimilarity between different groups of points and can allow the user to choose a ``natural'' level of clustering \cite{ESL}.
+In the hierarchical clustering setting, we do not specify a number of clusters.
+Instead, we aim to find a hierarchy of clusters so that clusters at a given level are a merging of clusters at a lower level.
+This presents a larger picture about the similarity between groups of data points, and can allow the user to identify a ``natural'' level of clustering \cite{ESL}.
 
-There are two main basic methods for hierarchical clustering: agglomerative clustering and divisive clustering.
-Agglomerative clustering starts with each data point in its own cluster.
-It then repeatedly merge two clusters to form the next level until the data is in a single cluster.
-On the other hand, divisive clustering starts with all points in a single cluster and repeatedly divides a cluster into two parts either using a $k$-means or $k$-medoids algorithm with $k=2$ or by greedily separating the cluster so that all points are closer to their part of separated cluster.
-The cluster to split is chosen to be the cluster with highest diameter or average dissimilarity.
-This method has not been studied nearly as much as agglomerative clustering \cite{ESL} but has had relative success in classifying document datasets \cite{Zhao}.
+The most common method for hierarchical clustering is agglomerative clustering.
+Agglomerative clustering starts with each data point in its own cluster, and proceeds to recursively merge the two closest clusters to form the next clustering level in the hierarchy, repeating until the entire data set is a single cluster.
+There are several choices on how to measure distance in order to choose which clusters to merge.
+Single linkage (also called nearest neighbor) measures the distance between two clusters to be the minimum pairwise distance between the data points in each cluster.
+Complete linkage clustering instead defines the distance to be the largest pairwise distance.
+A method between these two extremes is group average linkage which looks at the average pairwise distance between clusters.
+This strikes a balance between the closeness and compactness of the two techniques above and gives more reasonable results in practice \cite{ESL}.
+Lastly, there is Ward's method which chooses the two clusters to merge so as to minimize the sum of each cluster's variance \cite{Ward}.
 
-There are four choices of criteria for choosing which clusters to merge. 
-In single linkage clustering (also called nearest neighbor), we measure the distance between two clusters to be the smallest pairwise distance and then merge the two closest clusters.
-The main drawback to this method is that two points within a cluster can be very far apart.
-Complete linkage clustering on the other hand measures the distance between two clusters to be the largest pairwise distance and then merges the two closest clusters.
-While points within a single cluster will be quite similar using this metric, points may not belong to their closest cluster.
-A method between these two extremes is group average linkage which looks at the average pairwise distance between clusters and merges the two closest clusters.
-This strikes a balance between the closeness and compactness of the two techniques above and gives more reasonable results in practice \cite{ESL}.
-Lastly, there is Ward's method which merges the two clusters so as to minimize the sum of each cluster's variance \cite{Ward}.
-All four of these methods can be implemented in the Lance-Williams recursive framework which updates the cluster distances efficiently \cite{Lance}.
-
-Hierarchical clustering has also been studied in approximation algorithms.
-In this problem, we need to identify a hierarchy of clusters with their corresponding cluster centers so as to minimize the ratio of the cost of our solution on any level with $k$ clusters to the optimal $k$-median solution.
-Here, the cost of our solution is given by the sum of the distance from each point to its cluster center (i.e.
-the $k$-median objective we described above).
+Approximation algorithms for hierarchical clustering have also been studied, where we aim to identify a hierarchy of clusters so as to minimize the ratio of the objective value on every level in the hierarchy of size $k$ to the optimal objective value of the partitional clustering problem with $k$ clusters.
 Note that it may not always be possible to find a solution with approximation ratio 1.
@@
 This was improved by Lin et al.~\cite{Lin} who gave an incremental approximation algorithm framework.
-In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to construct a hierarchical clustering by increasing the cost by at most a factor of $20.71$ (or a factor of $10.03$ when randomized).
-Using the $(3+\frac{2}{p})$-approximation algorithm as the black-box, this translates into a $20.71 (3+\frac{2}{p})$-approximation algorithm (or $10.03 (3+\frac{2}{p})$ when randomized).
-These ratios can be improved by using a Lagrangian multiplier preserving $2$-approximation algorithm but the local search algorithm performed comparably in experimental tests \cite{Nagarajan}.
+In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to construct a hierarchical clustering that increases the objective value by at most a factor of $20.71$ (or $10.03$ when randomized) at each level.
+The local search $(3+\frac{2}{p})$-approximation algorithm as the $k$-median solver performed well in experimental tests \cite{Nagarajan}.
 While the authors tested the variants of this algorithm against other $k$-median approximation algorithms, it has not been tested against traditional clustering algorithms or on any labeled data sets.
 
-Again, several algorithms have been built upon agglomerative clustering to incorporate non-spherical cluster shapes and categorical data such as the CURE, ROCK, and Chameleon algorithms \cite{Guha_CURE, Guha_ROCK, Karypis} but do not relate as much to our proposed algorithm in objective.
-We also choose not compare our proposed method with Bayesian hierarchical methods for the same reasons as above.
+Several algorithms have been built upon agglomerative clustering to incorporate non-spherical cluster shapes and categorical data such as the CURE, ROCK, and Chameleon algorithms \cite{Guha_CURE, Guha_ROCK, Karypis}, but do not relate as much to the above approximation algorithm in objective.
+We also choose not compare the approximation framework with Bayesian hierarchical methods for the same reasons as above.
 
@@
 
+We will focus our attention on the $k$-median problem, where we are given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$.
+The goal is to find a set $S \subseteq X$ such that $|S| \leq k$ so as to minimize
+\[ f(S) = \sum_{ i\in X} \min_{j \in S} d(i,j). \]
+
 \subsection{Local Search with Multiswaps}
-In the $k$-median problem, we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$. The goal is to find a set $S \subseteq X$ such that $|S| \leq k$ so as to minimize 
-\[ f(S) = \sum_{ i\in X} \min_{j \in S} d(i,j). \]
 Consider a local search algorithm which starts with a random subset $S$ of $k$ points.
-In any iteration, the algorithm looks for a set $A \subseteq S$ and a set $B \subseteq X -S$ such that $|A| = |B| \leq p$ and
+At every iteration, the algorithm looks for a set $A \subseteq S$ and a set $B \subseteq X -S$ such that $|A| = |B| \leq p$ and
 \[ f(S- B \cup A) \geq f(S). \]
@@
 If no such set exists, the algorithm outputs $S$.
-For $p$ independent of $k$ and $n$, let $\mathcal{N}$ be the size of the search neighborhood on each iteration.
-If we ensure that the algorithm improves by at least a factor of $(1- \delta) = (1- \varepsilon/\mathcal{N})$ in each iteration, this is a polynomial-time algorithm.
+At every iteration, for $p$ independent of $k$ and $n$, let $\mathcal{N}$ be the size of the search neighborhood.
+If we ensure that the algorithm improves the objective by at least a factor of $(1- \delta) = (1- \varepsilon/\mathcal{N})$, this is a polynomial-time algorithm.
 Furthermore, the local search algorithm outputs $S \subseteq X$ such that $|S| = k$ and
-\[ f(S) \leq \frac{1}{1-\varepsilon} \left (3 + \frac{2}{p}) \right ) f(O) \]
+\[ f(S) \leq \frac{1}{1-\varepsilon} \left (3 + \frac{2}{p} \right ) f(O) \]
 for all $O \subseteq X$. Therefore, there is a trade-off between the computation time of this algorithm and the guaranteed quality of the solution.
@@
 \subsection{Supermodular Function Minimization}
-Given a set $X$, a function $f:2^{X} \mapsto \mathbb{R}$ is supermodular, if for all $T \subseteq S \subseteq X$ and $j \notin S$, 
-\[ f(T \cup \{j\}) - f(T) \leq f(S \cup \{j\}) - f(S). \]
-Furthermore, $f$ is non-increasing if for all $T \subseteq S \subseteq X$, $f(S) \leq f(T)$.
+Center-based clustering in general can be viewed as minimizing a supermodular decreasing function subject to a cardinality constraint, a problem related to maximizing a submodular function that is well-studied in approximation algorithms.
+The approximation algorithms in this area use local-search or greedy methods and are therefore computationally tractable on large data.
+In addition, using supermodular functions means the distance measure does not have to satisfy the triangle inequality, which can be useful when considering non-vector data.
 
-Suppose we have a set of data $X$ of which we want to choose a subset $S$ such that $|S| \leq k$ so as to minimize 
-\[ f(S) = \sum_{i \in X} \min_{j \in S} d(i,j) ,\]
-where we can think of $d$ as some measure of dissimilarity (for example squared distance or absolute distance). For completeness, we define $f(\emptyset) = \max_{j \in S} f(\{j\})$. 
-Note that this generalizes the $k$-median problem and can capture a variant of $k$-means in which we are restricted to choosing data points as centers.
+Given a set $X$, a function $g:2^{X} \mapsto \mathbb{R}$ is supermodular, if for all $T \subseteq S \subseteq X$ and $j \notin S$, 
+\[ g(T \cup \{j\}) - g(T) \leq g(S \cup \{j\}) - g(S). \]
+Furthermore, $g$ is non-increasing if for all $T \subseteq S \subseteq X$, $g(S) \leq g(T)$.
 
-It is clear that $f$ is non-increasing since for all $T \subseteq S \subseteq X$ and $i \in X$, $\min_{j \in S} d(i,j) \leq \min_{j \in T} d(i,j)$.
-Furthermore, for $j \notin S$, adding $j$ to $T$ will decrease the sum of the distances more than adding $j$ to $S$ so $f$ is supermodular.
+Now considering the $k$-median objective defined above, for completeness we define $f(\emptyset) = \max_{j \in X} f(\{j\})$.
+It is then elementary to check that $f$ is supermodular and non-increasing.
 In general, there is no known approximation algorithm for minimizing a non-increasing supermodular function subject to a cardinality constraint.
-However, if we can place some properties on the ``steepnesss'' of $f$ then there is a known optimal approximation algorithm \cite{Sviridenko}.
+However, if we incorporate the ``steepness'' of $f$, there is a known optimal approximation algorithm \cite{Sviridenko}.
 
@@
 \[ f(S) \leq \left( 1 + \frac{c}{1-c} \cdot e^{-1} + \frac{1}{1-c} \cdot O(\varepsilon) \right) f(O) \]
-for all $O \subseteq X$ such that $|O| \leq k$, with high probability \cite{Sviridenko}.
-Note that $c$ will vary depending on our dataset so while in some cases we may have a very good approximation ratio it can worsen as our distances become non-smooth (for example distances that do not form a metric).
+for all $O \subseteq X$ such that $|O| \leq k$, with high probability \cite{Sviridenko}. Note that $c$ will vary depending on our dataset so the approximation ratio will vary.
 
@@
 We first draw a value $p \in [0,1]$ using the density function $\frac{{e}^p}{e-1}$.
-Then, we choose a subset $B \in A$ by selecting each $x \in A$ to be in $B$ independently with probability $p$.
+Then, we choose $B \subseteq A$ by selecting each $x \in A$ to be in $B$ independently with probability $p$.
 The expected value of $g(B)$ is given by
 \[  h(A) = \sum_{B \subseteq A} g(B) \cdot \int_{0}^1 \frac{e^p}{e-1} \cdot p^{|B|-1} (1-p)^{|A|-|B|} dp. \]
-Then, we define a potential function
+We then define the potential function
 \[ \psi(A) = (1-e^{-1}) h(A) + l(A). \]
 
-We are now ready to describe how the algorithm works.
-The algorithm sets $\delta = \frac{\varepsilon}{n}$ to be the minimum step size and starts with a set $S$ of size $n-k$.
-Then, while there exists $a \in S$ and $b \in X \setminus S$ such that
+Now, the algorithm sets $\delta = \frac{\varepsilon}{n}$ to be the minimum step size and starts with a set $S$ of size $n-k$. Then, while there exists $a \in S$ and $b \in X \setminus S$ such that 
 \[ \psi(S -a +b) \geq \psi(S) + \delta, \]
@@
 Note that $\psi$ requires summing an exponential number of terms.
-However, we can estimate $h(A)$ (and therefore $\psi(A)$) by randomly sampling $g(B)$ in the above fashion.
+However, we can estimate $h(A)$ (and therefore $\psi(A)$) by randomly sampling $g(B)$.
 If $\hat{\psi}(A)$ is an estimate of $\psi(A)$ sampled from $\Omega (\varepsilon^{-2} n^4 \ln^4 n \ln M)$ samples, then
 \[ Pr[ | \hat{\psi}(A) - \psi(A) | \geq \delta ] = O(M^{-1}). \]
-By setting $M$ large enough (polynomially large), the algorithm will converge in polynomial time and will return a solution satisying the above guarantee with high probability \cite{Sviridenko}.
+By setting $M$ large enough (polynomially large), the algorithm will converge in polynomial time and will return a solution satisying the above approximation ratio with high probability \cite{Sviridenko}.
 
@@
 
-We implemented the local search swapping algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compare against $k$-means, $k$-medoids and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
-We also implemented the $k$-medoids algorithm with a maximum iteration of 1000.
-Note that for $p=1$, the multiswaps algorithm is equivalent to the well-known PAM algorithm.
+We implemented the local search multiswap algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compared against $k$-means, $k$-medoids (with a maximum iteration count of 1000) and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
+Note that for $p=1$, the multiswap algorithm is equivalent to the well-known PAM algorithm.
 In order to calculate approximation ratios, Gurobi was used to compute the global optimum for each data set.
 
-To achieve reasonable computation times for the multiswaps and supermodular function minimization algorithms, at each iteration we first consider nearby swaps before sampling for other potential swaps.
-For the multiswaps algorithm, we sample $O(p\cdot n)$ potential swaps on each iteration, and for the supermodular function, sample $min(100,n^2)$ swaps where each potential function is calculated with by sampling 50 subsets.
+To achieve reasonable computation times for the multiswap and supermodular function minimization algorithms, at each iteration we first consider nearby swaps before sampling for other potential swaps.
+For the multiswaps algorithm, we sample $O(p\cdot n)$ potential swaps on each iteration, and for the supermodular function, sample $min(100,n^2)$ swaps where each potential function is estimated by sampling 50 subsets.
 We also ensure that some minimum level progress is achieved for both algorithms by setting $\varepsilon = 0.1$ for the multiswaps algorithm and $\varepsilon = 10$ for the submodular function.
@@
 This is motivated by the fact that, similar to Lloyd's algorithm, $k$-medoids will quickly converge to local minima.
-Since $p$-swaps avoids getting stuck in local minima, the hybrid algorithm should in theory achieve good results with minimal extra computation time over the regular local search algorithm.
+Since $p$-swaps avoids getting stuck in local minima, the hybrid algorithm should in theory achieve good results with minimal extra computation time over the multiswap algorithm.
 
-For each data set and each clustering method, we report objective function values, the approximation ratio, computation time, and adjusted mutual information (AMI).
+For each data set and each clustering method, we report objective function values, the approximation ratio, computation time, and where applicable, adjusted mutual information (AMI).
 AMI is a measure which scores accuracy of predicted clusters compared to a true labeling by measuring the agreement of two assignments.
@@
 In particular, Gauss\textunderscore d \textunderscore k \textunderscore e was generated by first choosing k random centers in $[-1,1]^d$.
-Around each center, we then generated $1000/k \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
+Around each center, we then generated $(1000/k) \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
 \item[OR $p$-median Library] From a collection of test data sets created by J. E. Beasley with data sets of size $100, 200, \ldots, 900$ and varying values of $k$ \cite{Beasley}. 
@@
 In terms of computation time, our proposed algorithms were much slower than $k$-means, the spectral algorithm, or $k$-medoids.
-In general, the multiswaps algorithm was around two orders of magnitude slower with a small multiplicative increase as $p$ increased, and the supermodular algorithm was around four orders of magnitude slower.
-This was probably due to the extra sampling the algorithm had to do.
+In general, the multiswaps algorithm was around two orders of magnitude slower with a small multiplicative increase as $p$ increased, and the supermodular algorithm was around four orders of magnitude slower, likely due to the extra sampling the supermodular algorithm must perform.
 
-In addition, the supermodular function did not perform as well in terms of AMI or approximation ratio compared to the other methods.
-Calculating the curvature, we saw that every  test set's curvature was almost equal to 1 so the guarantee of the algorithm was quite poor.
-Given that, the algorithm did maintain an approximation ratio below 2 for all instances and an AMI measure above 0.5 for almost all instances even averaging around 0.85 for the 7 dimensional Gaussian data.
+In addition, the supermodular function algorithm did not perform as well in terms of AMI or approximation ratio compared to the other methods.
+Calculating the curvature, we saw that every  test set's curvature was very close to 1, so the guarantee of the algorithm was quite poor.
+Despite this, the algorithm did maintain an approximation ratio below 2 for all instances and an AMI measure above 0.5 for almost all instances, averaging around 0.85 for the 7 dimensional Gaussian data.
 
-When increasing $p$ for the multiswaps algorithm, we did see a slight increase in time but the times remained on the same order of magnitude.
-The increase in $p$ did not seem to have any significant effect on the AMI measure or approximation ratio with the values being roughly equivalent.
-This may due to the fact the algorithm for $p=1$ performed quite well already with an approximation ratio very close to 1 that beat out $k$-medoids.
+When increasing $p$ for the multiswaps algorithm, we did observe slight increases in computation time, but the times remained on the same order of magnitude.
+The increase in $p$ did not seem to have any significant effect on the AMI measure nor approximation ratio.
+This may be due to the fact the algorithm for $p=1$ performed quite well already with an approximation ratio very close to 1 that beat out $k$-medoids.
 In addition, PAM had the highest AMI measures for the iris and soybean data sets.

### 2015-05-10 tex/project.tex

@@
 The algorithm repeats this until the solution becomes stable.
-Additionally, Velmurugan and Santhanam showed that on uniform and normally distributed data, $k$-medoids was empirically faster than $k$-means \cite{Velmurugan}.
 
@@
 \begin{description}
-\item[Gaussian] Synthetic $k$-center Gaussian data, generated in Python.
-In particular, Gauss\textunderscore d \textunderscore k \textunderscore e was generated by first choosing k random centers in $[-1,1]^d$.
-Around each center, we then generated $(1000/k) \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
-\item[OR $p$-median Library] From a collection of test data sets created by J. E. Beasley with data sets of size $100, 200, \ldots, 900$ and varying values of $k$ \cite{Beasley}. 
+\item[Gaussian] 27 synthetic $k$-center Gaussian data sets, generated in Python.
+In particular, Gauss\textunderscore d\textunderscore k\textunderscore e was generated by first choosing k random centers in $[-1,1]^d$.
+Around each center, we then generated $(1000/k) \cdot \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
+\item[OR $p$-median Library] From a collection of test data sets created by \cite{Beasley} with 40 test sets of size $100, 200, \ldots, 900$ and varying values of $k$. 
 \item[UCI Iris] \cite{Iris} Contains 4 attributes on 150 instances of flowers, each of with belongs to one of three species of iris.
@@
 In terms of computation time, our proposed algorithms were much slower than $k$-means, the spectral algorithm, or $k$-medoids.
-In general, the multiswaps algorithm was around two orders of magnitude slower with a small multiplicative increase as $p$ increased, and the supermodular algorithm was around four orders of magnitude slower, likely due to the extra sampling the supermodular algorithm must perform.
+In general, the multiswaps algorithm was around two orders of magnitude slower with a small multiplicative increase as $p$ increased, and the supermodular algorithm was around four orders of magnitude slower, likely due to the extra potential function sampling the supermodular algorithm must perform.
 
@@
 The increase in $p$ did not seem to have any significant effect on the AMI measure nor approximation ratio.
-This may be due to the fact the algorithm for $p=1$ performed quite well already with an approximation ratio very close to 1 that beat out $k$-medoids.
+This may be due to the fact the algorithm for $p=1$ performed quite well already with an approximation ratio very close to 1.
 In addition, PAM had the highest AMI measures for the iris and soybean data sets.
 The AMI measures for the Gaussian data set were higher than those of $k$-medoids in general but equivalent or only slightly weaker than those of the spectral method or $k$-means.
-Adding in a few rounds of $k$-medoids at the end of our multiswaps algorithm did not seem to have much of an effect on the AMI measure but did boost the approximation ratio, sometimes finding an optimal solution.
+The heuristic method of adding in a few rounds of $k$-medoids at the end of our multiswaps algorithm did not have much effect on the AMI measure but did boost the approximation ratio, sometimes finding an optimal solution.
 
-It is interesting to note that higher dimensional Gaussian data seemed easier for the algorithms to classify along with smaller values of $k$ and lower percentage of noisy data, which makes sense based on our intuition of clustering.
+It is interesting to note that higher dimensional Gaussian data, with smaller values of $k$ and a lower percentage of noisy data, was easier for the algorithms to classify, which makes sense based on our intuition of clustering.
 For the OR library instances, we saw that the approximation ratio tended to increase as $k$ grew larger but not as $n$ grew.
 
-Overall, the results of our experiments do not point to the fact that these proposed approximation algorithms can add any advantage over current methods.
-However, the results do point to the fact that we can use sampling to run PAM with lower computational cost while still maintaining good performance accuracy both in terms of AMI measure and approximation ratios.
-Adding in a few rounds of $k$-medoids seems also to be a good method to refine a given solution towards a local optimum.
+Overall, the results of our experiments do not suggest that the proposed approximation algorithms provide any advantage over current well known methods.
+However, the results do suggest that we can use sampling to run PAM with lower computational cost while still maintaining good performance accuracy in terms of AMI and approximation ratio.
 
 \section{Hierarchical Clustering Approximation Algorithms}
-Suppose we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$.
-In the hierarchical setting, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
+Turning now to the hierarchical clustering setting, given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
 Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$.
-We define
-\[ \mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i)) . \]
+We define $\mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i))$.
 
-Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if 
+Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if the following conditions hold.
 \begin{enumerate}
 	\item $S_1 \subset S_2$, 
-\item  $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, and 
+	\item $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, 
 	\item $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
@@
  A solution to the hierarchical clustering problem is therefore a chain of nested solutions.
-However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the algorithm.
-Consider any $k$-median algorithm $\mathcal{A}$, such as the algorithms above.
-After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce  augmentation step that will nest these clusterings so that they fit into the hierarchical framework and only increase the costs by a bounded factor.
+However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
+Consider any $k$-median algorithm $\mathcal{A}$.
+After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce an augmentation step that will nest these clusterings into the hierarchical framework and only increase costs by a bounded factor.
 
-First, we set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
+First, set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
 Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
@@
 From each non-empty bucket, pick the solution with the smallest number of cluster centers and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r = V_n$.
-We then inductively build up our nested solutions in the following manner.
-Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_{i+1}$ by iteratively adding the closest point $j \in S_{i-1}$ to $i$ for all $i \in \overline{V}_i$.
+We then inductively build our nested solutions.
+Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_i$ by iteratively adding the closest point $j \in S_{i-1}$ to $i$ for all $i \in \overline{V}_i$.
 Note that $|S_{i}| \leq |\overline{V}_i|$.
@@
 
-We continue this nesting until $|S_i| = 1$ and we have reached the top level of the hierarchy.
+We continue this nesting until $|S_i| = 1$, reaching the top level of the hierarchy.
 After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantee that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq n-k+1 \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).
 
-We test this hierarchical framework for both the multiswap and supermodular local search algorithms above.
-
 \subsection{Implementation and Results}
-We implemented the nesting hierarchical framework above in Python offering options for both the randomized and deterministic versions.
-As a black-box $k$-median solver, we used $k$-medoids and the 1-swap+ algorithms above.
+We implemented the nesting hierarchical framework above in Python, offering options for both the randomized and deterministic versions.
+As a black-box $k$-median solver, we used $k$-medoids and the 1-swap+ algorithms, discussed in the partitional clustering section above.
 We compared the results to hierarchical agglomerative clustering with both average linkage and Ward's linkage (implemented in the scipy package in Python).
 
-We also tested a more computationally friendly heuristic of the hierarchical framework which just buckets solutions for values of $k$ that are a power of two.
-This yields many fewer $k$-median instances to solve and reduced the run-time significantly.
-This was motivated by the fact that with the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
+We also tested a more computationally friendly heuristic of the hierarchical framework which only buckets solutions for values of $k$ that are a power of two.
+This yields many fewer $k$-median instances to solve, reducing the run-time significantly.
+This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
 
-For each data set and each clustering method, we report the approximation ratio, computation time, and FScore measure.
+For each data set discussed previously, and each clustering method, we report the approximation ratio, computation time, and FScore measure.
 FScore is a measure that looks at the clusters represented on each level and how they represent the true clustering in terms of precision and recall.
 See \cite{Larsen} for more information.
+Results are in Figure 3.
 
-In general, the hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure even for our logarithmic version.
-The FScore measures were very low for the agglomerative clustering methods, often around 0.25, but shot up to closer to the AMI measures we saw in flat clustering for our implemented algorithms.
+\begin{figure*}[h]
+\label{fig:hier_results}
+\small
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+	   & Time & Time & Time & Time & Time & Time \\ \hline
+	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 0.0002 & 0.0002 & 0.2 & 0 & 0.2 & 0.2 \\ 
+	Iris & 0.0023 & 0.003 & 3.2 & 0.2 & 3.8 & 5.1 \\ 
+	Gauss\_5\_5\_5 &  & & & & & \\
+	Gauss\_5\_10\_5 & & & & & & \\
+	%Gauss\_5\_50\_5 &  & & & & & \\
+	Gaussian Average & 0.0858 & 0.0853 & 354.5 & 7.5 & 369.9 & 423.3 \\ 
+	pmed5 & 1.2 & 0.1 & 1.3 & 1.7 & \  & \  \\ 
+	pmed25 & 279.3 & 5 & 306.7 & 312.5 & \  & \  \\ 
+	pmedian avg & 87.4 & 2 & 94.5 & 101.6 & \  & \  \\ \hline
+\end{tabular}
+\vspace{0.2 cm}
 
-For $k$-medoids in the non-logarithmic version and PAM in the logarithmic version, the hierarchical framework increased the computation time by three orders of magnitude, whereas this was decreased to two orders difference for $k$-medoids logarithmic version.
-Interestingly, the approximation factors did not seem to change between the randomized and deterministic versions for $k$-medoids.
-The approximation ratios seemed best as $n$ increased or as the dimension increased.
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+	   & F-Score & F-Score & F-Score & F-Score & F-Score & F-Score \\ \hline
+	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 0.6 & 0.5 & 0.8 & 0.6 & 0.7 & 0.6 \\ 
+	Iris & 0.5 & 0.5 & 0.9 & 0.8 & 0.8 & 0.7 \\ 
+	Gauss\_5\_5\_5 &  & & & & & \\
+	Gauss\_5\_10\_5 &  & & & & & \\
+	%Gauss\_5\_50\_5 &  & & & & & \\
+	Gaussian Average & 0.4 & 0.4 & 0.6 & 0.6 & 0.6 & 0.6 \\ \hline
+\end{tabular}
+\vspace{0.2 cm}
+
+\begin{tabular}{ | l | l | l | l | l | }
+\hline
+	   & Approx & Approx & Approx & Approx \\ \hline
+	Dataset & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 1.6 & 2.3 & 1.7 & 2 \\ 
+	Iris & 2560631.5 & 2550279.6 & 2787312 & 1623132.1 \\ 
+	Gauss\_5\_5\_5 &  & & &  \\
+	Gauss\_5\_10\_5 &  & & &  \\
+	%Gauss\_5\_50\_5 &  & & &  \\
+	Gaussian Average & 5.2 & 7.3 & 5.5 & 4.7 \\ 
+	pmed5 & 11.4 & 51.1 & 16.3 & 12 \\ 
+	pmed25 & 6.6 & 13.4 & 5.4 & 13.2 \\ 
+	pmedian avg & 10.3 & 19.8 & 10.5 & 8.6 \\ \hline
+\end{tabular}
+
+\caption{Average Computation Time (sec), FScore measure, and Approximation Coeffiecients for Hierarchical Clustering Algorithms}
+\end{figure*}
+
+In general, the hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure, even for our logarithmic version.
+The FScore measures were very low for the agglomerative clustering methods, often around 0.25, but were much closer to the AMI measures we saw in flat clustering for our implemented algorithms.
+
+For $k$-medoids in the non-logarithmic version and PAM in the logarithmic version, the hierarchical framework increased the computation time by three orders of magnitude, and increased by two orders of magnitude for $k$-medoids in the logarithmic version.
+Interestingly, the approximation factors did not change significantly between the randomized and deterministic versions for $k$-medoids.
+The approximation ratios were strongest as $n$ increased or as the dimension increased.
 In particular, for the Gaussian data in 7 dimensions, the approximation ratio was often around 3.
-Recall that this approximation measures the constructed $k$-median solution on every level against an optimal one so we expect the ratio to not come close to 1 in many cases.
+The high approximation ratio for the iris data set was due to the fact that the optimal solution for $k = n-1, n-2$ was very close to zero, and our heuristic $k$-median algorithm only sometimes picked up on these solutions.
+The approximation factor at lower levels was more consistent with the other results.
 
 \section{Conclusion and Future Work}
-For partitional clustering, it seems that the approximation algorithms in the literature are in general slower and cannot improve much on the current heuristics used despite their approximation guarantees.
-This was not true in the hierarchical clustering case where we saw significant improvement in the FScore measures for our implemented methods and overall good approximation ratios where none were had before.
-In particular, our logarithmic nesting algorithm seems to provide the best balance between performance and computation.
+For partitional clustering, we found that the approximation algorithms in the literature are in general slower and cannot improve much on current heuristics used despite their approximation guarantees.
+This was not true in the hierarchical clustering case where we saw significant improvement in FScore measures for our implemented methods and overall good approximation ratios where none were had before.
+In particular, our logarithmic nesting algorithm seems to provide the best balance between performance and computational efficiency.
 
-In the future, we hope to explore how linear programming based methods perform in this context as well and to implement our own agglomerative clustering so we can test it on the OR library instances.
-Lastly, when nesting solutions in the logarithmic framework, we could use agglomerative clustering to make smarter local merges.
+In the future, we hope to explore how linear programming based methods perform in both the flat and hierarchical context as well and to implement our own agglomerative clustering to test on the OR library instances and help make smarter local merges in the nesting framework.
 
-\section{Acknowledgements}
-We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
+%\section{Acknowledgements}
+%We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
 

### 2015-05-10 tex/project.tex
updated tables

--- 2015-05-10 tex/project.tex
@@
 	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
-	Soybean & 0.0002 & 0.0002 & 0.2 & 0 & 0.2 & 0.2 \\ 
-	Iris & 0.0023 & 0.003 & 3.2 & 0.2 & 3.8 & 5.1 \\ 
-	Gauss\_5\_5\_5 &  & & & & & \\
-	Gauss\_5\_10\_5 & & & & & & \\
+	Soybean & 0.0002 & 0.0002 & 0.286 & 0 & 0.294 & 0.167 \\ 
+	Iris & 0.002 & 0.003 & 3.161 & 0.226 & 3.794 & 5.122 \\ 
+	Gauss\_5\_5\_5 & 0.038 & 0.060 & 106.619 & 2.735 &120.893 & 147.137 \\
+	Gauss\_5\_10\_5 & 0.099 & 0.096 & 383.854 & 10.259 & 383.298 & 494.554 \\
 	%Gauss\_5\_50\_5 &  & & & & & \\
-	Gaussian Average & 0.0858 & 0.0853 & 354.5 & 7.5 & 369.9 & 423.3 \\ 
-	pmed5 & 1.2 & 0.1 & 1.3 & 1.7 & \  & \  \\ 
-	pmed25 & 279.3 & 5 & 306.7 & 312.5 & \  & \  \\ 
-	pmedian avg & 87.4 & 2 & 94.5 & 101.6 & \  & \  \\ \hline
+	Gaussian Average & 0.0858 & 0.0853 & 354.459 & 7.492 & 369.967 & 423.293 \\ 
+	pmed5 &  N/A & N/A & 1.184 & 0.102 & 1.288 & 1.673   \\ 
+	pmed25 & N/A & N/A & 279.334 & 5.011 & 306.684 & 312.503   \\ 
+	pmedian avg & N/A & N/A &  87.355 & 1.957 & 94.548 & 101.576   \\ \hline
 \end{tabular}
@@
 	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
-	Soybean & 0.6 & 0.5 & 0.8 & 0.6 & 0.7 & 0.6 \\ 
-	Iris & 0.5 & 0.5 & 0.9 & 0.8 & 0.8 & 0.7 \\ 
-	Gauss\_5\_5\_5 &  & & & & & \\
-	Gauss\_5\_10\_5 &  & & & & & \\
+	Soybean & 0.615 & 0.510 & 0.772 & 0.643 & 0.742 & 0.633 \\ 
+	Iris & 0.5 & 0.5 & 0.888 & 0.767 & 0.762 & 0.732 \\ 
+	Gauss\_5\_5\_5 & 0.519 & 0.516 & 0.591 & 0.816 & 0.761 & 0.785 \\
+	Gauss\_5\_10\_5 & 0.298 & 0.289 & 0.589 & 0.584 & 0.576 & 0.446 \\
 	%Gauss\_5\_50\_5 &  & & & & & \\
-	Gaussian Average & 0.4 & 0.4 & 0.6 & 0.6 & 0.6 & 0.6 \\ \hline
+	Gaussian Average & 0.393 & 0.375 & 0.649 & 0.616 & 0.624 & 0.624 \\ \hline
 \end{tabular}
@@
 	Dataset & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
-	Soybean & 1.6 & 2.3 & 1.7 & 2 \\ 
-	Iris & 2560631.5 & 2550279.6 & 2787312 & 1623132.1 \\ 
-	Gauss\_5\_5\_5 &  & & &  \\
-	Gauss\_5\_10\_5 &  & & &  \\
+	Soybean & 1.585 & 2.334 & 1.711 & 2.001 \\ 
+	Iris & 2560631.459 & 2550279.566 & 2787312.009 & 1623132.096 \\ 
+	Gauss\_5\_5\_5 & 6.459 & 13.394 & 8.624 & 7.791 \\
+	Gauss\_5\_10\_5 &  3.420 & 3.756 & 4.196 & 2.309  \\
 	%Gauss\_5\_50\_5 &  & & &  \\
-	Gaussian Average & 5.2 & 7.3 & 5.5 & 4.7 \\ 
-	pmed5 & 11.4 & 51.1 & 16.3 & 12 \\ 
-	pmed25 & 6.6 & 13.4 & 5.4 & 13.2 \\ 
-	pmedian avg & 10.3 & 19.8 & 10.5 & 8.6 \\ \hline
+	Gaussian Average & 5.182 & 7.262  & 5.537 & 4.739 \\ 
+	pmed5 & 11.367 & 51.141 & 16.261 & 11.950 \\ 
+	pmed25 & 6.641 & 13.372 & 5.378 & 13.200 \\ 
+	pmedian avg & 10.330 & 19.822 & 10.524 & 8.599 \\ \hline
 \end{tabular}

### 2015-05-10 tex/project.tex
just some typo from daves edit

@@
 Often these problems are solved using heuristic algorithms which have no guarantee on the solution produced but are typically fast and easy to implement.
-We will look at how local search approximation algorithms, which do have some approximation guarantee on the constructed solution, and are less computationally expensive than other approximation methods, compare in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
+We will look at how local search approximation algorithms, which do have some approximation guarantee on the constructed solution and are less computationally expensive than other approximation methods, compare in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
 
@@
 
-Arguably the most common partitional clustering method is $k$-means, in which we want to choose $k$ cluster centers to minimize the sum of squared distances from each point to its cluster center \cite{Jain}.
+Arguably the most common partitional clustering method is $k$-means, in which we want to choose $k$ cluster centers to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
 $k$-means is usually solved using Lloyd's algorithm (also referred to simply as the $k$-means algorithm), which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}.
-Lloyd's algorithm is only guaranteed to reach a local optimum and can be very sensitive to outliers and the initial points chosen \cite{Kanungo}.
+Lloyd's algorithm is only guaranteed to reach a locally optimum solution and can be very sensitive to outliers and the initial points chosen \cite{Kanungo}.
 
@@
 
-Various approximation algorithms for he $k$-median problem have been proposed, with a best known approximation ratio of 2.592 \cite{Wu}.
+Various approximation algorithms for the $k$-median problem have been proposed, with a best known approximation ratio of 2.592 \cite{Wu}.
 However, this algorithm uses linear programming and is computationally intensive compared to local search and greedy algorithms.
 It is known that a local search algorithm that considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
-Note that for $p=1$, this equivalent to PAM.
+Note that for $p=1$, this is equivalent to PAM.
 As far as we know, this algorithm has not been implemented or tested on clustering data, so the relative trade-off of solution quality and computation time as $p$ increases is unknown.
@@
 There are several choices on how to measure distance in order to choose which clusters to merge.
-Single linkage (also called nearest neighbor) measures the distance between two clusters to be the minimum pairwise distance between the data points in each cluster.
+Single linkage (also called nearest neighbor) defines the distance between two clusters to be the minimum pairwise distance between the data points in each cluster.
 Complete linkage clustering instead defines the distance to be the largest pairwise distance.
@@
 
-Approximation algorithms for hierarchical clustering have also been studied, where we aim to identify a hierarchy of clusters so as to minimize the ratio of the objective value on every level in the hierarchy of size $k$ to the optimal objective value of the partitional clustering problem with $k$ clusters.
+Approximation algorithms for hierarchical clustering have also been studied, where we aim to identify a hierarchy of clusters so as to minimize the ratio of the objective value on every level in the hierarchy of size $k$ to the optimal objective value of the flat partitional clustering problem with $k$ clusters.
 Note that it may not always be possible to find a solution with approximation ratio 1.
@@
 	Dataset Name & Time & Time & Time & Time & Time & Time & Time & Time & Time \\ \hline
-	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap  \\ \hline
+	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap  \\ \hline
 	Soybean & 0.015 & 0.061 & 0.002 & 1.179 & 0.026 & 0.016 & 0.048 & 0.049 & 0.081 \\ 
@@
 	Dataset Name & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
-	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap   \\ \hline
+	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap   \\ \hline
 	Soybean & 0.686 & 0.178 & 0.55 & 0.612 & 0.708 & 0.732 & 0.717 & 0.708 & 0.712 \\ 
@@
 	Dataset Name & Approx & Approx & Approx & Approx & Approx & Approx & Approx  \\ \hline
-	& k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap   \\ \hline
+	& k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap   \\ \hline
 	Soybean & 1.167 & 1.199 & 1.027 & 1.012 & 1.022 & 1.012 & 1.024 \\ 
@@
 %	Dataset &Time & Time & Time & Time & Time & Time & Time & Time & Time & Time   \\ \hline
-%	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-swaps+ & 3-Swap & 4-swaps   \\ 
+%	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap & 4-swaps   \\ 
 %	Gauss\_3\_5\_0 & 0.024 & 0.056 & 0.066 & 131.305 & 2.707 & 3.195 & 4.779 & 4.138 & 7.683 & 11.325 \\ 
@@
 %	Dataset&AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
-%	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap & 4-swaps   \\ \hline
+%	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-Swaps + & 3-Swap & 4-swaps   \\ \hline
 %	Gauss\_3\_5\_0 & 0.884 & 0.887 & 0.729 & 0.678 & 0.88 & 0.876 & 0.884 & 0.876 & 0.881 & 0.882 \\ 
@@
 %	Dataset & Time & Time & Time & Time & Time & Time & Time \\ \hline
-%	&k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps + & 3-Swap   \\ \hline
+%	&k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-Swaps + & 3-Swap   \\ \hline
 %	pmed1 & 0.0032 & 5 & 0.1 & 0.1 & 0.2 & 0.2 & 0.3 \\ 
@@
 %	Dataset & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
-%	  & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-swaps+ & 3-Swap \\ \hline
+%	  & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-Swaps+ & 3-Swap \\ \hline
 %	pmed1 & 1.113 & 1.409 & 1.03 & 1.013 & 1.029 & 1.005 & 1.018 \\ 

### 2015-05-10 tex/project.tex
Last edits from me

@@
 \twocolumn[
-\icmltitle{Local Search Approximation Algorithms for Partitional and Hierarchical Clustering}
+\icmltitle{Simple Approximation Algorithms for Partitional and Hierarchical Clustering}
 
@@
 Often these problems are solved using heuristic algorithms which have no guarantee on the solution produced but are typically fast and easy to implement.
-We will look at how local search approximation algorithms, which do have some approximation guarantee on the constructed solution and are less computationally expensive than other approximation methods, compare in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
+We will look at how local search and greedy approximation algorithms, which do have some approximation guarantee on the constructed solution and are less computationally expensive than other approximation methods, compare in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
 
@@
 These algorithms construct a graph based on the similarity matrix and compute the eigenvectors of the normalized graph Laplacian matrix \cite{Luxburg}.
-These eigenvectors are used to split the current cluster in half using the second eigenvector (continuing recursively until there are $k$ clusters).
+These eigenvectors are used to split the current cluster in half using the eigenvector associated with the second highest eigenvalue (continuing recursively until there are $k$ clusters).
 Another method is to use the eigenvectors to project the data points into a lower dimensional space on which the $k$-means algorithm is run \cite{Verma}.
@@
 
-Various approximation algorithms for the $k$-median problem have been proposed, with a best known approximation ratio of 2.592 \cite{Wu}.
-However, this algorithm uses linear programming and is computationally intensive compared to local search and greedy algorithms.
+The $k$-median problem is also a fundamental problem in approximation algorithms.
+However, these algorithms are often not tested in practice despite their improved guarantees on the solution.
+The best known approximation ratio for this problem is 2.592 \cite{Wu}.
+This algorithm uses linear programming and is computationally intensive compared to local search and greedy algorithms.
 It is known that a local search algorithm that considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
@@
 This was improved by Lin et al.~\cite{Lin} who gave an incremental approximation algorithm framework.
-In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to construct a hierarchical clustering that increases the objective value by at most a factor of $20.71$ (or $10.03$ when randomized) at each level.
-The local search $(3+\frac{2}{p})$-approximation algorithm as the $k$-median solver performed well in experimental tests \cite{Nagarajan}.
+In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to greedily construct a hierarchical clustering that increases the objective value by at most a factor of $20.71$ (or $10.03$ when randomized) at each level.
+Using the local $p$-swap $(3+\frac{2}{p})$-approximation algorithm mentioned above as the black-box $k$-median solver performed well in experimental tests \cite{Nagarajan}.
 While the authors tested the variants of this algorithm against other $k$-median approximation algorithms, it has not been tested against traditional clustering algorithms or on any labeled data sets.
@@
 Consider a local search algorithm which starts with a random subset $S$ of $k$ points.
-At every iteration, the algorithm looks for a set $A \subseteq S$ and a set $B \subseteq X -S$ such that $|A| = |B| \leq p$ and
+At every iteration, the algorithm looks for a set $B \subseteq S$ and a set $A \subseteq X -S$ such that $|A| = |B| \leq p$ and
 \[ f(S- B \cup A) \geq f(S). \]
-If the algorithm finds such a set, it swaps out $B$ from $S$ and adds $A$.
+If the algorithm finds such sets, it swaps out $B$ from $S$ and adds $A$.
 If no such set exists, the algorithm outputs $S$.
@@
 By setting $M$ large enough (polynomially large), the algorithm will converge in polynomial time and will return a solution satisying the above approximation ratio with high probability \cite{Sviridenko}.
+As far as we are aware, this algorithm has not been tested for any clustering problems.
 
@@
 
-It is interesting to note that higher dimensional Gaussian data, with smaller values of $k$ and a lower percentage of noisy data, was easier for the algorithms to classify, which makes sense based on our intuition of clustering.
+It is interesting to note that, for the Gaussian data, an increase in dimension, reduction in $k$, or reduction in the percentage of noisy data made it easier for the algorithms to classify the data set, which makes sense based on our intuition of clustering.
 For the OR library instances, we saw that the approximation ratio tended to increase as $k$ grew larger but not as $n$ grew.
@@
 
-From each non-empty bucket, pick the solution with the smallest number of cluster centers and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r = V_n$.
+From each non-empty bucket, pick the solution with the smallest number of cluster centers, and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r$.
 We then inductively build our nested solutions.
-Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_i$ by iteratively adding the closest point $j \in S_{i-1}$ to $i$ for all $i \in \overline{V}_i$.
+Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_i$ by iteratively adding the closest point $j \in S_{i-1}$ to $k$ for all $k \in \overline{V}_i$.
 Note that $|S_{i}| \leq |\overline{V}_i|$.
@@
 	   & F-Score & F-Score & F-Score & F-Score & F-Score & F-Score \\ \hline
-	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Dataset & Ave Linking & Ward Linking & K-med & K-med logn rand & K-med rand & PAM logn rand \\ \hline
 	Soybean & 0.615 & 0.510 & 0.772 & 0.643 & 0.742 & 0.633 \\ 
@@
 	   & Approx & Approx & Approx & Approx \\ \hline
-	Dataset & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Dataset & K-med & K-med logn rand & K-med rand & PAM logn rand \\ \hline
 	Soybean & 1.585 & 2.334 & 1.711 & 2.001 \\ 
@@
 
-\caption{Average Computation Time (sec), FScore measure, and Approximation Coeffiecients for Hierarchical Clustering Algorithms}
+\caption{Average Computation Time (sec), FScore measure, and Approximation Coeffiecients for Hierarchical Clustering Algorithms.
+K-med and PAM represent the black-box $k$-median solvers used and rand and logn indicate whether the randomized and logarithmic setting were used, respectively.}
 \end{figure*}
@@
 
-%\section{Acknowledgements}
-%We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
+\section{Acknowledgements}
+We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
 

### 2015-05-11 tex/project.tex
Last results and computer specs in report
Added sentence to conclusion

@@
 
-Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core Intel processor. Results are given in Figure 2  below with some results averaged due to space constraints.
+Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core 3.4 Ghz Intel i7 processor with 16 GB physical memory.
+Results are given in Figure 2  below with some results averaged due to space constraints.
 
@@
 \section{Conclusion and Future Work}
+We implemented various simple approximation algorithms for partitional clustering, and a framework for hierarchical clustering.
+We compared these algorithms, along with heuristic modifications, against the most common clustering algorithms used in practice.
+
 For partitional clustering, we found that the approximation algorithms in the literature are in general slower and cannot improve much on current heuristics used despite their approximation guarantees.

### 2015-05-11 tex/project.tex
Updated project write-up

@@
 	   & Time & Time & Time & Time & Time & Time \\ \hline
-	Dataset & Ave Linking & Ward Linking & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
-	Soybean & 0.0002 & 0.0002 & 0.286 & 0 & 0.294 & 0.167 \\ 
-	Iris & 0.002 & 0.003 & 3.161 & 0.226 & 3.794 & 5.122 \\ 
-	Gauss\_5\_5\_5 & 0.038 & 0.060 & 106.619 & 2.735 &120.893 & 147.137 \\
-	Gauss\_5\_10\_5 & 0.099 & 0.096 & 383.854 & 10.259 & 383.298 & 494.554 \\
+	Dataset & Ave Linkage & Ward Linkage & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
+	Soybean & 0.0004 & 0.0002 & 0.286 & 0 & 0.294 & 0.167 \\ 
+	Iris & 0.0021 & 0.0011 & 3.161 & 0.226 & 3.794 & 5.122 \\ 
+	Gauss\_5\_5\_5 & 0.0097 & 0.0109 & 106.619 & 2.735 &120.893 & 147.137 \\
+	Gauss\_5\_10\_5 & 0.0237 & 0.0256 & 383.854 & 10.259 & 383.298 & 494.554 \\
 	%Gauss\_5\_50\_5 &  & & & & & \\
-	Gaussian Average & 0.0858 & 0.0853 & 354.459 & 7.492 & 369.967 & 423.293 \\ 
-	pmed5 &  N/A & N/A & 1.184 & 0.102 & 1.288 & 1.673   \\ 
-	pmed25 & N/A & N/A & 279.334 & 5.011 & 306.684 & 312.503   \\ 
-	pmedian avg & N/A & N/A &  87.355 & 1.957 & 94.548 & 101.576   \\ \hline
+	Gaussian Average & 0.0229 & 0.0239 & 354.459 & 7.492 & 369.967 & 423.293 \\ 
+	pmed5 &  0.0004 & N/A & 1.184 & 0.102 & 1.288 & 1.673   \\ 
+	pmed25 & 0.0174 & N/A & 279.334 & 5.011 & 306.684 & 312.503   \\ 
+	pmedian avg & 0.0067 & N/A &  87.355 & 1.957 & 94.548 & 101.576   \\ \hline
 \end{tabular}
@@
 	   & F-Score & F-Score & F-Score & F-Score & F-Score & F-Score \\ \hline
-	Dataset & Ave Linking & Ward Linking & K-med & K-med logn rand & K-med rand & PAM logn rand \\ \hline
-	Soybean & 0.615 & 0.510 & 0.772 & 0.643 & 0.742 & 0.633 \\ 
+	Dataset & Ave Linkage & Ward Linkage & K-med & K-med logn rand & K-med rand & PAM logn rand \\ \hline
+	Soybean & 0.615 & 0.554 & 0.772 & 0.643 & 0.742 & 0.633 \\ 
 	Iris & 0.5 & 0.5 & 0.888 & 0.767 & 0.762 & 0.732 \\ 
-	Gauss\_5\_5\_5 & 0.519 & 0.516 & 0.591 & 0.816 & 0.761 & 0.785 \\
-	Gauss\_5\_10\_5 & 0.298 & 0.289 & 0.589 & 0.584 & 0.576 & 0.446 \\
+	Gauss\_5\_5\_5 & 0.519 & 0.520 & 0.591 & 0.816 & 0.761 & 0.785 \\
+	Gauss\_5\_10\_5 & 0.298 & 0.281 & 0.589 & 0.584 & 0.576 & 0.446 \\
 	%Gauss\_5\_50\_5 &  & & & & & \\
-	Gaussian Average & 0.393 & 0.375 & 0.649 & 0.616 & 0.624 & 0.624 \\ \hline
+	Gaussian Average & 0.393 & 0.378 & 0.649 & 0.616 & 0.624 & 0.624 \\ \hline
 \end{tabular}
@@
 In general, the hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure, even for our logarithmic version.
-The FScore measures were very low for the agglomerative clustering methods, often around 0.25, but were much closer to the AMI measures we saw in flat clustering for our implemented algorithms.
+The FScore measures were very low for the agglomerative clustering methods, sometimes around 0.25, but were much closer to the AMI measures we saw in flat clustering for our implemented algorithms.
 
@@
 
-In the future, we hope to explore how linear programming based methods perform in both the flat and hierarchical context as well and to implement our own agglomerative clustering to test on the OR library instances and help make smarter local merges in the nesting framework.
+In the future, we hope to explore how linear programming based methods perform in both the flat and hierarchical context as well and to implement our own agglomerative clustering to help make smarter local merges in the nesting framework.
 

### 2015-05-13 data/LoadData.py
Started new hier framework
- Greedy LP alg
- Smart local merging
- Hier solns computed in respective files

@@ -1,55 +1,0 @@
-import scipy.sparse as sp
-import numpy as np
-
-# LoadData takes in a file path (string, a boolean cluster_loc
-# representing whether the cluster variable is the first (1)
-# or last entry (0) of each file line, and a boolean split
-# representing whether the file is space delimited (1) or
-# comma delimited (0)
-
-def LoadData(filepath, cluster_loc = 1, split = 1):
-    f = open(filepath)
-
-    # First find number of samples and dimension
-    n = sum([1 for line in f])
-    f.seek(0)
-
-    if split:
-        dim = len(f.readline().split())-1
-    else:
-        dim = len(f.readline().split(','))-1
-    f.seek(0)
-        
-    y = np.zeros(n)
-    M = np.zeros((n,dim))
-    clusterDict = dict()
-
-    # Iterate through each sample
-    i = 0
-    k = 0    
-    while True:
-        if split:
-            line = f.readline().split()
-        else:
-            line = f.readline().split(',')
-            
-        if len(line) < dim+1:
-            break
-        
-        M[i] = [float(line[j+cluster_loc]) for j in range(dim)]
-        cluster = line[cluster_loc-1]
-        if cluster[-1:] == "\n":
-            cluster = cluster[:-1]
-
-        if clusterDict.has_key(cluster):
-            y[i] = clusterDict[cluster]
-        else:
-            y[i] = k
-            clusterDict[cluster] = k
-            k += 1
-            
-        i += 1
-        
-    X =  sp.csr_matrix(M)
-    X = X.toarray()
-    return X, y, n, k

### 2015-05-13 data/LoadpMedian.py

@@ -1,33 +1,0 @@
-import numpy as np
-import networkx as nx
-
-def LoadpMedian(filepath):
-    f = open(filepath)
-    dims = f.readline().split()
-    n = int(dims[0])
-    m = int(dims[1])
-    k = int(dims[2])
-
-    G = nx.Graph()
-    G.add_nodes_from([i+1 for i in range(n)])
-
-    for j in range(m):
-        line = f.readline().split()
-        u = int(line[0])
-        v = int(line[1])
-        w = int(line[2])
-        G.add_edge(u, v, weight = w)
-
-    DistDict = nx.floyd_warshall(G)
-    Distances = np.zeros((n,n))
-
-    for i in range(n):
-        for j in range(n):
-            d = DistDict[i+1][j+1]
-            Distances[i][j] = d
-            Distances[j][i] = d
-
-    return Distances, n, k
-    
-
-        

### 2015-05-13 src/LP_greedy.py

@@ -1,0 +1,90 @@
+import numpy as np
+from gurobipy import *
+import LoadData
+from sklearn.metrics.pairwise import pairwise_distances
+
+def assign_points_to_clusters(medoids, distances):
+    distances_to_medoids = distances[:,medoids]
+    clusters = medoids[np.argmin(distances_to_medoids, axis=1)]
+    clusters[medoids] = medoids
+    return clusters
+
+def convert_to_int(y_vals, k, random):
+    n = len(y_vals)
+    y_val_sort = np.argsort(y_vals)[::-1]
+    if random:
+        medoids = np.zeros(shape=(k))
+        l = 0
+        num_k = 0
+        while num_k < k:
+            if (np.random.uniform(0,1) < y_vals[y_val_sort[l]]) or \
+               (n-l <= k-num_k):
+                medoids[num_k] = y_val_sort[l]
+                num_k += 1
+            l += 1
+        medoids = medoids.astype(int)
+    else:
+        medoids = y_val_sort[0:k]
+    return medoids 
+    
+
+def cluster_hier(distances, kvals, random=False):
+    model = Model("k-median")
+    n = np.shape(distances)[0]
+    y,x = {}, {}
+
+    # Add variables
+    var_type = GRB.CONTINUOUS
+    for j in range(n):
+        y[j] = model.addVar(obj=0, vtype=var_type, name="y[%s]"%j)
+        for i in range(n):
+            x[i,j] = model.addVar(obj=distances[i,j], vtype=var_type, \
+                                  name="x[%s,%s]"%(i,j))
+    model.update()
+
+    # Add assignment constraints
+    for i in range(n):
+        coef = [1 for j in range(n)]
+        var = [x[i,j] for j in range(n)]
+        model.addConstr(LinExpr(coef,var), "=", 1, name="Assign[%s]"%i)
+
+    # Add feasibility constraints   
+    for j in range(n):
+        for i in range(n):
+            model.addConstr(x[i,j], "<", y[j], name="Strong[%s,%s]"%(i,j))
+    model.setParam( 'OutputFlag', False )
+    model.update()
+
+    m = len(kvals)
+    all_costs = [0 for i in range(m)]
+    all_centers = [[0 for j in range(kvals[i])] for i in range(m)]
+    for i in range(m):
+        k = kvals[i]
+        # Update k constraint & solve
+        if i == 0:
+            coef = [1 for j in range(n)]
[…]

### 2015-05-13 src/hierarchical_approx.py

@@ -1,12 +1,13 @@
 import numpy as np
-#from sklearn import metrics
+from sklearn import metrics
 import kmedoids
+import LP_greedy
 import multiswaps
 import math
 import LoadData
 
 #from sklearn.datasets.samples_generator import make_blobs
-#from sklearn.metrics.pairwise import pairwise_distances
+from sklearn.metrics.pairwise import pairwise_distances
 
 def next_assignment(Z, k, assignment):
     n= len(assignment)
@@ -17,46 +18,28 @@ def next_assignment(Z, k, assignment):
             assignment[j] = int(c1)
     return assignment
 
-# Calls k-median algorithm
-def k_cluster_alg(k, distances, alg, warm_start = None):
-    if alg[0] == "multiswaps":
-        clusters, centers = multiswaps.cluster(distances, k, alg[1])
-    elif alg[0] == "multiswaps-plus":
-        clusters, centers = multiswaps.cluster2(distances, k, alg[1])
-    else:
-        clusters, centers = kmedoids.cluster(distances, k, 1000, warm_start)
-    distances_to_centers = distances[:,centers]
-    cost = sum( np.min(distances_to_centers, axis=1))
-              
-    return centers, cost
 
 # Finds the solutions with costs in buckets [lb, ub] with the fewest
 # cluster centers
 def bucketed_solns(distances, beta, beta_0, alg, logn = False):
     n = distances.shape[0]
-    all_centers = []
-    all_costs = []
+    if logn:
+        p = int(math.floor(math.log(n,2)))
+        kvals = [2**i for i in range(p+1)]
+    else:
+        kvals = [i for i in range(1,n)]
 
-    k = 1
-    while k < n:
-        if k == 1:
-            k_centers, k_cost = k_cluster_alg(k, distances, alg)
-        else:
-            warm_start = k_centers
-            non_centers = [i for i in range(n) if i not in k_centers]
-            warm_start = np.append(warm_start,np.random.choice(\
-                non_centers,size=k-len(k_centers), replace=False))
-            k_centers, k_cost = k_cluster_alg(k, distances, alg, warm_start)
+    if alg[0] == "multiswaps-plus":
+        all_centers, all_costs = multiswaps.cluster_hier(distances, kvals, alg[1])
+    elif alg[0] == "k-medoids":
+        all_centers, all_costs = kmedoids.cluster_hier(distances, kvals, 1000)
[…]

### 2015-05-13 src/kmedoids.py

@@ -45,6 +45,30 @@ def compute_new_medoid(cluster, distances):
     costs = cluster_distances.sum(axis=1)
     return costs.argmin(axis=0, fill_value=10e9)
 
+def cluster_hier(distances, kvals, max_iter=1000, warm_start=None, epsilon=0.1):
+    m = len(kvals)
+    n = distances.shape[0]
+    all_costs = [0 for i in range(m)]
+    all_centers = [[0 for j in range(kvals[i])] for i in range(m)]
+
+    for i in range(m):
+        k = kvals[i]
+        if k == 1:
+            k_clusters, k_centers = cluster(distances, k, max_iter)
+            k_cost = sum([distances[j][k_clusters[j]] for j in range(n)])
+        else:
+            warm_start = k_centers
+            non_centers = [l for l in range(n) if l not in k_centers]
+            warm_start = np.append(warm_start,np.random.choice(\
+                non_centers,size=k-len(k_centers), replace=False))
+            k_clusters, k_centers = cluster(distances, k, max_iter, warm_start)
+            k_cost = sum([distances[j][k_clusters[j]] for j in range(n)])
+            
+        all_costs[i] = k_cost
+        all_centers[i] = k_centers
+        
+    return all_centers, all_costs
+
 
 
 

### 2015-05-13 src/multiswaps.py

@@ -81,7 +81,7 @@ def predict(medoids, distances):
     n = distances.shape[0]
     return [ find_closest_median(i, medoids, distances) for i in range(n) ]
 
-def cluster(distances, k, p, warm_start = None, epsilon=0):
+def cluster(distances, k, p, warm_start = None, epsilon=0.1):
     n = distances.shape[0]
     p = min(k,p)
     delta = epsilon/float(binom(n-k,p)*binom(k,p))
@@ -101,7 +101,7 @@ def cluster(distances, k, p, warm_start = None, epsilon=0):
 
     return predict(medoids, distances), list(medoids)
 
-def cluster2(distances, k, p, warm_start = None, epsilon=0):
+def cluster2(distances, k, p, warm_start = None, epsilon=0.1):
     n = distances.shape[0]
     p = min(k,p)
     delta = epsilon/float(binom(n-k,p)*binom(k,p))
@@ -120,6 +120,32 @@ def cluster2(distances, k, p, warm_start = None, epsilon=0):
                                                     p, delta)
     
     return kmedoids.cluster(distances, k, max_iter=10, warm_start = list(medoids))
+
+def cluster_hier(distances, kvals, p, warm_start=None, epsilon=0.1):
+    m = len(kvals)
+    print kvals
+    n = distances.shape[0]
+    all_costs = [0 for i in range(m)]
+    all_centers = [[0 for j in range(kvals[i])] for i in range(m)]
+
+    for i in range(m):
+        k = kvals[i]
+        if k == 1:
+            k_clusters, k_centers = cluster2(distances, k, p)
+            k_cost = sum([distances[j][k_clusters[j]] for j in range(n)])
+        else:
+            warm_start = k_centers
+            non_centers = [l for l in range(n) if l not in k_centers]
+            warm_start = np.append(warm_start,np.random.choice(\
+                non_centers,size=k-len(k_centers), replace=False))
+            k_clusters, k_centers = cluster2(distances, k, p, warm_start)
+            k_cost = sum([distances[j][k_clusters[j]] for j in range(n)])
+            
+        all_costs[i] = k_cost
+        all_centers[i] = k_centers
+
+    return all_centers, all_costs
+        
 
 ##if __name__ == '__main__':
 ##    centers = [[1, 1], [-1, -1], [1, -1]]

### 2015-05-22 icdm/submission.tex
ICDM paper
Wrote first draft for ICDM

--- 2015-05-11 tex/project.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%% ICML 2014 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[conference, 10pt, final]{IEEEtran}
 
-% Use the following line _only_ if you're still using LaTeX 2.09.
-%\documentstyle[icml2014,epsf,natbib]{article}
-% If you rely on Latex2e packages, like most moden people use this:
-\documentclass{article}
-
-% use Times
-\usepackage{times}
-% For figures
-\usepackage{graphicx} % more modern
-%\usepackage{epsfig} % less modern
-\usepackage{subfigure}
-
-% For citations
-\usepackage{natbib}
-
-% For algorithms
-\usepackage{algorithm}
+\usepackage{cite}
+\usepackage[cmex10]{amsmath}
+\usepackage{amssymb}
+\usepackage[pdftex]{graphicx}
 \usepackage{algorithmic}
-
-\usepackage{amssymb,amsthm,amsmath}
-
-% As of 2011, we use the hyperref package to produce hyperlinks in the
-% resulting PDF.  If this breaks your system, please commend out the
-% following usepackage line and replace \usepackage{icml2014} with
-% \usepackage[nohyperref]{icml2014} above.
-\usepackage{hyperref}
-
-% Packages hyperref and algorithmic misbehave sometimes.  We can fix
-% this with the following command.
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-% Employ the following version of the ``usepackage'' statement for
-% submitting the draft version of the paper for review.  This will set
-% the note in the first column to ``Under review.  Do not distribute.''
-\usepackage[accepted]{icml2014} 
-% Employ this version of the ``usepackage'' statement after the paper has
-% been accepted, when creating the final version.  This will set the
-% note in the first column to ``Proceedings of the...''
-%\usepackage[accepted]{icml2014}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-%\icmltitlerunning{Project Proposal for CS6780}
+\usepackage{array}
+\usepackage[caption=false,font=footnotesize]{subfig}
+\usepackage{stfloats}
+\usepackage{url}
 
@@
 
-\twocolumn[
-\icmltitle{Simple Approximation Algorithms for Partitional and Hierarchical Clustering}
+\title{Improved Hierarchical Clustering Through Nesting}
 
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2014
-% package.
-\icmlauthor{Alice Paul (ajp336), Calvin Wylie (cjw278), David Lingenbrink (dal299)}{}
+\author{\IEEEauthorblockN{Alice Paul, Calvin Wylie, and David Lingenbrink}
+\IEEEauthorblockA{Operations Research and Information Engineering, Cornell University, Ithaca, NY.  \\
+Email: \{ ajp336, cjw278, dal299 \} @cornell.edu}}
 
-% You may provide any keywords that you 
-% find helpful for describing your paper; these are used to populate 
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{boring formatting information, machine learning, ICML}
+\maketitle
 
-\vskip 0.3in
-]
+\begin{abstract}
+Partitional clustering algorithms can organize large data into smaller, more informative groups.
+Hierarchical clustering, in addition, offers a more global view of the data structure by producing a hierarchy of clusters starting from individual data points and merging towards the cluster of all data.
+Typically, hierarchical clustering is done by greedily either merging or dividing clusters, called agglomerative and divisive clustering, respectively.
+Lin et al.~\cite{Lin} proposed a nesting framework that produces a hierarchy from partitional clusterings for every possible number of clusters.
+This algorithm was introduced as an approximation algorithm for the hierarchical $k$-median problem and guarantees the resulting hierarchical layers are not too far away from the original clusters.
+However, this algorithm has not been previously placed in a clustering context.
+We test this algorithm with various methods for producing the inputted partitional clusterings versus the traditional hierarchical clustering methods and show both the approximation ratio and FScore measure improve significantly.
+We also propose a heuristic for this nesting framework which improves the running time and only suffers a slight decrease in approximation and FScore.
+\end{abstract}
 
-\section{Introduction}
+\section{Introduction and Notation}
 As the data we have access to grows in variety and size and becomes cheaper to store, we need methods to efficiently learn from this data. 
@@
 There are two main branches of clustering: partitional (or flat) and hierarchical.
-In partitional clustering we want to partition the data into disjoint sets whereas in hierarchical clustering we are interested in finding a nested tree of partitions.
-Often these problems are solved using heuristic algorithms which have no guarantee on the solution produced but are typically fast and easy to implement.
-We will look at how local search and greedy approximation algorithms, which do have some approximation guarantee on the constructed solution and are less computationally expensive than other approximation methods, compare in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
+In partitional clustering, we often specify a number $k$ and pairwise distances between the data points, representing a measure of dissimilarity such as Euclidean distance.
+The clustering algorithm then partitions the data into $k$ disjoint sets where the sets represent relatively similar data.
+In hierarchical clustering, on the other hand, we are interested in finding a nested tree of partitions.
+This offers the user a global view of the data and allows the user to choose a ``natural'' level of clustering \cite{ESL}.
+For example, a retailer could use a hierarchical clustering of customers to first find regions of customers in order to tailor the products offered in each region.
+The retailer could then also offer specialized discounts to subgroups in each region based on each cluster's shopping records.
+
+Clustering algorithms date back more than 50 years, and there have been many variations introduced since \cite{Jain}.
+However, clustering algorithms are often formulated as NP-hard problems with non-convex objective functions, which are inherently hard to solve \cite{Jain}.
+One common objective function for partitional clustering is to minimize the sum of distances from each point to the median point of its cluster, called $k$-median or $k$-medoids clustering.
+More specifically, suppose we are given a set $X$ and a distance (dissimilarity) metric $d(i,j)$ for all $i,j \in X$.
+The goal is to find a set of centers $S \subseteq X$ such that $|S| \leq k$ so as to minimize
+\[ f(S) = \sum_{ i\in X} \min_{j \in S} d(i,j). \]
+The clustering is obtained from the set of centers $S$ by assigning each data point to its closest center.
+An $\alpha$-approximation algorithm for this problem produces a set $S$ such that $f(S) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
+
+[ADD FIGURES SHOWING BOTH TYPES OF CLUSTERING]
+
+This objective function can then be extended to the hierarchical $k$-median problem which represents an objective function relating to the hierarchical clustering problem.
+Suppose we represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
+Then, for any $i \in S$, $a^{-1}(i)$ is the cluster centered at $i$.
+We define $\mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i))$.
+This is the same objective function as above except the assignment is given rather than implied.
+
+Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ (denoted $(S_1, a_1) \subseteq (S_2, a_2)$) if the following conditions hold:
+\begin{enumerate}[\IEEEsetlabelwidth{3)}]
+	\item $S_1 \subset S_2$, 
+	\item $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, 
+	\item $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
+\end{enumerate}
+A solution to the hierarchical $k$-median problem is a nested chain of solutions and the hierarchical clustering given by looking at the clusters on each level.
+However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
+An $\alpha$-approximation algorithm for this problem ensures that for all $k = 1, \ldots, |X|$ there exists $(S_i, a_i)$ in our chain of solutions such that $|S_i| \leq k$ and $\mathrm{cost}(S_i, a_i) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
+
+While some flat clustering methods use the $k$-median objective function to inform their decisions and sometimes offer an approximation guarantee, popular methods for hierarchical clustering do not and rely on greedy algorithms which have no guarantee on the solution produced with respect to the objective function but are typically fast and easy to implement.
+This paper will look at how using this objective function can vastly improve the hierarchical clusterings produced.
+In particular, we will show how an approximation framework for this problem which takes in flat clusterings for all values of $k$ compares in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
+The algorithms we will use to produce these flat clusterings will include both approximation algorithms for the $k$-median problem, which will offer an overall approximation guarantee for the hierarchical $k$-median problem, and some heuristic algorithms that have been shown to perform well in practice.
 
@@
 
-Clustering algorithms date back more than 50 years, and there have been many variations introduced since\cite{Jain}.
-However, clustering algorithms are often formulated as NP-hard problems with non-convex objective functions, which are inherently hard to solve \cite{Jain}.
-Below, we will review the widely used heuristics and known approximation algorithms for both partitional and hierarchical clustering.
-
-\subsection{Partitional Clustering}
-
-For partitional clustering, we often specify a number $k$, representing the number of clusters, and pairwise distances between the data points, representing a measure of dissimilarity, such as Euclidean distance.
-The objective is then to identify $k$ clusters so as to minimize some function of the distances.
-
-Arguably the most common partitional clustering method is $k$-means, in which we want to choose $k$ cluster centers to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
+Arguably the most common partitional clustering method is $k$-means, in which we want to choose $k$ cluster centers points in space to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
 $k$-means is usually solved using Lloyd's algorithm (also referred to simply as the $k$-means algorithm), which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}.
@@
 
-Several variations on $k$-means attempt to reduce the sensitivity and worst-case behavior, such as to minimize the maximum distance from a point to its cluster center, called $k$-center clustering.
-Here, a simple greedy algorithm is guaranteed to produce a solution within a factor of $2$ of the optimum, as long as the given distances satisfy the triangle inequality \cite{Gonzalez}.
-Furthermore, this is best approximation ratio possible unless P=NP \cite{Shmoys}.
-However, $k$-center does not perform well in practice compared to other methods \cite{Murphy}.
-
-More recently, spectral based methods have become popular for clustering.
-These algorithms construct a graph based on the similarity matrix and compute the eigenvectors of the normalized graph Laplacian matrix \cite{Luxburg}.
-These eigenvectors are used to split the current cluster in half using the eigenvector associated with the second highest eigenvalue (continuing recursively until there are $k$ clusters).
-Another method is to use the eigenvectors to project the data points into a lower dimensional space on which the $k$-means algorithm is run \cite{Verma}.
-We will base our comparisons against the algorithm introduced by \cite{Meila} which uses the latter method and performed best in practice on several datasets \cite{Verma}.
-
-An objective function that is somewhat in between $k$-means and $k$-center is to minimize the sum of distances from each point to the median point of its cluster, called $k$-median or $k$-medoids clustering.
-There are two common heuristic algorithms for this problem.
+The $k$-median objective function removes some worst-case behavior of $k$-means by restricting the cluster centers to be data points.
+There are two common algorithms for this problem.
 The first is a local search algorithm called partition around medoids (PAM), which starts with $k$ randomly chosen data points as centers, and looks to swap one of the centers with another data point to improve the objective value.
 This repeats until no more improving swaps exist \cite{ESL}.
-Another common heuristic (which we shall call $k$-medoids throughout this paper), which is more computationally efficient than PAM \cite{Park}, is a varitation on Lloyd's $k$-means algorithm.
+The generalization of this algorithm which considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
+Another common heuristic (which we shall call $k$-medoids throughout this paper), which is more computationally efficient than PAM \cite{Park}, is a variation on Lloyd's $k$-means algorithm.
 Starting with randomly chosen centers, we assign each data point to its closest center, and then update the center within each cluster \cite{Park}.
@@
 
-The $k$-median problem is also a fundamental problem in approximation algorithms.
-However, these algorithms are often not tested in practice despite their improved guarantees on the solution.
-The best known approximation ratio for this problem is 2.592 \cite{Wu}.
-This algorithm uses linear programming and is computationally intensive compared to local search and greedy algorithms.
-It is known that a local search algorithm that considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
-Note that for $p=1$, this is equivalent to PAM.
-As far as we know, this algorithm has not been implemented or tested on clustering data, so the relative trade-off of solution quality and computation time as $p$ increases is unknown.
-
-Other partitional clustering methods include MeanShift and DBSCAN, which focus on finding clusters of high density \cite{Comaniciu, Ester}.
-We choose not to compare against density based methods since their objective differs widely from the center-based methods above.
-We also choose not to compare our proposed methods against Bayesian clustering methods such as Gaussian mixture models since these methods require many more parameters (representing the prior information).
-Lastly, if we relax the constraint to identify exactly $k$ clusters, and instead allow the user to input ``preferences'' on the data points being clusters, a recent method called affinity propagation has been shown to perform particularly well \cite{Frey}.
-However, again, because of the difference in available information, we do not to test against this algorithm.
+The $k$-median problem is also a fundamental problem in approximation algorithms where is often presented in the context of opening service centers or factories to serve clients.
+However, these algorithms are often not implemented for clustering despite their improved guarantees on the solution.
+The best known approximation algorithm for this problem has an approximation guarantee of 2.592 and is produced using linear programming \cite{Wu}.
+A simpler LP rounding algorithm by Charikar and Li has a slightly worse guarantee of 3.25 but is easier to implement in practice \cite{Charikar}.
 
 \subsection{Hierarchical Clustering}
-
-In the hierarchical clustering setting, we do not specify a number of clusters.
-Instead, we aim to find a hierarchy of clusters so that clusters at a given level are a merging of clusters at a lower level.
-This presents a larger picture about the similarity between groups of data points, and can allow the user to identify a ``natural'' level of clustering \cite{ESL}.
-
 The most common method for hierarchical clustering is agglomerative clustering.
@@
 Using the local $p$-swap $(3+\frac{2}{p})$-approximation algorithm mentioned above as the black-box $k$-median solver performed well in experimental tests \cite{Nagarajan}.
-While the authors tested the variants of this algorithm against other $k$-median approximation algorithms, it has not been tested against traditional clustering algorithms or on any labeled data sets.
+While the authors tested the variants of this algorithm for different $k$-median approximation algorithms, it has not been tested against traditional clustering algorithms or on any labeled data sets.
 
-Several algorithms have been built upon agglomerative clustering to incorporate non-spherical cluster shapes and categorical data such as the CURE, ROCK, and Chameleon algorithms \cite{Guha_CURE, Guha_ROCK, Karypis}, but do not relate as much to the above approximation algorithm in objective.
-We also choose not compare the approximation framework with Bayesian hierarchical methods for the same reasons as above.
+Several algorithms have been built upon agglomerative clustering that we choose not to test our proposed algorithms against.
+Divisive clustering, similar to agglomerative, starts with the data in one cluster and repeatedly splits a cluster on each level.
+This method has not been studied nearly as much as agglomerative clustering \cite{ESL} but has had relative success in classifying document datasets \cite{Zhao}.
+Algorithms such as the CURE, ROCK, and Chameleon \cite{Guha_CURE, Guha_ROCK, Karypis} incorporate non-spherical cluster shapes and cluster density, but do not relate as much to the above approximation algorithm in objective.
+We also choose not compare the approximation framework with Bayesian hierarchical methods for the same reasons since these methods require many more parameters and add additional prior information that our algorithms do not use.
 
-\section{Partitional Clustering Approximation Algorithms}
+\section{Proposed Algorithms}
 
-We will focus our attention on the $k$-median problem, where we are given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$.
-The goal is to find a set $S \subseteq X$ such that $|S| \leq k$ so as to minimize
-\[ f(S) = \sum_{ i\in X} \min_{j \in S} d(i,j). \]
+Recall the hierarchical $k$-median problem: given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
+ A solution to the hierarchical $k$-median problem is a chain of nested solutions $(S_1, a_1) \subseteq (S_2, a_2) \subseteq \ldots \subseteq (S_l, a_l)$.
+However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
+Consider any $k$-median algorithm $\mathcal{A}$.
+After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce an augmentation step that will nest these clusterings into the hierarchical framework and only increase costs by a bounded factor.
 
-\subsection{Local Search with Multiswaps}
-Consider a local search algorithm which starts with a random subset $S$ of $k$ points.
-At every iteration, the algorithm looks for a set $B \subseteq S$ and a set $A \subseteq X -S$ such that $|A| = |B| \leq p$ and
-\[ f(S- B \cup A) \geq f(S). \]
-If the algorithm finds such sets, it swaps out $B$ from $S$ and adds $A$.
-If no such set exists, the algorithm outputs $S$.
-At every iteration, for $p$ independent of $k$ and $n$, let $\mathcal{N}$ be the size of the search neighborhood.
-If we ensure that the algorithm improves the objective by at least a factor of $(1- \delta) = (1- \varepsilon/\mathcal{N})$, this is a polynomial-time algorithm.
-Furthermore, the local search algorithm outputs $S \subseteq X$ such that $|S| = k$ and
-\[ f(S) \leq \frac{1}{1-\varepsilon} \left (3 + \frac{2}{p} \right ) f(O) \]
-for all $O \subseteq X$. Therefore, there is a trade-off between the computation time of this algorithm and the guaranteed quality of the solution.
+First, set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
+Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
+The cost of these solutions is determined by assigning each point to the nearest cluster center and is scaled so that the minimum cost for $k<n$ is at least one.
+We then order these solutions according to their cost into buckets of the form $[0,0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
 
-\subsection{Supermodular Function Minimization}
-Center-based clustering in general can be viewed as minimizing a supermodular decreasing function subject to a cardinality constraint, a problem related to maximizing a submodular function that is well-studied in approximation algorithms.
-The approximation algorithms in this area use local-search or greedy methods and are therefore computationally tractable on large data.
-In addition, using supermodular functions means the distance measure does not have to satisfy the triangle inequality, which can be useful when considering non-vector data.
+From each non-empty bucket, pick the solution with the smallest number of cluster centers, and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r$.
+We then inductively build our nested solutions.
+Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_i$ by iteratively adding the closest point $j \in S_{i-1}$ to $l$ for all $l \in \overline{V}_i$.
+Note that $|S_{i}| \leq |\overline{V}_i|$.
 
-Given a set $X$, a function $g:2^{X} \mapsto \mathbb{R}$ is supermodular, if for all $T \subseteq S \subseteq X$ and $j \notin S$, 
-\[ g(T \cup \{j\}) - g(T) \leq g(S \cup \{j\}) - g(S). \]
-Furthermore, $g$ is non-increasing if for all $T \subseteq S \subseteq X$, $g(S) \leq g(T)$.
+We then define an assignment $a_i$.
+For each $j \in X$ such that $a_{i-1}(j) \in S_{i}$, we set $a_i(j) = a_{i-1}(j)$.
+Otherwise, we set $a_i(j)$ to be the facility $i \in S_i$ such that assigning all points in $a_{i-1}^{-1}(j)$ (i.e.
+all points in $j$'s cluster in $S_{i-1}$) would have minimum cost.
+Using this assignment, $(a_i, S_i)$ is nested in $(a_{i-1}, S_{i-1})$.
 
-Now considering the $k$-median objective defined above, for completeness we define $f(\emptyset) = \max_{j \in X} f(\{j\})$.
-It is then elementary to check that $f$ is supermodular and non-increasing.
-In general, there is no known approximation algorithm for minimizing a non-increasing supermodular function subject to a cardinality constraint.
-However, if we incorporate the ``steepness'' of $f$, there is a known optimal approximation algorithm \cite{Sviridenko}.
+We continue this nesting until $|S_i| = 1$, reaching the top level of the hierarchy.
+After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantee that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq n-k+1 \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).
 
-As in \cite{Sviridenko}, we define the \emph{total curvature} $c$ of a non-increasing supermodular function $f$ as 
-\[ c = 1 - \min_{j \in X}  \frac{ f(X) - f(X- \{j\})}{ f(\{j\}) - f(\emptyset)}.\]
-Then, for every $\varepsilon >0$ and $c \in [0,1)$, there exists a local search algorithm that produces a set $S \subseteq X$ such that $|S| \leq k$ in polynomial time such that 
-\[ f(S) \leq \left( 1 + \frac{c}{1-c} \cdot e^{-1} + \frac{1}{1-c} \cdot O(\varepsilon) \right) f(O) \]
-for all $O \subseteq X$ such that $|O| \leq k$, with high probability \cite{Sviridenko}. Note that $c$ will vary depending on our dataset so the approximation ratio will vary.
+\subsection{Flat Clustering Algorithms}
+This nesting framework allows us to choose the flat clustering algorithm $\mathcal{A}$.
+As mentioned above, PAM and $k$-medoids are two potential algorithms.
+PAM has the added benefit of being a 5-approximation algorithm for the flat $k$-median problem which yields to an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
+We also considered the size of the local swaps in PAM to improve the approximation ratio but found that the approximation ratio of constructed solutions for given values of $k$ did not significantly improve while the computation time increased greatly.
 
-This algorithm first defines linear and submodular functions on $X$:
-\begin{align*}
-l(A) &= \sum_{j \in A} \left [ f(\{j\}) - f(\emptyset) \right], \\
-g(A) & = - l(A) - f(X \setminus A).
-\end{align*}
-Furthermore, we let $\hat{v}_g = \max_{j \in X} g(\{j\})$, $\hat{v}_l = \max_{j \in X} |l(\{j\})|$, and $\hat{v} = \max(\hat{v}_g , \hat{v}_l)$.
+We also consider two linear programming based algorithms, one of which has an approximation guarantee.
+We first present the linear programming formulation for the $k$-median problem.
+Suppose we have a set $X$ and we let $s_i \in \{0,1\}$ be the variable representing whether or not we want to make $i$ a center for all $i \in X$.
+We also create a variable $a_{i,j} \in \{0, 1\}$ which represents whether or not we assign $i \in X$ to the cluster with center $j \in X$.
+Given this set and assignment, the $k$-median objective function is
+\begin{equation}
+\sum_{i \in X} \sum_{j \in X} a_{i,j} d(i,j), 
+\end{equation}
+which we want to minimize.
+However, we are restricted by the fact that we can choose at most $k$ centers, every point must be assigned to a single cluster center, and that we can only assign a point to a center that is actually in $S$.
+This is represented by the three equations:
+\begin{eqnarray}
+\sum_{i \in X} s_i   \leq k & ,\\
+\sum_{j \in X} a_{i,j}  = 1 & \forall i \in X, \\
+a_{i,j}   \leq s_i & \forall i, j \in X.
+\end{eqnarray}
+Therefore, solving the following integer program would directly find the optimal set of centers and assignment for the $k$-median problem.
+\begin{eqnarray}
+\text{Min } & \sum_{i \in X} \sum_{j \in X} a_{i,j} d(i,j) \\
+\text{s.t. } & \sum_{i \in X} s_i   \leq k  \\
+& \sum_{j \in X} a_{i,j}  = 1 \quad \forall i \in X \\
+& a_{i,j}   \leq s_i \quad \forall i, j \in X \\
+& s_i \in \{0,1\} \quad \forall i \in X \\
+& a_{i,j} \in \{0,1\} \quad \forall i, j \in X
+\end{eqnarray}
+Relaxing the constraints so that $s_i \in [0,1]$ for all $i \in X$ and $a_{i,j} \in [0,1]$ for all $i,j \in X$ yields the linear programming (LP) relaxation for the $k$-median problem, which we can solve in polynomial time.
 
-Given a set $A \subseteq X$, we consider a two-step random sampling.
-We first draw a value $p \in [0,1]$ using the density function $\frac{{e}^p}{e-1}$.
-Then, we choose $B \subseteq A$ by selecting each $x \in A$ to be in $B$ independently with probability $p$.
-The expected value of $g(B)$ is given by
-\[  h(A) = \sum_{B \subseteq A} g(B) \cdot \int_{0}^1 \frac{e^p}{e-1} \cdot p^{|B|-1} (1-p)^{|A|-|B|} dp. \]
-We then define the potential function
-\[ \psi(A) = (1-e^{-1}) h(A) + l(A). \]
+Interestingly, the solution to the linear programming is often integer valued [NEED CITATION].
+Based on this fact, we proposed a flat clustering algorithm that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values.
+We refer to this algorithm as LP-greedy from now on.
+Charikar and Li \cite{charikar} introduce a 3.25-approximation algorithm that uses this LP solution to smartly construct a set of centers.
+We refer to this algorithm as LP-Charikar.
+We choose to implement this LP algorithm rather than the LP-based approximation algorithm with best approximation ratio because the algorithm is simpler to implement (and will most likely have a lower running time).
 
-Now, the algorithm sets $\delta = \frac{\varepsilon}{n}$ to be the minimum step size and starts with a set $S$ of size $n-k$. Then, while there exists $a \in S$ and $b \in X \setminus S$ such that 
-\[ \psi(S -a +b) \geq \psi(S) + \delta, \]
-the algorithm removes $a$ from $S$ and adds $b$. If there is no swap that increases $\psi$, then the algorithm outputs $X \setminus S$.
+The algorithm works by first choosing a subset of data points $C \subseteq X$ to represent the demand.
+For each data point $i$, let $d_{av}(i) = \sum_{j \in X : a_{i,j} > 0} s_i \cdot d(i,j)$.
+We start with $C = \emptyset$.
+At each step, we choose a point $j \in X$ to add to $C$, and all points $j'$ such that $d(j,j') \leq 4 d_{av}(j') $ are removed from $X$.
+This is continued until $X$ is empty.
 
-Note that $\psi$ requires summing an exponential number of terms.
-However, we can estimate $h(A)$ (and therefore $\psi(A)$) by randomly sampling $g(B)$.
-If $\hat{\psi}(A)$ is an estimate of $\psi(A)$ sampled from $\Omega (\varepsilon^{-2} n^4 \ln^4 n \ln M)$ samples, then
-\[ Pr[ | \hat{\psi}(A) - \psi(A) | \geq \delta ] = O(M^{-1}). \]
-By setting $M$ large enough (polynomially large), the algorithm will converge in polynomial time and will return a solution satisying the above approximation ratio with high probability \cite{Sviridenko}.
-As far as we are aware, this algorithm has not been tested for any clustering problems.
+$C$ now represents data points that are relatively far apart.
+These points then choose a bundle of points $U_j$ that represent potential centers they would like to be assigned to.
+In particular, for each point $j \in C$, we look at all points $i \in X$ such that $0 < d(i,j) < 1.5 R_j$, where $R_j$ is half the distance from $j$ to it's nearest point in $C$, and bundle these points into $U_j$ (if a point is in multiple bundles you choose one arbitrarily).
+Now these bundles of points will have a sum of $s_i$ values at least one half.
+Lastly, these bundles are matched together greedily by distance.
+For each $U_j$, we let $\mathrm{vol}(U_j) = \sum_{i \in U_j} s_i$.
 
-\subsection{Implementation and Results}
+Now that the algorithm has grouped together points so the volume is at least 1, the algorithm starts to choosing points from these various groups through sampling.
+Here, opening $U_j$ refers to selecting a point $i \in U_j$ with probabilities relative to the $s_i$ values and adding that point to $S$.
+\begin{enumerate}[\IEEEsetlabelwidth{3)}]
+\item For each matched $U_j$ and $U_{j'}$, the algorithm opens $U_j$ with probability $1-\mathrm{vol}(U_{j'})$, opens $U_{j'}$ with probability $1- \mathrm{vol}(U_j)$, and opens both $U_{j}$ and $U_{j'}$ with probability $\mathrm{vol}(U_j) + \mathrm{vol}(U_{j'}) -1$.
+\item For any unmatched $U_j$, it is opened with probability $\mathrm{vol}(U_j)$.
+\item For any $i$ not in any $U_j$ such that $s_i > 0$, add $i$ to $S$ with probability $s_i$. 
+\end{enumerate}
+This sampling process returns a set of centers of size $k$  in expectation. For our purposes, we sample until we obtain a solution with exactly $k$ centers.
 
-We implemented the local search multiswap algorithm described above for $p=1, 2, 3$ as well as the supermodular function minimization algorithm in Python, and compared against $k$-means, $k$-medoids (with a maximum iteration count of 1000) and Shi and Meila's spectral algorithm (implemented in the scikit-learn package for Python).
-Note that for $p=1$, the multiswap algorithm is equivalent to the well-known PAM algorithm.
-In order to calculate approximation ratios, Gurobi was used to compute the global optimum for each data set.
+\section{Implementation and Results}
+We implemented the nesting hierarchical framework above in Python 2.7, offering options for both the randomized and deterministic versions.
+For the black-box $k$-median algorithm, we implemented $k$-medoids, PAM, LP-greedy, and LP-charikar in Python using Gurobi 6.0.3 to solve the linear programming instances.
+We compared the results to hierarchical agglomerative clustering with both average linkage and Ward's linkage (implemented in the scipy package in Python).
 
-To achieve reasonable computation times for the multiswap and supermodular function minimization algorithms, at each iteration we first consider nearby swaps before sampling for other potential swaps.
-For the multiswaps algorithm, we sample $O(p\cdot n)$ potential swaps on each iteration, and for the supermodular function, sample $min(100,n^2)$ swaps where each potential function is estimated by sampling 50 subsets.
-We also ensure that some minimum level progress is achieved for both algorithms by setting $\varepsilon = 0.1$ for the multiswaps algorithm and $\varepsilon = 10$ for the submodular function.
-These values were chosen by looking at the relative performance and computation time on random Gaussian data not tested below.
+To avoid cycling, we limited the number of iterations in $k$-medoids to 1000.
+For PAM, we also needed to limit the number of swaps to achieve a reasonable runtime.
+Instead of searching over all potential swaps, we sample at most $n$ potential ones.
+We also ensure that some minimum level progress is achieved by ensuring the algorithm improves by at least a factor of 0.9 on each iteration.
+At the end of PAM, we run 10 iterations of $k$-medoids to ensure the algorithm finds a local optimum.
+These choices were made by looking at the relative performance and computation time on random Gaussian data not tested below.
+For PAM and $k$-medoids, when solving for multiple values of $k$, we use the solution with $k$ centers as a warm start for $k+l$ centers by choosing an additional random $l$ centers.
+For the two LP algorithms, Gurobi uses a warm start from solving the LP for $k+l$ to solve it for $k$ because we found this reverse direction to be faster in practice.
 
-We also considered a hybrid heuristic algorithm (labelled $p$-swap+ in the results) where we ran $k$-medoids for 10 iterations on the result of the multiswap search algorithm.
-This is motivated by the fact that, similar to Lloyd's algorithm, $k$-medoids will quickly converge to local minima.
-Since $p$-swaps avoids getting stuck in local minima, the hybrid algorithm should in theory achieve good results with minimal extra computation time over the multiswap algorithm.
+We also tested a more computationally friendly heuristic of the hierarchical framework which only buckets solutions for values of $k$ that are a power of two.
+This yields many fewer $k$-median instances to solve, reducing the run-time significantly.
+This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
 
-For each data set and each clustering method, we report objective function values, the approximation ratio, computation time, and where applicable, adjusted mutual information (AMI).
-AMI is a measure which scores accuracy of predicted clusters compared to a true labeling by measuring the agreement of two assignments.
-The measure is normalized to reflect chance, so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
-See \cite{Vinh} for further information.
+For each data set and each clustering method, we report the approximation ratio, computation time, and FScore measure.
+FScore is a measure that looks at the clusters represented on each level and how they represent the true clustering in terms of precision and recall.
+See \cite{Larsen} for more information.
+[ADD DEFINITION OF FSCORE AND MORE DETAIL]
+
+[MAKE CODE AVAILABLE PUBLICLY]
+
+Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core 3.4 Ghz Intel i7 processor with 16 GB physical memory.
+Results are given in Tables 2-4  below with some results averaged due to space constraints.
 The results reported are average values across 10 trials.
@@
 The data sets tested against were:
-\begin{description}
+\begin{LaTeXdescription}
 \item[Gaussian] 27 synthetic $k$-center Gaussian data sets, generated in Python.
-In particular, Gauss\textunderscore d\textunderscore k\textunderscore e was generated by first choosing k random centers in $[-1,1]^d$.
+In particular, Gauss\textunderscore d\textunderscore k\textunderscore e was generated by first choosing $k$ random centers in $[-1,1]^d$.
 Around each center, we then generated $(1000/k) \cdot \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
-\item[OR $p$-median Library] From a collection of test data sets created by \cite{Beasley} with 40 test sets of size $100, 200, \ldots, 900$ and varying values of $k$. 
+\item[OR $p$-median Library] \cite{Beasley} From a collection of test data sets created by Beasley with 40 test sets of size $100, 200, \ldots, 900$ and varying values of $k$. 
 \item[UCI Iris] \cite{Iris} Contains 4 attributes on 150 instances of flowers, each of with belongs to one of three species of iris.
@@
 \item[UCI Soybean (Small)] \cite{Soybean} 47 instances, with 35 attributes data set of soybean disease from UCI.
-\end{description}
+\end{LaTeXdescription}
 
-\begin{figure}[h]
+\begin{table}[!t]
+\caption{Some specifics about our datasets}
+\label{table_specifics}
+\centering
 \begin{tabular}{ | l | l | l | l | l | }
@@
  \end{tabular}
-\caption{Some specifics about our datasets}
-\end{figure}
+\end{table}
 
-Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core 3.4 Ghz Intel i7 processor with 16 GB physical memory.
-Results are given in Figure 2  below with some results averaged due to space constraints.
-
-\begin{figure*}[h]
-\label{fig:flat_results}
-\small
-\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | }
-\hline
-	Dataset Name & Time & Time & Time & Time & Time & Time & Time & Time & Time \\ \hline
-	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap  \\ \hline
-	Soybean & 0.015 & 0.061 & 0.002 & 1.179 & 0.026 & 0.016 & 0.048 & 0.049 & 0.081 \\ 
-	Iris & 0.013 & 0.023 & 0.003 & 9.788 & 0.125 & 0.102 & 0.212 & 0.229 & 0.414 \\ 
-	Gauss\_5\_5\_5 & 0.025 & 0.056 & 0.032 & 75.262 & 1.159 & 1.506 & 2.188 & 2.382 & 3.71 \\ 
-	Gauss\_5\_10\_5 & 0.054 & 0.095 & 0.109 & 134.137 & 6.834 & 5.407 & 9.243 & 9.428 & 15.96 \\ 
-	%Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 \\ 
-	Gaussian avg & 0.073 & 0.128 & 0.182 & 144.455 & 13.289 & 13.021 & 18.749 & 18.696 & 24.592 \\ 
-	pmed5 & 0.038 & 8.298 & 0.481 & 0.622 & 0.653 & 0.748 & 0.939 & \  & \  \\ 
-	pmed20 & 1.795 & 174.951 & 52.208 & 44.979 & 68.554 & 68.231 & 79.288 & \  & \  \\ 
-	% pmed40 & 3.527 & 849.852 & 251.259 & 264.867 & 250.412 & 391.738 & 350.235 & \  & \  \\ 
-	pmedian avg & 0.807 & 203.308 & 40.386 & 40.347 & 48.264 & 56.76 & 56.19 & \  & \  \\ \hline
-\end{tabular}
-\vspace{0.2 cm}
-
-\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
-\hline
-	Dataset Name & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
-	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap   \\ \hline
-	Soybean & 0.686 & 0.178 & 0.55 & 0.612 & 0.708 & 0.732 & 0.717 & 0.708 & 0.712 \\ 
-	Iris & 0.748 & 0.757 & 0.69 & 0.7 & 0.772 & 0.760 & 0.763 & 0.757 & 0.761 \\ 
-	Gauss\_5\_5\_5 & 0.741 & 0.713 & 0.687 & 0.657 & 0.742 & 0.749 & 0.741 & 0.744 & 0.737  \\ 
-	Gauss\_5\_10\_5 & 0.783 & 0.736 & 0.677 & 0.575 & 0.729 & 0.725 & 0.737 & 0.737& 0.729  \\ 
-	%Gauss\_5\_50\_5 & 0.646 & 0.596 & 0.621 & 0.570 & 0.622 & 0.619 & 0.627 & 0.631 & 0.625  \\ 
-	Gaussian avg & 0.777 & 0.744 & 9.112 & 0.623 & 0.76 & 0.761 & 0.762 & 0.761 & 0.76 \\ \hline
-\end{tabular}
-\vspace{0.2 cm}
-
-\begin{tabular}{ | l | l | l | l | l | l | l | l | l | }
-\hline
-	Dataset Name & Approx & Approx & Approx & Approx & Approx & Approx & Approx  \\ \hline
-	& k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap   \\ \hline
-	Soybean & 1.167 & 1.199 & 1.027 & 1.012 & 1.022 & 1.012 & 1.024 \\ 
-	Iris & 1.081 & 1.375 & 1.016 & 1.004 & 1.013 & 1.004 & 1.009 \\ 
-	Gauss\_5\_5\_5 & 1.070 & 1.436 & 1.030 & 1.002 & 1.027 & 1.005 & 1.026  \\ 
-	Gauss\_5\_10\_5 & 1.080 & 1.389 & 1.017 & 1.011 & 1.018 & 1.009 & 1.017  \\ 
-	%Gauss\_5\_50\_5 & 1.0887 & 1.272 & 1.030 & 1.020 & 1.027 & 1.014 & 1.033  \\ 
-	Gaussian avg & 1.064 & 1.416 & 1.025 & 1.009 & 1.023 & 1.008 & 1.024 \\ 
-	pmed5 & 1.373 & 1.681 & 1.06 & 1.026 & 1.071 & 1.033 & 1.051 \\ 
-	pmed20 & 1.501 & 1.73 & 1.079 & 1.032 & 1.076 & 1.027 & 1.073 \\ 
-	%pmed40 & 1.187 & 1.423 & 1.046 & 1.017 & 1.043 & 1.019 & 1.039 \\ 
-	pmedian avg & 1.203 & 1.456 & 1.041 & 1.017 & 1.039 & 1.016 & 1.037 \\ \hline
-\end{tabular}
-
-\caption{Average Computation Time (sec), AMI measure, and Approximation Coeffiecients for Partitional Clustering Algorithms}
-\end{figure*}
-
-In terms of computation time, our proposed algorithms were much slower than $k$-means, the spectral algorithm, or $k$-medoids.
-In general, the multiswaps algorithm was around two orders of magnitude slower with a small multiplicative increase as $p$ increased, and the supermodular algorithm was around four orders of magnitude slower, likely due to the extra potential function sampling the supermodular algorithm must perform.
-
-In addition, the supermodular function algorithm did not perform as well in terms of AMI or approximation ratio compared to the other methods.
-Calculating the curvature, we saw that every  test set's curvature was very close to 1, so the guarantee of the algorithm was quite poor.
-Despite this, the algorithm did maintain an approximation ratio below 2 for all instances and an AMI measure above 0.5 for almost all instances, averaging around 0.85 for the 7 dimensional Gaussian data.
-
-When increasing $p$ for the multiswaps algorithm, we did observe slight increases in computation time, but the times remained on the same order of magnitude.
-The increase in $p$ did not seem to have any significant effect on the AMI measure nor approximation ratio.
-This may be due to the fact the algorithm for $p=1$ performed quite well already with an approximation ratio very close to 1.
-In addition, PAM had the highest AMI measures for the iris and soybean data sets.
-The AMI measures for the Gaussian data set were higher than those of $k$-medoids in general but equivalent or only slightly weaker than those of the spectral method or $k$-means.
-The heuristic method of adding in a few rounds of $k$-medoids at the end of our multiswaps algorithm did not have much effect on the AMI measure but did boost the approximation ratio, sometimes finding an optimal solution.
-
-It is interesting to note that, for the Gaussian data, an increase in dimension, reduction in $k$, or reduction in the percentage of noisy data made it easier for the algorithms to classify the data set, which makes sense based on our intuition of clustering.
-For the OR library instances, we saw that the approximation ratio tended to increase as $k$ grew larger but not as $n$ grew.
-
-Overall, the results of our experiments do not suggest that the proposed approximation algorithms provide any advantage over current well known methods.
-However, the results do suggest that we can use sampling to run PAM with lower computational cost while still maintaining good performance accuracy in terms of AMI and approximation ratio.
-
-\section{Hierarchical Clustering Approximation Algorithms}
-Turning now to the hierarchical clustering setting, given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
-Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$.
-We define $\mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i))$.
-
-Given solution pairs $(S_1, a_1)$ and $(S_2, a_2)$, we say that $(S_1, a_1)$ is \emph{nested} in $(S_2, a_2)$ if the following conditions hold.
-\begin{enumerate}
-	\item $S_1 \subset S_2$, 
-	\item $\forall  j \in X$, if $a_2(j) \in S_1$, then $a_1(j) = a_2(j)$, 
-	\item $\forall j, k \in X$, if $a_2(j) = a_2(k)$, then $a_1(j) = a_1(k)$.
-\end{enumerate}
- A solution to the hierarchical clustering problem is therefore a chain of nested solutions.
-However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
-Consider any $k$-median algorithm $\mathcal{A}$.
-After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce an augmentation step that will nest these clusterings into the hierarchical framework and only increase costs by a bounded factor.
-
-First, set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
-Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
-The cost of these solutions is determined by assigning each point to the nearest cluster center and is scaled so that the minimum cost for $k<n$ is at least one.
-We then order these solutions according to their cost into buckets of the form $[0,0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
-
-From each non-empty bucket, pick the solution with the smallest number of cluster centers, and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r$.
-We then inductively build our nested solutions.
-Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_i$ by iteratively adding the closest point $j \in S_{i-1}$ to $k$ for all $k \in \overline{V}_i$.
-Note that $|S_{i}| \leq |\overline{V}_i|$.
-
-We then define an assignment $a_i$.
-For each $j \in X$ such that $a_{i-1}(j) \in S_{i}$, we set $a_i(j) = a_{i-1}(j)$.
-Otherwise, we set $a_i(j)$ to be the facility $i \in S_i$ such that assigning all points in $a_{i-1}^{-1}(j)$ (i.e.
-all points in $j$'s cluster in $S_{i-1}$) would have minimum cost.
-Using this assignment, $(a_i, S_i)$ is nested in $(a_{i-1}, S_{i-1})$.
-
-We continue this nesting until $|S_i| = 1$, reaching the top level of the hierarchy.
-After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantee that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq n-k+1 \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).
-
-\subsection{Implementation and Results}
-We implemented the nesting hierarchical framework above in Python, offering options for both the randomized and deterministic versions.
-As a black-box $k$-median solver, we used $k$-medoids and the 1-swap+ algorithms, discussed in the partitional clustering section above.
-We compared the results to hierarchical agglomerative clustering with both average linkage and Ward's linkage (implemented in the scipy package in Python).
-
-We also tested a more computationally friendly heuristic of the hierarchical framework which only buckets solutions for values of $k$ that are a power of two.
-This yields many fewer $k$-median instances to solve, reducing the run-time significantly.
-This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
-
-For each data set discussed previously, and each clustering method, we report the approximation ratio, computation time, and FScore measure.
-FScore is a measure that looks at the clusters represented on each level and how they represent the true clustering in terms of precision and recall.
-See \cite{Larsen} for more information.
-Results are in Figure 3.
-
-\begin{figure*}[h]
-\label{fig:hier_results}
-\small
+\begin{table*}[!t]
+\label{fig:hier_results_comp}
+\caption{Average Computation Time (sec), FScore measure, and Approximation Coeffiecients for Hierarchical Clustering Algorithms.
+K-med and PAM represent the black-box $k$-median solvers used and rand and logn indicate whether the randomized and logarithmic setting were used, respectively.}
+\centering
 \begin{tabular}{ | l | l | l | l | l | l | l | }
@@
 \end{tabular}
-\vspace{0.2 cm}
+\end{table*}
 
+\begin{table*}[!t]
+\label{fig:hier_results_FScore}
+\caption{Average FScore measure for Hierarchical Clustering Algorithms.}
+\centering
 \begin{tabular}{ | l | l | l | l | l | l | l | }
@@
 \end{tabular}
-\vspace{0.2 cm}
+\end{table*}
 
+\begin{table*}[!t]
+\label{fig:hier_results_approx}
+\caption{Average Approximation Coeffiecients for Hierarchical Clustering Algorithms.}
+\centering
 \begin{tabular}{ | l | l | l | l | l | }
@@
 \end{tabular}
-
-\caption{Average Computation Time (sec), FScore measure, and Approximation Coeffiecients for Hierarchical Clustering Algorithms.
-K-med and PAM represent the black-box $k$-median solvers used and rand and logn indicate whether the randomized and logarithmic setting were used, respectively.}
-\end{figure*}
+\end{table*}
 
@@
 
-\section{Acknowledgements}
-We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback on this project.
-
-\label{submission}
+\section*{Acknowledgements}
+We would like to acknowledge David P. Williamson and Karthik Sridharan for their valuable feedback.
 
@@
 
-\bibliography{projectBib}
-\bibliographystyle{icml2014}
-
-%\appendix
-%
-%\section{Appendix 1: Full Results}
-%
-%\subsection{Gaussian Data}
-%\begin{figure*}[h]
-%\begin{center}
-%\begin{tabular}{ | l | l | l | l | l | }
-%\hline
-%	Dataset & NameOpt Obj & k & n & dim  \\ \hline
-%	Gauss\_3\_5\_0 & 195 & 5 & 542 & 3 \\ 
-%	Gauss\_3\_5\_5 & 225 & 5 & 620 & 3 \\ 
-%	Gauss\_3\_5\_10 & 158 & 5 & 441 & 3 \\
-%	Gauss\_3\_10\_0 & 147 & 10 & 472 & 3 \\
-%	Gauss\_3\_10\_5 & 142 & 10 & 480 & 3 \\ 
-%	Gauss\_3\_10\_10 & 201 & 10 & 557 & 3 \\ 
-%	Gauss\_3\_50\_0 & 108 & 50 & 471 & 3 \\
-%	Gauss\_3\_50\_5 & 94 & 50 & 422 & 3 \\ 
-%	Gauss\_3\_50\_10 & 95 & 50 & 408 & 3 \\ 
-%	Gauss\_5\_5\_0 & 352 & 5 & 720 & 5 \\ 
-%	Gauss\_5\_5\_5 & 195 & 5 & 388 & 5 \\ 
-%	Gauss\_5\_5\_10 & 214 & 5 & 394 & 5 \\ 
-%	Gauss\_5\_10\_0 & 346 & 10 & 703 & 5 \\ 
-%	Gauss\_5\_10\_5 & 270 & 10 & 544 & 5 \\ 
-%	Gauss\_5\_10\_10 & 292 & 10 & 572 & 5 \\ 
-%	Gauss\_5\_50\_0 & 188 & 50 & 445 & 5 \\ 
-%	Gauss\_5\_50\_5 & 178 & 50 & 408 & 5 \\ 
-%	Gauss\_5\_50\_10 & 229 & 50 & 480 & 5 \\ 
-%	Gauss\_7\_5\_0 & 349 & 5 & 567 & 7 \\ 
-%	Gauss\_7\_5\_5 & 368 & 5 & 579 & 7 \\ 
-%	Gauss\_7\_5\_10 & 356 & 5 & 527 & 7 \\ 
-%	Gauss\_7\_10\_0 & 351 & 10 & 573 & 7 \\ 
-%	Gauss\_7\_10\_5 & 302 & 10 & 478 & 7 \\ 
-%	Gauss\_7\_10\_10 & 354 & 10 & 534 & 7 \\ 
-%	Gauss\_7\_50\_0 & 257 & 50 & 451 & 7 \\ 
-%	Gauss\_7\_50\_5 & 256 & 50 & 430 & 7 \\ 
-%	Gauss\_7\_50\_10 & 298 & 50 & 480 & 7 \\ \hline
-%\end{tabular}
-%\end{center}
-%
-%\caption{Details For Gaussian Datasets}
-%\end{figure*}
-%
-%\begin{figure*}[h]
-%\begin{center}
-%\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
-%\hline
-%	Dataset &Time & Time & Time & Time & Time & Time & Time & Time & Time & Time   \\ \hline
-%	&K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swaps+ & 3-Swap & 4-swaps   \\ 
-%	Gauss\_3\_5\_0 & 0.024 & 0.056 & 0.066 & 131.305 & 2.707 & 3.195 & 4.779 & 4.138 & 7.683 & 11.325 \\ 
-%	Gauss\_3\_5\_5 & 0.032 & 0.076 & 0.099 & 195.013 & 3.682 & 3.764 & 6.639 & 5.32 & 9.627 & 14.575 \\ 
-%	Gauss\_3\_5\_10 & 0.033 & 0.055 & 0.047 & 80.439 & 2.244 & 1.802 & 2.806 & 3.024 & 5.107 & 7.457 \\ 
-%	Gauss\_3\_10\_0 & 0.041 & 0.083 & 0.095 & 121.551 & 5.133 & 3.576 & 7.452 & 6.3 & 10.157 & 11.872 \\ 
-%	Gauss\_3\_10\_5 & 0.048 & 0.1 & 0.117 & 101.072 & 5.026 & 4.96 & 8.57 & 6.952 & 11.948 & 12.225 \\ 
-%	Gauss\_3\_10\_10 & 0.05 & 0.106 & 0.153 & 149.352 & 7.101 & 6.456 & 11.273 & 10.646 & 18.21 & 17.414 \\ 
-%	Gauss\_3\_50\_0 & 0.154 & 0.324 & 0.435 & 175.827 & 29.201 & 34.228 & 43.65 & 49.964 & 48.259 & 66.413 \\ 
-%	Gauss\_3\_50\_5 & 0.151 & 0.253 & 0.376 & 127.64 & 26.318 & 21.541 & 38.004 & 33.213 & 45.862 & 57.262 \\ 
-%	Gauss\_3\_50\_10 & 0.153 & 0.265 & 0.301 & 132.119 & 21.341 & 22.556 & 35.389 & 32.139 & 44.361 & 50.143 \\ 
-%	Gauss\_5\_5\_0 & 0.026 & 0.079 & 0.08 & 230.345 & 4.997 & 4.21 & 9.428 & 8.063 & 12.811 & 19.337 \\ 
-%	Gauss\_5\_5\_5 & 0.025 & 0.056 & 0.032 & 75.262 & 1.159 & 1.506 & 2.188 & 2.382 & 3.71 & 5.797 \\ 
-%	Gauss\_5\_5\_10 & 0.025 & 0.042 & 0.027 & 69.349 & 1.383 & 1.458 & 2.403 & 2.575 & 3.811 & 5.748 \\ 
-%	Gauss\_5\_10\_0 & 0.044 & 0.1 & 0.18 & 227.747 & 11.633 & 12.129 & 17.724 & 18.654 & 26.563 & 33.646 \\ 
-%	Gauss\_5\_10\_5 & 0.054 & 0.095 & 0.109 & 134.137 & 6.834 & 5.407 & 9.243 & 9.428 & 15.96 & 21.795 \\ 
-%	Gauss\_5\_10\_10 & 0.045 & 0.092 & 0.117 & 127.614 & 7.639 & 6.456 & 11.807 & 12.186 & 18.339 & 23.554 \\ 
-%	Gauss\_5\_50\_0 & 0.142 & 0.211 & 0.389 & 139.818 & 30.155 & 27.431 & 37.251 & 43.421 & 50.192 & 68.329 \\ 
-%	Gauss\_5\_50\_5 & 0.137 & 0.196 & 0.302 & 97.336 & 27.856 & 21.446 & 32.407 & 39.614 & 44.69 & 53.546 \\ 
-%	Gauss\_5\_50\_10 & 0.149 & 0.219 & 0.417 & 170.431 & 37.132 & 35.119 & 45.341 & 47.002 & 56.683 & 69.712 \\ 
-%	Gauss\_7\_5\_0 & 0.019 & 0.054 & 0.052 & 185.865 & 2.975 & 3.565 & 4.954 & 5.005 & 8.531 & 11.867 \\ 
-%	Gauss\_7\_5\_5 & 0.031 & 0.078 & 0.047 & 154.1 & 2.603 & 3.04 & 5.45 & 5.738 & 8.056 & 12.505 \\ 
-%	Gauss\_7\_5\_10 & 0.018 & 0.052 & 0.043 & 130.185 & 2.949 & 2.478 & 4.668 & 4.849 & 7.324 & 11.036 \\ 
-%	Gauss\_7\_10\_0 & 0.045 & 0.071 & 0.097 & 185.521 & 8.015 & 7.338 & 12.495 & 10.474 & 16.174 & 20.671 \\ 
-%	Gauss\_7\_10\_5 & 0.04 & 0.063 & 0.073 & 136.189 & 5.315 & 5.365 & 9.287 & 6.87 & 11.768 & 16.072 \\ 
-%	Gauss\_7\_10\_10 & 0.048 & 0.09 & 0.1 & 168.962 & 7.047 & 6.538 & 9.783 & 9.322 & 13.581 & 19.41 \\ 
-%	Gauss\_7\_50\_0 & 0.143 & 0.199 & 0.401 & 152.958 & 32.592 & 37.747 & 39.026 & 42.999 & 60.483 & 69.667 \\ 
-%	Gauss\_7\_50\_5 & 0.147 & 0.212 & 0.327 & 136.496 & 27.8 & 26.288 & 38.683 & 34.335 & 47.779 & 59.414 \\ 
-%	Gauss\_7\_50\_10 & 0.149 & 0.231 & 0.426 & 163.651 & 37.968 & 41.976 & 55.514 & 50.175 & 56.302 & 79.819 \\ \hline
-%\end{tabular}
-%
-%\end{center}
-%
-%\caption{Running Times For Gaussian Datasets}
-%\end{figure*}
-%
-%\begin{figure*}[h]
-%\begin{center}
-%\begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | }
-%\hline
-%	Dataset&AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI & AMI  \\ \hline
-%	& K-means & Spectral & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-Swaps + & 3-Swap & 4-swaps   \\ \hline
-%	Gauss\_3\_5\_0 & 0.884 & 0.887 & 0.729 & 0.678 & 0.88 & 0.876 & 0.884 & 0.876 & 0.881 & 0.882 \\ 
-%	Gauss\_3\_5\_5 & 0.7 & 0.62 & 0.637 & 0.545 & 0.683 & 0.68 & 0.685 & 0.69 & 0.682 & 0.686 \\ 
-%	Gauss\_3\_5\_10 & 0.68 & 0.618 & 0.62 & 0.504 & 0.663 & 0.665 & 0.659 & 0.659 & 0.664 & 0.668 \\ 
-%	Gauss\_3\_10\_0 & 0.693 & 0.641 & 0.699 & 0.619 & 0.692 & 0.692 & 0.688 & 0.696 & 0.685 & 0.685 \\ 
-%	Gauss\_3\_10\_5 & 0.534 & 0.466 & 0.516 & 0.446 & 0.534 & 0.533 & 0.535 & 0.532 & 0.531 & 0.536 \\ 
-%	Gauss\_3\_10\_10 & 0.669 & 0.56 & 0.616 & 0.516 & 0.651 & 0.653 & 0.652 & 0.657 & 0.657 & 0.653 \\ 
-%	Gauss\_3\_50\_0 & 0.504 & 0.498 & 0.498 & 0.475 & 0.501 & 0.497 & 0.491 & 0.497 & 0.495 & 0.501 \\ 
-%	Gauss\_3\_50\_5 & 0.407 & 0.412 & 0.402 & 0.396 & 0.392 & 0.385 & 0.389 & 0.386 & 0.387 & 0.385 \\ 
-%	Gauss\_3\_50\_10 & 0.421 & 0.424 & 0.406 & 0.39 & 0.413 & 0.414 & 0.408 & 0.404 & 0.405 & 0.41 \\ 
-%	Gauss\_5\_5\_0 & 0.99 & 0.99 & 0.858 & 0.719 & 0.988 & 0.994 & 0.989 & 0.994 & 0.989 & 0.985 \\ 
-%	Gauss\_5\_5\_5 & 0.741 & 0.713 & 0.687 & 0.657 & 0.742 & 0.749 & 0.741 & 0.744 & 0.737 & 0.747 \\ 
-%	Gauss\_5\_5\_10 & 0.833 & 0.83 & 227.742 & 0.663 & 0.76 & 0.779 & 0.806 & 0.747 & 0.789 & 0.77 \\ 
-%	Gauss\_5\_10\_0 & 0.974 & 0.972 & 0.858 & 0.732 & 0.96 & 0.964 & 0.964 & 0.965 & 0.96 & 0.964 \\ 
-%	Gauss\_5\_10\_5 & 0.783 & 0.736 & 0.677 & 0.575 & 0.729 & 0.725 & 0.737 & 0.737 & 0.729 & 0.725 \\ 
-%	Gauss\_5\_10\_10 & 0.878 & 0.829 & 0.795 & 0.704 & 0.873 & 0.875 & 0.873 & 0.872 & 0.875 & 0.868 \\ 
-%	Gauss\_5\_50\_0 & 0.723 & 0.669 & 0.668 & 0.622 & 0.695 & 0.692 & 0.692 & 0.7 & 0.696 & 0.692 \\ 
-%	Gauss\_5\_50\_5 & 0.646 & 0.596 & 0.621 & 0.57 & 0.622 & 0.619 & 0.627 & 0.631 & 0.625 & 0.626 \\
-%	Gauss\_5\_50\_10 & 0.669 & 0.639 & 0.626 & 0.546 & 0.64 & 0.641 & 0.639 & 0.65 & 0.643 & 0.637 \\ 
-%	Gauss\_7\_5\_0 & 1 & 1 & 0.852 & 0.773 & 0.999 & 1 & 0.999 & 1 & 0.999 & 0.998 \\ 
-%	Gauss\_7\_5\_5 & 0.847 & 0.832 & 0.86 & 0.72 & 0.852 & 0.85 & 0.849 & 0.847 & 0.847 & 0.847 \\ 
-%	Gauss\_7\_5\_10 & 0.979 & 0.963 & 0.839 & 0.744 & 0.981 & 0.979 & 0.98 & 0.979 & 0.978 & 0.978 \\ 
-%	Gauss\_7\_10\_0 & 0.995 & 0.995 & 0.869 & 0.785 & 0.988 & 0.984 & 0.987 & 0.984 & 0.982 & 0.976 \\ 
-%	Gauss\_7\_10\_5 & 0.988 & 0.994 & 0.853 & 0.732 & 0.988 & 0.994 & 0.99 & 0.994 & 0.993 & 0.994 \\ 
-%	Gauss\_7\_10\_10 & 0.89 & 0.81 & 0.845 & 0.724 & 0.851 & 0.856 & 0.854 & 0.854 & 0.85 & 0.851 \\ 
-%	Gauss\_7\_50\_0 & 0.867 & 0.834 & 0.784 & 0.693 & 0.832 & 0.835 & 0.84 & 0.833 & 0.835 & 0.846 \\ 
-%	Gauss\_7\_50\_5 & 0.852 & 0.808 & 0.727 & 0.647 & 0.813 & 0.815 & 0.814 & 0.815 & 0.812 & 0.807 \\ 
-%	Gauss\_7\_50\_10 & 0.829 & 0.761 & 0.742 & 0.645 & 0.807 & 0.799 & 0.807 & 0.804 & 0.804 & 0.799 \\ \hline
-%\end{tabular}
-%
-%\end{center}
-%
-%\caption{AMI For Gaussian Datasets}
-%\end{figure*}
-%
-%
-%\begin{figure*}[h]
-%\begin{center}
-%\begin{tabular}{ | l | l | l | l | l | l | l | l | l | }
-%\hline
-%	Dataset & Approx & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
-%	& k-medoids & Supermodular & PAM & 1-Swap+ & 2-Swap & 2-Swap+ & 3-Swap & 4-swaps   \\ \hline
-%	Gauss\_3\_5\_0 & 1.159 & 1.481 & 1.013 & 1 & 1.011 & 1 & 1.013 & 1.009 \\ 
-%	Gauss\_3\_5\_5 & 1.057 & 1.4 & 1.005 & 1.001 & 1.006 & 1.001 & 1.01 & 1.007 \\ 
-%	Gauss\_3\_5\_10 & 1.047 & 1.377 & 1.007 & 1.004 & 1.014 & 1.007 & 1.009 & 1.008 \\ 
-%	Gauss\_3\_10\_0 & 1.029 & 1.404 & 1.018 & 1.01 & 1.013 & 1.01 & 1.022 & 1.019 \\ 
-%	Gauss\_3\_10\_5 & 1.063 & 1.329 & 1.025 & 1.013 & 1.021 & 1.012 & 1.02 & 1.026 \\ 
-%	Gauss\_3\_10\_10 & 1.06 & 1.369 & 1.017 & 1.009 & 1.018 & 1.007 & 1.016 & 1.016 \\ 
-%	Gauss\_3\_50\_0 & 1.112 & 1.354 & 1.05 & 1.026 & 1.037 & 1.025 & 1.049 & 1.044 \\ 
-%	Gauss\_3\_50\_5 & 1.124 & 1.327 & 1.045 & 1.025 & 1.041 & 1.019 & 1.044 & 1.04 \\ 
-%	Gauss\_3\_50\_10 & 1.121 & 1.316 & 1.044 & 1.023 & 1.037 & 1.021 & 1.043 & 1.043 \\ 
-%	Gauss\_5\_5\_0 & 1.105 & 1.576 & 1.013 & 1 & 1.011 & 1 & 1.016 & 1.015 \\ 
-%	Gauss\_5\_5\_5 & 1.07 & 1.436 & 1.03 & 1.002 & 1.028 & 1.005 & 1.026 & 1.024 \\ 
-%	Gauss\_5\_5\_10 & 0.003 & 1.369 & 1.021 & 1.004 & 1.016 & 1.007 & 1.016 & 1.015 \\ 
-%	Gauss\_5\_10\_0 & 1.148 & 1.619 & 1.029 & 1 & 1.026 & 1.001 & 1.028 & 1.029 \\ 
-%	Gauss\_5\_10\_5 & 1.08 & 1.389 & 1.017 & 1.011 & 1.018 & 1.009 & 1.017 & 1.019 \\ 
-%	Gauss\_5\_10\_10 & 1.152 & 1.458 & 1.022 & 1.006 & 1.023 & 1.007 & 1.021 & 1.026 \\
-%	Gauss\_5\_50\_0 & 1.096 & 1.271 & 1.031 & 1.017 & 1.033 & 1.015 & 1.031 & 1.029 \\ 
-%	Gauss\_5\_50\_5 & 1.089 & 1.272 & 1.03 & 1.02 & 1.027 & 1.014 & 1.033 & 1.032 \\ 
-%	Gauss\_5\_50\_10 & 1.09 & 1.299 & 1.029 & 1.017 & 1.036 & 1.016 & 1.028 & 1.036 \\ 
-%	Gauss\_7\_5\_0 & 1.214 & 1.59 & 1.016 & 1 & 1.018 & 1 & 1.014 & 1.016 \\ 
-%	Gauss\_7\_5\_5 & 1.046 & 1.446 & 1.017 & 1.009 & 1.014 & 1.004 & 1.014 & 1.019 \\ 
-%	Gauss\_7\_5\_10 & 1.153 & 1.53 & 1.01 & 1 & 1.007 & 1 & 1.01 & 1.009 \\ 
-%	Gauss\_7\_10\_0 & 1.132 & 1.504 & 1.032 & 1 & 1.023 & 1 & 1.032 & 1.034 \\ 
-%	Gauss\_7\_10\_5 & 1.159 & 1.569 & 1.031 & 1 & 1.028 & 1 & 1.027 & 1.024 \\ 
-%	Gauss\_7\_10\_10 & 1.074 & 1.56 & 1.024 & 1.003 & 1.019 & 1.005 & 1.02 & 1.02 \\ 
-%	Gauss\_7\_50\_0 & 1.104 & 1.321 & 1.029 & 1.01 & 1.031 & 1.011 & 1.027 & 1.026 \\ 
-%	Gauss\_7\_50\_5 & 1.132 & 1.326 & 1.033 & 1.013 & 1.031 & 1.013 & 1.026 & 1.036 \\ 
-%	Gauss\_7\_50\_10 & 1.107 & 1.345 & 1.027 & 1.014 & 1.025 & 1.012 & 1.031 & 1.026 \\ \hline
-%\end{tabular}
-%
-%\end{center}
-%
-%\caption{Approximation Constants For Gaussian Datasets}
-%\end{figure*}
-%
-%\subsection{P-Median}
-%
-%\begin{figure*}[h]
-%\begin{center}
-%\begin{tabular}{ | l | l | l | l | }
-%\hline
-%	Dataset & Optimal Objective & $k$ & $n$ \\ \hline
-%	pmed1 & 5819 & 5 & 100 \\ 
-%	pmed2 & 4093 & 10 & 100 \\ 
-%	pmed3 & 4250 & 10 & 100 \\ 
-%	pmed4 & 3034 & 20 & 100 \\
-%	pmed5 & 1355 & 33 & 100 \\ 
-%	pmed6 & 7824 & 5 & 200 \\ 
-%	pmed7 & 5631 & 10 & 200 \\ 
-%	pmed8 & 4445 & 20 & 200 \\ 
-%	pmed9 & 2734 & 40 & 200 \\ 
-%	pmed10 & 1255 & 67 & 200 \\ 
-%	pmed11 & 7696 & 5 & 300 \\ 
-%	pmed12 & 6634 & 10 & 300 \\ 
-%	pmed13 & 4374 & 30 & 300 \\ 
-%	pmed14 & 2968 & 60 & 300 \\ 
-%	pmed15 & 1729 & 100 & 300 \\
-%	pmed16 & 8162 & 5 & 400 \\ 
-%	pmed17 & 6999 & 10 & 400 \\
-%	pmed18 & 4809 & 40 & 400 \\ 
-%	pmed19 & 2845 & 80 & 400 \\ 
-%	pmed20 & 1789 & 133 & 400 \\ 
-%	pmed21 & 9138 & 5 & 500 \\ 
-%	pmed22 & 8579 & 10 & 500 \\ 
-%	pmed23 & 4619 & 50 & 500 \\ 
-%	pmed24 & 2961 & 100 & 500 \\ 
-%	pmed25 & 1828 & 167 & 500 \\ 
-%	pmed26 & 9917 & 5 & 600 \\ 
-%	pmed27 & 8307 & 10 & 600 \\ 
-%	pmed28 & 4498 & 60 & 600 \\ 
-%	pmed29 & 3033 & 120 & 600 \\
-%	pmed30 & 1989 & 200 & 600 \\ 
-%	pmed31 & 10086 & 5 & 700 \\ 
-%	pmed32 & 9297 & 10 & 700 \\ 
-%	pmed33 & 4700 & 60 & 700 \\ 
-%	pmed34 & 3013 & 140 & 700 \\ 
-%	pmed35 & 10400 & 5 & 800 \\ 
-%	pmed36 & 9934 & 10 & 800 \\ 
-%	pmed37 & 5057 & 80 & 800 \\ 
-%	pmed38 & 11060 & 5 & 900 \\ 
-%	pmed39 & 9423 & 10 & 900 \\ 
-%	pmed40 & 5128 & 90 & 900 \\ \hline
-%\end{tabular}
-%
-%
-%\end{center}
-%
-%\caption{Details for P-Median Datasets}
-%\end{figure*}
-%
-%\begin{figure*}[h]
-%\begin{center}
-%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
-%\hline
-%	Dataset & Time & Time & Time & Time & Time & Time & Time \\ \hline
-%	&k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-Swaps + & 3-Swap   \\ \hline
-%	pmed1 & 0.0032 & 5 & 0.1 & 0.1 & 0.2 & 0.2 & 0.3 \\ 
-%	pmed2 & 0.0069 & 6.4 & 0.2 & 0.2 & 0.3 & 0.3 & 0.4 \\ 
-%	pmed3 & 0.0067 & 7.5 & 0.2 & 0.2 & 0.3 & 0.3 & 0.4 \\ 
-%	pmed4 & 0.0125 & 8 & 0.3 & 0.4 & 0.5 & 0.6 & 0.6 \\ 
-%	pmed5 & 0.0381 & 8.3 & 0.5 & 0.6 & 0.7 & 0.7 & 0.9 \\ 
-%	pmed6 & 0.0065 & 19.3 & 0.3 & 0.4 & 0.5 & 0.6 & 0.8 \\ 
-%	pmed7 & 0.0117 & 23.9 & 0.7 & 0.9 & 0.9 & 1.3 & 1.5 \\ 
-%	pmed8 & 0.027 & 27.2 & 1.4 & 1.8 & 2.2 & 2.6 & 3.1 \\ 
-%	pmed9 & 0.0642 & 29 & 3.4 & 3.5 & 3.3 & 4.8 & 4.7 \\ 
-%	pmed10 & 0.7329 & 39.8 & 4.4 & 5.8 & 6.2 & 7.3 & 8.2 \\ 
-%	pmed11 & 0.0124 & 42 & 0.6 & 0.8 & 1.4 & 1.2 & 2.2 \\ 
-%	pmed12 & 0.0186 & 45.7 & 1.7 & 2 & 2.3 & 2.4 & 4.3 \\ 
-%	pmed13 & 0.0685 & 69.1 & 6.6 & 5.3 & 7 & 8.4 & 11 \\ 
-%	pmed14 & 0.2238 & 76.3 & 13.6 & 14.2 & 17.1 & 17.4 & 25 \\ 
-%	pmed15 & 1.0969 & 100.2 & 18.4 & 19.7 & 23.1 & 28.7 & 35.5 \\ 
-%	pmed16 & 0.0251 & 87.1 & 1.3 & 1.2 & 2.3 & 2.4 & 3.8 \\ 
-%	pmed17 & 0.0581 & 97.1 & 3.5 & 2.5 & 3.9 & 4.8 & 6.8 \\ 
-%	pmed18 & 0.2596 & 134.7 & 15.3 & 16.5 & 23.1 & 23.3 & 31.3 \\ 
-%	pmed19 & 1.0193 & 153.5 & 34.9 & 34.1 & 40.3 & 47.9 & 57.8 \\ 
-%	pmed20 & 1.7947 & 175 & 52.2 & 45 & 68.6 & 68.2 & 79.3 \\ 
-%	pmed21 & 0.0327 & 109.4 & 1.8 & 2.1 & 4.1 & 4.2 & 6.1 \\ 
-%	pmed22 & 0.0786 & 113.1 & 4.3 & 5.6 & 8.3 & 7.5 & 12.7 \\ 
-%	pmed23 & 0.5123 & 199.8 & 35.9 & 34.8 & 46.9 & 52.3 & 62.2 \\ 
-%	pmed24 & 1.7861 & 245.9 & 66.6 & 90.3 & 99.2 & 88.5 & 125.5 \\ 
-%	pmed25 & 3.1433 & 451.9 & 103 & 118 & 126.8 & 146.4 & 138.1 \\ 
-%	pmed26 & 0.0503 & 159.8 & 2.8 & 2.4 & 4.7 & 5.6 & 6.9 \\ 
-%	pmed27 & 0.1265 & 183.4 & 7.1 & 6.1 & 11.7 & 10.7 & 13 \\ 
-%	pmed28 & 0.8738 & 263.3 & 60.3 & 71.3 & 95.6 & 98.5 & 89.3 \\ 
-%	pmed29 & 2.8418 & 424.3 & 152.6 & 146.8 & 196.5 & 172.8 & 167 \\ 
-%	pmed30 & 5.2891 & 571 & 175 & 207.5 & 254.2 & 249.7 & 255.7 \\ 
-%	pmed31 & 0.0728 & 225.6 & 4.9 & 4.4 & 8.2 & 6.6 & 9.2 \\ 
-%	pmed32 & 0.1528 & 249.5 & 11 & 7.8 & 14.3 & 12.9 & 19 \\ 
-%	pmed33 & 1.4723 & 440.1 & 112.8 & 91.9 & 143.9 & 158.5 & 154.1 \\ 
-%	pmed34 & 3.6669 & 591.7 & 234.8 & 217.8 & 233.1 & 287.8 & 256.8 \\ 
-%	pmed35 & 0.1133 & 298.3 & 4.8 & 5.9 & 7.5 & 9.3 & 11.9 \\ 
-%	pmed36 & 0.2221 & 259.8 & 13 & 11.4 & 15.9 & 24 & 23.6 \\ 
-%	pmed37 & 2.4409 & 571 & 190.5 & 149.7 & 172.8 & 282.3 & 224.7 \\ 
-%	pmed38 & 0.1326 & 349.9 & 7.1 & 6.5 & 9.8 & 12.2 & 14.5 \\ 
-%	pmed39 & 0.2676 & 419.8 & 16.5 & 13.6 & 22.6 & 25.5 & 29.4 \\ 
-%	pmed40 & 3.5269 & 849.9 & 251.3 & 264.9 & 250.4 & 391.7 & 350.2 \\ \hline
-%\end{tabular}
-%
-%
-%\end{center}
-%
-%\caption{Running Time for P-Median Datasets}
-%\end{figure*}
-%
-%\begin{figure*}[h]
-%\begin{center}
-%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
-%\hline
-%	Dataset & Approx & Approx & Approx & Approx & Approx & Approx & Approx \\ \hline
-%	  & k-medoids & Supermodular & PAM & 1-Swap + & 2-Swap & 2-Swaps+ & 3-Swap \\ \hline
-%	pmed1 & 1.113 & 1.409 & 1.03 & 1.013 & 1.029 & 1.005 & 1.018 \\ 
-%	pmed2 & 1.158 & 1.486 & 1.044 & 1.011 & 1.038 & 1.011 & 1.044 \\ 
-%	pmed3 & 1.161 & 1.442 & 1.053 & 1.01 & 1.037 & 1.01 & 1.046 \\ 
-%	pmed4 & 1.16 & 1.5 & 1.07 & 1.028 & 1.061 & 1.019 & 1.056 \\ 
-%	pmed5 & 1.373 & 1.681 & 1.06 & 1.026 & 1.071 & 1.033 & 1.051 \\ 
-%	pmed6 & 1.09 & 1.286 & 1.015 & 1.008 & 1.016 & 1.008 & 1.012 \\ 
-%	pmed7 & 1.126 & 1.367 & 1.027 & 1.011 & 1.023 & 1.01 & 1.025 \\ 
-%	pmed8 & 1.179 & 1.489 & 1.064 & 1.017 & 1.051 & 1.021 & 1.048 \\ 
-%	pmed9 & 1.249 & 1.559 & 1.065 & 1.031 & 1.063 & 1.024 & 1.063 \\ 
-%	pmed10 & 1.462 & 1.839 & 1.076 & 1.037 & 1.076 & 1.039 & 1.076 \\ 
-%	pmed11 & 1.071 & 1.369 & 1.014 & 1.004 & 1.008 & 1.005 & 1.011 \\ 
-%	pmed12 & 1.144 & 1.368 & 1.025 & 1.007 & 1.028 & 1.007 & 1.025 \\ 
-%	pmed13 & 1.209 & 1.446 & 1.047 & 1.019 & 1.05 & 1.021 & 1.05 \\ 
-%	pmed14 & 1.243 & 1.504 & 1.056 & 1.026 & 1.053 & 1.025 & 1.049 \\ 
-%	pmed15 & 1.61 & 1.658 & 1.067 & 1.032 & 1.068 & 1.03 & 1.06 \\ 
-%	pmed16 & 1.062 & 1.333 & 1.012 & 1.004 & 1.012 & 1.005 & 1.008 \\ 
-%	pmed17 & 1.149 & 1.359 & 1.023 & 1.014 & 1.028 & 1.013 & 1.025 \\ 
-%	pmed18 & 1.169 & 1.414 & 1.044 & 1.016 & 1.033 & 1.016 & 1.04 \\ 
-%	pmed19 & 1.261 & 1.537 & 1.056 & 1.022 & 1.059 & 1.02 & 1.055 \\ 
-%	pmed20 & 1.501 & 1.73 & 1.079 & 1.032 & 1.076 & 1.027 & 1.073 \\ 
-%	pmed21 & 1.096 & 1.384 & 1.027 & 1.005 & 1.015 & 1.003 & 1.021 \\ 
-%	pmed22 & 1.128 & 1.371 & 1.026 & 1.015 & 1.018 & 1.007 & 1.021 \\ 
-%	pmed23 & 1.151 & 1.42 & 1.041 & 1.019 & 1.045 & 1.016 & 1.041 \\ 
-%	pmed24 & 1.271 & 1.519 & 1.058 & 1.019 & 1.055 & 1.019 & 1.052 \\ 
-%	pmed25 & 1.449 & 1.711 & 1.07 & 1.033 & 1.077 & 1.035 & 1.073 \\ 
-%	pmed26 & 1.073 & 1.317 & 1.018 & 1.008 & 1.012 & 1.003 & 1.014 \\ 
-%	pmed27 & 1.113 & 1.36 & 1.017 & 1.008 & 1.018 & 1.009 & 1.017 \\ 
-%	pmed28 & 1.168 & 1.386 & 1.046 & 1.011 & 1.037 & 1.015 & 1.04 \\ 
-%	pmed29 & 1.263 & 1.497 & 1.055 & 1.026 & 1.057 & 1.024 & 1.057 \\ 
-%	pmed30 & 1.456 & 1.643 & 1.067 & 1.033 & 1.064 & 1.032 & 1.061 \\ 
-%	pmed31 & 1.082 & 1.336 & 1.01 & 1.003 & 1.006 & 1.001 & 1.013 \\ 
-%	pmed32 & 1.12 & 1.344 & 1.02 & 1.011 & 1.018 & 1.011 & 1.017 \\ 
-%	pmed33 & 1.182 & 1.432 & 1.043 & 1.018 & 1.044 & 1.017 & 1.04 \\ 
-%	pmed34 & 1.323 & 1.539 & 1.061 & 1.024 & 1.059 & 1.021 & 1.056 \\ 
-%	pmed35 & 1.083 & 1.292 & 1.011 & 1.005 & 1.015 & 1.007 & 1.01 \\ 
-%	pmed36 & 1.138 & 1.334 & 1.016 & 1.009 & 1.016 & 1.011 & 1.016 \\ 
-%	pmed37 & 1.157 & 1.439 & 1.042 & 1.02 & 1.05 & 1.017 & 1.047 \\ 
-%	pmed38 & 1.063 & 1.288 & 1.007 & 1.009 & 1.008 & 1.006 & 1.01 \\ 
-%	pmed39 & 1.127 & 1.42 & 1.02 & 1.007 & 1.014 & 1.007 & 1.015 \\ 
-%	pmed40 & 1.187 & 1.423 & 1.046 & 1.017 & 1.043 & 1.019 & 1.039 \\ \hline
-%\end{tabular}
-%
-%
-%\end{center}
-%
-%\caption{Approximation Constants for P-Median Datasets}
-%\end{figure*}
+\bibliographystyle{IEEEtran}
+\bibliography{submissionBib}
 
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was
-% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
-% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
-% slightly modified from the 2009 version by Kiri Wagstaff and 
-% Sam Roweis's 2008 version, which is slightly modified from 
-% Prasad Tadepalli's 2007 version which is a lightly 
-% changed version of the previous year's version by Andrew Moore, 
-% which was in turn edited from those of Kristian Kersting and 
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2015-05-22 src/LP_charikar.py

@@ -2,6 +2,10 @@ import numpy as np
 from gurobipy import *
 import LoadData
 from sklearn.metrics.pairwise import pairwise_distances
+import timeit
+
+from itertools import izip
+argmin = lambda array: min(izip(array, xrange(len(array))))[1]
 
 def assign_points_to_clusters(medoids, distances):
     distances_to_medoids = distances[:,medoids]
@@ -66,46 +70,66 @@ def bundle(C_sub, x, y, distances):
             if (x[i,j].X > 0) and (distances[i][j] <= 1.5*Ri) and \
                (y_assign[j] == n):
                 y_assign[j] = i
-                vols[j] += y[i].X
-                U[j].append(i)
-                
+                vols[i] += y[j].X
+                U[i].append(j)
+
     return y_assign, U, vols
 
 # Greedily match clients based on distance
 def match(C_sub, distances):
     unmatched = set(C_sub)
     matching = {}
-
+    m = len(C_sub)
+    sort_dists = [ np.argsort([distances[C_sub[i]][C_sub[j]] for j in range(i+1, m)]) \
+                   for i in range(m)]
+    min_index = [0 for _ in range(m-1)]
+    mins = [distances[C_sub[i]][C_sub[sort_dists[i][0]+i+1]] for i in range(m-1)]
+    mins.append(float('inf'))
+    
     while len(unmatched) > 0:
         if len(unmatched) == 1:
             j = unmatched.pop()
             matching[j] = j
         else:
             # Find closest unmatched pair
-            min_dist = float('inf')
-            min_pair = None
-            for i in unmatched:
-                for j in unmatched:
-                    if i != j:
-                        if distances[i][j] < min_dist:
-                            min_dist = distances[i][j]
-                            min_pair = [i,j]
+            j = argmin(mins)
+            k = sort_dists[j][min_index[j]]+j+1
+            x = C_sub[j]
+            y = C_sub[k]
 
             # match that pair
-            matching[min_pair[0]] = min_pair[1]
-            matching[min_pair[1]] = min_pair[0]
-            unmatched.remove(min_pair[0])
-            unmatched.remove(min_pair[1])
+            matching[x] = y
+            matching[y] = x
+            unmatched.remove(x)
+            unmatched.remove(y)
+            mins[j] = float('inf')
+            mins[k] = float('inf')
 
+            # Update mins
+            for i in range(m-1):
+                if (C_sub[i] in unmatched):
+                    while True:
[…]

### 2015-05-22 src/hierarchical_approx.py

@@ -2,6 +2,7 @@ import numpy as np
 from sklearn import metrics
 import kmedoids
 import LP_greedy
+import LP_charikar
 import multiswaps
 import math
 import LoadData
@@ -30,15 +31,21 @@ def bucketed_solns(distances, beta, beta_0, alg, logn = False):
         kvals = [i for i in range(1,n)]
 
     if alg[0] == "multiswaps-plus":
-        all_centers, all_costs = multiswaps.cluster_hier(distances, kvals, alg[1])
+        all_centers, all_costs = multiswaps.cluster_hier(distances, kvals, 1)
     elif alg[0] == "k-medoids":
         all_centers, all_costs = kmedoids.cluster_hier(distances, kvals, 1000)
     elif alg[0] == "LP-greedy-rand":
         kvals.reverse()
         all_centers, all_costs = LP_greedy.cluster_hier(distances, kvals, 1)
-    else:
+    elif alg[0] == "LP-greedy":
         kvals.reverse()
         all_centers, all_costs = LP_greedy.cluster_hier(distances, kvals, 0)
+    elif alg[0] == "LP-charikar":
+        kvals.reverse()
+        all_centers, all_costs = LP_charikar.cluster_hier(distances, kvals, 1)
+    else:
+        kvals.reverse()
+        all_centers, all_costs = LP_charikar.cluster_hier(distances, kvals, 1000)
             
 
     min_cost = (1.0/np.min([all_costs[i] for i in range(len(all_costs)) if all_costs[i] != 0]))

### 2015-05-22 src/multiswaps.py

@@ -59,7 +59,7 @@ def search_for_swap(medoids, obj, distances, FDict, p=2, delta=0):
         #            new_objective = objective(new_medoids, distances, FDict)
         #            if new_objective < (1.0-delta)*prev_objective:
         #                improvement = True
-        #                #print("Improvement: %f" % (prev_objective - new_objective))
+        #                #print"Improvement: %f" % (prev_objective - new_objective))
         #                return new_medoids, new_objective, improvement
         #else:
         #print q
@@ -123,7 +123,6 @@ def cluster2(distances, k, p, warm_start = None, epsilon=0.1):
 
 def cluster_hier(distances, kvals, p, warm_start=None, epsilon=0.1):
     m = len(kvals)
-    print kvals
     n = distances.shape[0]
     all_costs = [0 for i in range(m)]
     all_centers = [[0 for j in range(kvals[i])] for i in range(m)]

### 2015-05-29 icdm/submission2.tex
New version for icdm

--- 2015-05-22 icdm/submission.tex
@@
 \begin{abstract}
-Partitional clustering algorithms can organize large data into smaller, more informative groups.
-Hierarchical clustering, in addition, offers a more global view of the data structure by producing a hierarchy of clusters starting from individual data points and merging towards the cluster of all data.
+Hierarchical clustering organizes large data into smaller, more informative groups and offers a global view of the data structure by producing a hierarchy of clusters starting from individual data points and merging towards the cluster of all data.
 Typically, hierarchical clustering is done by greedily either merging or dividing clusters, called agglomerative and divisive clustering, respectively.
-Lin et al.~\cite{Lin} proposed a nesting framework that produces a hierarchy from partitional clusterings for every possible number of clusters.
-This algorithm was introduced as an approximation algorithm for the hierarchical $k$-median problem and guarantees the resulting hierarchical layers are not too far away from the original clusters.
+Lin et al.~\cite{Lin} proposed a nesting framework that produces a hierarchal clustering from partitional clusterings for every possible number of clusters.
+This algorithm was introduced as an approximation algorithm for the hierarchical $k$-median problem and guarantees the resulting hierarchical layers are a limited distance from the original partitional clusters.
 However, this algorithm has not been previously placed in a clustering context.
-We test this algorithm with various methods for producing the inputted partitional clusterings versus the traditional hierarchical clustering methods and show both the approximation ratio and FScore measure improve significantly.
-We also propose a heuristic for this nesting framework which improves the running time and only suffers a slight decrease in approximation and FScore.
+We test this algorithm using various methods for producing the partitional clusterings against the traditional hierarchical clustering methods and show both the approximation ratio and FScore measure improve significantly.
+We also propose a heuristic for this nesting framework which improves the run time and only suffers a slight decrease in approximation and FScore.
 \end{abstract}
@@
 The retailer could then also offer specialized discounts to subgroups in each region based on each cluster's shopping records.
+Examples of both types of clustering are given in Fig.~\ref{fig_part}-\ref{fig_hier}.
 
-Clustering algorithms date back more than 50 years, and there have been many variations introduced since \cite{Jain}.
+Clustering algorithms date back more than 50 years, and there have been many variations introduced since then \cite{Jain}.
 However, clustering algorithms are often formulated as NP-hard problems with non-convex objective functions, which are inherently hard to solve \cite{Jain}.
@@
 The clustering is obtained from the set of centers $S$ by assigning each data point to its closest center.
-An $\alpha$-approximation algorithm for this problem produces a set $S$ such that $f(S) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
+An $\alpha$-approximation algorithm for this problem produces a set $S$ in polynomial time such that $f(S) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
 
-[ADD FIGURES SHOWING BOTH TYPES OF CLUSTERING]
+\begin{figure*}[!t]
+\centerline{\subfloat[Partitional Clustering]{\includegraphics[width =1.75in]{partitional_clustering}
+\label{fig_part}}
+\hfil
+\subfloat[Hierarchical Clustering]{\includegraphics[width=2.5in]{hier_clustering}
+\label{fig_second_case}}}
+\caption{A partitional clustering for $k=3$ and a hierarchical clustering of the data where the clustering on the third level down (represented by the red line) corresponds to the partitional clustering on the left.}
+\label{fig_hier}
+\end{figure*}
 
@@
 Then, for any $i \in S$, $a^{-1}(i)$ is the cluster centered at $i$.
-We define $\mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i))$.
+We define $f(S,a) := \sum_{i \in X} d(i, a(i))$.
 This is the same objective function as above except the assignment is given rather than implied.
@@
 \end{enumerate}
-A solution to the hierarchical $k$-median problem is a nested chain of solutions and the hierarchical clustering given by looking at the clusters on each level.
+A solution to the hierarchical $k$-median problem is a nested chain of solutions, and the hierarchical clustering it generates is found by looking at the clusters given by each $(S_i, a_i)$.
 However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
-An $\alpha$-approximation algorithm for this problem ensures that for all $k = 1, \ldots, |X|$ there exists $(S_i, a_i)$ in our chain of solutions such that $|S_i| \leq k$ and $\mathrm{cost}(S_i, a_i) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
+An $\alpha$-approximation algorithm for this problem generates a nested chain of solutions in polynomial time such that for all $k = 1, \ldots, |X|$ there exists $(S_i, a_i)$ in the chain such that $|S_i| \leq k$ and $f(S_i, a_i) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
 
-While some flat clustering methods use the $k$-median objective function to inform their decisions and sometimes offer an approximation guarantee, popular methods for hierarchical clustering do not and rely on greedy algorithms which have no guarantee on the solution produced with respect to the objective function but are typically fast and easy to implement.
-This paper will look at how using this objective function can vastly improve the hierarchical clusterings produced.
-In particular, we will show how an approximation framework for this problem which takes in flat clusterings for all values of $k$ compares in practice to some of the most widely used clustering algorithms both in the quality of the solution and in computation time.
-The algorithms we will use to produce these flat clusterings will include both approximation algorithms for the $k$-median problem, which will offer an overall approximation guarantee for the hierarchical $k$-median problem, and some heuristic algorithms that have been shown to perform well in practice.
+While some partitional clustering methods use the $k$-median objective function to inform their decisions, popular methods for hierarchical clustering do not and rely on greedy algorithms which have no guarantee on the solution produced with respect to the objective function but are typically fast and easy to implement.
+This paper looks at how using this objective function can vastly improve the hierarchical clusterings produced.
+In particular, we show how an approximation framework for this problem, which takes in partitional clusterings for all values of $k$, compares in practice to some of the most widely used clustering algorithms, both in the quality of the solution and in computation time.
+The algorithms we use to produce these partitional clusterings include both approximation algorithms for the $k$-median problem, which offer an overall approximation guarantee for the hierarchical $k$-median problem, and some heuristic algorithms that have been shown to perform well in practice.
 
@@
 
-Arguably the most common partitional clustering method is $k$-means, in which we want to choose $k$ cluster centers points in space to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
+\subsection{Partitional Clustering}
+Arguably the most common partitional clustering method is $k$-means, which chooses $k$ cluster centers points in space to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
 $k$-means is usually solved using Lloyd's algorithm (also referred to simply as the $k$-means algorithm), which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}.
@@
 The generalization of this algorithm which considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
-Another common heuristic (which we shall call $k$-medoids throughout this paper), which is more computationally efficient than PAM \cite{Park}, is a variation on Lloyd's $k$-means algorithm.
+Another common heuristic (called $k$-medoids throughout this paper), which is more computationally efficient than PAM \cite{Park}, is a variation on Lloyd's $k$-means algorithm.
 Starting with randomly chosen centers, we assign each data point to its closest center, and then update the center within each cluster \cite{Park}.
@@
 
-The $k$-median problem is also a fundamental problem in approximation algorithms where is often presented in the context of opening service centers or factories to serve clients.
+The $k$-median problem is also a fundamental problem in approximation algorithms where it is often presented in the context of opening service centers or factories to serve clients.
 However, these algorithms are often not implemented for clustering despite their improved guarantees on the solution.
@@
 
-Approximation algorithms for hierarchical clustering have also been studied, where we aim to identify a hierarchy of clusters so as to minimize the ratio of the objective value on every level in the hierarchy of size $k$ to the optimal objective value of the flat partitional clustering problem with $k$ clusters.
+Approximation algorithms for hierarchical clustering have also been studied, where we aim to identify a hierarchy of clusters so as to minimize the ratio of the objective value on every level in the hierarchy of size $k$ to the optimal objective value of the partitional clustering problem with $k$ clusters.
 Note that it may not always be possible to find a solution with approximation ratio 1.
@@
 In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to greedily construct a hierarchical clustering that increases the objective value by at most a factor of $20.71$ (or $10.03$ when randomized) at each level.
-Using the local $p$-swap $(3+\frac{2}{p})$-approximation algorithm mentioned above as the black-box $k$-median solver performed well in experimental tests \cite{Nagarajan}.
+Using the local $p$-swap $(3+\frac{2}{p})$-approximation algorithm mentioned above as the black-box $k$-median solver performed well in experimental hierarchical $k$-median test problems \cite{Nagarajan}.
 While the authors tested the variants of this algorithm for different $k$-median approximation algorithms, it has not been tested against traditional clustering algorithms or on any labeled data sets.
 
-Several algorithms have been built upon agglomerative clustering that we choose not to test our proposed algorithms against.
+Several algorithms have been built upon agglomerative clustering that we choose not to compare with our proposed algorithms.
 Divisive clustering, similar to agglomerative, starts with the data in one cluster and repeatedly splits a cluster on each level.
 This method has not been studied nearly as much as agglomerative clustering \cite{ESL} but has had relative success in classifying document datasets \cite{Zhao}.
-Algorithms such as the CURE, ROCK, and Chameleon \cite{Guha_CURE, Guha_ROCK, Karypis} incorporate non-spherical cluster shapes and cluster density, but do not relate as much to the above approximation algorithm in objective.
-We also choose not compare the approximation framework with Bayesian hierarchical methods for the same reasons since these methods require many more parameters and add additional prior information that our algorithms do not use.
+Algorithms such as the CURE, ROCK, and Chameleon \cite{Guha_CURE, Guha_ROCK, Karypis} incorporate non-spherical cluster shapes and cluster density, but do not relate as much to the above approximation algorithm in their objective.
+We also choose not compare the approximation framework with Bayesian hierarchical methods since these methods require many more parameters and add additional prior information that our algorithms do not use.
 
@@
 
-Recall the hierarchical $k$-median problem: given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
+\subsection{Hierarchical Algorithm}
+
+Recall the hierarchical $k$-median problem: given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$, we represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
  A solution to the hierarchical $k$-median problem is a chain of nested solutions $(S_1, a_1) \subseteq (S_2, a_2) \subseteq \ldots \subseteq (S_l, a_l)$.
@@
 Consider any $k$-median algorithm $\mathcal{A}$.
-After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce an augmentation step that will nest these clusterings into the hierarchical framework and only increase costs by a bounded factor.
+After running $\mathcal{A}$ for all potential values of $k$, Lin et al.~\cite{Lin} introduce an augmentation step that nests these clusterings into the hierarchical framework and only increase costs by a bounded factor.
 
-First, set $i=1$, $S_0 = \emptyset$, $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
+First, set $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
 Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
 The cost of these solutions is determined by assigning each point to the nearest cluster center and is scaled so that the minimum cost for $k<n$ is at least one.
-We then order these solutions according to their cost into buckets of the form $[0,0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
+We then bucket these solutions by cost into buckets of the form $[0,0], (\beta_0, \beta_0 \beta], (\beta_0 \beta, \beta_0 \beta^2], \ldots $.
 
@@
 We then inductively build our nested solutions.
-Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we define $S_i$ by iteratively adding the closest point $j \in S_{i-1}$ to $l$ for all $l \in \overline{V}_i$.
+Set $S_1 = \overline{V}_1$ and $a_1$ to be the assignment that assigns each point to the closest point in $S_1$.
+Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we construct $S_i$ by  adding the closest point $j \in S_{i-1}$ to $l$ for each $l \in \overline{V}_i$.
 Note that $|S_{i}| \leq |\overline{V}_i|$.
 
-We then define an assignment $a_i$.
+\begin{figure*}[!t]
+\centering
+\subfloat[Nested solution $(S_{i-1}, a_{i-1})$ ]{\includegraphics[width =1.75in]{Si-1}
+\label{fig_Si-1}} \hspace{15mm}
+\subfloat[Next set $\bar{V}_i$]{\includegraphics[width=1.75in]{Vi}
+\label{fig_Vi}} \\
+\subfloat[Next nested centers $S_i$]{\includegraphics[width=1.75in]{Si}
+\label{fig_Si}}  \hspace{15mm}
+\subfloat[Next nested solution $(S_i, a_i)$]{\includegraphics[width=1.75in]{Siai}
+\label{fig_Siai}}
+\caption{An example of nesting $\bar{V}_i$ with $(S_{i-1}, a_{i-1})$.
+Dark  nodes represent chosen centers and edges represent assignments to that center's cluster.
+First, node 2 chooses node from 1 from $S_{i-1}$ as its closest center and node 5 chooses node 4.
+Then, each cluster in $(S_{i-1}, a_{i-1})$ is assigned to the cheapest (lowest sum of distances) center.}
+\label{fig_sim}
+\end{figure*}
+
+We then need to define an assignment $a_i$.
 For each $j \in X$ such that $a_{i-1}(j) \in S_{i}$, we set $a_i(j) = a_{i-1}(j)$.
@@
 Using this assignment, $(a_i, S_i)$ is nested in $(a_{i-1}, S_{i-1})$.
+An example of this nesting is given in Fig.~\ref{fig_Si-1}-\ref{fig_Siai}.
 
-We continue this nesting until $|S_i| = 1$, reaching the top level of the hierarchy.
-After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantee that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq n-k+1 \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).
+We continue this nesting until $|S_i| = 1$, reaching the top level of the hierarchy where all data points are in the same cluster.
+After forming this hierarchical clustering, for any $k \in 1, \ldots, n$ \cite{Lin} guarantees that the cost of $S_i$ is at most $20.71$ times the cost of $V_k$ where $|S_{i-1}| \geq |V_k| \geq |S_{i}|$ (in the randomized case the expected cost of $S_i$ is at most $10.03$ times the cost of $V_k$).
 
-\subsection{Flat Clustering Algorithms}
-This nesting framework allows us to choose the flat clustering algorithm $\mathcal{A}$.
+\subsection{Partitional Clustering Algorithms}
+This nesting framework allows us to choose the partitional clustering algorithm $\mathcal{A}$.
 As mentioned above, PAM and $k$-medoids are two potential algorithms.
-PAM has the added benefit of being a 5-approximation algorithm for the flat $k$-median problem which yields to an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
+PAM has the added benefit of being a 5-approximation algorithm for the partitional $k$-median problem which yields to an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
 We also considered the size of the local swaps in PAM to improve the approximation ratio but found that the approximation ratio of constructed solutions for given values of $k$ did not significantly improve while the computation time increased greatly.
@@
 \sum_{j \in X} a_{i,j}  = 1 & \forall i \in X, \\
-a_{i,j}   \leq s_i & \forall i, j \in X.
+a_{i,j}   \leq s_j & \forall i, j \in X.
 \end{eqnarray}
@@
 & \sum_{j \in X} a_{i,j}  = 1 \quad \forall i \in X \\
-& a_{i,j}   \leq s_i \quad \forall i, j \in X \\
+& a_{i,j}   \leq s_j \quad \forall i, j \in X \\
 & s_i \in \{0,1\} \quad \forall i \in X \\
@@
 
-Interestingly, the solution to the linear programming is often integer valued [NEED CITATION].
-Based on this fact, we proposed a flat clustering algorithm that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values.
-We refer to this algorithm as LP-greedy from now on.
-Charikar and Li \cite{charikar} introduce a 3.25-approximation algorithm that uses this LP solution to smartly construct a set of centers.
+Interestingly, the solution to the linear programming is often integer valued \cite{Nagarajan}.
+Based on this fact, we propose two partitional clustering algorithms: one that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values and one that solves the LP relaxation and constructs $S$ by choosing $i$ to be in $S$ with probability $s_i$, sampling for each $i \in X$ until $k$ centers are chosen.
+We refer to these algorithms as LP-greedy and LP-greedy-rand, respectively.
+Charikar and Li \cite{Charikar} introduce a 3.25-approximation algorithm that uses this LP solution to smartly construct a set of centers.
 We refer to this algorithm as LP-Charikar.
-We choose to implement this LP algorithm rather than the LP-based approximation algorithm with best approximation ratio because the algorithm is simpler to implement (and will most likely have a lower running time).
+We choose to implement this LP algorithm rather than the LP-based approximation algorithm with best approximation ratio because the algorithm is simpler to implement (and we expect it to have a lower running time).
 
 The algorithm works by first choosing a subset of data points $C \subseteq X$ to represent the demand.
-For each data point $i$, let $d_{av}(i) = \sum_{j \in X : a_{i,j} > 0} s_i \cdot d(i,j)$.
+For each data point $i$, let $d_{av}(i) = \sum_{j \in X : a_{i,j} > 0} s_j \cdot d(i,j)$.
 We start with $C = \emptyset$.
@@
 These points then choose a bundle of points $U_j$ that represent potential centers they would like to be assigned to.
-In particular, for each point $j \in C$, we look at all points $i \in X$ such that $0 < d(i,j) < 1.5 R_j$, where $R_j$ is half the distance from $j$ to it's nearest point in $C$, and bundle these points into $U_j$ (if a point is in multiple bundles you choose one arbitrarily).
-Now these bundles of points will have a sum of $s_i$ values at least one half.
+In particular, for each point $j \in C$, we look at all points $i \in X$ such that $0 < d(i,j) < 1.5 R_j$, where $R_j$ is half the distance from $j$ to its nearest point in $C$, and bundle these points into $U_j$ (if a point is in multiple bundles, choose one arbitrarily).
+Now these bundles of points have a sum of $s_i$ values of at least one half.
+For each $U_j$, we define the volume of $U_j$ to be $\mathrm{vol}(U_j) = \sum_{i \in U_j} s_i$.
 Lastly, these bundles are matched together greedily by distance.
-For each $U_j$, we let $\mathrm{vol}(U_j) = \sum_{i \in U_j} s_i$.
 
-Now that the algorithm has grouped together points so the volume is at least 1, the algorithm starts to choosing points from these various groups through sampling.
-Here, opening $U_j$ refers to selecting a point $i \in U_j$ with probabilities relative to the $s_i$ values and adding that point to $S$.
+Now that the algorithm has grouped and matched together points so the volume is at least 1, the algorithm starts choosing points from these various groups through sampling.
+Here, opening $U_j$ refers to selecting a point $i \in U_j$ with probabilities proportional to the $s_i$ values and adding that point to $S$.
 \begin{enumerate}[\IEEEsetlabelwidth{3)}]
@@
 \end{enumerate}
-This sampling process returns a set of centers of size $k$  in expectation. For our purposes, we sample until we obtain a solution with exactly $k$ centers.
+This sampling process returns a set of centers of size $k$  in expectation and produces a set $S$ such that $\mathbb{E}[f(S)] \leq 3.25 \cdot f(O)$ for all $O \subseteq X$ such that $|O| = k$.
+For our purposes, we sample until we obtain a solution with exactly $k$ centers.
 
 \section{Implementation and Results}
-We implemented the nesting hierarchical framework above in Python 2.7, offering options for both the randomized and deterministic versions.
-For the black-box $k$-median algorithm, we implemented $k$-medoids, PAM, LP-greedy, and LP-charikar in Python using Gurobi 6.0.3 to solve the linear programming instances.
-We compared the results to hierarchical agglomerative clustering with both average linkage and Ward's linkage (implemented in the scipy package in Python).
 
-To avoid cycling, we limited the number of iterations in $k$-medoids to 1000.
-For PAM, we also needed to limit the number of swaps to achieve a reasonable runtime.
-Instead of searching over all potential swaps, we sample at most $n$ potential ones.
-We also ensure that some minimum level progress is achieved by ensuring the algorithm improves by at least a factor of 0.9 on each iteration.
-At the end of PAM, we run 10 iterations of $k$-medoids to ensure the algorithm finds a local optimum.
-These choices were made by looking at the relative performance and computation time on random Gaussian data not tested below.
-For PAM and $k$-medoids, when solving for multiple values of $k$, we use the solution with $k$ centers as a warm start for $k+l$ centers by choosing an additional random $l$ centers.
-For the two LP algorithms, Gurobi uses a warm start from solving the LP for $k+l$ to solve it for $k$ because we found this reverse direction to be faster in practice.
+\subsection{Implementation Specifics}
 
-We also tested a more computationally friendly heuristic of the hierarchical framework which only buckets solutions for values of $k$ that are a power of two.
-This yields many fewer $k$-median instances to solve, reducing the run-time significantly.
-This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
-
-For each data set and each clustering method, we report the approximation ratio, computation time, and FScore measure.
-FScore is a measure that looks at the clusters represented on each level and how they represent the true clustering in terms of precision and recall.
-See \cite{Larsen} for more information.
-[ADD DEFINITION OF FSCORE AND MORE DETAIL]
-
-[MAKE CODE AVAILABLE PUBLICLY]
-
-Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on a 4 core 3.4 Ghz Intel i7 processor with 16 GB physical memory.
-Results are given in Tables 2-4  below with some results averaged due to space constraints.
-The results reported are average values across 10 trials.
-
-The data sets tested against were:
+All tests were conducted using Python 2.7.9 on a 64-bit Windows platform with a 4 core 3.4 Ghz Intel i7 processor and 16 GB physical memory.
+All code is available at \url{https://github.com/apaul29/Hierarchical-Code}.
+We chose to test our methods on a variety of data:
 \begin{LaTeXdescription}
@@
 Around each center, we then generated $(1000/k) \cdot \mathrm{Unif}[0,1]$ Gaussian points of which $e$ percent had standard deviation 0.25 (noisy) and $1-e$ percent had standard deviation $0.05$.
-\item[OR $p$-median Library] \cite{Beasley} From a collection of test data sets created by Beasley with 40 test sets of size $100, 200, \ldots, 900$ and varying values of $k$. 
+\item[OR $p$-median Library] \cite{Beasley} From a collection of classic test data sets for approximation algorithms created by Beasley with 40 test sets of size $100, 200, \ldots, 900$ and varying values of $k$.
+This data is unlabeled.
 \item[UCI Iris] \cite{Iris} Contains 4 attributes on 150 instances of flowers, each of with belongs to one of three species of iris.
@@
 \end{table}
+
+\subsection{Partitional Algorithms}
+ For the black-box $k$-median algorithm, we implemented $k$-medoids, PAM, LP-greedy, LP-greedy-rand, and LP-Charikar in Python using Gurobi 6.0.3 to solve the linear programming instances. 
+To avoid cycling, we limited the number of iterations in $k$-medoids to 1000.
+For PAM, we also needed to limit the number of swaps to achieve a reasonable runtime.
+Instead of searching over all potential swaps, we sample at most $n$ potential ones.
+We also ensure that some minimum progress is achieved by ensuring the algorithm decreases in objective by at least a factor of 0.9 in each iteration.
+At the end of PAM, we run 10 iterations of $k$-medoids to ensure the algorithm finds a local optimum.
+These choices were made by looking at the relative performance and computation time on random Gaussian data not in the data sets above.
+
+Before implementing these data sets in the hiearchical framework, we tested them as partitional algorithms on the data above which had given values of $k$.
+For comparison, we also included the results from running $k$-means (implemented in the scikit-learn package in Python).
+
+For each data set and each partitional clustering method, we recorded the approximation ratio (for the flat $k$-median objective function), computation time, and, where applicable, adjusted mutual information (AMI).
+AMI is a measure which scores accuracy of predicted clusters compared to a true labeling by measuring the agreement of two assignments.
+The measure is normalized to reflect chance, so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
+See \cite{Vinh} for further information.
+Results are given in Tables [ADD REF]  below with some results averaged due to space constraints.
+The results reported are average values across 10 trials.The results reported are average values across 10 trials.
+
+[ADD TABLES]
+
+[ADD FLAT RESULT ANAYLIS]
+
+It is interesting to note that, for the Gaussian data, an increase in dimension, reduction in $k$, or reduction in the percentage of noisy data made it easier for the algorithms to classify the data set, which makes sense based on our intuition of clustering.
+For the OR library instances, we saw that the approximation ratio tended to increase as $k$ grew larger but not as $n$ grew.
+
+\subsection{Hierarchical Results}
+After testing the partitional algorithms above, we implemented the nesting hierarchical framework, offering options for both the randomized and deterministic versions.
+For PAM and $k$-medoids, when solving for multiple values of $k$, we use the solution with $k$ centers as a warm start for $k+l$ centers by choosing an additional random $l$ centers.
+For the LP algorithms, Gurobi uses a warm start from solving the LP for $k+l$ to solve it for $k$ because we found this reverse direction to be faster in practice.
+
+We also tested a more computationally friendly heuristic of the hierarchical framework which only buckets solutions for values of $k$ that are a power of two.
+This yields many fewer $k$-median instances to solve, reducing the run-time significantly.
+This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
+For PAM, we only tested the logarithmic heuristic since, as also indicated by the partitional results, the run time for the full version proved impractical.
+
+We compared the results for all versions to hierarchical agglomerative clustering with both average linkage and Ward's linkage (implemented in the scipy package in Python). 
+For each data set and each clustering method, we report the approximation ratio, computation time, and FScore measure.
+FScore is a measure that looks at the clusters represented on each level and how they represent the true clustering in terms of precision and recall.
+
+Let $n$ be the number of data points, let $L_1, L_2 \ldots, L_c$ be the true classes of the data, and let $\mathcal{C}$ all clusters represented on any level of our hierarchical clustering.
+For any class $L_r$ and cluster $C_i$, let $n_r$ be the number of data points in $L_r$ and $n_i$ be the number of data points in $C_i$.
+Then, let $n_{ri}$ be the number of data points in $C_i$ from class $L_r$.
+We define the precision of $C_i$ to be $P(L_r, C_i) = n_{ri}/n_r$ and the recall of $C_i$ to be $R(L_r, C_i) = n_{ri}/n_r$.
+The harmonic mean of these two metrics gives a balanced measure of how accurately $S_i$ represents $L_r$.
+The FScore of $L_r$ is then the best measure over all clusters
+\begin{equation}
+ F(L_r) = \max_{C_i \in \mathcal{C}}  \frac{ 2 P(L_r, C_i)  R(L_r, C_i)}{ P(L_r, C_i) + R(L_r, C_i)} 
+\end{equation}
+and the overall FScore is the sum of weighted FScores for each class
+\begin{equation}
+\mathrm{FScore} = \sum_{r=1}^c \frac{n_r}{n} F(L_r).
+\end{equation}
+See \cite{Larsen} for more information.
+
+Results are given in Tables~\ref{fig:hier_results_comp}-\ref{fig:hier_results_approx}  below with some results averaged due to space constraints.
+The results reported are average values across 10 trials.
 
@@
 
+[ADD MORE RESULTS AND UPDATE TABLES]
+
 In general, the hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure, even for our logarithmic version.
-The FScore measures were very low for the agglomerative clustering methods, sometimes around 0.25, but were much closer to the AMI measures we saw in flat clustering for our implemented algorithms.
+The FScore measures were very low for the agglomerative clustering methods, sometimes around 0.25, but were much closer to the AMI measures we saw in partitional clustering for our implemented algorithms.
 
-For $k$-medoids in the non-logarithmic version and PAM in the logarithmic version, the hierarchical framework increased the computation time by three orders of magnitude, and increased by two orders of magnitude for $k$-medoids in the logarithmic version.
+For $k$-medoids in the non-logarithmic version and PAM in the logarithmic version, the hierarchical framework increased the computation time by three orders of magnitude, and increased by two orders of magnitude for $k$-medoids in the logarithmic version over the run times of agglomerative clustering.
 Interestingly, the approximation factors did not change significantly between the randomized and deterministic versions for $k$-medoids.
@@
 \section{Conclusion and Future Work}
-We implemented various simple approximation algorithms for partitional clustering, and a framework for hierarchical clustering.
-We compared these algorithms, along with heuristic modifications, against the most common clustering algorithms used in practice.
-
-For partitional clustering, we found that the approximation algorithms in the literature are in general slower and cannot improve much on current heuristics used despite their approximation guarantees.
-This was not true in the hierarchical clustering case where we saw significant improvement in FScore measures for our implemented methods and overall good approximation ratios where none were had before.
+We implemented the nesting framework introduced by Lin et al.~\cite{Lin}, which focuses on minimizing the hierarchical $k$-median objective function, to produce hierarchical clusterings using various methods for constructing the inputted partitional clusterings.
+Overall, we saw significant improvement in FScore measures for our implemented methods compared to agglomerative clustering and overall good approximation ratios where none were had before.
 In particular, our logarithmic nesting algorithm seems to provide the best balance between performance and computational efficiency.
 
-In the future, we hope to explore how linear programming based methods perform in both the flat and hierarchical context as well and to implement our own agglomerative clustering to help make smarter local merges in the nesting framework.
+The reason the logarithmic version improved the run time was because the algorithm did not need to run the partitional clustering algorithm for each value of $k$.
+However, even in the full version of the algorithm, the bucketing of solutions leads to an overall small amount of representatives.
+If we could efficiently bound the costs for each value of $k$, these solutions could be partially bucketed ahead of time and we would only need to solve for a small fraction of $k$ values.
+Future work in this area could maintain the high FScores and approximations of the full version while reducing the run time towards that of the logarithmic version.
+Additionally, this nesting framework could be applied to other hierarchical objective functions such as those that reflect the density of the given clusters.
 

### 2015-05-30 icdm/submission.tex
Results and updated submission

@@
 
-Interestingly, the solution to the linear programming is often integer valued [NEED CITATION].
-Based on this fact, we proposed a flat clustering algorithm that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values.
-We refer to this algorithm as LP-greedy from now on.
-Charikar and Li \cite{charikar} introduce a 3.25-approximation algorithm that uses this LP solution to smartly construct a set of centers.
+Interestingly, the solution to the linear programming is often integer valued \cite{Nagarajan}.
+Based on this fact, we propose two partitional clustering algorithms: one that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values and one that solves the LP relaxation and constructs $S$ by choosing $i$ to be in $S$ with probability $s_i$, sampling for each $i \in X$ until $k$ centers are chosen.
+We refer to these algorithms as LP-greedy and LP-rand, respectively.
+Charikar and Li \cite{Charikar} introduce a 3.25-approximation algorithm that uses this LP solution to smartly construct a set of centers.
 We refer to this algorithm as LP-Charikar.
-We choose to implement this LP algorithm rather than the LP-based approximation algorithm with best approximation ratio because the algorithm is simpler to implement (and will most likely have a lower running time).
+We choose to implement this LP algorithm rather than the LP-based approximation algorithm with best approximation ratio because the algorithm is simpler to implement (and we expect it to have a lower running time).
 
 The algorithm works by first choosing a subset of data points $C \subseteq X$ to represent the demand.
-For each data point $i$, let $d_{av}(i) = \sum_{j \in X : a_{i,j} > 0} s_i \cdot d(i,j)$.
+For each data point $i$, let $d_{av}(i) = \sum_{j \in X : a_{i,j} > 0} s_j \cdot d(i,j)$.
 We start with $C = \emptyset$.
@@
 % pmed40 & 5128 & 90 & 900 & \  \\ 
-pmedian avg & 5535 & 46 & 460 & \  \\ \hline
+$p$-median avg & 5535 & 46 & 460 & \  \\ \hline
  \end{tabular}
@@
 
+\subsection{Partitional Algorithms}
+ For the black-box $k$-median algorithm, we implemented $k$-medoids, PAM, LP-greedy, LP-rand, and LP-Charikar in Python using Gurobi 6.0.3 to solve the linear programming instances. 
+To avoid cycling, we limited the number of iterations in $k$-medoids to 1000.
+For PAM, we also needed to limit the number of swaps to achieve a reasonable runtime.
+Instead of searching over all potential swaps, we sample at most $n$ potential ones.
+We also ensure that some minimum progress is achieved by ensuring the algorithm decreases in objective by at least a factor of 0.9 in each iteration.
+At the end of PAM, we run 10 iterations of $k$-medoids to ensure the algorithm finds a local optimum.
+These choices were made by looking at the relative performance and computation time on random Gaussian data not in the data sets above.
+
+Before implementing these data sets in the hiearchical framework, we tested them as partitional algorithms on the data above which had given values of $k$.
+For comparison, we also included the results from running $k$-means (implemented in the scikit-learn package in Python) and the extension of PAM called 2-swaps that also looks at local swaps of size 2 (which improves the $k$-median approximation ratio to 4 from 5).
+
+For each data set and each partitional clustering method, we recorded the approximation ratio (for the flat $k$-median objective function), computation time, and, where applicable, adjusted mutual information (AMI).
+AMI is a measure which scores accuracy of predicted clusters compared to a true labeling by measuring the agreement of two assignments.
+The measure is normalized to reflect chance, so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.
+See \cite{Vinh} for further information.
+Results are given in Tables~\ref{fig:flat_hier_comp}-\ref{fig:flat_hier_approx}  below with some results averaged due to space constraints.
+The results reported are average values across 10 trials.
+
+\begin{table*}[!t]
+\label{fig:flat_results_comp}
+\caption{Average Computation Time (sec) for Partitional Clustering Algorithms.}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l | l | l|}
+\hline
+Dataset	&	k-means	&	k-medoids	&	PAM	&	2-swaps	&	LP-Charikar	&	LP-greedy	&	LP-rand	\\ \hline
+Soybean	&	0.0151	&	0.0020	&	0.0159	&	0.0493	&	0.1180	&		&		\\
+Iris	&	0.0129	&	0.0030	&	0.1018	&	0.2294	&	1.4644	&		&		\\
+Gauss\_5\_5\_5	&	0.0249	&	0.0318	&	1.5058	&	2.3824	&	18.1701	&		&		\\
+Gauss\_5\_10\_5	&	0.0540	&	0.1085	&	5.4071	&	9.4277	&	47.7316	&		&		\\
+Gauss\_5\_50\_5	&	0.1371	&	0.3019	&	21.4457	&	39.6144	&	17.5115	&		&		\\
+Gaussian avg	&	0.0730	&	0.1817	&	13.0214	&	18.6957	&	37.2315	&		&		\\
+pmed5	&	N/A	&	0.0381	&	0.6223	&	0.7479	&	0.5357	&		&		\\
+pmed20	&	N/A	&	1.7947	&	44.9791	&	68.2306	&	18.2055	&		&		\\
+pmed35	&	N/A	&	0.1133	&	5.8814	&	9.2881	&	117.4038	&		&		\\
+$p$-median avg	&	N/A	&	0.8072	&	40.3474	&	56.7595	&	46.1541	&		&		\\ \hline
+\end{tabular}
+\end{table*}
+
+\begin{table*}[!t]
+\label{fig:flat_results_AMI}
+\caption{Average AMI measure for Partitional Clustering Algorithms.}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l | l | l| }
+\hline
+Dataset	&	k-means	&	k-medoids	&	PAM	&	2-swaps	&	LP-Charikar	&	LP-greedy	&	LP-rand	\\ \hline
+Soybean	&	0.686	&	0.550	&	0.732	&	0.708	&	0.721	&		&		\\
+Iris	&	0.748	&	0.690	&	0.760	&	0.757	&	0.647	&		&		\\
+Gauss\_5\_5\_5	&	0.741	&	0.687	&	0.749	&	0.744	&	0.754	&		&		\\
+Gauss\_5\_10\_5	&	0.783	&	0.677	&	0.725	&	0.737	&	0.729	&		&		\\
+Gauss\_5\_50\_5	&	0.646	&	0.621	&	0.619	&	0.631	&	0.631	&		&		\\
+Gaussian avg	&	0.777	&	9.112	&	0.761	&	0.761	&	0.767	&		&		\\ \hline
+\end{tabular}
+\end{table*}
+
+\begin{table*}[!t]
+\label{fig:flat_results_approx}
+\caption{Average Approximation Coeffiecients for Partitional Clustering Algorithms.}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l| l|}
+\hline
+Dataset	&	k-medoids	&	PAM	&	2-swaps	&	LP-Charikar	&	LP-greedy	&	LP-rand	\\ \hline
+Soybean	&	1.167	&	1.012	&	1.012	&	1.000	&		&		\\
+Iris	&	1.081	&	1.004	&	1.004	&	0.873	&		&		\\
+Gauss\_5\_5\_5	&	1.070	&	1.002	&	1.005	&	1.002	&		&		\\
+Gauss\_5\_10\_5	&	1.080	&	1.011	&	1.009	&	1.020	&		&		\\
+Gauss\_5\_50\_5	&	1.089	&	1.020	&	1.014	&	1.001	&		&		\\
+Gaussian avg	&	1.064	&	1.009	&	1.008	&	1.004	&		&		\\
+pmed5	&	1.373	&	1.026	&	1.033	&	1.000	&		&		\\
+pmed20	&	1.501	&	1.032	&	1.027	&	1.000	&		&		\\
+pmed35	&	1.083	&	1.005	&	1.007	&	1.027	&		&		\\
+$p$-median avg	&	1.187	&	1.017	&	1.019	&	1.000	&		&		\\ \hline
+\end{tabular}
+\end{table*}
+
+[ADD FLAT RESULT ANAYLIS]
+
+It is interesting to note that, for the Gaussian data, an increase in dimension, reduction in $k$, or reduction in the percentage of noisy data made it easier for the algorithms to classify the data set, which makes sense based on our intuition of clustering.
+For the OR library instances, we saw that the approximation ratio tended to increase as $k$ grew larger but not as $n$ grew.
+
+\subsection{Hierarchical Results}
+After testing the partitional algorithms above, we implemented the nesting hierarchical framework, offering options for both the randomized and deterministic versions.
+For PAM and $k$-medoids, when solving for multiple values of $k$, we use the solution with $k$ centers as a warm start for $k+l$ centers by choosing an additional random $l$ centers.
+For the LP algorithms, Gurobi uses a warm start from solving the LP for $k+l$ to solve it for $k$ because we found this reverse direction to be faster in practice.
+
+We also tested a more computationally friendly heuristic of the hierarchical framework which only buckets solutions for values of $k$ that are a power of two.
+This yields many fewer $k$-median instances to solve, reducing the run-time significantly.
+This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
+For PAM, we only tested the logarithmic heuristic since, as also indicated by the partitional results, the run time for the full version proved impractical.
+
+We compared the results for all versions to hierarchical agglomerative clustering with both average linkage and Ward's linkage (implemented in the scipy package in Python). 
+For each data set and each clustering method, we report the approximation ratio, computation time, and FScore measure.
+FScore is a measure that looks at the clusters represented on each level and how they represent the true clustering in terms of precision and recall.
+
+Let $n$ be the number of data points, let $L_1, L_2 \ldots, L_c$ be the true classes of the data, and let $\mathcal{C}$ all clusters represented on any level of our hierarchical clustering.
+For any class $L_r$ and cluster $C_i$, let $n_r$ be the number of data points in $L_r$ and $n_i$ be the number of data points in $C_i$.
+Then, let $n_{ri}$ be the number of data points in $C_i$ from class $L_r$.
+We define the precision of $C_i$ to be $P(L_r, C_i) = n_{ri}/n_r$ and the recall of $C_i$ to be $R(L_r, C_i) = n_{ri}/n_r$.
+The harmonic mean of these two metrics gives a balanced measure of how accurately $S_i$ represents $L_r$.
+The FScore of $L_r$ is then the best measure over all clusters
+\begin{equation}
+ F(L_r) = \max_{C_i \in \mathcal{C}}  \frac{ 2 P(L_r, C_i)  R(L_r, C_i)}{ P(L_r, C_i) + R(L_r, C_i)} 
+\end{equation}
+and the overall FScore is the sum of weighted FScores for each class
+\begin{equation}
+\mathrm{FScore} = \sum_{r=1}^c \frac{n_r}{n} F(L_r).
+\end{equation}
+See \cite{Larsen} for more information.
+
+Results are given in Tables~\ref{fig:hier_results_comp}-\ref{fig:hier_results_approx} for the full hierarchical version and in Tables~\ref{fig:hier_log_results_compt}-\ref{fig:hier_log_results_approx} for the logarithmic version.
+The results reported are average values across 10 trials.
+
 \begin{table*}[!t]
 \label{fig:hier_results_comp}
-\caption{Average Computation Time (sec), FScore measure, and Approximation Coeffiecients for Hierarchical Clustering Algorithms.
-K-med and PAM represent the black-box $k$-median solvers used and rand and logn indicate whether the randomized and logarithmic setting were used, respectively.}
+\caption{Average Computation Time (sec) for Hierarchical Clustering Algorithms.
+Besides the agglomerative methods, the algorithm is denoted by the $k$-median solver used and [R] and [D] are used to denote the randomized and determinisitic versions, respectively.}
 \centering
-\begin{tabular}{ | l | l | l | l | l | l | l | }
+\begin{tabular}{ | l | l | l | l | l | l | l | l| l| }
 \hline
-	   & Time & Time & Time & Time & Time & Time \\ \hline
-	Dataset & Ave Linkage & Ward Linkage & K-med & K-med logn ran & k-med ran & PAM logn ran \\ \hline
-	Soybean & 0.0004 & 0.0002 & 0.286 & 0 & 0.294 & 0.167 \\ 
-	Iris & 0.0021 & 0.0011 & 3.161 & 0.226 & 3.794 & 5.122 \\ 
-	Gauss\_5\_5\_5 & 0.0097 & 0.0109 & 106.619 & 2.735 &120.893 & 147.137 \\
-	Gauss\_5\_10\_5 & 0.0237 & 0.0256 & 383.854 & 10.259 & 383.298 & 494.554 \\
-	%Gauss\_5\_50\_5 &  & & & & & \\
-	Gaussian Average & 0.0229 & 0.0239 & 354.459 & 7.492 & 369.967 & 423.293 \\ 
-	pmed5 &  0.0004 & N/A & 1.184 & 0.102 & 1.288 & 1.673   \\ 
-	pmed25 & 0.0174 & N/A & 279.334 & 5.011 & 306.684 & 312.503   \\ 
-	pmedian avg & 0.0067 & N/A &  87.355 & 1.957 & 94.548 & 101.576   \\ \hline
+Dataset	&	Ward Linkage	&	Average Linkage	&	$k$-medoids [R]	&	$k$-medoids [D]	&	LP-Charikar [R]	&	LP-rand [R]	&	LP-greedy [R]	&	LP-greedy [D]	\\ \hline
+Soybean	&	0.0002	&	0.0002	&	0.2132	&	0.2086	&	0.9903	&	0.4068	&	0.2817	&	0.4708	\\
+Iris	&	0.0011	&	0.0009	&	3.2308	&	3.4530	&	32.3171	&	11.5128	&	8.8181	&	14.0168	\\
+Gauss\_5\_5\_5	&	0.0123	&	0.0105	&	165.3442	&	106.6187	&	712.6960	&	273.1223	&	267.1039	&	222.1711	\\
+Gauss\_5\_10\_5	&	0.0284	&	0.0260	&	556.8065	&	383.8541	&	2088.8757	&	793.7941	&	622.8916	&	623.8670	\\
+Gauss\_5\_50\_5	&	0.0121	&	0.0112	&	176.8477	&	128.2697	&	764.3297	&	269.5327	&	253.2762	&	222.7860	\\
+Gaussian avg	&	0.0250	&	0.0217	&	521.3331	&	354.4588	&	1695.7987	&	666.5034	&	566.4133	&	567.0086	\\
+pmed5	&	N/A	&	0.0004	&	1.4797	&	1.1839	&		&	3.8239	&	4.6917	&		\\
+pmed20	&	N/A	&	0.0098	&	153.5358	&	121.9013	&	550.8604	&	240.8364	&	193.8426	&	189.6854	\\
+pmed35	&	N/A	&	0.0666	&	3225.4469	&	1957.6519	&		&	1939.0170	&	2326.1185	&	1801.2031	\\
+$p$-median avg	&	N/A	&	0.0243	&	1021.8706	&	652.1928	&		&	723.1452	&	857.3750	&		\\ \hline
 \end{tabular}
@@
 \centering
-\begin{tabular}{ | l | l | l | l | l | l | l | }
+\begin{tabular}{ | l | l | l | l | l | l | l | l| l| }
 \hline
-	   & F-Score & F-Score & F-Score & F-Score & F-Score & F-Score \\ \hline
-	Dataset & Ave Linkage & Ward Linkage & K-med & K-med logn rand & K-med rand & PAM logn rand \\ \hline
-	Soybean & 0.615 & 0.554 & 0.772 & 0.643 & 0.742 & 0.633 \\ 
-	Iris & 0.5 & 0.5 & 0.888 & 0.767 & 0.762 & 0.732 \\ 
-	Gauss\_5\_5\_5 & 0.519 & 0.520 & 0.591 & 0.816 & 0.761 & 0.785 \\
-	Gauss\_5\_10\_5 & 0.298 & 0.281 & 0.589 & 0.584 & 0.576 & 0.446 \\
-	%Gauss\_5\_50\_5 &  & & & & & \\
-	Gaussian Average & 0.393 & 0.378 & 0.649 & 0.616 & 0.624 & 0.624 \\ \hline
+Dataset	&	Ward	&	Average	&	$k$-medoids [R]	&	$k$-medoids [D]	&	LP-Charikar [R]	&	LP-rand [R]	&	LP-greedy [R]	&	LP-greedy [D]	\\ \hline
+Soybean	&	0.554	&	0.557	&	0.653	&	0.772	&	0.645	&	0.699	&	0.677	&	0.842	\\
+Iris	&	0.500	&	0.500	&	0.772	&	0.789	&	0.804	&	0.739	&	0.829	&	0.720	\\
+Gauss\_5\_5\_5	&	0.520	&	0.520	&	0.765	&	0.591	&	0.698	&	0.791	&	0.754	&	0.775	\\
+Gauss\_5\_10\_5	&	0.281	&	0.285	&	0.463	&	0.589	&	0.575	&	0.536	&	0.506	&	0.466	\\
+Gauss\_5\_50\_5	&	0.509	&	0.519	&	0.554	&	0.539	&	0.607	&	0.579	&	0.560	&	0.594	\\
+Gaussian avg	&	0.378	&	0.391	&	0.592	&	0.649	&	0.603	&	0.619	&	0.593	&	0.614	\\ \hline
 \end{tabular}
@@
 \centering
-\begin{tabular}{ | l | l | l | l | l | }
+\begin{tabular}{ | l | l | l | l | l | l| l|}
 \hline
-	   & Approx & Approx & Approx & Approx \\ \hline
-	Dataset & K-med & K-med logn rand & K-med rand & PAM logn rand \\ \hline
-	Soybean & 1.585 & 2.334 & 1.711 & 2.001 \\ 
-	Iris & 2560631.459 & 2550279.566 & 2787312.009 & 1623132.096 \\ 
-	Gauss\_5\_5\_5 & 6.459 & 13.394 & 8.624 & 7.791 \\
-	Gauss\_5\_10\_5 &  3.420 & 3.756 & 4.196 & 2.309  \\
-	%Gauss\_5\_50\_5 &  & & &  \\
-	Gaussian Average & 5.182 & 7.262  & 5.537 & 4.739 \\ 
-	pmed5 & 11.367 & 51.141 & 16.261 & 11.950 \\ 
-	pmed25 & 6.641 & 13.372 & 5.378 & 13.200 \\ 
-	pmedian avg & 10.330 & 19.822 & 10.524 & 8.599 \\ \hline
+Dataset	&	$k$-medoids [R]	&	$k$-medoids [D]	&	LP-Charikar [R]	&	LP-rand [R]	&	LP-greedy [R]	&	LP-greedy [D]	\\ \hline
+Soybean	&	1.925	&	1.585	&	1.700	&	1.664	&	1.685	&	1.372	\\
+Iris	&	2269336.817	&	2093263.257	&	2.455	&	545260.030	&	2.307	&	2.708	\\
+Gauss\_5\_5\_5	&	5.601	&	6.459	&	2.327	&	2.153	&	2.210	&	1.858	\\
+Gauss\_5\_10\_5	&	3.775	&	3.420	&	2.055	&	2.232	&	2.212	&	2.145	\\
+Gauss\_5\_50\_5	&	2.840	&	3.210	&	1.872	&	1.993	&	2.063	&	1.626	\\
+Gaussian avg	&	5.502	&	5.182	&	2.292	&	2.282	&	2.290	&	2.075	\\
+pmed5	&	28.151	&	11.367	&		&	1.710	&	1.759	&		\\
+pmed20	&	11.606	&	7.210	&	1.756	&	1.757	&	1.776	&	1.680	\\
+pmed35	&	5.476	&	3.850	&		&	1.817	&	1.834	&	1.690	\\
+$p$-median avg	&	12.005	&	8.260	&		&	1.758	&	1.762	&		\\ \hline
 \end{tabular}
@@
 
-In general, the hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure, even for our logarithmic version.
-The FScore measures were very low for the agglomerative clustering methods, sometimes around 0.25, but were much closer to the AMI measures we saw in flat clustering for our implemented algorithms.
+\begin{table*}[!t]
+\label{fig:hier_log_results_comp}
+\caption{Average Computation Time (sec) for the Logarithmic Hierarchical Clustering Algorithms.
+The algorithm is denoted by the $k$-median solver used and [R] indicates that the randomized version was used.}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l | }
+\hline
+Dataset	&	PAM	&	$k$-medoids [R]	&	LP-Charikar [R]	&	LP-rand [R]	&	LP-greedy [R]	\\ \hline
+Soybean	&	0.2083	&	0.0332	&	0.2156	&	0.1370	&	0.1565	\\
+Iris	&	6.9149	&	0.2393	&	3.4825	&	2.5000	&	2.6544	\\
+Gauss\_5\_5\_5	&	145.7625	&	3.3253	&	77.6025	&	222.1711	&	61.0401	\\
+Gauss\_5\_10\_5	&	446.8337	&	11.5259	&	210.5056	&	623.8670 	&	162.6597	\\
+Gauss\_5\_50\_5	&	176.3119	&	3.3950	&	76.7538	&	222.7860 	&	58.5883	\\
+Gaussian avg	&	445.8375	&	8.3335	&	181.0512	&	380.8876 	&	146.9885	\\
+pmed5	&	1.4092	&	0.1142	&	1.3029	&		&	0.9092	\\
+pmed20	&	149.3397	&	3.5036	&	53.7326	&	36.1935	&	43.3070	\\
+pmed35	&	1351.1987	&	28.6012	&	554.3661	&		&	488.2756	\\
+$p$-median avg	&	492.7771	&	10.5447	&	204.3356	&		&	173.2638	\\ \hline
+\end{tabular}
+\end{table*}
 
-For $k$-medoids in the non-logarithmic version and PAM in the logarithmic version, the hierarchical framework increased the computation time by three orders of magnitude, and increased by two orders of magnitude for $k$-medoids in the logarithmic version.
-Interestingly, the approximation factors did not change significantly between the randomized and deterministic versions for $k$-medoids.
-The approximation ratios were strongest as $n$ increased or as the dimension increased.
-In particular, for the Gaussian data in 7 dimensions, the approximation ratio was often around 3.
-The high approximation ratio for the iris data set was due to the fact that the optimal solution for $k = n-1, n-2$ was very close to zero, and our heuristic $k$-median algorithm only sometimes picked up on these solutions.
+\begin{table*}[!t]
+\label{fig:hier_log_results_FScore}
+\caption{Average FScore measure for Logarithmic Hierarchical Clustering Algorithms.}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l | }
+\hline
+Dataset	&	PAM	&	$k$-medoids [R]	&	LP-Charikar [R]	&	LP-rand [R]	&	LP-greedy [R]	\\ \hline
+Soybean	&	0.664	&	0.645	&	0.664	&	0.633	&	0.626	\\
+Iris	&	0.790	&	0.754	&	0.777	&	0.784	&	0.714	\\
+Gauss\_5\_5\_5	&	0.727	&	0.692	&	0.747	&		&	0.724	\\
+Gauss\_5\_10\_5	&	0.605	&	0.506	&	0.557	&		&	0.546	\\
+Gauss\_5\_50\_5	&	0.509	&	0.557	&	0.578	&		&	0.551	\\
+Gaussian avg	&	0.595	&	0.585	&	0.576	&		&	0.590	\\ \hline
+\end{tabular}
+\end{table*}
+
+\begin{table*}[!t]
+\label{fig:hier_log_results_approx}
+\caption{Average Approximation Coeffiecients for Logarithmic Hierarchical Clustering Algorithms.}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l| }
+\hline
+Dataset	&	PAM	&	$k$-medoids [R]	&	LP-Charikar [R]	&	LP-rand [R]	&	LP-greedy [R]	\\ \hline
+Soybean	&	2.201	&	2.533	&	2.142	&	2.000	&	2.053	\\
+Iris	&	1734568.640	&	4969423.160	&	1779492.480	&	1779492.480	&	1779492.480	\\
+Gauss\_5\_5\_5	&	13.297	&	34.739	&	10.408	&	1.858	&	8.702	\\
+Gauss\_5\_10\_5	&	3.066	&	6.145	&	2.913	&	2.145	&	2.925	\\
+Gauss\_5\_50\_5	&	4.694	&	9.615	&	5.114	&	1.626	&	5.525	\\
+Gaussian avg	&	8.494	&	18.565	&	7.282	&	5.000	&	7.004	\\
+pmed5	&	26.550	&	73.950	&	14.000	&		&	21.800	\\
+pmed20	&	11.850	&	38.863	&	10.000	&	7.000	&	8.000	\\
+pmed35	&	6.017	&	25.900	&	4.207	&		&	6.000	\\
+$p$-median avg	&	12.739	&	40.316	&	9.158	&		&	10.066	\\ \hline
+\end{tabular}
+\end{table*}
+
+In general, the full hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure.
+The FScore measures were very low for the agglomerative clustering methods, consistently under 0.5 and sometimes around 0.25, but averaged around 0.6 for our nesting algorithms.
+Surprisingly, the algorithms all had similar FScore measures.
+However, the approximation ratios for the LP algorithms, which were all around 2 for the Gaussian data and slightly under 2 for the pmedian instances, outperformed the $k$-medoids algorithm, which often had approximation coefficients around 5 for the Gaussian data and around 10 for the pmedian data.
+The high approximation ratio for some of the $k$-medoids algorithms on the iris data set was due to the fact that the optimal solution for $k = n-1, n-2$ was very close to zero, and some heuristic algorithms only sometimes picked up on these solutions.
 The approximation factor at lower levels was more consistent with the other results.
 
+There were several interesting trends in the performance.
+First, the deterministic versions had higher FScore measures and lower approximation coefficients on average compared to their randomized counterparts so we would advise to just use the deterministic version despite its higher approximation guarantee.
+Similarly, our simple LP heuristics performed comparably to the more advanced LP-Charikar algorithm.
+Also worth noting is the fact that the approximation ratio for $k$-medoids improved as $n$ increased in the $p$-median instances.
+
+In terms of the computation time, the agglomerative clustering methods were significantly faster than all of nesting algorithms, which we would expect given how many $k$-median instances they have to solve.
+$k$-medoids was the fastest method, but $k$-medoids, LP-greedy, and LP-rand all performed around 4 orders of magnitude slower than the agglomerative methods (this order increased to 5 for a few of the very large $p$-median instances).
+LP-Charikar was the slowest method and was often at least 5 orders of magnitude slower than the agglomerative methods.
+
+In the logarithmic version, the computation times all improved.
+For $k$-medoids, the algorithm became 2-3 orders of magnitude slower than agglomerative clustering while the LP methods all decreased by 1 order of magnitude.
+However, LP-rand now had the worst running time whereas LP-Charikar improved to be similar to LP-greedy.
+PAM, on the other hand, was 4-5 orders of magnitude higher, similar to the methods used in the full hierarchical framework.
+
+Using the logarithmic version did decrease the FScores slightly, but the averages still remained close to 0.6 and above the agglomerative methods.
+Again, the algorithms performed similarly in FScore.
+The approximation ratios also suffered with the $k$-medoids algorithm suffering the most.
+PAM had an average Gaussian approximation ratio of around 8.5, LP-Charikar and LP-greedy around 7, and LP-rand around 5 (the ratios did improve as the dimension increased).
+These averages all increased slightly for the $p$-median instances.
+
+Given these performances, we would recommend using the deterministic, logarithmic framework with a simple LP heuristic as the black-box $k$-median solver.
+If one does not want to use an LP solver, then $k$-medoids is a good choice given the low running times but the approximation ratio might suffer.
+Switching to the full hierarchical version does increase the run time by around one order of magnitude, but offers better FScore and approximation ratios.
+Using PAM seems to be the least beneficial option since the full hierarchical version with $k$-medoids as a solver has better overall performance.
+
 \section{Conclusion and Future Work}
-We implemented various simple approximation algorithms for partitional clustering, and a framework for hierarchical clustering.
-We compared these algorithms, along with heuristic modifications, against the most common clustering algorithms used in practice.
-
-For partitional clustering, we found that the approximation algorithms in the literature are in general slower and cannot improve much on current heuristics used despite their approximation guarantees.
-This was not true in the hierarchical clustering case where we saw significant improvement in FScore measures for our implemented methods and overall good approximation ratios where none were had before.
+We implemented the nesting framework introduced by Lin et al.~\cite{Lin}, which focuses on minimizing the hierarchical $k$-median objective function, to produce hierarchical clusterings using various methods for constructing the inputted partitional clusterings.
+Overall, we saw significant improvement in FScore measures for our implemented methods compared to agglomerative clustering and overall good approximation ratios where none were had before.
 In particular, our logarithmic nesting algorithm seems to provide the best balance between performance and computational efficiency.
 
-In the future, we hope to explore how linear programming based methods perform in both the flat and hierarchical context as well and to implement our own agglomerative clustering to help make smarter local merges in the nesting framework.
+The reason the logarithmic version improved the run time was because the algorithm did not need to run the partitional clustering algorithm for each value of $k$.
+However, even in the full version of the algorithm, the bucketing of solutions leads to an overall small amount of representatives.
+If we could efficiently bound the costs for each value of $k$, these solutions could be partially bucketed ahead of time and we would only need to solve for a small fraction of $k$ values.
+Future work in this area could maintain the high FScores and approximations of the full version while reducing the run time towards that of the logarithmic version.
+Additionally, this nesting framework could be applied to other hierarchical objective functions such as those that reflect the density of the given clusters.
 

### 2015-05-30 icdm/submission.tex
Typos

@@
 
-\begin{figure*}[!t]
+\begin{figure*}[!]
 \centerline{\subfloat[Partitional Clustering]{\includegraphics[width =1.75in]{partitional_clustering}
@@
 \subfloat[Hierarchical Clustering]{\includegraphics[width=2.5in]{hier_clustering}
-\label{fig_second_case}}}
-\caption{A partitional clustering for $k=3$ and a hierarchical clustering of the data where the clustering on the third level down (represented by the red line) corresponds to the partitional clustering on the left.}
-\label{fig_hier}
+\label{fig_hier}}}
+\caption{A partitional clustering for $k=3$, where the colors represent the different partitions, and a dendrogram representation of a hierarchical clustering of the data where edges represent merges of clusters and the clustering on the third level down (represented by the red line) corresponds to the partitional clustering on the left.}
 \end{figure*}
@@
 A solution to the hierarchical $k$-median problem is a nested chain of solutions, and the hierarchical clustering it generates is found by looking at the clusters given by each $(S_i, a_i)$.
+An example is given in Fig.~\ref{fig_chain1}-\ref{fig_chain4}.
 However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
 An $\alpha$-approximation algorithm for this problem generates a nested chain of solutions in polynomial time such that for all $k = 1, \ldots, |X|$ there exists $(S_i, a_i)$ in the chain such that $|S_i| \leq k$ and $f(S_i, a_i) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
+
+\begin{figure*}[b]
+\centering
+\subfloat[$(S_1, a_1)$]{\includegraphics[width =1.75in]{chain1}
+\label{fig_chain1}} \hspace{15mm}
+\subfloat[$(S_2, a_2)$]{\includegraphics[width=1.75in]{chain2}
+\label{fig_chain2}} \\
+\subfloat[$(S_3, a_3)$]{\includegraphics[width=1.75in]{chain3}
+\label{fig_chain3}}  \hspace{15mm}
+\subfloat[$(S_4, a_4)$]{\includegraphics[width=1.75in]{chain4}
+\label{fig_chain4}}
+\caption{An example of a nested chain of solutions.
+$S_i$ is indicated by the colored nodes and $a_i$ by the edges.
+The clusters on that level are given by the connected components.
+Note that $S_{i-1}$ is a subset of $S_i$ and the assignments stay consistent.
+}
+\label{fig_sim}
+\end{figure*}
 
@@
 In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to greedily construct a hierarchical clustering that increases the objective value by at most a factor of $20.71$ (or $10.03$ when randomized) at each level.
-Using the local $p$-swap $(3+\frac{2}{p})$-approximation algorithm mentioned above as the black-box $k$-median solver performed well in experimental hierarchical $k$-median test problems \cite{Nagarajan}.
+Using the local $p$-swap $(3+\frac{2}{p})$-approximation algorithm mentioned above as the black-box $k$-median solver performed well in experiments on hierarchical $k$-median test problems \cite{Nagarajan}.
 While the authors tested the variants of this algorithm for different $k$-median approximation algorithms, it has not been tested against traditional clustering algorithms or on any labeled data sets.
@@
 Algorithms such as the CURE, ROCK, and Chameleon \cite{Guha_CURE, Guha_ROCK, Karypis} incorporate non-spherical cluster shapes and cluster density, but do not relate as much to the above approximation algorithm in their objective.
-We also choose not compare the approximation framework with Bayesian hierarchical methods since these methods require many more parameters and add additional prior information that our algorithms do not use.
+We also choose not to compare our algorithms with Bayesian hierarchical methods since these methods require many more parameters and add additional prior information that our algorithms do not use.
 
@@
 From each non-empty bucket, pick the solution with the smallest number of cluster centers, and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r$.
-We then inductively build our nested solutions.
-Set $S_1 = \overline{V}_1$ and $a_1$ to be the assignment that assigns each point to the closest point in $S_1$.
+We then inductively build our nested solutions: Set $S_1 = \overline{V}_1$ and $a_1$ to be the assignment that assigns each point to the closest point in $S_1$.
 Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we construct $S_i$ by  adding the closest point $j \in S_{i-1}$ to $l$ for each $l \in \overline{V}_i$.
@@
 PAM has the added benefit of being a 5-approximation algorithm for the partitional $k$-median problem which yields to an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
-We also considered the size of the local swaps in PAM to improve the approximation ratio but found that the approximation ratio of constructed solutions for given values of $k$ did not significantly improve while the computation time increased greatly.
+We looked at increasing the size of the local swaps in PAM to improve the approximation ratio but found that the approximation ratio of constructed solutions for given values of $k$ did not significantly improve while the computation time increased greatly.
 
-We also consider two linear programming based algorithms, one of which has an approximation guarantee.
+We also consider three linear programming based algorithms, one of which has an approximation guarantee.
 We first present the linear programming formulation for the $k$-median problem.
-Suppose we have a set $X$ and we let $s_i \in \{0,1\}$ be the variable representing whether or not we want to make $i$ a center for all $i \in X$.
+Suppose we have a set $X$.
+Let $s_i \in \{0,1\}$ be the variable representing whether or not we want to make $i$ a center for all $i \in X$.
 We also create a variable $a_{i,j} \in \{0, 1\}$ which represents whether or not we assign $i \in X$ to the cluster with center $j \in X$.
@@
 However, we are restricted by the fact that we can choose at most $k$ centers, every point must be assigned to a single cluster center, and that we can only assign a point to a center that is actually in $S$.
-This is represented by the three equations:
+This is represented by the following three equations, respectively:
 \begin{eqnarray}
@@
 Interestingly, the solution to the linear programming is often integer valued \cite{Nagarajan}.
-Based on this fact, we propose two partitional clustering algorithms: one that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values and one that solves the LP relaxation and constructs $S$ by choosing $i$ to be in $S$ with probability $s_i$, sampling for each $i \in X$ until $k$ centers are chosen.
+Based on this fact, we propose two LP-based clustering algorithms: one that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values and one that solves the LP relaxation and constructs $S$ by choosing $i$ to be in $S$ with probability $s_i$, sampling for each $i \in X$ until $k$ centers are chosen.
 We refer to these algorithms as LP-greedy and LP-rand, respectively.
@@
 All tests were conducted using Python 2.7.9 on a 64-bit Windows platform with a 4 core 3.4 Ghz Intel i7 processor and 16 GB physical memory.
-All code is available at \url{https://github.com/apaul29/Hierarchical-Code}.
-We chose to test our methods on a variety of data:
+All code and the full computational results are available at \url{https://github.com/apaul29/Hierarchical-Code}.
+We chose to test our methods on data from a few different contexts:
 \begin{LaTeXdescription}
@@
  Dataset Name & Opt Obj & $k$ & $n$ & dim \\ \hline
-Soybean & 114 & 4 & 47 & 35 \\ 
-Iris & 98 & 3 & 150 & 4 \\ 
-Gauss\_5\_5\_5 & 195.4 & 5 & 388 & 5 \\ 
-Gauss\_5\_10\_5 & 270 & 10 & 544 & 5 \\ 
-Gaussian avg & 241 & 22 & 507 & 5 \\ 
+Soybean & 113.805 & 4 & 47 & 35 \\ 
+Iris & 98.214 & 3 & 150 & 4 \\ 
+Gauss\_5\_5\_5 & 195.414 & 5 & 388 & 5 \\ 
+Gauss\_5\_10\_5 & 270.066 & 10 & 544 & 5 \\ 
+Gauss\_5\_50\_5 & 178.122 & 50 & 408 & 5 \\ 
+Gaussian avg & 241.497 & 21.7 & 507 & 5 \\ 
 pmed5 & 1355 & 33 & 100 & \  \\ 
 pmed20 & 1789 & 133 & 400 & \  \\ 
-% pmed40 & 5128 & 90 & 900 & \  \\ 
-$p$-median avg & 5535 & 46 & 460 & \  \\ \hline
+pmed35 & 10400 & 5 & 800 & \  \\ 
+$p$-median avg & 5535.3 & 45.9 & 460 & \  \\ \hline
  \end{tabular}
@@
 
-Before implementing these data sets in the hiearchical framework, we tested them as partitional algorithms on the data above which had given values of $k$.
-For comparison, we also included the results from running $k$-means (implemented in the scikit-learn package in Python) and the extension of PAM called 2-swaps that also looks at local swaps of size 2 (which improves the $k$-median approximation ratio to 4 from 5).
+Before implementing these data sets in the hierarchical framework, we compared them as partitional algorithms on the data above which had given values of $k$.
+For context, we also included the results from running $k$-means (implemented in the scikit-learn package in Python) and the extension of PAM called 2-swaps that also looks at local swaps of size 2 (which improves the $k$-median approximation ratio to 4 from 5).
 
@@
 See \cite{Vinh} for further information.
-Results are given in Tables~\ref{fig:flat_hier_comp}-\ref{fig:flat_hier_approx}  below with some results averaged due to space constraints.
+Results are given in Tables~\ref{fig:flat_results_comp}-\ref{fig:flat_results_approx}  below with some results averaged due to space constraints.
 The results reported are average values across 10 trials.
@@
 \begin{table*}[!t]
-\label{fig:flat_results_comp}
 \caption{Average Computation Time (sec) for Partitional Clustering Algorithms.}
@@
 \end{tabular}
+\label{fig:flat_results_comp}
 \end{table*}
@@
 \begin{table*}[!t]
-\label{fig:flat_results_AMI}
 \caption{Average AMI measure for Partitional Clustering Algorithms.}
@@
 \end{tabular}
+\label{fig:flat_results_AMI}
 \end{table*}
@@
 \begin{table*}[!t]
-\label{fig:flat_results_approx}
-\caption{Average Approximation Coeffiecients for Partitional Clustering Algorithms.}
+\caption{Average Approximation Coefficients for Partitional Clustering Algorithms.}
 \centering
@@
 \end{tabular}
+\label{fig:flat_results_approx}
 \end{table*}
@@
 
-Results are given in Tables~\ref{fig:hier_results_comp}-\ref{fig:hier_results_approx} for the full hierarchical version and in Tables~\ref{fig:hier_log_results_compt}-\ref{fig:hier_log_results_approx} for the logarithmic version.
+Results are given in Tables~\ref{fig:hier_results_comp}-\ref{fig:hier_results_approx} for the full hierarchical version and in Tables~\ref{fig:hier_log_results_comp}-\ref{fig:hier_log_results_approx} for the logarithmic version.
 The results reported are average values across 10 trials.
@@
 \begin{table*}[!t]
-\label{fig:hier_results_comp}
 \caption{Average Computation Time (sec) for Hierarchical Clustering Algorithms.
-Besides the agglomerative methods, the algorithm is denoted by the $k$-median solver used and [R] and [D] are used to denote the randomized and determinisitic versions, respectively.}
+Besides the agglomerative methods, the algorithm is denoted by the $k$-median solver used and [R] and [D] are used to denote the randomized and deterministic versions, respectively.}
 \centering
@@
 Gaussian avg	&	0.0250	&	0.0217	&	521.3331	&	354.4588	&	1695.7987	&	666.5034	&	566.4133	&	567.0086	\\
-pmed5	&	N/A	&	0.0004	&	1.4797	&	1.1839	&		&	3.8239	&	4.6917	&		\\
+pmed5	&	N/A	&	0.0004	&	1.4797	&	1.1839	&		&	3.8239	&	4.6917	&		4.290 \\
 pmed20	&	N/A	&	0.0098	&	153.5358	&	121.9013	&	550.8604	&	240.8364	&	193.8426	&	189.6854	\\
 pmed35	&	N/A	&	0.0666	&	3225.4469	&	1957.6519	&		&	1939.0170	&	2326.1185	&	1801.2031	\\
-$p$-median avg	&	N/A	&	0.0243	&	1021.8706	&	652.1928	&		&	723.1452	&	857.3750	&		\\ \hline
+$p$-median avg	&	N/A	&	0.0243	&	1021.8706	&	652.1928	&		&	723.1452	&	857.3750	&	704.8966	\\ \hline
 \end{tabular}
+\label{fig:hier_results_comp}
 \end{table*}
@@
 \begin{table*}[!t]
-\label{fig:hier_results_FScore}
 \caption{Average FScore measure for Hierarchical Clustering Algorithms.}
@@
 \end{tabular}
+\label{fig:hier_results_FScore}
 \end{table*}
@@
 \begin{table*}[!t]
-\label{fig:hier_results_approx}
-\caption{Average Approximation Coeffiecients for Hierarchical Clustering Algorithms.}
+\caption{Average Approximation Coefficients for Hierarchical Clustering Algorithms.}
 \centering
@@
 Gaussian avg	&	5.502	&	5.182	&	2.292	&	2.282	&	2.290	&	2.075	\\
-pmed5	&	28.151	&	11.367	&		&	1.710	&	1.759	&		\\
+pmed5	&	28.151	&	11.367	&		&	1.710	&	1.759	&	1.5791	\\
 pmed20	&	11.606	&	7.210	&	1.756	&	1.757	&	1.776	&	1.680	\\
 pmed35	&	5.476	&	3.850	&		&	1.817	&	1.834	&	1.690	\\
-$p$-median avg	&	12.005	&	8.260	&		&	1.758	&	1.762	&		\\ \hline
+$p$-median avg	&	12.005	&	8.260	&		&	1.758	&	1.762	&	1.634	\\ \hline
 \end{tabular}
+\label{fig:hier_results_approx}
 \end{table*}
@@
 \begin{table*}[!t]
-\label{fig:hier_log_results_comp}
 \caption{Average Computation Time (sec) for the Logarithmic Hierarchical Clustering Algorithms.
@@
 pmed20	&	149.3397	&	3.5036	&	53.7326	&	36.1935	&	43.3070	\\
-pmed35	&	1351.1987	&	28.6012	&	554.3661	&		&	488.2756	\\
+pmed35	&	1351.1987	&	28.6012	&	554.3661	&	375.4926	&	488.2756	\\
 $p$-median avg	&	492.7771	&	10.5447	&	204.3356	&		&	173.2638	\\ \hline
 \end{tabular}
+\label{fig:hier_log_results_comp}
 \end{table*}
@@
 \begin{table*}[!t]
-\label{fig:hier_log_results_FScore}
 \caption{Average FScore measure for Logarithmic Hierarchical Clustering Algorithms.}
@@
 Iris	&	0.790	&	0.754	&	0.777	&	0.784	&	0.714	\\
-Gauss\_5\_5\_5	&	0.727	&	0.692	&	0.747	&		&	0.724	\\
-Gauss\_5\_10\_5	&	0.605	&	0.506	&	0.557	&		&	0.546	\\
-Gauss\_5\_50\_5	&	0.509	&	0.557	&	0.578	&		&	0.551	\\
-Gaussian avg	&	0.595	&	0.585	&	0.576	&		&	0.590	\\ \hline
+Gauss\_5\_5\_5	&	0.727	&	0.692	&	0.747	&	0.775	&	0.724	\\
+Gauss\_5\_10\_5	&	0.605	&	0.506	&	0.557	&	0.466	&	0.546	\\
+Gauss\_5\_50\_5	&	0.509	&	0.557	&	0.578	&	0.594	&	0.551	\\
+Gaussian avg	&	0.595	&	0.585	&	0.576	&	0.602	&	0.590	\\ \hline
 \end{tabular}
+\label{fig:hier_log_results_FScore}
 \end{table*}
@@
 \begin{table*}[!t]
-\label{fig:hier_log_results_approx}
 \caption{Average Approximation Coeffiecients for Logarithmic Hierarchical Clustering Algorithms.}
@@
 pmed20	&	11.850	&	38.863	&	10.000	&	7.000	&	8.000	\\
-pmed35	&	6.017	&	25.900	&	4.207	&		&	6.000	\\
+pmed35	&	6.017	&	25.900	&	4.207	&	3.634	&	6.000	\\
 $p$-median avg	&	12.739	&	40.316	&	9.158	&		&	10.066	\\ \hline
 \end{tabular}
+\label{fig:hier_log_results_approx}
 \end{table*}
@@
 In general, the full hierarchical framework performed significantly better than both methods of agglomerative clustering in terms of FScore measure.
-The FScore measures were very low for the agglomerative clustering methods, consistently under 0.5 and sometimes around 0.25, but averaged around 0.6 for our nesting algorithms.
-Surprisingly, the algorithms all had similar FScore measures.
-However, the approximation ratios for the LP algorithms, which were all around 2 for the Gaussian data and slightly under 2 for the pmedian instances, outperformed the $k$-medoids algorithm, which often had approximation coefficients around 5 for the Gaussian data and around 10 for the pmedian data.
-The high approximation ratio for some of the $k$-medoids algorithms on the iris data set was due to the fact that the optimal solution for $k = n-1, n-2$ was very close to zero, and some heuristic algorithms only sometimes picked up on these solutions.
+The FScore measures were very low for the agglomerative clustering methods, consistently under 0.5 and averaging around 0.38, but averaged around 0.6 for our nesting algorithms.
+Surprisingly, the algorithms all had similar FScore measures with no method consistently scoring higher than the others.
+However, the approximation ratios for the LP algorithms, which were all around 2 for the Gaussian data and slightly under 2 for the $p$-median instances, outperformed the $k$-medoids algorithm, which often had approximation coefficients around 5 for the Gaussian data and around 10 for the $p$-median data.
+The high approximation ratio for some of the $k$-medoids algorithms on the Iris data set was due to the fact that the optimal solution for $k = n-1, n-2$ was very close to zero, and some heuristic algorithms only sometimes picked up on these solutions.
 The approximation factor at lower levels was more consistent with the other results.
@@
 Similarly, our simple LP heuristics performed comparably to the more advanced LP-Charikar algorithm.
-Also worth noting is the fact that the approximation ratio for $k$-medoids improved as $n$ increased in the $p$-median instances.
+Also worth noting is the fact that the approximation ratio for $k$-medoids improved as $n$ increased in the $p$-median instances, which is surprising the values of $k$ for some of the instances increased proporitional to $n$ so we would expect the overall difficulty to be higher.
 
@@
 Again, the algorithms performed similarly in FScore.
-The approximation ratios also suffered with the $k$-medoids algorithm suffering the most.
+The approximation ratios, on the other hand, did change significantly in the logarithmic version with the $k$-medoids algorithm having the highest increases in ratio.
 PAM had an average Gaussian approximation ratio of around 8.5, LP-Charikar and LP-greedy around 7, and LP-rand around 5 (the ratios did improve as the dimension increased).
 These averages all increased slightly for the $p$-median instances.
+The algorithms also seemed to do worse on the Iris data set and not pick up on the optimal solution for the higher levels.
+This indicates that if a data set has a very steep change in objective function between neighboring solutions that the full hierarchical version should be used.
 
 Given these performances, we would recommend using the deterministic, logarithmic framework with a simple LP heuristic as the black-box $k$-median solver.
-If one does not want to use an LP solver, then $k$-medoids is a good choice given the low running times but the approximation ratio might suffer.
+If one does not want to use an LP solver, then $k$-medoids is a good choice given the low running times but the approximation ratio might increase.
 Switching to the full hierarchical version does increase the run time by around one order of magnitude, but offers better FScore and approximation ratios.

### 2015-06-01 icdm/submission.tex
typos

@@
 
-While some partitional clustering methods use the $k$-median objective function to inform their decisions, popular methods for hierarchical clustering do not and rely on greedy algorithms which have no guarantee on the solution produced with respect to the objective function but are typically fast and easy to implement.
+While some partitional clustering methods use the $k$-median objective function to inform their decisions, popular methods for hierarchical clustering do not and instead rely on greedy algorithms which have no guarantee on the solution produced with respect to the objective function but are typically fast and easy to implement.
 This paper looks at how using this objective function can vastly improve the hierarchical clusterings produced.
@@
 \subsection{Partitional Clustering}
-Arguably the most common partitional clustering method is $k$-means, which chooses $k$ cluster centers points in space to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
+Arguably the most common partitional clustering method is $k$-means, which chooses $k$ cluster centers (which are points in space, not necessarily points from the sample) to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
 $k$-means is usually solved using Lloyd's algorithm (also referred to simply as the $k$-means algorithm), which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}.
@@
 As mentioned above, PAM and $k$-medoids are two potential algorithms.
-PAM has the added benefit of being a 5-approximation algorithm for the partitional $k$-median problem which yields to an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
+PAM has the added benefit of being a 5-approximation algorithm for the partitional $k$-median problem which yields an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
 We looked at increasing the size of the local swaps in PAM to improve the approximation ratio but found that the approximation ratio of constructed solutions for given values of $k$ did not significantly improve while the computation time increased greatly.
@@
 
-Now that the algorithm has grouped and matched together points so the volume is at least 1, the algorithm starts choosing points from these various groups through sampling.
+Now that the algorithm has grouped and matched together points so that the volume is at least 1, the algorithm starts choosing points from these various groups through sampling.
 Here, opening $U_j$ refers to selecting a point $i \in U_j$ with probabilities proportional to the $s_i$ values and adding that point to $S$.

### 2015-06-02 icdm/submission.tex
Edits

--- 2015-05-30 icdm/submission.tex
@@
 \begin{abstract}
+[TODO]
 Hierarchical clustering organizes large data into smaller, more informative groups and offers a global view of the data structure by producing a hierarchy of clusters starting from individual data points and merging towards the cluster of all data.
@@
 \section{Introduction and Notation}
-As the data we have access to grows in variety and size and becomes cheaper to store, we need methods to efficiently learn from this data. 
-Partitioning data (clustering) into similar groups (clusters) is one of the key tools to learn about a set of unlabeled data such as a set of customers who have visited a website or a set of handwritten documents.
-Clustering in these situations helps us learn underlying structure and classifications of our data, and it allows large data sets to be presented in a simplified and compressed view \cite{Jain}.
+% As the data we have access to grows in variety and size and becomes cheaper to store, we need methods to efficiently learn from this data. 
+% Partitioning data (clustering) into similar groups (clusters) is one of the key tools to learn about a set of unlabeled data such as a set of customers who have visited a website or a set of handwritten documents.
+% Clustering in these situations helps us learn underlying structure and classifications of our data, and it allows large data sets to be presented in a simplified and compressed view \cite{Jain}.
+
+[INTRODUCE HIERARCHICAL CLUSTERING]
 
@@
 \[ f(S) = \sum_{ i\in X} \min_{j \in S} d(i,j). \]
-The clustering is obtained from the set of centers $S$ by assigning each data point to its closest center.
+The clustering is obtained from the set of centers $S$ by assigning each data point to its nearest center.
 An $\alpha$-approximation algorithm for this problem produces a set $S$ in polynomial time such that $f(S) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
@@
 \label{fig_hier}}}
-\caption{A partitional clustering for $k=3$, where the colors represent the different partitions, and a dendrogram representation of a hierarchical clustering of the data where edges represent merges of clusters and the clustering on the third level down (represented by the red line) corresponds to the partitional clustering on the left.}
+\caption{A partitional clustering for $k=3$, where the colors represent the different partitions, and a dendrogram representation of a hierarchical clustering of the data, where edges represent mergeing of clusters.
+Note the clustering on the third level (represented by the red line) corresponds to the partitional clustering on the left.}
 \end{figure*}
@@
 \subsection{Partitional Clustering}
-Arguably the most common partitional clustering method is $k$-means, which chooses $k$ cluster centers points in space to minimize the sum of squared distances from each point to its closest cluster center \cite{Jain}.
-$k$-means is usually solved using Lloyd's algorithm (also referred to simply as the $k$-means algorithm), which iteratively greedily updates the centers of each cluster (assigning points to the closest center) until the solution becomes stable \cite{Jain}.
+Arguably the most common partitional clustering method is $k$-means, which chooses $k$ cluster centers in space so as to minimize the sum of squared distances from each point to its nearest cluster center \cite{Jain}.
+$k$-means is usually solved using Lloyd's algorithm (also referred to simply as the $k$-means algorithm), which iteratively greedily updates the centers of each cluster (assigning points to the nearest center) until the solution becomes stable \cite{Jain}.
 Lloyd's algorithm is only guaranteed to reach a locally optimum solution and can be very sensitive to outliers and the initial points chosen \cite{Kanungo}.
@@
 The $k$-median objective function removes some worst-case behavior of $k$-means by restricting the cluster centers to be data points.
-There are two common algorithms for this problem.
-The first is a local search algorithm called partition around medoids (PAM), which starts with $k$ randomly chosen data points as centers, and looks to swap one of the centers with another data point to improve the objective value.
+Two common algorithms for this problem are partition around medoids (PAM), and $k$-medoids.
+PAM is a local seach algorithm which starts with $k$ randomly chosen data points as centers, and looks to swap one of the centers with another data point to improve the objective value.
 This repeats until no more improving swaps exist \cite{ESL}.
 The generalization of this algorithm which considers swaps of size $\leq p$ is a $(3 + \frac{2}{p})$-approximation algorithm \cite{Arya}.
-Another common heuristic (called $k$-medoids throughout this paper), which is more computationally efficient than PAM \cite{Park}, is a variation on Lloyd's $k$-means algorithm.
-Starting with randomly chosen centers, we assign each data point to its closest center, and then update the center within each cluster \cite{Park}.
+$k$-medoids, which is more computationally efficient than PAM \cite{Park}, is a variation on Lloyd's $k$-means algorithm.
+Starting with randomly chosen centers, we assign each data point to its nearest center, and then update the center within each cluster \cite{Park}.
 The algorithm repeats this until the solution becomes stable.
 
-The $k$-median problem is also a fundamental problem in approximation algorithms where it is often presented in the context of opening service centers or factories to serve clients.
+The $k$-median problem is also a fundamental problem in approximation algorithms, where it is often presented in the context of opening service centers or factories to serve clients.
 However, these algorithms are often not implemented for clustering despite their improved guarantees on the solution.
@@
 The most common method for hierarchical clustering is agglomerative clustering.
-Agglomerative clustering starts with each data point in its own cluster, and proceeds to recursively merge the two closest clusters to form the next clustering level in the hierarchy, repeating until the entire data set is a single cluster.
-There are several choices on how to measure distance in order to choose which clusters to merge.
+The agglomerative clustering procedure starts with each data point in its own cluster, and recursively merges the two nearest clusters to form the next clustering level in the hierarchy, repeating until the entire data set is a single cluster.
+There are several choices on how to measure distance in order to choose the nearest clusters to merge.
 Single linkage (also called nearest neighbor) defines the distance between two clusters to be the minimum pairwise distance between the data points in each cluster.
@@
 This was improved by Lin et al.~\cite{Lin} who gave an incremental approximation algorithm framework.
-In particular, their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to greedily construct a hierarchical clustering that increases the objective value by at most a factor of $20.71$ (or $10.03$ when randomized) at each level.
+Their approach relies on a black box $k$-median solver and then uses these solutions for $k=1, \ldots, n$ to greedily construct a hierarchical clustering that increases the objective value by at most a factor of $20.71$ (or $10.03$ when randomized) at each level.
 Using the local $p$-swap $(3+\frac{2}{p})$-approximation algorithm mentioned above as the black-box $k$-median solver performed well in experiments on hierarchical $k$-median test problems \cite{Nagarajan}.
@@
 Recall the hierarchical $k$-median problem: given a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$, we represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
- A solution to the hierarchical $k$-median problem is a chain of nested solutions $(S_1, a_1) \subseteq (S_2, a_2) \subseteq \ldots \subseteq (S_l, a_l)$.
+ A solution to the hierarchical $k$-median problem is a chain of nested solutions $(S_1, a_1) \subseteq (S_2, a_2) \subseteq \ldots \subseteq (S_n, a_n)$.
 However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
@@
 
-First, set $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we switch $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is chosen from a uniform distribution on $[0,1]$).
+First, set $\beta = 3+\sqrt{3}$, and $\beta_0 = 1$ (for the randomized algorithm we instead let $\beta = 6.355$ and set $\beta_0 = \beta^X$, where $X$ is drawn uniformly from $[0,1]$).
 Let $V_i$ be the solutions generated from running $\mathcal{A}$ for $k=n-i+1$ for $i=1, \ldots, n$.
@@
 From each non-empty bucket, pick the solution with the smallest number of cluster centers, and let these solutions be $\overline{V}_1, \overline{V}_2, \ldots \overline{V}_r$.
-We then inductively build our nested solutions: Set $S_1 = \overline{V}_1$ and $a_1$ to be the assignment that assigns each point to the closest point in $S_1$.
-Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we construct $S_i$ by  adding the closest point $j \in S_{i-1}$ to $l$ for each $l \in \overline{V}_i$.
+We then inductively build our nested solutions: Set $S_1 = \overline{V}_1$ and $a_1$ to be the assignment that assigns each point to the nearest point in $S_1$.
+Given $(S_{i-1}, a_{i-1})$ and $\overline{V}_i$, we construct $S_i$ by  adding the nearest point $j \in S_{i-1}$ to $l$ for each $l \in \overline{V}_i$.
 Note that $|S_{i}| \leq |\overline{V}_i|$.
@@
 \label{fig_Siai}}
-\caption{An example of nesting $\bar{V}_i$ with $(S_{i-1}, a_{i-1})$.
-Dark  nodes represent chosen centers and edges represent assignments to that center's cluster.
-First, node 2 chooses node from 1 from $S_{i-1}$ as its closest center and node 5 chooses node 4.
-Then, each cluster in $(S_{i-1}, a_{i-1})$ is assigned to the cheapest (lowest sum of distances) center.}
+\caption{An example of nesting $\bar{V}_i$ with $(S_{i-1}, a_{i-1})$. Dark  nodes represent chosen centers and edges represent assignments to that center's cluster. }
 \label{fig_sim}
 \end{figure*}
+% First, node 2 chooses node from 1 from $S_{i-1}$ as its nearest center and node 5 chooses node 4.
+Then, each cluster in $(S_{i-1}, a_{i-1})$ is assigned to the cheapest (lowest sum of distances) center.
 
@@
 For each $j \in X$ such that $a_{i-1}(j) \in S_{i}$, we set $a_i(j) = a_{i-1}(j)$.
-Otherwise, we set $a_i(j)$ to be the facility $i \in S_i$ such that assigning all points in $a_{i-1}^{-1}(j)$ (i.e.
+Otherwise, we set $a_i(j)$ to be the point $i \in S_i$ such that assigning all points in $a_{i-1}^{-1}(j)$ (i.e.
 all points in $j$'s cluster in $S_{i-1}$) would have minimum cost.
@@
 \subsection{Partitional Clustering Algorithms}
-This nesting framework allows us to choose the partitional clustering algorithm $\mathcal{A}$.
-As mentioned above, PAM and $k$-medoids are two potential algorithms.
-PAM has the added benefit of being a 5-approximation algorithm for the partitional $k$-median problem which yields to an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
+Since the described nesting framework allows us to choose the partitional clustering algorithm $\mathcal{A}$, the choice of this algorithm is critical.
+PAM has the added benefit of being a 5-approximation algorithm for the partitional $k$-median problem, which yields to an overall approximation ratio of $20.71 \cdot 5 = 103.55$ for the hierarchical $k$-median problem ($10.03 \cdot 5 = 50.15$ randomized).
 We looked at increasing the size of the local swaps in PAM to improve the approximation ratio but found that the approximation ratio of constructed solutions for given values of $k$ did not significantly improve while the computation time increased greatly.
@@
 We also consider three linear programming based algorithms, one of which has an approximation guarantee.
-We first present the linear programming formulation for the $k$-median problem.
+We first present the integer programming formulation for the $k$-median problem.
 Suppose we have a set $X$.
-Let $s_i \in \{0,1\}$ be the variable representing whether or not we want to make $i$ a center for all $i \in X$.
-We also create a variable $a_{i,j} \in \{0, 1\}$ which represents whether or not we assign $i \in X$ to the cluster with center $j \in X$.
-Given this set and assignment, the $k$-median objective function is
+Let for all $i \in X$, $s_i \in \{0,1\}$ be variables representing whether or not we want to make $i$ a center.
+We also create variables $a_{i,j} \in \{0, 1\}$ which represent whether or not we assign $i \in X$ to the cluster with center $j \in X$.
+Given this set and assignment, and distance metric $d$, the $k$-median objective function is
 \begin{equation}
@@
 which we want to minimize.
-However, we are restricted by the fact that we can choose at most $k$ centers, every point must be assigned to a single cluster center, and that we can only assign a point to a center that is actually in $S$.
-This is represented by the following three equations, respectively:
+However, we are restricted to choosing at most $k$ centers, every point must be assigned to a single cluster center, and we can only assign a point to another point that has actually been chosen as a center.
+These restrictions are represented by the following constraints, respectively:
 \begin{eqnarray}
@@
 & s_i \in \{0,1\} \quad \forall i \in X \\
-& a_{i,j} \in \{0,1\} \quad \forall i, j \in X
+& a_{i,j} \in \{0,1\} \quad \forall i, j \in X.
 \end{eqnarray}
-Relaxing the constraints so that $s_i \in [0,1]$ for all $i \in X$ and $a_{i,j} \in [0,1]$ for all $i,j \in X$ yields the linear programming (LP) relaxation for the $k$-median problem, which we can solve in polynomial time.
+Relaxing the constraints to $s_i \in [0,1]$ for all $i \in X$ and $a_{i,j} \in [0,1]$ for all $i,j \in X$ yields the linear programming (LP) relaxation for the $k$-median problem, which can be solved in polynomial time.
 
-Interestingly, the solution to the linear programming is often integer valued \cite{Nagarajan}.
-Based on this fact, we propose two LP-based clustering algorithms: one that solves the LP relaxation and constructs $S$ by taking the points in $X$ with the $k$ highest $s_i$ values and one that solves the LP relaxation and constructs $S$ by choosing $i$ to be in $S$ with probability $s_i$, sampling for each $i \in X$ until $k$ centers are chosen.
+Interestingly, the solution to the LP relaxation is often integer valued \cite{Nagarajan}.
+In light of this, we propose two LP-based clustering algorithms that both first solve the LP relaxation.
+The first proposed algorithm constructs the set of centers $S$ by taking the points in $X$ with the $k$ highest $s_i$ values.
+The second constructs $S$ by choosing $i$ to be in $S$ with probability $s_i$, sampling for each $i \in X$ until $k$ centers are chosen.
 We refer to these algorithms as LP-greedy and LP-rand, respectively.
-Charikar and Li \cite{Charikar} introduce a 3.25-approximation algorithm that uses this LP solution to smartly construct a set of centers.
+
+Charikar and Li \cite{Charikar} introduce a 3.25-approximation algorithm that uses LP relaxation solution to smartly construct a set of centers.
 We refer to this algorithm as LP-Charikar.
-We choose to implement this LP algorithm rather than the LP-based approximation algorithm with best approximation ratio because the algorithm is simpler to implement (and we expect it to have a lower running time).
+We chose to implement this LP algorithm, rather than the LP-based approximation algorithm with best approximation ratio, because LP-Charikar is simpler to implement, and expected to have a lower running time.
 
-The algorithm works by first choosing a subset of data points $C \subseteq X$ to represent the demand.
+[MAYBE CUT THE ALGORITHM DESCRIPTION AND SAY TO REFER TO THE ORIGINAL PAPER?]
+The algorithm first chooses a subset of data points $C \subseteq X$ to represent the demand.
 For each data point $i$, let $d_{av}(i) = \sum_{j \in X : a_{i,j} > 0} s_j \cdot d(i,j)$.
@@
 All code and the full computational results are available at \url{https://github.com/apaul29/Hierarchical-Code}.
-We chose to test our methods on data from a few different contexts:
+We tested the algorithms on the following data sets:
 \begin{LaTeXdescription}
@@
 \subsection{Partitional Algorithms}
- For the black-box $k$-median algorithm, we implemented $k$-medoids, PAM, LP-greedy, LP-rand, and LP-Charikar in Python using Gurobi 6.0.3 to solve the linear programming instances. 
+For $k$-median algorithms, we implemented $k$-medoids, PAM, LP-greedy, LP-rand, and LP-Charikar in Python, using Gurobi 6.0.3 to solve the linear programs. 
 To avoid cycling, we limited the number of iterations in $k$-medoids to 1000.
-For PAM, we also needed to limit the number of swaps to achieve a reasonable runtime.
-Instead of searching over all potential swaps, we sample at most $n$ potential ones.
+For PAM, we also limit the number of swaps to achieve a reasonable runtime.
+Instead of searching over all potential swaps, we sample at most $n$ potential swaps.
 We also ensure that some minimum progress is achieved by ensuring the algorithm decreases in objective by at least a factor of 0.9 in each iteration.
 At the end of PAM, we run 10 iterations of $k$-medoids to ensure the algorithm finds a local optimum.
-These choices were made by looking at the relative performance and computation time on random Gaussian data not in the data sets above.
+These choices were made by looking at the relative performance and computation time on random Gaussian data not included in the data sets described above.
 
@@
 \subsection{Hierarchical Results}
-After testing the partitional algorithms above, we implemented the nesting hierarchical framework, offering options for both the randomized and deterministic versions.
+We implemented both the determinisic and randomized versions of the nesting hierarchical framework.
 For PAM and $k$-medoids, when solving for multiple values of $k$, we use the solution with $k$ centers as a warm start for $k+l$ centers by choosing an additional random $l$ centers.
@@
 This yields many fewer $k$-median instances to solve, reducing the run-time significantly.
-This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small so we expect our sampling to still represent much of the underlying structure.
-For PAM, we only tested the logarithmic heuristic since, as also indicated by the partitional results, the run time for the full version proved impractical.
+This was motivated by the fact that the difference between various $k$-median solutions will be larger when $k$ is small, so we expect our sampling to still represent much of the underlying structure.
+For PAM, we only tested the logarithmic heuristic since the run time for the full version proved impractical.
 
@@
 
-Let $n$ be the number of data points, let $L_1, L_2 \ldots, L_c$ be the true classes of the data, and let $\mathcal{C}$ all clusters represented on any level of our hierarchical clustering.
+Let $n$ be the number of data points, let $L_1, L_2 \ldots, L_c$ be the true classes of the data, and let $\mathcal{C}$ be all clusters represented on any level of our hierarchical clustering.
 For any class $L_r$ and cluster $C_i$, let $n_r$ be the number of data points in $L_r$ and $n_i$ be the number of data points in $C_i$.
 Then, let $n_{ri}$ be the number of data points in $C_i$ from class $L_r$.
-We define the precision of $C_i$ to be $P(L_r, C_i) = n_{ri}/n_r$ and the recall of $C_i$ to be $R(L_r, C_i) = n_{ri}/n_r$.
-The harmonic mean of these two metrics gives a balanced measure of how accurately $S_i$ represents $L_r$.
+We define the precision of $C_i$ to be $P(L_r, C_i) = n_{ri}/n_i$ and the recall of $C_i$ to be $R(L_r, C_i) = n_{ri}/n_r$.
+The harmonic mean of these two metrics gives a balanced measure of how accurately $C_i$ represents $L_r$.
 The FScore of $L_r$ is then the best measure over all clusters
@@
 There were several interesting trends in the performance.
-First, the deterministic versions had higher FScore measures and lower approximation coefficients on average compared to their randomized counterparts so we would advise to just use the deterministic version despite its higher approximation guarantee.
-Similarly, our simple LP heuristics performed comparably to the more advanced LP-Charikar algorithm.
-Also worth noting is the fact that the approximation ratio for $k$-medoids improved as $n$ increased in the $p$-median instances, which is surprising the values of $k$ for some of the instances increased proporitional to $n$ so we would expect the overall difficulty to be higher.
+First, the deterministic versions had higher FScore measures and lower approximation coefficients on average compared to their randomized counterparts, suggesting usage of deterministic version in practice despite its higher approximation guarantee.
+Similarly, our relatively simpler LP heuristics performed comparably to the more advanced LP-Charikar algorithm.
+Also worth noting is the fact that the approximation ratio for $k$-medoids improved as $n$ increased in the $p$-median instances, which is surprising, as the values of $k$ for some of the instances increased proporitional to $n$, so we would expect the overall difficulty to be higher.
 
-In terms of the computation time, the agglomerative clustering methods were significantly faster than all of nesting algorithms, which we would expect given how many $k$-median instances they have to solve.
-$k$-medoids was the fastest method, but $k$-medoids, LP-greedy, and LP-rand all performed around 4 orders of magnitude slower than the agglomerative methods (this order increased to 5 for a few of the very large $p$-median instances).
+In terms of computation time, the agglomerative clustering methods were significantly faster than all of nesting algorithms, which is to be expected given the numerous $k$-median instances the nesting algorithms must solve.
+$k$-medoids, LP-greedy, and LP-rand all performed around 4 orders of magnitude slower than the agglomerative methods (this order increased to 5 for a few of the very large $p$-median instances), with $k$-medoids being the fastest.
 LP-Charikar was the slowest method and was often at least 5 orders of magnitude slower than the agglomerative methods.
 
-In the logarithmic version, the computation times all improved.
+In the logarithmic version, computation times improved.
 For $k$-medoids, the algorithm became 2-3 orders of magnitude slower than agglomerative clustering while the LP methods all decreased by 1 order of magnitude.
-However, LP-rand now had the worst running time whereas LP-Charikar improved to be similar to LP-greedy.
-PAM, on the other hand, was 4-5 orders of magnitude higher, similar to the methods used in the full hierarchical framework.
+PAM was 4-5 orders of magnitude slower than agglomerative clustering, similar to the methods used in the full hierarchical framework.
 
-Using the logarithmic version did decrease the FScores slightly, but the averages still remained close to 0.6 and above the agglomerative methods.
-Again, the algorithms performed similarly in FScore.
-The approximation ratios, on the other hand, did change significantly in the logarithmic version with the $k$-medoids algorithm having the highest increases in ratio.
+Using the logarithmic version did decrease the FScores slightly, but the averages still remained close to 0.6, higher than the agglomerative methods.
+Again, the different partitional algorithms performed similarly in FScore.
+The approximation ratios, on the other hand, did change significantly in the logarithmic version, with the $k$-medoids algorithm having the highest increases in ratio.
 PAM had an average Gaussian approximation ratio of around 8.5, LP-Charikar and LP-greedy around 7, and LP-rand around 5 (the ratios did improve as the dimension increased).
@@
 The algorithms also seemed to do worse on the Iris data set and not pick up on the optimal solution for the higher levels.
-This indicates that if a data set has a very steep change in objective function between neighboring solutions that the full hierarchical version should be used.
+This indicates that if a data set has a very steep change in objective function between neighboring solutions, the full hierarchical version should be used.
 
-Given these performances, we would recommend using the deterministic, logarithmic framework with a simple LP heuristic as the black-box $k$-median solver.
+Given these performance results, we would recommend using the deterministic, logarithmic framework with a simple LP heuristic as the black-box $k$-median solver.
 If one does not want to use an LP solver, then $k$-medoids is a good choice given the low running times but the approximation ratio might increase.
-Switching to the full hierarchical version does increase the run time by around one order of magnitude, but offers better FScore and approximation ratios.
+Using the full hierarchical version does increase the run time by around one order of magnitude, but offers better FScore and approximation ratios.
 Using PAM seems to be the least beneficial option since the full hierarchical version with $k$-medoids as a solver has better overall performance.
@@
 \section{Conclusion and Future Work}
-We implemented the nesting framework introduced by Lin et al.~\cite{Lin}, which focuses on minimizing the hierarchical $k$-median objective function, to produce hierarchical clusterings using various methods for constructing the inputted partitional clusterings.
+We implemented the nesting framework introduced by Lin et al.~\cite{Lin} to produce hierarchical clusterings using various methods for constructing the inputted partitional clusterings.
 Overall, we saw significant improvement in FScore measures for our implemented methods compared to agglomerative clustering and overall good approximation ratios where none were had before.

### 2015-06-02 icdm/submission.tex
Simplified introduction

@@
 \section{Introduction and Notation}
-% As the data we have access to grows in variety and size and becomes cheaper to store, we need methods to efficiently learn from this data. 
-% Partitioning data (clustering) into similar groups (clusters) is one of the key tools to learn about a set of unlabeled data such as a set of customers who have visited a website or a set of handwritten documents.
-% Clustering in these situations helps us learn underlying structure and classifications of our data, and it allows large data sets to be presented in a simplified and compressed view \cite{Jain}.
+Clustering algorithms date back more than 50 years, and there have been many variations introduced since \cite{Jain}.
+The two main branches of clustering are partitional (or flat) and hierarchical.
+In partitional clustering, we often specify a number $k$ and pairwise distances between the data points, representing a measure of dissimilarity such as Euclidean distance.
+A clustering algorithm thus partitions the data into $k$ disjoint sets where the sets represent relatively similar data.
+For hierarchical clustering, we are interested in finding a nested tree of partitions.
+This offers the user a global view of the data and allows the user to choose a ``natural'' level of clustering \cite{ESL}.
+Examples of both partitional and hierarchical clustering are given in Fig.~\ref{fig_part}-\ref{fig_hier}.
 
-[INTRODUCE HIERARCHICAL CLUSTERING]
-
-There are two main branches of clustering: partitional (or flat) and hierarchical.
-In partitional clustering, we often specify a number $k$ and pairwise distances between the data points, representing a measure of dissimilarity such as Euclidean distance.
-The clustering algorithm then partitions the data into $k$ disjoint sets where the sets represent relatively similar data.
-In hierarchical clustering, on the other hand, we are interested in finding a nested tree of partitions.
-This offers the user a global view of the data and allows the user to choose a ``natural'' level of clustering \cite{ESL}.
-For example, a retailer could use a hierarchical clustering of customers to first find regions of customers in order to tailor the products offered in each region.
-The retailer could then also offer specialized discounts to subgroups in each region based on each cluster's shopping records.
-Examples of both types of clustering are given in Fig.~\ref{fig_part}-\ref{fig_hier}.
-
-Clustering algorithms date back more than 50 years, and there have been many variations introduced since then \cite{Jain}.
-However, clustering algorithms are often formulated as NP-hard problems with non-convex objective functions, which are inherently hard to solve \cite{Jain}.
 One common objective function for partitional clustering is to minimize the sum of distances from each point to the median point of its cluster, called $k$-median or $k$-medoids clustering.
@@
 
-This objective function can then be extended to the hierarchical $k$-median problem which represents an objective function relating to the hierarchical clustering problem.
+This objective function can then be extended to the hierarchical $k$-median problem.
 Suppose we represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$.
@@
 We define $f(S,a) := \sum_{i \in X} d(i, a(i))$.
-This is the same objective function as above except the assignment is given rather than implied.
+Note this is the same objective function as above, except the assignment is given rather than implied.
 
@@
 An example is given in Fig.~\ref{fig_chain1}-\ref{fig_chain4}.
-However, we want this solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
+
+We would like a nested hierachical solution to stay close to the optimal $k$-median solutions when looking at any level of the hierarchy.
 An $\alpha$-approximation algorithm for this problem generates a nested chain of solutions in polynomial time such that for all $k = 1, \ldots, |X|$ there exists $(S_i, a_i)$ in the chain such that $|S_i| \leq k$ and $f(S_i, a_i) \leq \alpha \cdot f(O)$ for all $O \subseteq X$ with $|O| = k$.
